### 2020-01-21 manuscript.tex
initial commit

\documentclass[12pt, b5paper,twoside]{tesi_upf}

%CODIFICATION
\usepackage[latin1]{inputenc}

%LENGUAGE
\usepackage[english,catalan]{babel}

%ONLY TO OBTAIN MARK BANK INDEX INDICATION B5
\usepackage[b5paper]{geometry}
\usepackage[cam,a4,center,frame]{crop}

%INCLUDE GRAPHICS AND THE LOGO OF THE UPF
\usepackage{graphicx}

%FONTS TIMES OR GARAMOND, 
\usepackage{times}
%\usepackage{garamond}

%WITHOUT HEADINGS: NO MODIFICATION
\pagestyle{plain}

%FOR THE INDEX OF SUBJECTS
\usepackage{makeidx}
\makeindex

%BIBLIOGRAPHY STYLE
\bibliographystyle{apalike}

%SELECT LANGUAGE
\selectlanguage{english}

%THE TABLE OF CONTENTS IS TITLE CONTENTS
\addto\captionscatalan
  {\renewcommand{\contentsname}{\Large \sffamily Sumari}}

%ADD YOUR DATA
\title{The title of the thesis:Required}
\subtitle{The subtitle of the thesis: Required}
\author{Author: Required}
\thyear{of the thesis: Required}
\department{Required}
\supervisor{Required}

% TODO
\newcommand\todo[1]{\textcolor{red}{#1}}

\begin{document}

\frontmatter

\maketitle

\cleardoublepage

%%%%%% Dedication

\noindent Write here your dedication

\cleardoublepage

%%%%%% End dedication

%%%%%% Thanks
\noindent {\Large \sffamily Thanks} thanks to....

\cleardoublepage

%%%%%% End of thanks

%ABSTRACT IN TWO LEGUAGES.
\selectlanguage{english}
\section*{\Large \sffamily Abstract}
This is the abstract of the thesis in English.  Please, use less
than 150 words.

\selectlanguage{catalan}
\vspace*{\fill}
\section*{\Large \sffamily  Resum}
Vet aquí el resum de la tesi en català.  Si us plau, utilitzeu
menys de 150 paraules.
\vspace*{\fill}

\cleardoublepage
%END OF ABSTRACT

%PREFACE. 
{\bf Preface}

\cleardoublepage
%END OF PREFACE

%TABLE OF CONTENTS: REQUIRED
\tableofcontents

%lIST OF FIGURES; ONLY IF THERE ARE FIGURES
\listoffigures
%TO APPER THE LIST OF FIGURES IN THE TABLE OF CONTENTS 
\addcontentsline{toc}{chapter}{List of figures}

%LIST OF TABLES; ONLY IF THERE ARE TABLES
\listoftables
%TO APPEAR THE LIST OF TABLES IN THE TABLE OF CONTENTS
\addcontentsline{toc}{chapter}{List of tables}

%START THE TEXT
\mainmatter
\chapter*{Introduction}

This is a template or provisional style to be used to present doctoral theses at the UPF in format B5. It has been tried to simplify it to the maximum so that experienced users can easily adapt their own styles and that less-used people to work with styles can easily use it.

Has
\begin{description}
\item[file tesi-upf.cls] It is a modification of the style {\tt book} with the following particularities
  \begin{enumerate}
  \item The cover is redesigned (the instruction\verb+\maketitle+).

  \item The titles of the chapters are written in capital letters.

  \item Is redesigned {\tt cleardoublepage} because the blank pages are not numbered.
  \end{enumerate}

\item[The preamble] Has diferent instructions
  \begin{enumerate}
  \item The combination of the packages {\tt geometry} i {\tt crop} It allows framing what will be the final page that will then be cut to size {\tt B5}. Before cutting it, it is advisable to remove the marks, commenting on the corresponding orders, unless the printing company asks us to do so.

  \item In the event that the thesis is in Catalan, the title of `` Index '' for `` Summary '' is redefined.

  \item The package is used {\tt times} to use this font.

  \item It indicates \verb+\pagestyle{plain}+ so that the headings do not appear.

  \item  To the orders {\tt title}, {\tt subtitle}, {\tt author}, {\tt thyear},{\tt department}, {\tt supervisor} the appropriate values ??must be put. \LaTeX will generate the cover from them.
  \end{enumerate}

\item[packages] For the convenience of the user the packages are incorporated {\tt crop} i {\tt geometry}.

\item[Bibliography] o check the operation, a small bibliographic database is incorporated, named  \verb+tesi_upf_bib.bib+

\item[Tables] It incorporates a \verb+figure+ i un \verb+tabular+ to check the operation of the indexes.

\end{description}

\section*{Index}

It is advisable to use the index with entries like the following, which refers to the inventor of \LaTeX:  \verb+\index{Lamport}+ \index{Lamport}. 
\begin{enumerate}
\item The orders \verb+\usepackage{makeidx} \makeindex+ in the preamble of the document and \verb+\printindex+ In the place where we want the references to appear, they cause the index to be generated. To generate it effectively you must use the program \verb+makeindex+ (it is done in a different way according to the distribution and the editor we use).
\end{enumerate}

\section*{The bibliography}
In order to be able to efficiently use bibliographic references, it is convenient to use the BibTeX program. To do it 
%\begin{enumerate}
%\item We write in the preamble \verb+\bibliographystyle{plain}+ where instead of \verb+plain+ It can be some of the usual styles, for example \verb+vancouver+\ldots Vancouver style is also included \verb+vancouver+.
%
%\item On the site of the document where we want the list of references to appear, we write \verb+ \bibliography{tesi_upf_bib}
%
%
%\end{enumerate}

\chapter{The first chapter}
Example of the first chapter using some section and subsection.\footnote{Example of footnote} Note that you can have the chapters in separate files and use them \verb+\input{}+.

\todo{THIS IS MY TODO}.

\section{A section}
Specifically, the first one, which contains a table

\begin{table}[h]
  \centering
  \begin{tabular}{|l|l|}
   \hline
    0 & 0 \\ \hline
    0 & 0 \\ \hline    
  \end{tabular}
  \caption{test}

\end{table}

\subsection{A subsection}
The first subsection of the first section.

\section{Another section}
the second one

It incorporates a figure with the logo of the UPF (Figura \ref{fig:logo}) {so that it goes down to the index of figures.
\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{logo_upf.png}
    \caption{Graphic example}
    \label{fig:logo}
\end{figure}

\chapter{The second chapter}

Lorem ipsum dicam tation maiorum vel ea, an nam minimum pertinax vituperatoribus, mea no illud tibique nostrum. Odio aperiri dolorem in sit, utinam labore adipiscing cu vel. Nam porro choro cu, nam virtute maluisset voluptatum in, per te vero iuvaret praesent. Eos ut volumus pertinacia interesset, tota debitis abhorreant mei et, clita electram pri at. Ea fabulas admodum theophrastus sea, id nostro omnium periculis sed. Vim id nulla aliquando. Id mel habeo nostrud mandamus, laudem legimus fastidii quo ea, ut iriure utamur quo.

Cu sit nobis omnesque scripserit, nec albucius perfecto salutatus ex. Novum debet propriae vis cu, sed ea suscipit accusata, nam minim salutandi voluptatum ut. Quaerendum definitiones est ut, no quo postea prompta scaevola. Nec at choro soleat causae, per ut docendi denique, discere veritus no sed. Ei sed sale delicatissimi mediocritatem, eligendi deterruisset ut mei, ignota offendit ocurreret at mel. Numquam blandit gubergren an pro, mei cu quot dolore recteque. Suscipit aliquyam principes eam ut, id error dicit perpetua qui.

Eam mentitum noluisse instructior cu, ad scripta nonummy aliquyam eos. Augue possit gloriatur mel an, mea et dolorum suavitate assueverit. At ius dolores persecuti moderatius, id qui justo tantas dissentiet. Integre invenire imperdiet ut mea, harum alienum officiis te nam, integre qualisque conceptam in vis. Viderer consetetur vim id, sumo tempor complectitur vim an. His te aeterno tamquam integre, nec agam aliquando repudiandae id, te fabellas verterem sea.

Per dicit veniam delicata ut, dicunt corpora mei eu. His no alienum accusata accommodare, eu nam clita audire invidunt. Mutat debitis mediocrem duo te, ex vis prima labore inermis. Usu eu dictas possim vocent. Suas meliore ancillae ad eum, eam ei cibo viderer similique.

Usu ne apeirian deserunt expetendis, sed in iisque abhorreant definiebas, quo nulla inermis appareat cu. Ei nam possim docendi pertinax, ut duis veritus detracto pri. Animal percipit ius ne, ne iisque quaeque philosophia pro, ut sed nobis nullam adolescens. Persius principes omittantur per an. Mel quot nostrud labores at, ad vis latine persius petentium. Cum partiendo neglegentur mediocritatem ut, per simul aliquando tincidunt no, falli nominavi molestiae ex nam.

Saperet inciderint in vim, vis quod civibus accusata in. Et ullum facilisi accommodare nec, decore utroque definitionem an vel, id duo kasd facer latine. Ea sumo ullamcorper per, ex pro nobis efficiendi concludaturque, vix eu verear luptatum ocurreret. Ocurreret intellegat te eum, ius ea cibo epicurei dignissim. Diam elit quaeque cum ut. Est dicat summo prodesset te, no vel harum saepe sensibus, et ignota altera doming qui.

Sit altera regione feugiat in, vim ad augue mediocritatem, duo reque choro theophrastus ea. Ipsum apeirian voluptatum duo no, nam ut ornatus accumsan percipitur, est no posse suscipiantur vituperatoribus. Errem legendos partiendo est cu, est clita signiferumque cu. Ea nulla scripta signiferumque eam, sea vidit dolores officiis ei, nibh saepe in nam. Ad vim justo dolor reprimique, eum ad labitur dissentias scriptorem, cu tempor mentitum assentior eum. Id vis mucius percipit invidunt.

Vel at ullum intellegam constituam, mea dicta affert perpetua ea. Eripuit blandit iracundia nec ea, tale timeam molestie cum te. Cum ei kasd mazim, per ei delenit ullamcorper, ad sale vocibus temporibus usu. An nonummy oporteat phaedrum sed, eu unum deserunt incorrupte mei. His an possim delicata, ei accumsan constituam mea. Ei viris constituam pro, cu inimicus argumentum duo.

Cu regione veritus persequeris vis, tation nusquam percipit in pri, vis no labitur deseruisse. Ius in reque fierent. Quo et erant recusabo honestatis. Iusto accusam liberavisse ex mei, ut cum nobis primis accommodare, viderer vocibus assueverit ut nec. Id odio oporteat instructior pri. Nam sint illud vituperatoribus no, zzril ubique dissentias id nec.

Cu est legimus reformidans. Nam cibo mucius dolores ad, prima posse similique vix te. Mea at nemore mandamus assueverit, sit viris option scaevola cu. At omnis decore nostro pro, aliquyam explicari salutatus ei mea, sea no falli incorrupte. Graeco bonorum lobortis ut sed, ut pro discere legimus, an sea volumus consetetur.

\bibliography{exemple_bibtex}

\backmatter
\printindex

\end{document}

%NUMBER OF THE EXTERNAL PAGE EXCEPT IN THE FIRST PAGE OF EACH CAPITAL
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot{}
\fancyfoot[RO]{\thepage}
\fancyfoot[LE]{\thepage}

%MULTIPLE INDEX
%In the preamble
\usepackage{multind}
\makeindex{authors}
%Introduction to form entries
\index{authors}{Einstein}
%Situation of the Index
\printindex{authors}{Author index}
%The \ usepakage {makeidx} \ makeindex \ printindex commands must be removed
%You need to exacute from the command line makeindex authors

### 2020-01-21 manuscript-encoding-backup.tex
update

--- 2020-01-21 manuscript.tex
@@
 %THE TABLE OF CONTENTS IS TITLE CONTENTS
-\addto\captionscatalan
-  {\renewcommand{\contentsname}{\Large \sffamily Sumari}}
+%\addto\captionscatalan
+  {\renewcommand{\contentsname}{\Large \sffamily Contents}}
 
 %ADD YOUR DATA
-\title{The title of the thesis:Required}
-\subtitle{The subtitle of the thesis: Required}
-\author{Author: Required}
-\thyear{of the thesis: Required}
-\department{Required}
-\supervisor{Required}
+\title{The title of the thesis:\todo{Required}}
+\subtitle{The subtitle of the thesis: \todo{Required}}
+\author{Author: Andrés Pérez López}
+\thyear{2020}
+\department{of Information and Communication Technologies}
+\supervisor{}
+\supervisor{Dr. Emilia Gómez Gutiérrez\\Dr. Adan Garriga Torres}
 
@@
 
-\noindent Write here your dedication
+\noindent Write here your dedication \todo{dedication}
 
@@
 %%%%%% Thanks
-\noindent {\Large \sffamily Thanks} thanks to....
+\noindent {\Large \sffamily Thanks} thanks to....  \todo{thanks}
 
@@
 This is the abstract of the thesis in English.  Please, use less
-than 150 words.
+than 150 words. \todo{abstract in english}
 
@@
 Vet aquí el resum de la tesi en català.  Si us plau, utilitzeu
-menys de 150 paraules.
+menys de 150 paraules. \todo{abstract in catalan}
 \vspace*{\fill}
@@
 %PREFACE. 
-{\bf Preface}
+{\bf Preface \todo{is that really needed?}}
 
@@
 
-It is advisable to use the index with entries like the following, which refers to the inventor of \LaTeX:  \verb+\index{Lamport}+ \index{Lamport}. 
-\begin{enumerate}
-\item The orders \verb+\usepackage{makeidx} \makeindex+ in the preamble of the document and \verb+\printindex+ In the place where we want the references to appear, they cause the index to be generated.
-To generate it effectively you must use the program \verb+makeindex+ (it is done in a different way according to the distribution and the editor we use).
-\end{enumerate}
+blablabla
 
 \section*{The bibliography}
-In order to be able to efficiently use bibliographic references, it is convenient to use the BibTeX program. To do it 
+In order to be able to efficiently use bibliographic references, it is convenient to use the BibTeX program. To do it \cite{sartre1984ser}.
 %\begin{enumerate}
@@
 Example of the first chapter using some section and subsection.\footnote{Example of footnote} Note that you can have the chapters in separate files and use them \verb+\input{}+.
-
-\todo{THIS IS MY TODO}.
 
@@
 %NUMBER OF THE EXTERNAL PAGE EXCEPT IN THE FIRST PAGE OF EACH CAPITAL
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\fancyfoot{}
-\fancyfoot[RO]{\thepage}
-\fancyfoot[LE]{\thepage}
-
-%MULTIPLE INDEX
-%In the preamble
-\usepackage{multind}
-\makeindex{authors}
-%Introduction to form entries
-\index{authors}{Einstein}
-%Situation of the Index
-\printindex{authors}{Author index}
-%The \ usepakage {makeidx} \ makeindex \ printindex commands must be removed
-%You need to exacute from the command line makeindex authors
+%\usepackage{fancyhdr}
+%\pagestyle{fancy}
+%\fancyfoot{}
+%\fancyfoot[RO]{\thepage}
+%\fancyfoot[LE]{\thepage}
+%
+%
+%%MULTIPLE INDEX
+%%In the preamble
+%\usepackage{multind}
+%\makeindex{authors}
+%%Introduction to form entries
+%\index{authors}{Einstein}
+%%Situation of the Index
+%\printindex{authors}{Author index}
+%%The \ usepakage {makeidx} \ makeindex \ printindex commands must be removed
+%%You need to exacute from the command line makeindex authors

### 2020-01-23 Chapters/Conclusions.tex
update

\chapter{Conclusions}

\section{Summary of Contributions}

\begin{itemize}
  \item Academic Contributions

  \begin{enumerate}
  	\item Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone \todo{ref}
  	\begin{itemize}
  		\item \textbf{Parameter estimation}: Contribution to the characterization of coherence with tetrahedral microphones (the most common spherical arrangement)
	\end{itemize}

  	\item Autoregressive B-Format Late IR Estimation \todo{ref}
  	\begin{itemize}
  		\item \textbf{Parameter estimation}: Novel technique for RT60 estimation from autoregressive models (subproduct of dereverberation)
		\item \textbf{Signal enhancement}: Novel methodology to re-reverberate sound scenes (include new elements in the scene using the reverb of the recorded scene)
	\end{itemize}

  	\item A hybrid parametric-deep learning approach for sound event localization and detection \todo{ref}
  	\begin{itemize}
  		\item \textbf{Scene Description}: ovel State-of-the-Art methodology for Sound Event Localization and Detection	
  	\end{itemize}

  \end{enumerate}

  \item Software Contributions

	\begin{enumerate}

  		\item Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes \todo{ref}
  		\begin{itemize}
  			\item \textbf{Data Generation}: Novel tool for reverberant ambisonic dataset generation
		\end{itemize}

  		\item  Ambisonics Directional Room Impulse Response as a New Convention of the Spatially Oriented Format for Acoustics \todo{ref}
		\begin{itemize}
  			\item \textbf{Recorded IRs}: File standard/convention proposal for storage of recorded ambisonic IRs
		\end{itemize}

		\item Multichannel Array Signal Processing library \todo{ref}
		\begin{itemize}
  			\item \textbf{Simulated IRs}: Library for acoustic simulation (IR generation, microphone array simulation, etc)
		\end{itemize}

  		\item pysofaconventions \todo{ref}
		\begin{itemize}
  			\item \textbf{Recorded IRs}: implementation of SOFA for python 
		\end{itemize}

	\end{enumerate}

\end{itemize}

\section{List of related publications}

### 2020-01-23 Chapters/DataGeneration.tex
update

\chapter{Data generation and storage}

Explain about mono files plus ambisonics IRs.

\section{Recorded IRs}

AmbisonicsDRIR: explain problematics, standard, etc.
also: pysofaconventions

\section{Simulated IRs}

explain different methods and libraries.
tell about masp.

\section{High-level scene description}

ambiscaper

### 2020-01-23 Chapters/Introduction.tex
update

\chapter{Introduction}

\section{Motivation}

\subsection{3D arrays}

Sound propagates in 3D: need for 3D mic arrays to capture spatial properties

\subsection{spherical microphone arrays}

\begin{itemize}
  \item even distribution of capsules
  \item mathematical convenience: spherical harmonics
\end{itemize}

\subsection{ambisonics}

advantages on the vr/ar context
\begin{itemize}
  \item device independent
  \item intermediate storage format
  \item signal-independent transformations are easy
  \item de-facto standard for vr
\end{itemize}

\subsection{Current limitations of vr/ar production}

\section{Goals}

Research question: 	How can we exploit the characteristics of ambisonic recordings in order to manipulate them more adequately?

\section{Context}

Different levels of application/contribution:

\begin{itemize}
	\item Acoustic Parameter Estimation (low level, audio2data)
	\begin{itemize}
		\item Direction of Arrival estimation
		\item Coherence analysis
		\item Acoustic description (RT60, etc)
		\item Source counting
	\end{itemize}

	\item Signal Enhancement  (high level, audio2audio)
	\begin{itemize}
		\item Source Separation
		\item Dereverberation / denoising
		\item IR estimation
	\end{itemize}

	\item Scene Description (high level, audio2data)
	\begin{itemize}
		\item Event Detection
		\item Acoustic Scene Classification
	\end{itemize}
\end{itemize}

\begin{figure}[hbt]
  \includegraphics[width=\textwidth]{Figures/context_scheme.png}
  \caption{\todo{todo caption}}
\end{figure}

\todo{introduce rest of chapters?}

### 2020-01-23 Chapters/ScientificBackground.tex
update

\chapter{Scientific Background}

\section{State of the Art}

### 2020-01-23 manuscript.tex
update

@@
 %THE TABLE OF CONTENTS IS TITLE CONTENTS
-\addto\captionscatalan
-  {\renewcommand{\contentsname}{\Large \sffamily Sumari}}
+%\addto\captionscatalan
+  {\renewcommand{\contentsname}{\Large \sffamily Contents}}
 
 %ADD YOUR DATA
-\title{The title of the thesis:Required}
-\subtitle{The subtitle of the thesis: Required}
-\author{Author: Required}
-\thyear{of the thesis: Required}
-\department{Required}
-\supervisor{Required}
+\title{Spherical harmonic domain parametric methods for 3D audio production enhancement\todo{(check capitalization)}}
+\subtitle{\todo{subtitle Required?}}
+\author{Author: Andrés Pérez López}
+\thyear{2020}
+\department{of Information and Communication Technologies}
+\supervisor{Dr. Emilia Gómez GutiérrezDr. Adan Garriga Torres}
 
@@
 \cleardoublepage
+\selectlanguage{english}
 %END OF ABSTRACT
@@
 \mainmatter
-\chapter*{Introduction}
 
-This is a template or provisional style to be used to present doctoral theses at the UPF in format B5.
-It has been tried to simplify it to the maximum so that experienced users can easily adapt their own styles and that less-used people to work with styles can easily use it.
+%INSERT CHAPTERS
+\input{Chapters/Introduction.tex}
+\input{Chapters/ScientificBackground.tex}
+\input{Chapters/CoherenceEstimation}
+\input{Chapters/AutoregressiveReverbEstimation}
+\input{Chapters/SELD}
+\input{Chapters/DataGeneration}
+\input{Chapters/Conclusions}
 
-Has
-\begin{description}
-\item[file tesi-upf.cls] It is a modification of the style {\tt book} with the following particularities
-  \begin{enumerate}
-  \item The cover is redesigned (the instruction\verb+\maketitle+).
-
-  \item The titles of the chapters are written in capital letters.
-
-  \item Is redesigned {\tt cleardoublepage} because the blank pages are not numbered.
-  \end{enumerate}
-
-\item[The preamble] Has diferent instructions
-  \begin{enumerate}
-  \item The combination of the packages {\tt geometry} i {\tt crop} It allows framing what will be the final page that will then be cut to size {\tt B5}.
-Before cutting it, it is advisable to remove the marks, commenting on the corresponding orders, unless the printing company asks us to do so.
-
-  \item In the event that the thesis is in Catalan, the title of `` Index '' for `` Summary '' is redefined.
-
-  \item The package is used {\tt times} to use this font.
-
-  \item It indicates \verb+\pagestyle{plain}+ so that the headings do not appear.
-
-  \item  To the orders {\tt title}, {\tt subtitle}, {\tt author}, {\tt thyear},{\tt department}, {\tt supervisor} the appropriate values ??must be put. \LaTeX will generate the cover from them.
-  \end{enumerate}
-
-\item[packages] For the convenience of the user the packages are incorporated {\tt crop} i {\tt geometry}.
-
-\item[Bibliography] o check the operation, a small bibliographic database is incorporated, named  \verb+tesi_upf_bib.bib+
-
-\item[Tables] It incorporates a \verb+figure+ i un \verb+tabular+ to check the operation of the indexes.
-
-\end{description}
-
-\section*{Index}
-
-It is advisable to use the index with entries like the following, which refers to the inventor of \LaTeX:  \verb+\index{Lamport}+ \index{Lamport}. 
-\begin{enumerate}
-\item The orders \verb+\usepackage{makeidx} \makeindex+ in the preamble of the document and \verb+\printindex+ In the place where we want the references to appear, they cause the index to be generated.
-To generate it effectively you must use the program \verb+makeindex+ (it is done in a different way according to the distribution and the editor we use).
-\end{enumerate}
-
-\section*{The bibliography}
-In order to be able to efficiently use bibliographic references, it is convenient to use the BibTeX program. To do it 
-%\begin{enumerate}
-%\item We write in the preamble \verb+\bibliographystyle{plain}+ where instead of \verb+plain+ It can be some of the usual styles, for example \verb+vancouver+\ldots Vancouver style is also included \verb+vancouver+.
-%
-%\item On the site of the document where we want the list of references to appear, we write \verb+ \bibliography{tesi_upf_bib}
-%
-%
-%\end{enumerate}
-
-\chapter{The first chapter}
-Example of the first chapter using some section and subsection.\footnote{Example of footnote} Note that you can have the chapters in separate files and use them \verb+\input{}+.
-
-\todo{THIS IS MY TODO}.
-
-\section{A section}
-Specifically, the first one, which contains a table
-
-\begin{table}[h]
-  \centering
-  \begin{tabular}{|l|l|}
-   \hline
-    0 & 0 \\ \hline
-    0 & 0 \\ \hline    
-  \end{tabular}
-  \caption{test}
-
-\end{table}
-
-\subsection{A subsection}
-The first subsection of the first section.
-
-\section{Another section}
-the second one
-
-It incorporates a figure with the logo of the UPF (Figura \ref{fig:logo}) {so that it goes down to the index of figures.
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.5]{logo_upf.png}
-    \caption{Graphic example}
-    \label{fig:logo}
-\end{figure}
-
-\chapter{The second chapter}
-
-Lorem ipsum dicam tation maiorum vel ea, an nam minimum pertinax vituperatoribus, mea no illud tibique nostrum.
-Odio aperiri dolorem in sit, utinam labore adipiscing cu vel.
-Nam porro choro cu, nam virtute maluisset voluptatum in, per te vero iuvaret praesent.
-Eos ut volumus pertinacia interesset, tota debitis abhorreant mei et, clita electram pri at.
-Ea fabulas admodum theophrastus sea, id nostro omnium periculis sed.
-Vim id nulla aliquando.
-Id mel habeo nostrud mandamus, laudem legimus fastidii quo ea, ut iriure utamur quo.
-
-Cu sit nobis omnesque scripserit, nec albucius perfecto salutatus ex.
-Novum debet propriae vis cu, sed ea suscipit accusata, nam minim salutandi voluptatum ut.
-Quaerendum definitiones est ut, no quo postea prompta scaevola.
-Nec at choro soleat causae, per ut docendi denique, discere veritus no sed.
-Ei sed sale delicatissimi mediocritatem, eligendi deterruisset ut mei, ignota offendit ocurreret at mel.
-Numquam blandit gubergren an pro, mei cu quot dolore recteque.
-Suscipit aliquyam principes eam ut, id error dicit perpetua qui.
-
-Eam mentitum noluisse instructior cu, ad scripta nonummy aliquyam eos.
-Augue possit gloriatur mel an, mea et dolorum suavitate assueverit.
-At ius dolores persecuti moderatius, id qui justo tantas dissentiet.
-Integre invenire imperdiet ut mea, harum alienum officiis te nam, integre qualisque conceptam in vis.
-Viderer consetetur vim id, sumo tempor complectitur vim an.
-His te aeterno tamquam integre, nec agam aliquando repudiandae id, te fabellas verterem sea.
-
-Per dicit veniam delicata ut, dicunt corpora mei eu.
-His no alienum accusata accommodare, eu nam clita audire invidunt.
-Mutat debitis mediocrem duo te, ex vis prima labore inermis.
-Usu eu dictas possim vocent.
-Suas meliore ancillae ad eum, eam ei cibo viderer similique.
-
-Usu ne apeirian deserunt expetendis, sed in iisque abhorreant definiebas, quo nulla inermis appareat cu.
-Ei nam possim docendi pertinax, ut duis veritus detracto pri.
-Animal percipit ius ne, ne iisque quaeque philosophia pro, ut sed nobis nullam adolescens.
-Persius principes omittantur per an.
-Mel quot nostrud labores at, ad vis latine persius petentium.
-Cum partiendo neglegentur mediocritatem ut, per simul aliquando tincidunt no, falli nominavi molestiae ex nam.
-
-Saperet inciderint in vim, vis quod civibus accusata in.
-Et ullum facilisi accommodare nec, decore utroque definitionem an vel, id duo kasd facer latine.
-Ea sumo ullamcorper per, ex pro nobis efficiendi concludaturque, vix eu verear luptatum ocurreret.
-Ocurreret intellegat te eum, ius ea cibo epicurei dignissim.
-Diam elit quaeque cum ut.
-Est dicat summo prodesset te, no vel harum saepe sensibus, et ignota altera doming qui.
-
-Sit altera regione feugiat in, vim ad augue mediocritatem, duo reque choro theophrastus ea.
-Ipsum apeirian voluptatum duo no, nam ut ornatus accumsan percipitur, est no posse suscipiantur vituperatoribus.
-Errem legendos partiendo est cu, est clita signiferumque cu.
-Ea nulla scripta signiferumque eam, sea vidit dolores officiis ei, nibh saepe in nam.
-Ad vim justo dolor reprimique, eum ad labitur dissentias scriptorem, cu tempor mentitum assentior eum.
-Id vis mucius percipit invidunt.
-
-Vel at ullum intellegam constituam, mea dicta affert perpetua ea.
-Eripuit blandit iracundia nec ea, tale timeam molestie cum te.
-Cum ei kasd mazim, per ei delenit ullamcorper, ad sale vocibus temporibus usu.
-An nonummy oporteat phaedrum sed, eu unum deserunt incorrupte mei.
-His an possim delicata, ei accumsan constituam mea.
-Ei viris constituam pro, cu inimicus argumentum duo.
-
-Cu regione veritus persequeris vis, tation nusquam percipit in pri, vis no labitur deseruisse.
-Ius in reque fierent.
-Quo et erant recusabo honestatis.
-Iusto accusam liberavisse ex mei, ut cum nobis primis accommodare, viderer vocibus assueverit ut nec.
-Id odio oporteat instructior pri.
-Nam sint illud vituperatoribus no, zzril ubique dissentias id nec.
-
-Cu est legimus reformidans.
-Nam cibo mucius dolores ad, prima posse similique vix te.
-Mea at nemore mandamus assueverit, sit viris option scaevola cu.
-At omnis decore nostro pro, aliquyam explicari salutatus ei mea, sea no falli incorrupte.
-Graeco bonorum lobortis ut sed, ut pro discere legimus, an sea volumus consetetur.
-
+%ADD BIBLIOGRAPHY
 \bibliography{exemple_bibtex}

### 2020-03-04 Chapters/CoherenceEstimation.tex
update

\chapter{Coherence Estimation}

\section{Introduction}

A number of practical applications benefit of the knowledge about the diffuseness of a sound field, including speech enhancement and dereverberation \cite{p_habets_dual-microphone_2006}, noise suppression \cite{ito_designing_2010}, source separation \cite{duong_under-determined_2009} or background estimation \cite{stefanakis_foreground_2015}. In the field of spatial audio, diffuseness estimation is often used for parametrization \cite{pulkki_directional_2006, politis_compass_2018}, Direction-of-Arrival estimation \cite{thiergart_localization_2009} or source separation \cite{motlicek_real-time_2013}.

\todo{In this paper}, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
The motivation for this work is, first, that tetrahedral arrays are a well known type of microphone arrays, which have today become popular for applications related to Virtual and Augmented Reality. 
Second, the spherical isotropic sound field is known to be a good approximation to the reverberant part of the sound field in a room \cite{elko_spatial_2001, mccowan_microphone_2003}, and therefore it would be interesting to investigate how different microphone arrays behave under such conditions.

\subsection{\label{subsec:3:1} Coherence analysis}

Diffuseness is commonly estimated through the
\textit{Magnitude Squared Coherence} (MSC) \cite{elko_spatial_2001} between two frequency-domain signals $S_1$ and $S_2$, as a function of the
\textit{wavenumber} $k$ and the microphone distance $r$:
\begin{equation}
    \text{MSC}_{12}(k r) =
	\frac{|\left\langle S_1(k r) S_2(k r)^* \right\rangle|^2}
	{\left\langle|S_1(k r)|^2\right\rangle \left\langle|S_2(k r)|^2\right\rangle},
    \label{eq:MSC}
\end{equation}
where the $\left\langle \cdot \right\rangle$ operator represents the temporal
expected value, and $^*$ defines the complex conjugate operator. In the case of spherical isotropic noise fields, Eq.~(\ref{eq:MSC})
can be expressed in terms of microphone directivity patterns
$T(\phi,\theta,k r)$ as \cite{elko_spatial_2001}:

\begin{equation}
	\begin{aligned}
&\text{MSC}_{12}(k r) = \frac{|N_{12}(k r)|^2}{|D_{12}(kr)|^2} \\
&= \frac{|\int_{0}^{\pi} \int_{0}^{2\pi} T_1(\phi,\theta,k r) T_2^*(\phi,\theta,k r) e^{-jk r cos\theta} sin\theta d\theta d\phi|^2}{|\sqrt{ \int_{0}^{\pi} \int_{0}^{2\pi} |T_1(\phi,\theta,k r)|^2 sin\theta d\theta d\phi } \sqrt{\int_{0}^{\pi} \int_{0}^{2\pi}|T_2(\phi,\theta,k r)|^2 sin\theta d\theta d\phi}|^2}.
\label{eq:MSCdir}
    \end{aligned}
\end{equation}

Moreover, the general expression of the directivity of a first-order differential microphone is given by the following relationship:

\begin{equation}
	\begin{aligned}
	T_i(\psi_i) = \alpha_i + (1 - \alpha_i) \cos{\psi_i},
	\end{aligned}
\end{equation}

where $i \in [1,2]$ is the microphone index, $\psi_i$ is the angle between wave incidence and microphone orientation axis, and $\alpha_i \in [0,1]$ is the directivity parameter of the microphone $i$, which ranges from bidirectional ($\alpha_i = 0$) to omnidirectional ($\alpha_i = 1$). \\

For first-order differential microphones, there is a closed-form expression for the numerator and denominator of Eq.~(\ref{eq:MSCdir}):
\begin{equation}
	\begin{aligned}
    &N_{12}(k r) =  \frac{\alpha_1 \alpha_2 sin(kr)}{kr} 
    + \frac{(1-\alpha_2)(1-\alpha_2)(x_1x_2+y_1y_2)}{(kr)^3}(sin(kr)-kr cos(kr)) \\
    &+ \frac{z_1 z_2}{kr^3}[ ( (kr)^2 sin(kr) + 2kr cos(kr) )(1-\alpha_1)(1-\alpha_2) 
    + 2 sin(kr)(1-\alpha_1)(1-\alpha_2) ] \\
    &+ \frac{z_1}{(kr)^3}[ j(kr)^2 \alpha_2 cos(kr)(\alpha_1-1) + jkr \alpha_2 sin(kr)(1+\alpha_1) ] \\
    &+ \frac{z_2}{(kr)^3}[ j(kr)^2 \alpha_1 cos(kr)(\alpha_2-1) + jkr \alpha_1 sin(kr)(1+\alpha_2) ],\\
    &D_{12}(kr) =  \frac{\sqrt{3 \alpha_1^2+(1-\alpha_1)^2}\sqrt{3 \alpha_2^2+(1-\alpha_2)^2}}{3},
    \label{eq:closedform_msc}
    \end{aligned}
\end{equation}
where: $ x_i = cos(\phi_i) sin(\theta_i); y_i = sin(\phi_i) sin(\theta_i); z_i = cos(\theta_i)$
refers to the wave incidence angle $\psi_i$ expressed in spherical coordinates (with azimuth $\phi$ and inclination $\theta$).

\subsection{\label{subsec:3:2}Diffuseness estimation in ambisonics}
Let us consider a sound field captured with a spherical microphone array,
which contains $Q$ microphones distributed around a spherical surface of radius $R$
at the angular positions given by the azimuth-inclination pairs
$\Omega_q = (\phi_q,\theta_q)$.
The captured frequency-domain signals $ X_q(k)$ can be represented as the spherical harmonic domain signals $X_{mn}(k)$ through the spherical harmonic transform of order $L$ \cite{bertet_3d_2006}:
\begin{equation}
        X_{mn}(k) = \sum^{Q} X_q(k) Y_{mn}(\Omega_q) \Gamma_m(kR),
    \label{eq:ambisonics_encoding}
\end{equation}
where $Y_{mn}(\Omega_q)$ are the \textit{real-valued spherical harmonics}, and $\Gamma_m(kR)$ are the
\textit{radial filters} or equalization terms of order $m$, with $m \in [0,L]$ and $n \in [-m,m]$.\\

Due to a number of practical reasons, it is desirable to distribute the microphone capsules in a uniform manner along the sphere, with the the regular tetrahedron being the simplest possible configuration
 \cite{gerzon_design_1975}. Capsule signals recorded with such topology receive the name of \textit{A-Format} signals. Conversely, the term \textit{B-Format} (\textit{ambisonics}) describes the application of Eq.~(\ref{eq:ambisonics_encoding}) (\textit{ambisonic encoding}) to the \textit{A-Format} signals.
One of the most common coherence estimators for first-order ambisonic frequency-domain signals $X_{mn}(k)$ is the \textit{diffuseness} $\Psi$ as defined in \textit{DirAC} \cite{pulkki_directional_2006}:

\begin{equation}
    \Psi(k) = 1 - \frac{2 ||\left\langle \mathbb{R}\{\bm{X_1}(k)X_0(k)^* \} \right\rangle|| }{ \left\langle||\bm{X_1}(k)||^2 + |X_0(k)|^2\right\rangle},
    \label{eq:psi}
\end{equation}
where $X_0(k) = X_{00}(k)$ 
and $\bm{X}_1(k) = [X_{1-1}(k), X_{10}(k), X_{11}(k)]^\intercal$ are \textit{SN3D}-normalized.
For the sake of clarity, we will further define the \textit{B-Format coherence} estimator $\Delta$ as:
\begin{equation}
	\Delta(k) = 1 - \Psi(k).
	\label{eq:delta}
\end{equation}

Under spherical isotropic noise, the theoretical coherence between any pair of zeroth and first order ambisonic virtual microphones is equal to 0 for all frequencies, due to orthogonality and symmetry of the spherical harmonics \cite{elko_spatial_2001}. . This result can be also assessed by Eq.~(\ref{eq:closedform_msc}).

However, there are several practical factors that might corrupt the coherence estimation, such as the approximation of the temporal expectation by time averaging \cite{thiergart_diffuseness_2011} in Eq.~(\ref{eq:psi}), or the non-ideal implementation of the radial filters $\Gamma_m(kR)$ \cite{schorkhuber_ambisonic_2017}.
In the following sections, we present several experiments that illustrate the behavior of different coherence estimators applied on the signals captured with a tetrahedral microphone subjected to spherical isotropic noise, using both simulated and real sound recordings.

\section{Methods}

\subsection{Simulation}
Spherical isotropic noise has been generated following the \textit{geometrical method} \cite{habets_generating_2007, habets_comments_2010}, using $N = 1024$ plane waves. The resulting \textit{A-Format} signals correspond to a virtual tetrahedral microphone array mimicking the Ambeo
\footnote{Sennheiser Ambeo VR Mic. 
\todo{https://en-us.sennheiser.com/microphone-3d-audio-ambeo-vr-mic}}
characteristics ($R=0.015$ meter, $\alpha=0.5$). 
The generated audio has a duration of 60 seconds.

\subsection{Recording}
Spherical isotropic noise has been rendered to a spherical loudspeaker layout with 25 \textit{Genelec 8040}. The loudspeakers are arranged into three azimuth-equidistant 8-speaker rings at inclinations $\theta = [\pi/4, \pi/2, 3\pi/4]$, plus one speaker at the zenith ($\theta=0$).
The different speaker distances to the center are delay- and gain-corrected, and the signal feeds are equalized to compensate for speaker coloration. The room has an approximate $T_{60}$ of 300 ms measured at the 1 kHz third-band octave. 
The spherical isotropic noise has been again created following the \textit{geometrical method}, encoding a number of uncorrelated noise plane waves in ambisonics with varying orders $L \in [1,5]$. Due to practical limitations related with the software, the minimum number of sources $N = 256$ for an accurate sound field reconstruction \cite{habets_comments_2010} could not be reached - instead, the analysis has been performed parametrically with $N = [8, 16, 32, 64]$.
For each value of $L$ and $N$, approximately 15 seconds of audio have been recorded with an Ambeo microphone located at the center of the speaker array.
Ambisonics decoding uses the AllRAD method \cite{zotter_all-round_2012}, passing through a spherical 64-point 10-design virtual speaker layout, and includes an imaginary speaker at the nadir ($\theta=\pi$). The decoding matrix uses \textit{in-phase} weights.

\subsection{Data processing and metrics}

The sampling rate of all signals is 48 kHz.
All frequency-domain results have been obtained by averaging their time-frequency representations over time.  
Ambisonics conversion is performed using \textit{Ambeo A-B converter} AU plugin, version 1.2.1.

Two error metrics are considered: the frequency-dependent squared error $\varepsilon(k)$, and the mean squared error $\bar{\varepsilon}$:

\begin{equation}
    \varepsilon(k) = |X_1(k) - X_2(k)|^2; \text{         }\bar{\varepsilon} = \frac{1}{K}{\sum_{k=1}^{K} |X_1(k) - X_2(k)|^2}
    \label{nmse}
\end{equation}

\section{Results and discussion}
\subsection{\label{subsec:results_aformat}A-Format}

\begin{figure}
	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure1}
    \caption{\label{fig:Fig1}\textit{A-Format} coherence between microphone signals. Left: $\text{MSC}$ as a function of the frequency of theoretical, simulated and recorded \textit{((BLD,BRU)},  $L=5, N=64$) signals. Right: mean error $\bar{\varepsilon}$ of the recorded signals' $MSC$ \textit{(BLD,BRU)} compared to the simulated values, for all values of $L$ and $N$.}
\end{figure}

The coherence of the generated \textit{A-Format} signals is exemplified in Fig.~\ref{fig:Fig1} (left), which shows the $MSC$ between the capsule pair (\textit{BLD,BRU}) for the theoretical, simulated and recorded cases.
The theoretical coherence is derived from Eq.~(\ref{eq:closedform_msc}), while simulated and recorded MSC have been computed by Welch's method, using a \textit{hanning} window of 256 samples and 1/2 overlap.
The difference between theoretical and simulated coherence is negligible for practical applications.
However, there is a noticeable difference when compared to the recorded coherence. 
In general, the recorded $\text{MSC}$ follows the tendency of the simulated curve up to around 5 kHz.
Above this frequency, the recorded $MSC$ presents several spectral peaks, which might be partially explained by the interference of the microphone itself in the recorded sound field, and by the non-ideal directivity of the capsules.
The squared error $\varepsilon(k)$ with respect to the simulated curve is shown in Fig.~\ref{fig:Fig1} (left), while Fig.~\ref{fig:Fig1} (right) represents the same error averaged over frequency $\bar{\varepsilon}$ for different spatial resolution values of the diffuse field reproduction algorithm.
As expected, $\bar{\varepsilon}$ decreases with increasing values of $L$ and $N$.

\subsection{B-Format} 
\begin{figure}
	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure2}
	\caption{\label{fig:Fig2} Estimated \textit{B-Format} coherence ($\Delta$) of a simulated diffuse sound field, as a function of the temporal averaging vicinity radius $r$. Left: $\Delta(k)$ for different values of $r$, with (coarse) and without (fine) application of radial filters. Right: mean and standard deviation of $\Delta(k)$ as a function of $r$.}
\end{figure}

In order to evaluate the dependency of $\Delta$ on the number of time frames used for averaging, the following procedure is presented.
The simulated \textit{A-Format} sound field has been transformed into the spherical harmonic domain, with and without the application of radial filters $\Gamma_m(kR)$. Then, $\Delta$ has been computed with Eq.~(\ref{eq:delta}) for exponentially growing values of $r$ between 1 (8 ms) and 2048 (10.92 s), where $r$ is the vicinity radius used for time averaging, and the number of time windows is given by $T = 2r+1$.
The time-frequency representation is derived by applying the STFT with the same window parameters as in Subsection \ref{subsec:results_aformat}.\\

Figure~\ref{fig:Fig2} (left) shows the great dependence of $\Delta$ on $r$.  The estimated coherence tends to the theoretical values with increasing values of $r$. This tendency is better appreciated in Fig.~\ref{fig:Fig2} (right): the curve asymptotically decreases to a value $\Delta_{min}\approx0$.
Another interesting observation comes from the frequency response of the curves. For all values of $r$, the coherence of the compensated \textit{B-Format} signal (with $\Gamma_m(kR)$) is roughly flat up to around 7 kHz, which approximately corresponds to the operational spatial frequency range of the microphone \cite{gerzon_design_1975}.
Above this value, the coherence response looses the flatness due to spatial aliasing. The response above the maximum frequency could be stabilized, if needed, by alternative diffuseness estimation methods \cite{politis_direction--arrival_2015}.
The coherence level differences along frequency are inversely proportional to $r$ --- the effect is better depicted by the standard deviation values (right).
The effect of the radial filters in the coherence measurement is also shown: for a given $r$, the shape of the coherence is always less flat if no filters are applied. Conversely, in this case, coherence values are always smaller for the same $r$. This effect might be explained taking into account the inter-channel coherence introduced by microphone and encoder imperfections in real scenarios \cite{schorkhuber_ambisonic_2017}.
As a remark, the comparison between Figs.~\ref{fig:Fig1} and \ref{fig:Fig2} provides evidence that the application of the spherical harmonic transform might be able to yield more accurate diffuseness estimations, due to a better signal conditioning \cite{epain_spherical_2016}.\\

\begin{figure}
	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure3}
	\caption{\label{fig:Fig3}\textit{B-Format} coherence between microphone signals. Left: $\Delta$ of simulated and recorded ($L=5, N=64$) signals. Right: $\bar{\varepsilon}$ of the recorded signals coherence across all values of $L$ and $N$.}
\end{figure}

Figure~\ref{fig:Fig3} (left) shows the estimated coherence for the recorded sound field with $L=5$ and $N=64$, using a vicinity radius of $r=1024$ ($\approx$ 5 s).
The curve is centred around $\Delta=0.25$ and presents several spectral peaks, as in the \textit{A-Format} case. 
It is important to notice here that the deviations between the coherence of the simulated and the recorded sound fields are much stronger compared to those of Fig.~\ref{fig:Fig1}. 
This effect can be also appreciated in Fig.~\ref{fig:Fig3} (right): the mean squared error is around two orders of magnitude higher in \textit{B-Format}.
Nevertheless, similar as in Fig.~\ref{fig:Fig1} (right), $\bar{\varepsilon}$ decreases with increasing values of $L$ and $N$.
This behavior suggests that the deviations between the recorded and the simulated coherence can be to a large degree explained by the low spatial resolution of the reproduction system; given a higher number of loudspeakers, we expect that the reproduced diffuseness will tend to the theoretical expression.

\section{Conclusions}
The diffuseness of a sound field is an important parameter for several applications. In this work, two different metrics of diffuseness have been defined and measured with a tetrahedral microphone subjected to spherical isotropic noise.
The analysis shows, first, the impact of the time-averaging window length on the \textit{B-Format} diffuseness estimator.
This result might be useful for designing coherence estimators that are parametrized with respect to the length of the analysis window \cite{thiergart_diffuseness_2011}.
Second, the feasibility of diffuse sound field reproduction by a spherical loudspeaker array using ambisonics plane-wave encoding and the \textit{geometrical method} is studied. 
Results suggest that this approach is viable, given a sufficient spatial resolution; a quantification of the impact of the number of loudspeakers remains for future work.

### 2020-03-04 Chapters/DataGeneration.tex
update

@@
 
-AmbisonicsDRIR: explain problematics, standard, etc.
-also: pysofaconventions
+Impulse Responses (IRs) measurements constitute a compact way of representing the acoustic propertiesof a linear time-invariant system.
+When such measurements are performed in a specific room or enclosure, the so-called Room Impulse Responses (RIRs) are able to capture the intrinsic reverberation and acoustic characteristics of the enclosure, for which several methods have been developed these past years \cite{stan2002comparison}.
+Furthermore, it is possible to account for different emitter/receiver positions in the measurement, usually performing the measurement with a microphone array.
+In that case, this kind of measurements is referred to as Directional Room Impulse Responses (DRIRs) \cite{embrechts2005computation}.
+DRIRs have a wide range of applications: auralization \cite{embrechts2005computation}, room acoustics analysis \cite{embrechts2015measurement,clapp2011investigations} and modelling \cite{romblom2017diffuse}, spatial audio synthesis \cite{coleman2017object}, source separation and dereverberation \cite{baque2016separation}, acoustic heritage preservation \cite{gerzon1975recording,murphy2005multi}, etc.
+
+\subsection{Ambisonics Recording}
+
+It is also possible to capture Ambisonics audio scenes by using specific recording devices.
+Spherical microphone arrays (also known as \textit{Ambisonics microphones}) are a type of microphone arrays in which the capsules are located around a spherical surface, presenting rotational symmetry.
+Such geometric arrangement allows the recorded signals to be transformed to the Ambisonics domain, by means of projection into the Spherical Harmonics basis, and further equalization with radial filters \cite{Bertet2006}.
+
+This process is known in the audio production domain as \textit{A-B Conversion}.
+All major spherical microphone array manufacturers provide tools for achieving this transformation \cite{soundfield,ambeo,zylia,tetra,em32}.
+
+\subsection{Ambisonics DRIRs}
+
+The intrinsic spatial capabilities of Ambisonics microphones might be applied to DRIR recording, as originally proposed by Gerzon in the context of acoustical heritage preservation \cite{gerzon1975recording}.
+In recent years, several datasets of Ambisonics DRIRs have been publicly released, such as the OpenAIR database \cite{murphy2010openair} or the set of  measurements performed in the scope of the S3A project \cite{coleman2015s3a,openair}.
+
+\subsection{The SOFA Conventions}
+
+In general, the Ambisonics DRIR databases show a common approach for describing the measurements: given a specific room, usually represented as a folder, IR data consist of several multichannel audio files, with one audio channel per spherical harmonic, and one file per emitter/receiver combination.
+Furthermore, it is also usual to provide a \textit{metadata} file, describing the different emitter and receiver positions, and eventually some information about the measurement setup, methodology, etc.
+Such files might be formatted as plain text or delimiter-separated value files.
+
+Despite the common approach, it can be foreseen that each database generated by a different individual or institution might potentially have a different naming convention, folder structure, file format, and so on.
+This situation hinders data manipulation and exchange, and forces users to write ad-hoc parsers and algorithms for each specific database.
+
+The \textit{Spatially Oriented Format for Acoustics} (SOFA) is a file format designed for a consistent, standardized storage and manipulation of IR data \cite{majdak2015aes69}.
+The need for such standard arose from dealing with different databases of \textit{Head-Related Transfer Functions} (HRTFs), in a similar manner as the one mentioned for Ambisonics DRIRs.
+
+There are several SOFA conventions, each one addressing a particular type of IR measurement.
+In the case of Ambisonics DRIRs, given the data representation in existing databases, one could outline the following specificities:
+
+\begin{itemize}
+    \item Presence of Ambisonics-related information (Ambisonics order, channel ordering and normalization)
+    \item Audio stored in the Ambisonics (spherical harmonics) domain  
+    \item Data structure support for different combinations of source and receiver positions
+\end{itemize}
+
+However, none of the existing SOFA conventions meets those requisites.
+The potentially first candidate by name, \textit{SingleRoomDRIR}, is limited to one source position per file.
+On the other hand, \textit{MultiSpeakerBRIR} allows for multiple sound sources, but restricts the number of receivers (microphone capsules) to two, as expected in a binaural recording.
+\textit{GeneralFIRE} is intended for "data which are too general to store in more specific conventions" \cite{generalfire}.
+
+\subsection{Convention Proposal}
+
+\subsubsection{Considerations}
+
+The SOFA specification defines some criteria that must be fulfilled in order to propose a new convention \cite{sofaconventions}. These criteria are:
+\begin{enumerate}
+    \item Data must exist.
+    \item Data can not be described by existing SOFA conventions.
+    \item Relevant information about the data must be available.
+\end{enumerate}
+
+As we already mentioned, existing databases of Ambisonics DRIRs can be found in OpenAIR and S3A databases (criterium 1).
+Criterium 2 has been discussed in the previous paragraph.
+All available Ambisonics DRIR datasets are accompanied by explanations, pictures, diagrams and related information (criterium 3).
+
+\subsubsection{Specifications}
+
+We therefore propose a new SOFA convention, \textit{AmbisonicsDRIR}, designed for DRIRs measured with spherical microphone arrays, and presented in the Ambisonics domain.
+In other words, we propose to store the instantaneous $B_{mn}$ values for a given Ambisonics order $L$, provided that $S$ is a unit impulse (Eq.
+\ref{eq1}).
+
+The \textit{Listener} - as defined by the SOFA specification - is embodied by the spherical microphone array.
+The different \textit{Receivers} are the different Ambisonics Components $B_{mn}$, so that their number is fixed provided $L$ (more precisely, by following the relationship $R = (L+1)^2$), and their positions are not applicable.
+Furthermore, there might be many different \textit{Sources}, all of them omnidirectional and consisting of one only \textit{Emitter}.
+In that sense, \textit{M} represents the number of different \textit{Source} positions.
+
+The proposed convention is based on \textit{GeneralFIRE}, with the following modifications:
+\begin{itemize}
+
+    \item Mandatory field \textit{GLOBAL:AmbisonicsOrder} (type \textit{double}, dimension \textit{I}, default \textit{1}). Indicates the order of the Spherical Harmonic expansion.
+
+    \item Mandatory field \textit{DATA:ChannelOrdering} (type \textit{attribute}, default \textit{acn}).
+Describes the ordering of the different Ambisonics Channels.
+Must be one of: \textit{acn} or \textit{fuma}.
+
+    \item Mandatory field \textit{DATAChannelNormalization} (type \textit{attribute}, default \textit{sn3d}).
+Describes the Ambisonics normalization convention used in the data.
+Must be one of: \textit{sn3d}, \textit{n3d}, \textit{fuma} or \textit{maxn}.
+
+    \item Fields \textit{ReceiverPosition}, \textit{ReceiverPosition:Type} and \textit{ReceiverPosition:Units} are not needed, since it is assumed that \textit{Receivers} represent the Ambisonics channels.
+Furthermore, the values of \textit{R} (number of receivers) are defined by $R = (L+1)^2$.
+Accordingly, the values of \textit{GLOBAL:AmbisonicsOrder} and \textit{R} are valid only if they follow the given equation.
+
+    \item Field \textit{DATA:Delay} is not mandatory.
+
+    \item Relevant information about the microphone model and brand, Ambisonics encoding methodology, software used, etc, is recommended to be added into the field \textit{GLOBAL:Comment}.
+
+\end{itemize}
+
+\subsection{Results}
+
+As a preliminary result, we have extended the current SOFA C++  \cite{sofacpp} and Matlab \cite{sofamo} APIs to be fully compatible with the AmbisonicsDRIR convention. \todo{pysofaconventions}
+
+Furthermore, as a use-case, a selection of existing Ambisonics DRIRs have been transcoded to the proposed convention: \textit{Emmanuel Main Church} from S3A database \cite{coleman2015s3a}, and \textit{Heslington Church} and \textit{York Guildhall Council Chamber} from the OpenAIRlib \cite{openair}.
+Figure \ref{guildhall} shows a schematic diagram of the different \textit{Emitter/Receiver} positions at the \textit{York Guildhall Council Chamber} recordings.
+The data, as well as the tools used to perform the conversion, are available online under an open source license \cite{ambisonicsdrirexamples}.
+
+Finally, we must remark that the \textit{AmbisonicsDRIR} SOFA convention proposal is currently under discussion, and the described specifications are subject to change with upcoming versions.
+
+% TODO: include some examples!!! about data dimensions etc
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{Figures/DataGeneration/guildhall}
+	\caption{Source and Listener position diagram for the "York Guildhall Council Chamber" Ambisonics DRIR (original diagram attribution to \cite{guildhall} }
+	\label{guildhall}
+\end{figure}
+
+\subsection{Summary}
+
+This document addresses the lack of compatibility among different databases of Ambisonics DRIRs.
+The present proposal consists in defining a new SOFA convention, specifically designed for Ambisonics DRIRs.
+That way we contribute to improve the ease of data manipulation and database interoperability.
+Software implementations and tools for automatic conversion are provided.
 
@@
 
+% ==================================================================
 \section{High-level scene description}
 
-ambiscaper
+\subsection{Motivation}
+
+\todo{introduce the topic}
+
+\todo{ARRANGE this introduction in a meaningful way}\\
+The work by Moore \cite{Moore2015} was the first to combine Intensity Vector statistics with DPD-test preprocessing.
+A similar approach is found on the \textit{Single Source Zones (SSZs)} algorithm, first presented by Pavlidi \cite{Pavlidi2015}.
+Other recent proposals, as for example the work by He \cite{He2017}, consider the local DOA variance as an estimator of reliability.
+
+Source Localization based on IV statistics has been also successfully used as a preprocessing step for Source Separation tasks.
+The proposal of G\"unel \cite{Gunel2008}, which uses beamforming over the DOA estimation, was a pioneer work on this scope.
+Other similar approaches, which rely on DOA-based Time-Frequency (TF) masking, can be found on the works by Shujau \cite{Shujau2011}, Riaz \cite{Riaz2015} or Chen \cite{Chen2015}.
+
+\subsection{Evaluation data}
+\label{subsec:evaluationdata}
+
+Audio data used for evaluation in the works presented in Section \ref{subsec:algorithms} can be classified in  three main categories:
+
+\begin{itemize}
+
+    \item \textit{IR simulations:} Impulse Responses simulated by numerical methods, using software tools such as \textit{SMIR Generator} \cite{smir}.
+IR simulation provides the most flexible approach for evaluation, since it allows custom design and parametric analysis (algorithm evaluation as a function of reverberation time, source(s) distance(s), etc).
+However, the resulting audio data realism is restricted by the simulation algorithm, which is usually limited to empty \textit{shoebox} rooms.
+
+    \item \textit{IR recordings:} Impulse Responses (IRs) of specific rooms recorded with an Ambisonics microphone, as first presented by Gerzon in the context of acoustic heritage preservation \cite{gerzon1975recording}.
+IR recordings do not provide the flexibility of IR simulations, but allow realistic captures of the room's acoustic properties, while maintaining the potential of using different source contents and positions.
+OpenAIRlib \cite{murphy2010openair} is the reference public repository for IRs, even including a dedicated Ambisonics IR section.
+
+    \item \textit{Audio recordings:} a sound scene recorded in a specific room, using an Ambisonics microphone and real sound sources.
+This approach  represents the  most  accurate  option  for algorithm evaluation in a real scenario.
+However, its cost and lack of flexibility makes this approach only attractive for last stages of algorithm  evaluation.
+
+\end{itemize}
+
+In the case of IR simulations and recordings, the Ambisonics IRs are convolved with anechoic monophonic recordings to produce the actual evaluation data.
+The anechoic recordings are usually taken from standard audio processing databases, such as TIMIT \cite{timit}.
+\\
+
+Table \ref{table:evaluationdata} summarizes the evaluation data types used in each work review in Section \ref{subsec:algorithms}, along with the audio content type and origin database (if any).
+
+%% DOA ESTIMATION
+\begin{table}[htpb]
+    \centering
+    \footnotesize
+
+	\begin{tabular}{ p{2cm} p{2.8cm} p{2.5cm} p{2.3cm} }
+    \toprule
+    % \cmidrule(r){1-2}
+
+    % \centering 
+    \textbf{Article} &
+    % \textbf{\textit{$L$}} &
+    \textbf{Evaluation Data Type} &
+    \textbf{Audio Content Type} &
+    \textbf{Database}\\
+    \midrule
+
+    Thiergart \cite{Thiergart2009}
+    % & 1 h
+    & Audio recording 
+    & Speech
+    & -\\
+
+    Tervo \cite{Tervo2009}
+    % & 1 h
+    & Audio recording
+    & Noise, music
+    & -\\
+
+    Jarret \cite{Jarrett2010} 
+    % & 4 
+    & IR simulation, Audio recording
+    & Noise
+    & -\\
+
+    Nadiri \cite{Nadiri2014}   
+    % & 4 
+    & IR simulation, Audio recording
+    & Speech
+    & TIMIT \\
+
+    Moore \cite{Moore2015}   
+    % & 4 
+    & IR simulation 
+    & Speech
+    & APLAWD \\
+
+    Pavlidi \cite{Pavlidi2015}
+    % & 4
+    & IR simulation 
+    & Noise, speech
+    & -\\
+
+    He \cite{He2017}   
+    % & 1 h
+    & IR simulation, Audio recording
+    & Speech
+    & TIMIT \\
+
+    % Ding \cite{Ding2017}   
+    % % & 1 h
+    % & IR simulation \& Audio recording \\
+
+    \midrule
+
+    Gunel \cite{Gunel2008}
+    & IR recording
+    & Speech, music
+    & Music for Archimedes \\
+
+    Shujau \cite{Shujau2011} 
+    & Audio recording
+    & Speech
+    & TIMIT \\
+
+    Riaz \cite{Riaz2015}
+    & IR recording
+    & Speech, music
+    & Music for Archimedes \\
+
+    Chen \cite{Chen2015}
+    & IR simulation, IR recording
+    & Speech
+    & TIMIT \\
+
+    \bottomrule
+    \end{tabular}
+    \caption{Summary of audio data used across Ambisonics-based Source Localization (above) and Source Separation (below)  algorithm proposals. \todo{update table? fix references}}
+    \label{table:evaluationdata}
+\end{table}
+
+Although algorithm evaluation in real scenarios provides the most realistic approach to the problem, this methodology has not been widely adopted due to its cost and lack of flexibility.
+As an example, none of the works using real recordings in Table \ref{table:evaluationdata} performs the evaluation with more than one  recording in each case.
+In contrast, when using IR-based  scenes, the number of audios evaluated are usually one or two magnitude orders greater.
+
+However, it is important to notice the lack of public availability of evaluation data.
+None of the analyzed articles provide a way to access neither the used audio dataset, nor the groundtruth (position annotations in the case of Sound Localization, and original sound sources for Sound Separation).
+Only in the case of simulated IRs it is possible to partially replicate the experimental setup, since some of the parameters used in the simulation software are usually provided.
+Furthermore, the process of creation of the custom datasets (selecting an anechoic audio dataset, convolving with custom IRs, or performing real recordings) seems to be performed \textit{ad-hoc} on each article.
+
+Taking into account the flexibility offered by IR-based scenes, it would be desirable to have a tool for automatic generation of reverberant Ambisonics scenes (and their associated groundtruth) for analysis purposes.
+Such tool would help the scientific community in several ways: reducing the amount of time dedicated to build custom datasets, reusing publicly available resources and recordings, and enhancing experiment reproducibility by making easier the exchange of datasets.
+Furthermore, the capacity of producing a big number of diverse audio scenes for analysis will help the algorithm design and early testing stages,and specially the training stage of machine learning-based algorithms.
+
+\subsection{AMBISCAPER}
+\label{sec:ambiscaper}
+
+\todo{this has been downgraded to subsection. maybe place it at the introduction or something to find a more meaningful explanation}
+
+AmbiScaper \cite{ambiscaper} is a tool designed to provide a flexible way of creating complex Ambisonics sound scenes and their associated groundtruth, to be used in the context of Source Localization and Source Separation algorithms.
+AmbiScaper offers a high level control of the sound scene parameters, and provides an easy way of creating large datasets with custom characteristics.
+AmbiScaper is based on Scaper, a framework designed to generate ground truth information to train Sound Event Detection models \cite{Salamon2017}.
+
+\subsection{Sound scene description}
+\label{subsec:description}
+
+One of the main features of AmbiScaper is that all parameters of the sound scene can be specified in a non-deterministic way.
+In that sense, the parameters for each \textit{event} (sound source) are actually generated through a two-step process.
+First, in the \textit{Event Specification}, all parameters related to an event are defined in terms of statistical distributions.
+During the \textit{Event Instanciation}, the actual values for each parameter are then sampled from the statistical distributions.
+This two-step process allows the user to describe abstract \textit{templates} of sound scenes, rather than manually assigning values to parameters.
+Therefore, a single \textit{event specification} might produce potentially infinite different sound scenes.
+
+\subsection{Architecture}
+\label{subsec:architecture}
+
+\begin{figure}
+\label{fig_architecture}
+  \centering
+    \includegraphics[width=\textwidth]{Figures/DataGeneration/figure_architecture_V2.png}
+    \caption{AmbiScaper architecture.}
+\end{figure}
+
+%%%%%%%%
+In order to generate a sound scene, AmbiScaper requires three different inputs: the original \textit{mono signals}, which will be the basis for the scene's audio content, an optional \textit{Ambisonics IR}, and the \textit{event specification}.
+
+The process of dataset creation starts with the \textit{event instanciation}, as described in Section \ref{subsec:description}.
+Once all values are sampled, three different types of output are generated: the \textit{Ambisonics scene}, the \textit{instanciated mono signals} (the original mono signals after data augmentation and duration changes), and the \textit{annotations}, in the form of a JAMS file \cite{humphrey2014jams}, containing all information about the scene specification and the instanciated values.
+
+AmbiScaper's architecture is depicted in Figure \ref{fig_architecture}.
+
+\subsection{Reverberation}
+
+When no reverberation is specified, AmbiScaper can generate anechoic sound scenes, which can be useful for baseline performance evaluation.
+In this case, there is no upper limit on the Ambisonics order of the rendered scene.
+Furthermore, the anoechic case allows for the specification of source \textit{spread} or apparent size through Ambisonics order downgrade \cite{carpentier2017ambisonic}.
+
+AmbiScaper supports the usage of recorded Ambisonics IRs, although it is currently limited to IRs from the S3A database \cite{coleman2015s3a}.
+The development of a standardized file format for Ambisonics IRs, which is being discussed at the moment of writing, will provide the flexibility to work with arbitrary Ambisonics IRs.
+
+Lastly, AmbiScaper features the possibility of using simulated Ambisonics IRs, through a wrapper to \textit{SMIR Generator}.
+In this case, the reverberation model specifications might be defined as well in statistical terms, and the generated IRs are stored for evaluation purposes.
+A working copy of Matlab is required to run this option.
+
+When a reverberant sound scene is created, the specific Ambisonics IRs used on the scene are also provided as an output.
+Other research problems such as dereverberation or room reflection modelling might therefore benefit from such data.
+
+\subsection{AmbiScaper and experiment reproducibility}
+
+As mentioned in Section \ref{subsec:evaluationdata}, there is a generalized lack of publicly available datasets for Source Localization and Source Separation in the Ambisonics domain.
+Even when using general purpose audio/speech datasets (such as TIMIT), the actual reverberant Ambisonics evaluation data is usually not available.
+In that sense, the compatibility of AmbiScaper with public Ambisonics IR databases is a key aspect for reproducibility, since it allows for the reutilization of acoustical measurements in a systematic way.
+
+Furthermore, the output of the AmbiScaper dataset generation process is not limited to the actual dataset.
+In fact, as explained in Section \ref{subsec:architecture}, the resulting annotation file does not only contain the \textit{instanciation} (the actual values of each parameter in the sound scene), but also the \textit{specification} (the statistical distributions from which the instanciated values are sampled).
+In the scope of experiment reproducibility, the exchange of \textit{specification files} instead of actual audio files greatly reduces the storage capacitiy and bandwith required to transfer big databases.
+
+AmbiScaper is implemented in the form of a Python package, publicly available through the \textit{Python Package Index} repository. 
+AmbiScaper is free software under the GPL license, easing the software adoption and the potential engagement of the scientific community with the development.
+
+\subsection{Sample Dataset}
+\label{ssec:dataset}
+
+As an example of the potential capabilities of AmbiScaper,  we have created and published a dataset for the evaluation of source localization algorithms \footnote{https://zenodo.org/record/1186907 \todo{link}}.
+
+The dataset contains 300 first order Ambisonics sound scenes, each one containing from one to three static sound sources (be they simultaneous or not), with different gains and SNRs, and placed at random positions around the sphere.
+Each sound scene has a duration between 1 and 2 seconds.
+
+The sources are randomly chosen from a subset of the Anechoic OpenAirlib database, which mostly contains recordings from baroque musical instruments.
+The AudioBooth IR database from S3A has been used for all scenes.
+It features SoundField recordings from an acoustically treated room with a geodesic dome structure, to which 17 speakers have been attached.
+
+\subsection{CONCLUSIONS AND FUTURE WORK}
+\label{sec:conclusions}
+
+AmbiScaper is a tool designed for easy dataset creation and exchange, in the context of reverberant Ambisonics Source Localization and Source Separation.
+It responds to the lack of public datasets for algorithm design, evaluation and reproducibility.
+An example dataset generated with AmbiScaper, and its analysis with state-of-the-art Source Localization algorithms are also provided.
+Further studies in Ambisonics IR position interpolation would allow creating non-static audio scenes, which might be an interesting feature. 
+Another feature in progress is the support for a standardized Ambisonics IR format \cite{perez2018ambisonics}, which will ease data reusability and provide the potential to cover a big variety of acoustic scenarios.

### 2020-03-04 Chapters/SELD.tex
update

\chapter{Sound Event Localization and Detection}

\section{Introduction}
\label{sec:intro}

Sound Event Localization and Detection (SELD) refers to the problem of identifying, for each individual  event present in a sound field, the temporal activity, spatial location, and sound class to which it belongs. SELD is a current research topic which deals with microphone array processing and sound classification, with potential applications in the fields of signal enhancement, autonomous navigation, acoustic scene description or surveillance, among others.

SELD arises from the combination of two different problems: Sound Event Detection (SED) and Direction of Arrival (DOA) estimation. The number of works in the literature which jointly address SED and DOA problems is relatively small. It is possible to classify them by the type of microphone arrays used: distributed \cite{grobler2017sound, butko2011two, chakraborty2014sound} or near-coincident \cite{hirvonen2015classification, lopatka2016detection, Adavanne2018_JSTSP}.
As mentioned in \cite{Adavanne2018_JSTSP}, the usage of near-coincident circular/spherical arrays enables the representation of the sound field in the spatial domain, using the spherical harmonic decomposition, also known as Ambisonics \cite{gerzon1973periphony, daniel2000representation}. Such spatial representation allows a flexible, device-independent comparison between methods. Furthermore, the number of commercially available ambisonic microphones has increased in recent years due to their suitability for immersive multimedia applications.  
Taking advantage of the compact spatial representation provided by the spherical harmonic decomposition, several methods for parametric analysis of the sound field in the ambisonic domain have been proposed  \cite{pulkki2006directional, berge2010high, Politis2018, pulkki2018parametric}.
These methods ease sound field segmentation into direct and diffuse components, and further localization of the direct sounds.
The advent of deep learning techniques for DOA estimation has also improved the results of traditional methods \cite{Adavanne2018_JSTSP}. However, none of the deep learning-based DOA estimation methods explicitly exploits the spatial parametric analysis. This situation is further extended to the SELD problem, with the exception of \cite{lopatka2016detection}, where DOAs are estimated from the \textit{active intensity vector}  \cite{pulkki2006directional}.

The motivation for the proposed methodology is two-fold.
First, we would like to check whether the usage of spatial parametric analysis in the ambisonic domain can improve the performance of SELD algorithms.
Second, temporal information derived by the parametric analysis could be further exploited to estimate event onsets and offsets, thus lightening the event classifier complexity; such reduction might positively impact algorithm's performance.

In what follows, we present the methodology and the architecture of the proposed system (Section~\ref{sec:method}). Then, we describe the design choices and the experimental setup (Section~\ref{sec:expe}), and discuss the results in the context of DCASE2019 Challenge - Task 3 (Section~\ref{sec:results}). A summary is presented in Section~\ref{sec:conclusion}. In order to support open access and reproducibility, all code is freely available at \cite{code}. \todo{code ref}.

\section{Method}
\label{sec:method}

The proposed method presents a solution for the SELD problem splitting the task into four different problems: \textit{DOA  estimation}, \textit{association}, \textit{beamforming} and \textit{classification}, which will be described in the following subsections.
The former three systems follow a heuristic approach---in what follows, they will be jointly referred to as the \textit{parametric front-end}. Conversely, the \textit{classification} system is data-driven, and will be referred to as the \textit{deep learning back-end}. The method architecture is depicted in Figure~\ref{fig:diagram_all}.

\begin{figure}[h]
	\centering
    \includegraphics[width=\columnwidth]{Figures//SELD/dcase_challenge_tech_report_all2.pdf}
    \caption{System architecture.}
    \label{fig:diagram_all}

\end{figure}

\subsection{DOA estimation}
\label{ssec:doa_estimation}

\begin{figure*}[h]

    \includegraphics[width=\textwidth]{Figures/SELD/DOA1.pdf}
    \caption{DOA estimation architecture.}
    \label{fig:doa}

\end{figure*}

The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency (TF) spatial audio analysis.
Let us consider a first-order ($L=1$) ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
\begin{equation}
    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
\end{equation}
From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
\begin{equation}
    \begin{aligned}
    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
    \end{aligned}
\end{equation}
where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.

It is desirable to identify the TF regions of $\pmb{\Omega}(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.

The first mask is the \textit{energy density mask}, which is used as an activity detector. The energy density $E(k,n)$ is defined as in \cite{stanzial1996reactive} :
\begin{equation}
    E(k,n) = \frac{|B_w(k,n)|^2
    + || [ B_x(k,n), B_y(k,n), B_z(k,n)] ||^2}{2Z_0c},
\end{equation}

with $c$ being the sound speed. A gaussian adaptive thresholding algorithm is then applied to $E(k,n)$, which selects TF bins with local maximum energy density, as expected from direct sounds.

The \textit{diffuseness mask} selects the TF bins with high energy propagation. Diffuseness $\Psi(k,n)$ is defined in \cite{merimaa2005spatial} as:
\begin{equation}
    \Psi(k,n) = 1 - ||\LL \pmb{I}(k,n) \RR|| / (c \LL E(k,n) \RR),
    \label{psi}
\end{equation}
where $\LL \cdot \RR$ represents the temporal expected value.

The third mask is the \textit{DOA variance mask}. It tries to select TF regions with small standard deviation\footnote{In this work, all statistical operators for angular position refer to the $2\pi$-\textit{periodic} operator for azimuth, and the standard operator for elevation.} with respect to their neighbor bins---a characteristic of sound fields with low diffuseness \cite{pulkki2018parametric}.

The three masks are then applied to the DOA estimation, obtaining the TF-filtered DOAs $\bar{\pmb{\Omega}}(k,n)$. Finally, a median filter is applied, with the aim of improving DOA estimation consistency and removing spurious TF bins.
The median filter is applied in a TF bin belonging to $\bar{\pmb{\Omega}}(k,n)$ only if the number of TF bins belonging to $\bar{\pmb{\Omega}}(k,n)$  in its vicinity is greater than a given threshold $B_{min}$.
The resulting filtered DOA estimation is referred to as $\breve{\pmb{\Omega}}(k,n)$.

\subsection{Association}
\label{ssec:association}

\begin{figure*}[h]
    \includegraphics[width=\textwidth]{Figures/SELD//ASS.pdf}
    \caption{Association architecture.}
    \label{fig:association}

\end{figure*}

The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\pmb{\Omega}}(k,n)$ to a set of events, each one having a specific onset, offset  and location.

First, DOA estimates are resampled into \textit{frames} of the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\pmb{\Omega}}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is greater than a threshold $K_{min}$.

Next, the standard deviation in azimuth ($\sigma_{\varphi}$) and elevation ($\sigma_{\theta} $) of the frame-based DOA estimates $\breve{\pmb{\Omega}}(k,m)$ are compared to a threshold value ($\sigma_{max}$), and the result is used to estimate the frame-based event overlapping amount $o(m)$ :

\begin{equation}
\begin{aligned}
	o(m) =  \begin{cases}
    		1, &\text{if } \sigma_{\varphi}/2 + \sigma_{\theta} < \sigma_{max},\\
    		2, &\text{otherwise}.
    \end{cases}    		
\end{aligned}
\end{equation}
The clustered values $\pmb{\Omega}_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\pmb{\Omega}}(k,m)$, using a modified version of K-Means which minimizes the central angle distance. Notice that, for $o(m)=1$, the operation is equivalent to the median.

The following step is the grouping of clustered DOA values into events.
Let us define $\pmb{\Omega}_S(m)$ as the frame-wise DOA estimations belonging to the event $S$. A given clustered DOA estimation $\pmb{\Omega}_{\text{cluster}}(m)$ belongs to the event $S$ if the following criteria are met:
\begin{itemize}
    \item The central angle between  $\pmb{\Omega}_{\text{cluster}}(m)$ and the median of $\pmb{\Omega}_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
    \item The frame distance between M and the closest frame of $\pmb{\Omega}_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
\end{itemize}

The resulting DOAs $\pmb{\Omega}_S(m)$ are subject to a postprocessing step with the purpose of 
delaying event onsets in frames where $o(m) > 2$, and discarding events shorter than a given minimum length. 
Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\pmb{\Lambda}_S = (\pmb{\Omega}_S, \text{onset}_S, \text{offset}_S)$.

\subsection{Beamforming}
\label{ssec:Beamforming}

The last step performed in the front-end is the input signal segmentation. The spatial and temporal information provided by the annotations $\pmb{\Lambda}_S$ are used to produce monophonic signal estimations of the events, $\tilde{b}_S(t)$, as the signals captured by a virtual hypercardioid:
\begin{equation}
\tilde{b}_S(t) = \pmb{Y}(\pmb{\Omega}_S) \pmb{b}^\intercal(t),
\end{equation}
where $ \pmb{Y}(\pmb{\Omega}_S) = [Y_w(\pmb{\Omega}_S), Y_x(\pmb{\Omega}_S), Y_y(\pmb{\Omega}_S), Y_z(\pmb{\Omega}_S)]$ is the set of real-valued spherical harmonics up to order $L=1$ evaluated at $\pmb{\Omega}_S$.

\subsection{Deep learning classification back-end}
\label{ssec:backend_method}

%intro - problem formulation==============
The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{b}_S(t)$.
Then, the back-end classifies the resulting signals as belonging to one of a target set of 11 classes.
Therefore, the multi-task nature of the front-end allows us to define the back-end classification task as a simple multi-class problem, even though the original SELD task is multi-label.
It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\pmb{\Lambda}_S$ are perfectly estimated.

%2 stages: context
The classification method is divided into two stages.
First, the incoming signal is transformed into the log-mel spectrogram and split into TF patches.
Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $k \in \left\lbrace 1 ... K \right\rbrace$, with $K=11$.
Then, we use a single model based on a Convolutional Recurrent Neural Network (CRNN), fed by the TF patches, and outputting probabilities for event classes $k \in \left\lbrace 1 ... K \right\rbrace$, with $K=11$.
Predictions are done at the event-level (not at the frame level), since the temporal activities have been already determined by the front-end.
Only one label is predicted for each incoming event, such that there is no binarization stage needed as in a standard SED task.

%archi=============
The proposed CRNN is depicted in Figure \ref{fig:archi}.
It presents three convolutional \textit{blocks} to extract local features from the input representation.
Each convolutional block consists of one convolutional layer, after which the resulting feature maps are passed through a ReLU non-linearity \cite{nair2010rectified}.
This is followed by a max-pooling operation to downsample the feature maps and add invariance along the frequency dimension.
The target classes vary to a large extent in terms of their temporal dynamics, with some of them being rather impulsive (e.g., \textit{Door slam}), while others being more stationary (e.g., \textit{Phone ringing}). 
Therefore, after stacking the feature maps resulting from the convolutional blocks, this representation is fed into one bidirectional recurrent layer in order to model discriminative temporal structures.
Specifically, 64 nodes of gated recurrent units (GRU) are used with \textit{tanh} activations.
The recurrent layer is followed by a Fully Connected (FC) layer, and finally a 11-way softmax classifier layer produces the event-level probabilities.
Dropout is applied extensively.
The loss function used is categorical cross-entropy.
The model has $\sim$175k weights.

\begin{figure}[ht]
  \vspace{-3mm}
  \centering
  \centerline{\includegraphics[width=\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
  \vspace{-2mm}
  \caption{Back-end architecture.}
  \label{fig:archi}
    \vspace{-5mm}
\end{figure}

%===============================================================================

\section{Experiments}
\label{sec:expe}

\subsection{Dataset, evaluation metrics and baseline system}
\label{ssec:dataset}
We use the TAU Spatial Sound Events 2019 - Ambisonic, which provides first-order ambisonic recordings.
Details about the recording format and dataset specifications can be found in \cite{Adavanne2019_DCASE}.
The dataset features a vocabulary of 11 classes encompassing human sounds and sound events typically found in indoor office environments.
The dataset is split into a development and evaluation sets. 
The development set consists of a four fold cross-validation setup.

The SELD task is evaluated with individual metrics for SED (F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments) and DOA estimation (DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise) \cite{Adavanne2018_JSTSP}.
The \textit{SELD score} is an averaged summary of the system performance.

The baseline system features a CRNN that jointly performs DOA and SED through multi-task learning \cite{Adavanne2018_JSTSP}.
Baseline results are shown in Table \ref{tab:results_real}.

\subsection{Parametric front-end}
\label{ssec:param_frontend}

Based on the method's exploratory analysis, we propose the following set of parameter values, which are shown in Table~\ref{tab:params_front}.

\begin{table}[!htbp]

\centering
\caption{Parameter values for the selected configuration. Top: \textit{DOA analysis} parameters. Bottom: \textit{Association} parameters.}
\begin{tabular}{cccc}
\toprule
Parameter & Unit & Value\\
\midrule
sampling rate & Hz  & 48000\\
STFT window size & sample  & 256\\
STFT window overlap & sample & 128\\
STFT window type & -  &  Hann\\
analysis frequency range & Hz & [0,8000]\\
time average vicinity radius $r$ & bin& 10 \\
diffuseness mask threshold $\Psi_{max}$ & - & 0.5 \\
energy density filter length & bin & 11\\
std mask vicinity radius & bin &  2\\
std mask normalized threshold & - & 0.15\\
median filter minimum ratio $B_{min}$ & - & 0.5 \\
median filter vicinity radius (k,n) & bin & (20, 20) \\
\midrule
frame size $h$ & s  & 0.02\\
resampling minimum valid bins $K_{min}$ & bin & 1 \\
overlapping std threshold $\sigma_{max}$ & degree & 10 \\
grouping maximum angle $d_{max}^\text{ANGLE}$ & degree & 20\\
grouping maximum distance $d_{max}^\text{FRAME}$ & frame & 20\\
event minimum length & frame & 8\\
\bottomrule
\end{tabular}
\label{tab:params_front}
\end{table}

\subsection{Deep learning classification back-end}
\label{ssec:backend_experiments}

We use the provided four fold cross-validation setup.
Training and validation stages use the outcome of an \textit{ideal} front-end, where the groundtruth DOA estimation and activation times are used to feed the beamformer for time-space segmentation.
Conversely, we test the trained models with the signals coming from the \textit{complete} front-end described in Section \ref{sec:method}.

We conducted a set of preliminary experiments with different types of networks including a VGG-like net, a less deep CNN \cite{Fonseca2019learning}, a Mobilenetv1 \cite{howard2017mobilenets} and a CRNN \cite{cakir2017convolutional}.
The latter was found to stand out, and we explore certain facets of the CRNN architecture and the learning pipeline.

Sound events in the dataset last from $\sim$ 0.2 to 3.3 s.
First, clips shorter than 2s are replicated to meet this length.
Then, we compute TF patches of log-mel spectrograms of $T=50$ frames (1 s) and $F=64$ bands.
The values come from the exploration of $T \in \left\lbrace 25,50,75,100\right\rbrace$ and $F \in \left\lbrace 40,64,96,128\right\rbrace$.
$T=50$ is the top performing value, roughly coinciding with the median event duration. In turn, more than 64 bands provide inconsistent improvements, at the cost of increasing the number of network weights.

Regarding the network structure, several variants of the CRNN architecture were explored until reaching the network of Figure \ref{fig:archi}.
This included a small grid search over number of CNN filters, CNN filter size and shape, number of GRU units, number of FC units, dropout \cite{srivastava2014dropout}, learning rate, and the usage of Batch Normalization (BN) \cite{ioffe2015batch}.
Network extensions (involving more weights) were considered only if providing major improvements, as a measure against overfitting.
The main takeaways are: \textit{i)} squared 3x3 filters provide better results than larger filters, \textit{ii)} dropout of 0.5 is critical for overfitting mitigation, \textit{iii)} more than one recurrent layer does not yield improvements, while slowing down training, and \textit{iv)} surprisingly, slightly better performance is attained without BN nor pre-activation \cite{fonseca2018simple}.

For all experiments, the batch size was 100 and Adam optimizer was used \cite{kingma2014adam} with initial learning rate of 0.001, halved each time the validation accuracy plateaus for 5 epochs.
Earlystopping was adopted with a patience of 15 epochs, monitoring validation accuracy.

Prediction for every event was obtained by computing predictions at the patch level, and aggregating them with the geometric mean to produce a clip-level prediction.

Finally, we apply \textit{mixup} \cite{zhang2017mixup} as data augmentation technique.
Mixup consists in creating virtual training examples through linear interpolations in the feature space, assuming that they correspond to linear interpolations in the label space.
Essentially, virtual TF patches are created on the fly as convex combinations of the input training patches, with a hyper-parameter $\alpha$ controlling the interpolation strength.
Mixup has been proven successful for sound event classification, even in adverse conditions of corrupted labels \cite{Fonseca2019model}.
It seems appropriate for this task since the front-end outcome can present leakage due to overlapping sources, effectively mixing two sources while only one training label is available, which can be understood as a form of label noise \cite{Fonseca2019learning}.
Experiments revealed that mixup with $\alpha=0.1$ boosted testing accuracy in $\sim1.5$\%.

%==============================================

\section{Results and Discussion}
\label{sec:results}
\begin{table}[!htbp]
\centering
\caption{Results for development (top) and evaluation (bottom) sets.}

\begin{tabular}{cccccc}
\toprule
Method & \textit{ER} & \textit{F} & \textit{DOA} & \textit{FR} & \textbf{\textit{SELD}}\\
\midrule
Baseline & 0.34 & 79.9\% & 28.5$^{\circ}$  & 85.4\% & 0.2113\\
Proposed & 0.32 & 79.7\% & 9.1$^{\circ}$   & 76.4\% & \textbf{0.2026}\\
Ideal front-end & 0.08 & 93.2\% & $\sim0^{\circ}$   &$ \sim100$\% & 0.0379\\
\midrule
Baseline & 0.28 & 85.4\% & 24.6$^{\circ}$  & 85.7\% &  0.1764\\
Proposed & 0.29 & 82.1\% & 9.3$^{\circ}$   & 75.8\% & \textbf{ 0.1907}\\
\bottomrule
\end{tabular}
\label{tab:results_real}
\end{table}

Table~\ref{tab:results_real} shows the results of the proposed method for both development and evaluation sets, compared to the baseline.
Focusing on evaluation results, our method and the baseline obtain similar performance in SED (\textit{ER} and \textit{F}). However, there is a clear difference in the DOA metrics:
in our method, \textit{DOA} error is reduced by a factor of 2.6, but \textit{FR}  is $\sim10$ points worst.
In terms of \textit{SELD score}, our method performs slightly worse than the baseline in evaluation mode, while marginally outperforming it in development mode.

The most relevant observation is the great improvement in \textit{DOA} error. Results suggest that using spatial audio parametric analysis as a preprocessing step can help to substantially improve localization. 
Figure \ref{fig:5a} provides further evidence for this argument: Challenge methods using some kind of parametric preprocessing (\textit{GCC-PHAT} with the microphone dataset, and \textit{Intensity Vector-Based} in ambisonics) obtained in average better \textit{DOA} error results.

\begin{figure}
  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{Figures/SELD//doaerror2.jpg}
    \caption{\textit{DOA error} across submissions. Hatched bars denote methods using parametric preprocessing. Horizontal lines depict average DOA error accross different subsets: all methods (solid), parametric methods (dashed), non-parametric methods (dotted).}
    \label{fig:5a}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{Figures/SELD//complexity_vs_seld2.jpg}
    \caption{\textit{SELD score} versus complexity.}
    \label{fig:5b}
  \end{subfigure}
\caption{DCASE2019 Challenge Task 3 results, evaluation set.}
\end{figure}

Conversely, the front-end fails regarding \textit{FR}. This is probably due to the complexity added by the association step \cite{Adavanne2018_JSTSP}, and its lack of robustness under highly reverberant scenarios. Including spectral information at the grouping stage might help to improve \textit{FR} --- such information could be provided by the classification back-end, in a similar approach to the baseline system. Another option would be the usage of more sophisticated source counting methods \cite{he2010detecting, Stefanakis2017}.

In order to gain a better insight of the classification back-end performance, Table~\ref{tab:results_real} shows the method results when the testing clips are obtained by feeding  the beamformer  with groundtruth annotations (\textit{ideal} front-end).
In this ideal scenario of DOA performance, the SED metrics show a significant boost. 
This result suggests that the low \textit{FR} given by the front-end has a severe impact on the back-end  performance.
Yet, the proposed system reaches similar performance to the baseline system in terms of SED metrics.

Finally, we would like to discuss algorithm complexity among Challenge methods. 
As depicted in Figure \ref{fig:5b}, there is a general trend towards architectures with very high number of weights, as a consequence of the usage of ensembles and large capacity networks.
Specifically, 66\% of submitted methods employ 1M weights or more, 30\% employ 10M or more, and 15\% employ 100M or more. Such complexities are several orders of magnitude greater than the baseline (150k weights) or the proposed method ($\sim$175k weights).
In this context, our method represents a low-complexity solution to the SELD problem, featuring a number of parameters and a performance comparable to the baseline method.

\section{Conclusion}
\label{sec:conclusion}

We present a novel approach for the SELD task. Our method relies on spatial parametric analysis for the computation of event DOAs, onsets and offsets. This information is used to filter the input signals in time and space, and the resulting event estimations are fed into a CRNN which predicts the class to which the events belong; the classification problem is thereby handled from a simple multi-class perspective.
The proposed method is able to obtain an overall performance comparable to the baseline system. 
The localization accuracy achieved by our method greatly improves the baseline performance, suggesting that spatial parametric analysis might enhance performance of SELD algorithms. Moreover, detection and classification performance in our method suffers from a low Frame Recall; improving this metric could lead to promising SELD scores.

### 2020-03-04 Chapters/ScientificBackground.tex
update

@@
 
-\section{State of the Art}
+\section{Reference system and conventions}
+
+In what follows, we will make use of a right-handed coordinate system, where the positive \textit{x}-axis points towards the \textit{front}, the positive \textit{y}-axis points towards the \textit{left}, and the positive \textit{z}-axis points towards the \textit{zenith} (North Pole).
+
+Any position in the unit sphere may be described in spherical coordinates by two angles: the \textit{inclination} angle $\vartheta$, which accounts for the aperture with respect to the \textit{z}-axis, and the \textit{azimuth} angle $\varphi$, which represents the counter-clockwise angle with respect to the \textit{x}-axis from the top-view.
+The value ranges are $0 \leq \vartheta \leq \pi$ for the inclination, and $0 \leq \varphi \leq 2\pi$ for the azimuth.
+
+Table~\ref{tab:cartesian} shows the spherical coordinate values for some reference points on the unit sphere.
+Notice that the poles ($\vartheta = ±\pi$) are a special case for the spherical coordinate system -- in that case, the azimuth angle is not defined. \\
+
+\begin{table}[!htbp]
+\centering
+\caption{Cartesian and spherical representation of characteristic points along the unit sphere.}
+  \begin{tabular}{cccc}
+  \toprule
+    Position & Cartesian & $\vartheta$ & $\varphi$ \\
+\midrule
+	front & $[1, 0, 0]$ & $\pi/2$ & $0$ \\
+	back & $[-1, 0, 0]$ & $\pi/2$ & $\pi$ \\
+	left & $[0, 1, 0]$ & $\pi/2$ & $\pi/2$ \\
+	right & $[0, -1, 0]$ & $\pi/2$ & $-\pi/2$ \\
+    \textit{zenith} & $[0, 0, 1]$ & $0$ & * \\
+    \textit{nadir}  & $[0, 0, -1]$ & $\pi$ & * \\
+    \bottomrule
+  \end{tabular}
+\label{tab:cartesian}
+\end{table}
+
+The transformation between spherical and cartesian coordinate systems is given by the following relationship:
+\begin{equation}
+	\begin{aligned}
+		x = & cos \varphi sin \vartheta\\
+		y = & sin \varphi sin \vartheta\\
+		z = & cos \vartheta\\
+	\end{aligned}
+\end{equation}
+
+The \textit{elevation} angle $\theta$ provides an alternative way of describing the relationship with respect to the \textit{z}-axis.
+$\theta$ is defined as  the aperture with respect to the \textit{xy}-plane, with positive values towards the positive \textit{z}-axis.
+The relationship between elevation and inclination angles is:
+\begin{equation}
+	\theta = \pi/2 - \vartheta
+\end{equation}
+
+For the sake of compactness, a point in the unit sphere will be often represented by $\pmb{\Omega} = (\vartheta, \varphi)$.
+
+\section{Spherical Harmonics}
+
+\subsection{Definition}
+
+Spherical harmonics are continuous functions defined on the sphere surface.
+Due to their mathematical properties, any spherical function can be decomposed as a combination of spherical harmonics, in what is known as the \textit{Spherical Harmonics Expansion} [Jarrett book, page 17].
+
+%The spherical harmonic of \textit{order} $\ell>0$ and \textit{degree} $m$, with $|m| <= \ell$  in the direction \Omega is defined as:
+Many different spherical harmonic definitions exist in the literature, with minor variations among them.
+In the following, we will use the real-valued, fully normalized spherical harmonics as defined by [Zotter]:
+
+\begin{equation}
+%ambisonics book 186
+	Y_n^m(\varphi, \vartheta) = N_n^{|m|} P_n^{|m|}\cos(\vartheta) \Phi_m(\varphi), 
+\end{equation}
+
+where the \textit{normalization factor} $N_n^m$ is:
+
+\begin{equation}
+%ambisonics book 186
+	N_n^m = (-1)^m \sqrt{\frac{2n+1}{2} \frac{(n-m)!}{(n+m)!}}
+\end{equation}
+
+the \textit{Legendre polynomials} $P_n^{m}$ are defined as:
+
+\begin{equation}
+%ambisonics book 185
+	P_{n+1}^{m} =
+	\begin{cases}
+ 		\frac{2n+1}{n-m+1} x P_n^{m},  &\text{for } n = m,  \\
+ 		\frac{2n+1}{n-m+1} x P_n^{m} - \frac{n+m}{n-m+1}P_{n-1}^{m} &\text{else}, \\
+ 	\end{cases}
+\end{equation}
+
+with $P_{n}^{n} = \frac{(-1)^n(2n)!}{2^nn!}\sqrt{1-x^{2^n}}$
+and the initial term $P_{0}^{0} = 1$,
+
+and $\Phi_m$ is the azimuthal part of the spherical harmonics:
+
+\begin{equation}
+%ambisonics book 176
+	\Phi_m(\varphi) = \frac{1}{\sqrt{2\pi}}
+	\begin{cases}
+    	\sqrt{2} \sin(|m|\varphi),  &\text{for } m < 0,  \\
+    	1,  & \text{for } m = 0,  \\
+    	\sqrt{2} \cos(m\varphi),  & \text{for } m > 0.  \\
+  	\end{cases}
+\end{equation}
+
+One of the properties of the spherical harmonics is orthonormality on the sphere surface:
+
+\begin{equation}
+	\int_{\mathbb{S}^2} Y_n^m(\varphi, \vartheta) Y_{n'}^{m'}(\varphi, \vartheta) \dif cos{\vartheta} \dif \varphi = \delta_{nn'} \delta_{mm'},
+\end{equation}
+
+where $\delta_{xy}$ represents the Kronecker delta operator:
+\begin{equation}
+	\delta_{xy} = \begin{cases}
+		1,  &\text{if } x = y,\\
+		0,  &\text{else}.
+	\end{cases}
+\end{equation}
+
+The spherical harmonics depend on the \textit{order} $n \geq 0$ and the \textit{degree} $m$,  $|m| \leq n$ for each value of $n$.
+In practice, the maximum order $N$, $n \leq N$ determines the spatial resolution of the sound field expansion.\\
+
+Through the spherical harmonic expansion, any sound field may be represented with a limited spatial resolution by the finite combination of all spherical harmonics up to order $N$. 
+For a given order $n$, the number of spherical harmonic functions is $2n+1$. With the accumulation of all orders up to $N$, the total number of spherical harmonics is given by $(N+1)^2$.
+Figure~\ref{fig:sphericalharmonics} depicts all spherical harmonics from orders 0 to 3.
+
+\begin{figure}[hbt]
+  \includegraphics[width=\textwidth]{Figures/ScientificBackground/Spherical_Harmonics_deg3.png}
+  \caption{Spherical harmonics up to order $N=3$. The rows correspond to the spherical harmonics of a given order $n$, and the columns span all possible degree values.}
+  \label{fig:sphericalharmonics}
+\end{figure}
+
+\subsection{Spherical array processing}
+
+Let us consider a sound field captured with a spherical microphone array, which contains $Q$ capsules distributed around a spherical surface of radius $R$ at the positions $\pmb{\Omega}_q, 1 \leq q \leq Q$.
+The captured frequency-domain signals $X_q(k)$ can be represented as the spherical harmonic domain signals $X_n^m(k)$ through the spherical harmonic transform of order $N$ (Moreau et al., 2006):
+
+\begin{equation}
+	X_n^m(k) = \sum_{q=1}^{Q} X_q(k) Y_n^m(\pmb{\Omega}_q) \Gamma_n(kR),
+\end{equation}
+\todo{is that actually valid? or Y should be the complex-valued spherical harmonics?}
+
+where the term $\Gamma_n(kR)$ models the radial transfer function, and depends on the microphone geometry[SOME REFS: moreau, rafaely].
+There are several possible sampling schemes of capsules along the sphere, each one having different properties; the reader is redirected to [rafaely] for a deeper insight.
+\\
+
+By using this model, the maximum spherical harmonic order $N$ that can be retrieved with negligible spatial aliasing depends on the number of microphone capsules [moreau]:
+\begin{equation}
+	N \geq (Q + 1)^2.
+\end{equation}
+
+Furthermore, the sphere radius $R$ has also an effect on the operational bandwidth of the microphone.
+More precisely, for a given spherical harmonic order, the maximum aliasing-free operational frequency is given by [moreau, rafaely]:
+\begin{equation}
+	f_{max} = \frac{N c} {2 \pi R},
+\end{equation}
+
+with $c$ being the sound speed.
+
+\todo{moreau has c/2Rgamma}.
+
+\section{Ambisonics}
+
+Ambisonics is a spatial sound recording and playback technology initially developed during the 1970's, and further expanded into its modern formulation around the 2000s [ZOTTER, page 53].  
+Ambisonics is based on the idea of decomposing a sound field into its spherical harmonic representation.
+
+Originally, the decomposition was limited to first-order spherical harmonics, mainly due to practical limitations [CITE GERZON], as the so-called \textit{First Order Ambisonics} (FOA).
+The technique was later formalized for arbitrary spherical harmonic orders, known as \textit{Higher Order Ambisonics} or HOA [CITE DANIEL].
+In general, with the term Ambisonics we will be referring to the latter definition.\\
+
+Let us consider a sound field composed of a point sound source $s(t)$ located in far-field at the angular position $\pmb{\Omega}_s$. The
+
+One of the main characteristics of Ambisonics as an audio technique is the
+
+%
+%
+%
+%sound field capturing
+%
+%spherical harmonics
+%
+%decoding/beamforming
+%
+%\subsection{normalization, channel ordering}
+%
+%\section{Parametric spatial analysis}
+%
+%sirr, dirac
+%
+%sound field pressure and velocity approximation
+%
+%intensity vector
+%direction of arrival
+%energy density
+%diffuseness (limits)
+
+%alternative formulations
+
+\section{Practical Considerations}
+
+ambisonics, normalization, etc

### 2020-03-04 manuscript.tex
update

@@
 
+%COHERENCE ESTIMATION
+\usepackage{graphicx}% Include figure files
+\usepackage{dcolumn}% Align table columns on decimal point
+\usepackage{bm}% bold math
+\usepackage{latexsym}% 
+\usepackage{lineno}   
+\usepackage{array}
+\usepackage{adjustbox}
+\usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{newpxtext} 
+\usepackage{cite}
+\usepackage{booktabs}
+
+%SELD
+\DeclareMathOperator{\LL}{\langle}
+\DeclareMathOperator{\RR}{\rangle}
+\usepackage{graphicx,url,times,booktabs, tabularx}
+\usepackage{array}
+\usepackage{booktabs}
+\usepackage{algorithm, algorithmic}
+\usepackage{color}
+\usepackage{subcaption}
+
 %WITHOUT HEADINGS: NO MODIFICATION
@@
 %ADD YOUR DATA
-\title{Spherical harmonic domain parametric methods for 3D audio production enhancement\todo{(check capitalization)}}
+\title{Spherical harmonic domain methods for 3D audio production enhancement \todo{(check capitalization)}}
 \subtitle{\todo{subtitle Required?}}
@@
 \department{of Information and Communication Technologies}
-\supervisor{Dr. Emilia Gómez GutiérrezDr. Adan Garriga Torres}
+\supervisor{Dr. Emilia Gómez Gutiérrez\\Dr. Adan Garriga Torres}
 
@@
 \newcommand\todo[1]{\textcolor{red}{#1}}
+
+% INTEGRAL D
+\newcommand{\dif}{\mathop{}\!\mathrm{d}}
 
@@
 %ADD BIBLIOGRAPHY
-\bibliography{exemple_bibtex}
+\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib}
 

### 2020-03-06 Chapters/CoherenceEstimation.tex
updated

@@
 
-\subsection{\label{subsec:3:1} Coherence analysis}
+\subsection{Problem definition}
 
-Diffuseness is commonly estimated through the
-\textit{Magnitude Squared Coherence} (MSC) \cite{elko_spatial_2001} between two frequency-domain signals $S_1$ and $S_2$, as a function of the
-\textit{wavenumber} $k$ and the microphone distance $r$:
-\begin{equation}
-    \text{MSC}_{12}(k r) =
-	\frac{|\left\langle S_1(k r) S_2(k r)^* \right\rangle|^2}
-	{\left\langle|S_1(k r)|^2\right\rangle \left\langle|S_2(k r)|^2\right\rangle},
-    \label{eq:MSC}
-\end{equation}
-where the $\left\langle \cdot \right\rangle$ operator represents the temporal
-expected value, and $^*$ defines the complex conjugate operator. In the case of spherical isotropic noise fields, Eq.~(\ref{eq:MSC})
-can be expressed in terms of microphone directivity patterns
-$T(\phi,\theta,k r)$ as \cite{elko_spatial_2001}:
+Under spherical isotropic noise, the theoretical coherence between any pair of zeroth- and first-order ambisonic virtual microphones is equal to 0 for all frequencies, due to the spherical harmonic orthogonality (Eq.~\ref{eq:orthonormality}) \cite{elko_spatial_2001}.
+This result can also be assessed by Eq.~(\ref{eq:closedform_msc}).
 
-\begin{equation}
-	\begin{aligned}
-&\text{MSC}_{12}(k r) = \frac{|N_{12}(k r)|^2}{|D_{12}(kr)|^2} \\
-&= \frac{|\int_{0}^{\pi} \int_{0}^{2\pi} T_1(\phi,\theta,k r) T_2^*(\phi,\theta,k r) e^{-jk r cos\theta} sin\theta d\theta d\phi|^2}{|\sqrt{ \int_{0}^{\pi} \int_{0}^{2\pi} |T_1(\phi,\theta,k r)|^2 sin\theta d\theta d\phi } \sqrt{\int_{0}^{\pi} \int_{0}^{2\pi}|T_2(\phi,\theta,k r)|^2 sin\theta d\theta d\phi}|^2}.
-\label{eq:MSCdir}
-    \end{aligned}
-\end{equation}
+However, there are several practical factors that might corrupt the coherence estimation, such as the approximation of the temporal expectation by time averaging \cite{thiergart_diffuseness_2011} in Eq.~(\ref{eq:psidefinition}), or the non-ideal implementation of the radial filters $\Gamma_n(kR)$ (Eq.~\ref{eq:a2b}) for the \textit{A-B conversion}\cite{schorkhuber_ambisonic_2017}.
 
-Moreover, the general expression of the directivity of a first-order differential microphone is given by the following relationship:
-
-\begin{equation}
-	\begin{aligned}
-	T_i(\psi_i) = \alpha_i + (1 - \alpha_i) \cos{\psi_i},
-	\end{aligned}
-\end{equation}
-
-where $i \in [1,2]$ is the microphone index, $\psi_i$ is the angle between wave incidence and microphone orientation axis, and $\alpha_i \in [0,1]$ is the directivity parameter of the microphone $i$, which ranges from bidirectional ($\alpha_i = 0$) to omnidirectional ($\alpha_i = 1$).
-\\
-
-For first-order differential microphones, there is a closed-form expression for the numerator and denominator of Eq.~(\ref{eq:MSCdir}):
-\begin{equation}
-	\begin{aligned}
-    &N_{12}(k r) =  \frac{\alpha_1 \alpha_2 sin(kr)}{kr} 
-    + \frac{(1-\alpha_2)(1-\alpha_2)(x_1x_2+y_1y_2)}{(kr)^3}(sin(kr)-kr cos(kr)) \\
-    &+ \frac{z_1 z_2}{kr^3}[ ( (kr)^2 sin(kr) + 2kr cos(kr) )(1-\alpha_1)(1-\alpha_2) 
-    + 2 sin(kr)(1-\alpha_1)(1-\alpha_2) ] \\
-    &+ \frac{z_1}{(kr)^3}[ j(kr)^2 \alpha_2 cos(kr)(\alpha_1-1) + jkr \alpha_2 sin(kr)(1+\alpha_1) ] \\
-    &+ \frac{z_2}{(kr)^3}[ j(kr)^2 \alpha_1 cos(kr)(\alpha_2-1) + jkr \alpha_1 sin(kr)(1+\alpha_2) ],\\
-    &D_{12}(kr) =  \frac{\sqrt{3 \alpha_1^2+(1-\alpha_1)^2}\sqrt{3 \alpha_2^2+(1-\alpha_2)^2}}{3},
-    \label{eq:closedform_msc}
-    \end{aligned}
-\end{equation}
-where: $ x_i = cos(\phi_i) sin(\theta_i); y_i = sin(\phi_i) sin(\theta_i); z_i = cos(\theta_i)$
-refers to the wave incidence angle $\psi_i$ expressed in spherical coordinates (with azimuth $\phi$ and inclination $\theta$).
-
-\subsection{\label{subsec:3:2}Diffuseness estimation in ambisonics}
-Let us consider a sound field captured with a spherical microphone array,
-which contains $Q$ microphones distributed around a spherical surface of radius $R$
-at the angular positions given by the azimuth-inclination pairs
-$\Omega_q = (\phi_q,\theta_q)$.
-The captured frequency-domain signals $ X_q(k)$ can be represented as the spherical harmonic domain signals $X_{mn}(k)$ through the spherical harmonic transform of order $L$ \cite{bertet_3d_2006}:
-\begin{equation}
-        X_{mn}(k) = \sum^{Q} X_q(k) Y_{mn}(\Omega_q) \Gamma_m(kR),
-    \label{eq:ambisonics_encoding}
-\end{equation}
-where $Y_{mn}(\Omega_q)$ are the \textit{real-valued spherical harmonics}, and $\Gamma_m(kR)$ are the
-\textit{radial filters} or equalization terms of order $m$, with $m \in [0,L]$ and $n \in [-m,m]$.\\
-
-Due to a number of practical reasons, it is desirable to distribute the microphone capsules in a uniform manner along the sphere, with the the regular tetrahedron being the simplest possible configuration
- \cite{gerzon_design_1975}.
-Capsule signals recorded with such topology receive the name of \textit{A-Format} signals.
-Conversely, the term \textit{B-Format} (\textit{ambisonics}) describes the application of Eq.~(\ref{eq:ambisonics_encoding}) (\textit{ambisonic encoding}) to the \textit{A-Format} signals.
-One of the most common coherence estimators for first-order ambisonic frequency-domain signals $X_{mn}(k)$ is the \textit{diffuseness} $\Psi$ as defined in \textit{DirAC} \cite{pulkki_directional_2006}:
-
-\begin{equation}
-    \Psi(k) = 1 - \frac{2 ||\left\langle \mathbb{R}\{\bm{X_1}(k)X_0(k)^* \} \right\rangle|| }{ \left\langle||\bm{X_1}(k)||^2 + |X_0(k)|^2\right\rangle},
-    \label{eq:psi}
-\end{equation}
-where $X_0(k) = X_{00}(k)$ 
-and $\bm{X}_1(k) = [X_{1-1}(k), X_{10}(k), X_{11}(k)]^\intercal$ are \textit{SN3D}-normalized.
-For the sake of clarity, we will further define the \textit{B-Format coherence} estimator $\Delta$ as:
-\begin{equation}
-	\Delta(k) = 1 - \Psi(k).
-	\label{eq:delta}
-\end{equation}
-
-Under spherical isotropic noise, the theoretical coherence between any pair of zeroth and first order ambisonic virtual microphones is equal to 0 for all frequencies, due to orthogonality and symmetry of the spherical harmonics \cite{elko_spatial_2001}.
-.
-This result can be also assessed by Eq.~(\ref{eq:closedform_msc}).
-
-However, there are several practical factors that might corrupt the coherence estimation, such as the approximation of the temporal expectation by time averaging \cite{thiergart_diffuseness_2011} in Eq.~(\ref{eq:psi}), or the non-ideal implementation of the radial filters $\Gamma_m(kR)$ \cite{schorkhuber_ambisonic_2017}.
 In the following sections, we present several experiments that illustrate the behavior of different coherence estimators applied on the signals captured with a tetrahedral microphone subjected to spherical isotropic noise, using both simulated and real sound recordings.
@@
 \subsection{Simulation}
-Spherical isotropic noise has been generated following the \textit{geometrical method} \cite{habets_generating_2007, habets_comments_2010}, using $N = 1024$ plane waves.
-The resulting \textit{A-Format} signals correspond to a virtual tetrahedral microphone array mimicking the Ambeo
-\footnote{Sennheiser Ambeo VR Mic. 
+Spherical isotropic noise has been generated following the \textit{geometrical method} \cite{habets_generating_2007, habets_comments_2010}, using $I = 1024$ plane waves.
+The resulting \textit{A-Format} signals correspond to a virtual tetrahedral microphone array mimicking the Ambeo\footnote{Sennheiser Ambeo VR Mic.
 \todo{https://en-us.sennheiser.com/microphone-3d-audio-ambeo-vr-mic}}
@@
 \subsection{Recording}
+
 Spherical isotropic noise has been rendered to a spherical loudspeaker layout with 25 \textit{Genelec 8040}.
-The loudspeakers are arranged into three azimuth-equidistant 8-speaker rings at inclinations $\theta = [\pi/4, \pi/2, 3\pi/4]$, plus one speaker at the zenith ($\theta=0$).
+The loudspeakers are arranged into three azimuth-equidistant 8-speaker rings at inclinations $\vartheta = [\pi/4, \pi/2, 3\pi/4]$, plus one speaker at the zenith.
 The different speaker distances to the center are delay- and gain-corrected, and the signal feeds are equalized to compensate for speaker coloration.
 The room has an approximate $T_{60}$ of 300 ms measured at the 1 kHz third-band octave.
-The spherical isotropic noise has been again created following the \textit{geometrical method}, encoding a number of uncorrelated noise plane waves in ambisonics with varying orders $L \in [1,5]$.
-Due to practical limitations related with the software, the minimum number of sources $N = 256$ for an accurate sound field reconstruction \cite{habets_comments_2010} could not be reached - instead, the analysis has been performed parametrically with $N = [8, 16, 32, 64]$.
-For each value of $L$ and $N$, approximately 15 seconds of audio have been recorded with an Ambeo microphone located at the center of the speaker array.
-Ambisonics decoding uses the AllRAD method \cite{zotter_all-round_2012}, passing through a spherical 64-point 10-design virtual speaker layout, and includes an imaginary speaker at the nadir ($\theta=\pi$).
+
+The spherical isotropic noise has been also created by the \textit{geometrical method}, encoding a number of uncorrelated noise plane waves in ambisonics with varying orders $N \in [1,5]$.
+Due to practical limitations related with the software, the minimum number of sources $I = 256$ for an accurate sound field reconstruction \cite{habets_comments_2010} could not be reached - instead, the analysis has been performed parametrically with $I = [8, 16, 32, 64]$.
+For each value of $N$ and $I$, approximately 15 seconds of audio have been recorded with an Ambeo microphone located at the center of the speaker array.
+
+Ambisonics decoding is performed with an AllRAD decoder, passing through a spherical 64-point 10-design virtual speaker layout, and includes an imaginary speaker at the nadir.
 The decoding matrix uses \textit{in-phase} weights.
@@
 All frequency-domain results have been obtained by averaging their time-frequency representations over time.  
-Ambisonics conversion is performed using \textit{Ambeo A-B converter} AU plugin, version 1.2.1.
+\textit{A-B conversion} has been computed using \textit{Ambeo A-B converter} AU plugin, version 1.2.1.
 
-Two error metrics are considered: the frequency-dependent squared error $\varepsilon(k)$, and the mean squared error $\bar{\varepsilon}$:
+Two error metrics are considered: the frequency-dependent squared error $\varepsilon(k)$:
+\begin{equation}
+	\varepsilon(k) = |X_1(k) - X_2(k)|^2,
+	\label{eq:mse}
+\end{equation}
+
+ and the mean squared error $\bar{\varepsilon}$:
 
 \begin{equation}
-    \varepsilon(k) = |X_1(k) - X_2(k)|^2; \text{         }\bar{\varepsilon} = \frac{1}{K}{\sum_{k=1}^{K} |X_1(k) - X_2(k)|^2}
-    \label{nmse}
+    \bar{\varepsilon} = \frac{1}{K}{\sum_{k=1}^{K} |X_1(k) - X_2(k)|^2}
+    \label{eq:nmse}
 \end{equation}
@@
     \caption{\label{fig:Fig1}\textit{A-Format} coherence between microphone signals.
-Left: $\text{MSC}$ as a function of the frequency of theoretical, simulated and recorded \textit{((BLD,BRU)},  $L=5, N=64$) signals.
-Right: mean error $\bar{\varepsilon}$ of the recorded signals' $MSC$ \textit{(BLD,BRU)} compared to the simulated values, for all values of $L$ and $N$.}
+Left: $\text{MSC}$ as a function of the frequency of theoretical, simulated and recorded \textit{((BLD,BRU)},  $N=5, I=64$) signals.
+Right: mean error $\bar{\varepsilon}$ of the recorded signals' $MSC$ \textit{(BLD,BRU)} compared to the simulated values, for all values of $N$ and $I$.\todo{REDO FIGURE WITH N AND I}}
 \end{figure}
@@
 The squared error $\varepsilon(k)$ with respect to the simulated curve is shown in Fig.~\ref{fig:Fig1} (left), while Fig.~\ref{fig:Fig1} (right) represents the same error averaged over frequency $\bar{\varepsilon}$ for different spatial resolution values of the diffuse field reproduction algorithm.
-As expected, $\bar{\varepsilon}$ decreases with increasing values of $L$ and $N$.
+As expected, $\bar{\varepsilon}$ decreases with increasing values of $N$ and $I$.
 
@@
 
-In order to evaluate the dependency of $\Delta$ on the number of time frames used for averaging, the following procedure is presented.
-The simulated \textit{A-Format} sound field has been transformed into the spherical harmonic domain, with and without the application of radial filters $\Gamma_m(kR)$.
+In order to evaluate the dependency of the \textit{B-Format coherence} $\Delta$ on the number of time frames used for averaging, the following procedure is presented.
+The simulated \textit{A-Format} sound field has been transformed into the spherical harmonic domain, with and without the application of radial filters $\Gamma_n(kR)$ (Eq.~\ref{eq:a2b}).
 Then, $\Delta$ has been computed with Eq.~(\ref{eq:delta}) for exponentially growing values of $r$ between 1 (8 ms) and 2048 (10.92 s), where $r$ is the vicinity radius used for time averaging, and the number of time windows is given by $T = 2r+1$.
@@
 This tendency is better appreciated in Fig.~\ref{fig:Fig2} (right): the curve asymptotically decreases to a value $\Delta_{min}\approx0$.
+
 Another interesting observation comes from the frequency response of the curves.
 For all values of $r$, the coherence of the compensated \textit{B-Format} signal (with $\Gamma_m(kR)$) is roughly flat up to around 7 kHz, which approximately corresponds to the operational spatial frequency range of the microphone \cite{gerzon_design_1975}.
-Above this value, the coherence response looses the flatness due to spatial aliasing.
+Above this value, the coherence response looses the flatness due to spatial aliasing (Eq.~\ref{eq:falias}).
 The response above the maximum frequency could be stabilized, if needed, by alternative diffuseness estimation methods \cite{politis_direction--arrival_2015}.
+
 The coherence level differences along frequency are inversely proportional to $r$ --- the effect is better depicted by the standard deviation values (right).
@@
 This effect might be explained taking into account the inter-channel coherence introduced by microphone and encoder imperfections in real scenarios \cite{schorkhuber_ambisonic_2017}.
+
 As a remark, the comparison between Figs.~\ref{fig:Fig1} and \ref{fig:Fig2} provides evidence that the application of the spherical harmonic transform might be able to yield more accurate diffuseness estimations, due to a better signal conditioning \cite{epain_spherical_2016}.\\
@@
 	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure3}
-	\caption{\label{fig:Fig3}\textit{B-Format} coherence between microphone signals.
-Left: $\Delta$ of simulated and recorded ($L=5, N=64$) signals.
-Right: $\bar{\varepsilon}$ of the recorded signals coherence across all values of $L$ and $N$.}
+	\caption{\label{fig:Fig3}\textit{B-Format coherence} between microphone signals.
+Left: $\Delta$ of simulated and recorded ($N=5, I=64$) signals.
+Right: $\bar{\varepsilon}$ of the recorded signals coherence across all values of $N$ and $I$.\todo{REDO FIGURE} }
 \end{figure}
 
-Figure~\ref{fig:Fig3} (left) shows the estimated coherence for the recorded sound field with $L=5$ and $N=64$, using a vicinity radius of $r=1024$ ($\approx$ 5 s).
+Figure~\ref{fig:Fig3} (left) shows the estimated coherence for the recorded sound field with $N=5$ and $I=64$, using a vicinity radius of $r=1024$ ($\approx$ 5 s).
 The curve is centred around $\Delta=0.25$ and presents several spectral peaks, as in the \textit{A-Format} case. 
 It is important to notice here that the deviations between the coherence of the simulated and the recorded sound fields are much stronger compared to those of Fig.~\ref{fig:Fig1}.
+
 This effect can be also appreciated in Fig.~\ref{fig:Fig3} (right): the mean squared error is around two orders of magnitude higher in \textit{B-Format}.
-Nevertheless, similar as in Fig.~\ref{fig:Fig1} (right), $\bar{\varepsilon}$ decreases with increasing values of $L$ and $N$.
+Nevertheless, similar as in Fig.~\ref{fig:Fig1} (right), $\bar{\varepsilon}$ decreases with increasing values of $N$ and $I$.
 This behavior suggests that the deviations between the recorded and the simulated coherence can be to a large degree explained by the low spatial resolution of the reproduction system; given a higher number of loudspeakers, we expect that the reproduced diffuseness will tend to the theoretical expression.
@@
 In this work, two different metrics of diffuseness have been defined and measured with a tetrahedral microphone subjected to spherical isotropic noise.
+
 The analysis shows, first, the impact of the time-averaging window length on the \textit{B-Format} diffuseness estimator.
 This result might be useful for designing coherence estimators that are parametrized with respect to the length of the analysis window \cite{thiergart_diffuseness_2011}.
+
 Second, the feasibility of diffuse sound field reproduction by a spherical loudspeaker array using ambisonics plane-wave encoding and the \textit{geometrical method} is studied. 

### 2020-03-06 Chapters/ScientificBackground.tex
updated

@@
 
-For the sake of compactness, a point in the unit sphere will be often represented by $\pmb{\Omega} = (\vartheta, \varphi)$.
+For the sake of compactness, a point in the unit sphere will be often represented by $\Omega = (\vartheta, \varphi)$.
 
@@
 	Y_n^m(\varphi, \vartheta) = N_n^{|m|} P_n^{|m|}\cos(\vartheta) \Phi_m(\varphi), 
+	\label{eq:sphericalharmonics}
 \end{equation}
@@
 	\int_{\mathbb{S}^2} Y_n^m(\varphi, \vartheta) Y_{n'}^{m'}(\varphi, \vartheta) \dif cos{\vartheta} \dif \varphi = \delta_{nn'} \delta_{mm'},
+	\label{eq:orthonormality}
 \end{equation}
@@
 Through the spherical harmonic expansion, any sound field may be represented with a limited spatial resolution by the finite combination of all spherical harmonics up to order $N$. 
-For a given order $n$, the number of spherical harmonic functions is $2n+1$. With the accumulation of all orders up to $N$, the total number of spherical harmonics is given by $(N+1)^2$.
+For a given order $n$, the number of spherical harmonic functions is $2n+1$. With the accumulation of all orders up to $N$, the total number of spherical harmonics is given by $M = (N+1)^2$.
 Figure~\ref{fig:sphericalharmonics} depicts all spherical harmonics from orders 0 to 3.
@@
 
-Let us consider a sound field captured with a spherical microphone array, which contains $Q$ capsules distributed around a spherical surface of radius $R$ at the positions $\pmb{\Omega}_q, 1 \leq q \leq Q$.
-The captured frequency-domain signals $X_q(k)$ can be represented as the spherical harmonic domain signals $X_n^m(k)$ through the spherical harmonic transform of order $N$ (Moreau et al., 2006):
+Let us consider a sound field captured with a spherical microphone array, which contains $Q$ capsules distributed around a spherical surface of radius $R$ at the positions $\Omega_q, 1 \leq q \leq Q$.
+The captured frequency-domain signals $X_q(k)$ can be represented as the spherical harmonic domain signals $X_n^m(k)$ through the spherical harmonic transform of order $n$ and degree $m$ (Moreau et al., 2006):
 
 \begin{equation}
-	X_n^m(k) = \sum_{q=1}^{Q} X_q(k) Y_n^m(\pmb{\Omega}_q) \Gamma_n(kR),
+	X_n^m(k) = \sum_{q=1}^{Q} X_q(k) Y_n^m(\Omega_q) \Gamma_n(kR),
+	\label{eq:a2b}
 \end{equation}
@@
 Furthermore, the sphere radius $R$ has also an effect on the operational bandwidth of the microphone.
-More precisely, for a given spherical harmonic order, the maximum aliasing-free operational frequency is given by [moreau, rafaely]:
+More precisely, for a given spherical harmonic order $n$, the maximum aliasing-free operational frequency is given by [moreau, rafaely]:
 \begin{equation}
-	f_{max} = \frac{N c} {2 \pi R},
+	f_{max} = \frac{n c} {2 \pi R},
+	\label{eq:falias}
 \end{equation}
@@
 
+\subsection{Ambisonics Theory}
+
 Ambisonics is a spatial sound recording and playback technology initially developed during the 1970's, and further expanded into its modern formulation around the 2000s [ZOTTER, page 53].  
@@
 
-Let us consider a sound field composed of a point sound source $s(t)$ located in far-field at the angular position $\pmb{\Omega}_s$. The
+\subsubsection{Ambisonic encoding}
 
-One of the main characteristics of Ambisonics as an audio technique is the
+Let us consider a sound field composed of a point sound source $S$ located in far-field at the angular position $\Omega_s$.
+The sound pressure at the coordinate origin can be expressed in terms of the spherical harmonic expansion of order $N$ as: \todo{check equation, find references, how to explain the domain?
+extend also to multiple sources by superposition}
+\begin{equation}
+	P = \sum_{n=0}^{N} \sum_{m=-n}^{n} Y_n^m(\Omega_s) S
+	\label{eq:encoding}
+\end{equation}
+
+The ordered set of values of all spherical harmonics up to order $N$, evaluated at the source position, is known as the \textit{ambisonic coefficients}:
+\begin{equation}
+	\pmb{Y}_n^m(\Omega_s) = [Y_0^0(\Omega_s), Y_1^{-1}(\Omega_s),  \ldots ,  Y_N^N(\Omega_s)]
+	\label{eq:sphericalharmonicvector}
+\end{equation}
+
+Furthermore, the process of multiplying the signal $S$ by the ambisonic coefficients is known in the literature as the \textit{ambisonic encoding}.
+The resulting signal vector is usually referred to as the \textit{ambisonic} (or \textit{B-Format}) signal $\pmb{S}_n^m$:
+\begin{equation}
+	\pmb{S}_n^m = \pmb{Y}_n^m(\Omega_s) S
+\end{equation}
+
+Although the term \textit{B-Format} was initially introduced as an alternative name for first-order ambisonic signals [gerzon, tesis de daniel], it is nowadays common to use it as a synonim of ambisonic signals, without any order restriction.
+We will use the latter acception in what follows.
+
+Historically, the name \textit{B-Format} was used as an opposite of \textit{A-Format}, which describes the signals recorded by a tetrahedral microphone array [cite gerzon?].
+The tetrahedron is the simplest and most common form of spherical microphone arrays (\textit{ambisonic microphones}) with uniform capsule distribution.
+Again, the term \textit{A-Format} is also currently employed for referring to the signals recorded by any spherical microphone array, regardless of the number or arrangement of capsules.
+
+Likewise, the process of signal conversion from the spatial domain (microphone capsules) to the spherical harmonic domain (ambisonic signals), as in Eq.~\ref{eq:a2b}, is known as \textit{A-B conversion}.
+A number of different approaches have been developed for this process, and the interested reader is referred to [\todo{find ref}] for more information.
+
+In practice, there are two alternative ways to generate ambisonic signals.
+The first one is the \textit{synthesis}, based on the direct application of ambisonics encoding (Eq.~\ref{eq:encoding}) to a monophonic signal.
+The second one is the \textit{recording} with a spherical microphone array, followed by the aforementioned domain conversion.
+\\
+
+\subsubsection{Ambisonic Decoding}
+Conversely, the sound field reconstruction is performed by the \textit{ambisonic decoding} operation.
+This process is equivalent to weight-and-sum beamforming in the spherical harmonic domain, and it is sometimes also referred to as the \textit{virtual microphone} technique [ambisonic book].
+
+\todo{change L for other name}
+Let us consider a loudspeaker located at the angular position $\Omega_\ell$. The signal feed $L$ is \textit{decoded} from the ambisonic signal as:
+\begin{equation}
+	L = \sum_{n=0}^{N} \sum_{m=-n}^{n} Y_n^m(\Omega_s) Y_n^m(\Omega_\ell)  S \alpha_n 
+	\label{eq:decoding}
+\end{equation}
+
+where $\alpha_n$ is a weighting factor which accounts for the beam directivity.
+Some of the weightings are widely used for its specific properties, such as \textit{max-rE} or \textit{in-phase} -- the reader is referred to [\todo{find ref.
+daniel?
+zotter?}] for more information.
+
+The decoding equation \ref{eq:decoding} can be written in vector form as:
+\begin{equation}
+	L =  \pmb{S}_n^m {\pmb{Y}_n^{m}(\Omega_\ell)}^T \alpha_n
+\end{equation}
+
+where the superscript $T$ represents the matrix transposition. 
+This equation can be extended to the regular case of decoding to a loudspeaker array, comprised of $L$ loudspeakers located at the positions $\pmb{\Omega}_l = [\Omega_{\ell_1}, \ldots, \Omega_{\ell_L}]$ .
+In such case, the loudspeaker feed vector $\pmb{L}$ can be written as:
+\begin{equation}
+	\pmb{L} = \pmb{S}_n^m \pmb{D},
+\end{equation}
+where 
+\begin{equation}
+	\pmb{D} = \text{diag}(\alpha_n) [{\pmb{Y}_n^{m}(\Omega_{\ell_1})}^T, \ldots, {\pmb{Y}_n^{m}(\Omega_{\ell_L})}^T]
+\end{equation}
+
+is a $M \times L$ matrix known as the \textit{decoding matrix}, and $\text{diag}(\alpha_n)$ is a diagonal matrix of size $M$ containing the values of $\alpha_n$ along the main diagonal. 
+$\pmb{D}$ is frequency-independent and depends solely on the loudspeaker array geometry.
+The reader is referred to [zotter?
+daniel?
+idhoa?] for more information about the vast field of study of ambisonic decoding.
+
+\todo{figure from daniel? to sum up the subsection). Explain the allrad method and stuff because it will appear later on}.
+
+\subsection{Practical considerations}
+
+Due to historical and practical reasons, there are two aspects that must be taking into account when working with ambisonic signals: \textit{channel normalization} and \textit{channel ordering}. 
+In the following, the term \textit{channels} will be used as a synonym for spherical harmonics, as they are usually referred to in sound engineering contexts\footnote{In fact, ambisonic signals are inherently multichannel, even though each channel corresponds to a spherical harmonic, and not to a loudspeaker feed as in traditional \textit{channel-based} audio.}.
+\\
+
+\subsubsection{Channel normalization}
+Let us consider the spherical harmonics $Y_n^m(\Omega)$ as defined in Eq.~\ref{eq:sphericalharmonics}.
+Due to the orthonormal property showed in Eq.~\ref{eq:orthonormality}, they follow the \textit{fully 3d normalized} or \textit{N3D} channel normalization convention.
+\todo{what about the 1/sqrt(4pi)???}.
+
+Alternatively, the \textit{Schmidt 3d semi-normalized} or \textit{SN3D} [daniel] convention is also of widespread usage.
+The conversion between \textit{N3D} and \textit{SN3D} is driven by the following expression:
+\begin{equation}
+	{Y_n^m(\Omega)}^{\text{(N3D)}} = \sqrt{2n+1} {Y_n^m(\Omega)}^{\text{(SN3D)}}
+\end{equation}
+
+\textit{MaxN} is another existing convention. It defines all spherical harmonics as having a maximum absolute value of 1: 
+\begin{equation}
+	\max_{\Omega} |{Y_n^m(\Omega)}^{\text{(MaxN)}}| = 1, \forall (n, m)
+\end{equation}
+
+Finally, the \textit{Furse-Malham} (or \textit{FuMa}) normalization only differs from \textit{Max-N} in the scaling of the zero-th order component: 
+\begin{equation}
+	{Y_n^m(\Omega)}^{\text{(FuMa)}} = \begin{cases}
+		1 / \sqrt{2},  &\text{if } n = 0,\\
+		{Y_n^m(\Omega)}^{\text{(MaxN)}},  &\text{else}.
+	\end{cases}	
+\end{equation}
+
+Each of the normalization schemes has its own particularities.
+For instance, \textit{N3D} is the most mathematically straightforward, and spherical harmonics defined in that way can be directly used for both encoding and decoding (as in Eqs~\ref{eq:encoding} and Eq.~\ref{eq:decoding}) -- however, from a sound engineer point of view, other normalization schemes with maximum values below the unity might be preferred, such as \textit{SN3D}.
+Besides this, \textit{FuMa} has been historically the default normalization \todo{[gerzon?]}, while the more modern \textit{N3D} and \textit{SN3D} were popularized after Daniel \todo{[daniel]}.
+\\
+
+\begin{figure}[hbt!]
+  \includegraphics[width=\textwidth]{Figures/ScientificBackground/normalization.png}
+  \caption{Maximum value of each ambisonic channel up to order 5, for all different normalization schemes. Image from T. Carpentier [\todo{cite}].}
+  \label{fig:normalization}
+\end{figure}
+
+As a summary, Figure~\ref{fig:normalization} displaysthe different normalization schemes. The reader is referred to \todo{[thibaut]} for an extensive review on the topic.
+
+\subsubsection{Channel ordering}
+
+Channel ordering refers to the manner in which spherical harmonics, inherently organized in the 2D space by dimensions $n$ and $m$, are sorted into a one-dimensional vector.
+
+The \textit{ACN} (from \textit{Ambisonic Channel Number}) scheme follows from the mathematical description given in Eq.~\ref{eq:sphericalharmonicvector}.
+The spherical harmonics are first ordered by ascending order $n$ and, inside each order, by ascending degree $m$.
+The index of a given channel $i \in [0 \ldots M-1]$ can be thus obtained by the following relationship:
+\begin{equation}
+	i = n^2 + n + m 
+\end{equation}
+
+Historically, first-order ambisonic audio has followed what it might be called \textit{traditional B-Format} channel ordering [ambisonics in multichannel broadcasting and video]. 
+By following this scheme, the four channels are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
+\begin{equation}
+	{\pmb{Y}_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
+\end{equation}
+with:
+\begin{equation}
+\begin{aligned}
+	&W = Y_0^0(\Omega) \\
+	&X = Y_1^1(\Omega) \\
+	&Y = Y_1^{-1}(\Omega) \\
+	&Z = Y_1^0(\Omega)
+\end{aligned}	
+\end{equation}
+
+This nomenclature was extended to second and third order by Malham, and is currently known as the \textit{Furse-Malham} or \textit{FuMa} channel ordering.
+The channel names use all english alphabet letters from K to Z in third order and, although there would be enough letters to go up to fourth order, the unconvenience of the system was evident [Higher order Ambisonic systems].
+Figure~\ref{fig:normalization} shows the equivalence between \textit{FuMa} (\textit{"letter code"}) and \textit{ACN} channel names.\\
+
+In practice, there exist two main combinations of channel normalization and ordering schemes:
+\begin{itemize}
+  \item The \textit{classical} approach, usually limited to first-order ambisonics, which uses \textit{FuMa} normalization and channel ordering\footnote{In general, it may be expected that \textit{early} ambisonic material follow these conventions without any explicit mention to them.}.
+  \item The \textit{modern} approach, inspired by the \textit{ambix} file format [cite ambix], with \textit{SN3D} normalization and \textit{ACN} channel ordering.
+\end{itemize}
+
+Anyhow, the \textit{classical B-Format} channel naming and ordering is still widely used when referring to first-order ambisonics.
+In what follows, we will use indistinctly both \textit{classical} and \textit{ACN} conventions.
+\todo{is that true?}
+
+\section{Parametric Spatial Audio Analysis}
+
+Trough parametric analysis, sound fields may be described in terms of a small amount of sound sources and associate parameters.
+Such representation might reduce to a great extent the complexity of processing methods [book jarrett].
+
+One of the most successful sound field parametric models is DirAC[pulkki2007 dirac].
+Originally conceived as a method for impulse response processing and spatial sound reproduction [merima2005], it has been widely used in many different audio-related problems [\todo{find some references!}].
+
+DirAC (acronym for \textit{Directional Audio Coding}) is a perceptually motivated time-frequency domain method, based on the assumption that any sound field may be reproduced with high perceptual quality by considering two parameters: the sound field diffuseness and the most prominent sound direction of arrivals (DOAs) [cite parametric book].
+\\
+
+Let us consider a \textit{N3D}-normalized first-order ambisonic signal in time-frequency domain $\pmb{B}_n^m(k, n)$, begin $k$ the frequency index and $n$ the time index.
+For the sake of clarity, we will use \textit{FuMa} channel notation and ordering:
+\begin{equation}
+	\pmb{B}_n^m(k, n) = [W(k, n), X(k, n), Y(k, n), Z(k, n)]
+\end{equation}
+
+Given this representation, we can express the \textit{pressure} $P(k,n)$ of the sound field as:
+\begin{equation}
+	P(k,n) = W(k,n)
+\end{equation}
+as well as the sound \textit{pressure-gradient} (or \textit{velocity}) $\pmb{U}(k,n)$ as: 
+\begin{equation}
+	\pmb{U}(k,n) = - \frac{1}{\rho_0 c} [X(k, n), Y(k, n), Z(k, n)], 
+\end{equation}
+where $\rho_0$ is the mean density of the medium, and $c$ is the speed of sound.\\
+
+The \textit{active intensity} $\pmb{I}(k,n)$, defined as the amount of transmitted acoustic energy, can be expressed in terms of sound pressure and velocity [fahy2002]:
+\begin{equation}
+	\begin{aligned}
+	\pmb{I}(k,n) &=  \Re\{P^*(k,n)\pmb{U}(k,n)\} \\
+	&= - \frac{1}{\rho_0 c}\Re\{W^*(k,n)[X(k,n),Y(k,n),Z(k,n)]\},
+	\end{aligned}
+\end{equation}
+where $^*$ represents the complex conjugate operator. \\
+
+An estimate of the instantaneous DOA $\Omega(k,n)$ can be extracted from the intensity vector, interpreting each of its time-frequency bins as a point in the cartesian space.
+Effectively, the sound propagation direction is the opposite to the observed arrival direction.
+\begin{equation}
+	\Omega(k,n) = \angle(-\pmb{I}(k,n)),
+\end{equation}
+with $\angle$ representing the spherical angle operator of a cartesian vector.
+The result of this computation must be understood as the direction of the net energy flow, which in the case of a single plane-wave will correspond to the source position.
+\\
+
+Another useful parameter is the \textit{energy density} $E(k,n)$ [stazial 1996]:
+\begin{equation}
+	\begin{aligned}
+		E(k,n) &= \frac{|P(k,n)|^2 + \|\pmb{U}(k,n)\|^2}{2\rho_0 c^2} \\ 
+		&=  \frac{|W(k,n)|^2 + \| [X(k,n), Y(k,n), Z(k,n)] \|^2}{2\rho_0 c^2}.
+	\end{aligned}
+\end{equation}\\
+
+Finally, the \textit{diffuseness} $\Psi(k,n)$ can be computed from the sound intensity and  energy density [merimaa2005]:
+\begin{equation}
+	\begin{aligned}
+		\Psi(k,n) &= 1 - \frac{ \| \langle \pmb{I}(k,n) \rangle \| }{ c \langle E(k,n) \rangle } \\
+		&= 1 - 2\frac{ \| \langle \Re\{W^*(k,n)[X(k,n),Y(k,n),Z(k,n)]\} \rangle \| }{ \langle |W(k,n)|^2 + \| [X(k,n), Y(k,n), Z(k,n)] \|^2 \rangle },
+	\end{aligned}
+	\label{eq:psidefinition}
+\end{equation}
+
+ where the symbols $\langle \cdot \rangle$ represent the expectation operator, which is usually implemented as time-domain averaging. \todo{check equation}
+
+Even though Eq.~\ref{eq:psidefinition} (known as \textit{DirAC's diffuseness}) is one of the most common ambisonic diffuseness estimators, several alternative formulations exist. 
+Other diffuseness estimation procedures include the \textit{coefficient of variation method} [ahonen, pulkki, del galdo] and the more recent \textit{COMEDIE} estimator [epain, craig].
+
+In any case, in what follows, the term \textit{diffuseness} and the symbol $\Psi$ will refer by default to Eq.~\ref{eq:psidefinition}.  \\
+
+As a mathematical convenience, we will define the \textit{B-Format coherence} as the complement of the diffuseness:
+\begin{equation}
+	\Delta(k,n) = 1 - \Psi(k,n) 
+	\label{eq:delta}
+\end{equation}
+
+\section{Spatial Coherence Analysis}
+\todo{put this chapter in context or something}\\
+
+In the context of microphone array signal processing, diffuseness is commonly estimated through the
+\textit{Magnitude Squared Coherence} (MSC) \cite{elko_spatial_2001} between two frequency-domain signals $S_1$ and $S_2$, as a function of the
+\textit{wavenumber} $k$ and the microphone distance $r$:
+\begin{equation}
+    \text{MSC}_{12}(k r) =
+	\frac{|\left\langle S_1(k r) S_2(k r)^* \right\rangle|^2}
+	{\left\langle|S_1(k r)|^2\right\rangle \left\langle|S_2(k r)|^2\right\rangle},
+    \label{eq:MSC}
+\end{equation}
+where the $\left\langle \cdot \right\rangle$ operator represents the temporal
+expected value, and $^*$ defines the complex conjugate operator. In the case of spherical isotropic noise fields, Eq.~(\ref{eq:MSC})
+can be expressed in terms of microphone directivity patterns
+$T(\phi,\theta,k r)$ as \cite{elko_spatial_2001}:
+
+\begin{equation}
+	\begin{aligned}
+&\text{MSC}_{12}(k r) = \frac{|N_{12}(k r)|^2}{|D_{12}(kr)|^2} \\
+&= \frac{|\int_{0}^{\pi} \int_{0}^{2\pi} T_1(\phi,\theta,k r) T_2^*(\phi,\theta,k r) e^{-jk r cos\theta} sin\theta d\theta d\phi|^2}{|\sqrt{ \int_{0}^{\pi} \int_{0}^{2\pi} |T_1(\phi,\theta,k r)|^2 sin\theta d\theta d\phi } \sqrt{\int_{0}^{\pi} \int_{0}^{2\pi}|T_2(\phi,\theta,k r)|^2 sin\theta d\theta d\phi}|^2}.
+\label{eq:MSCdir}
+    \end{aligned}
+\end{equation}
+
+Moreover, the general expression of the directivity of a first-order differential microphone is given by the following relationship:
+
+\begin{equation}
+	\begin{aligned}
+	T_i(\Omega_i) = \alpha_i + (1 - \alpha_i) \cos{\Omega_i},
+	\end{aligned}
+\end{equation}
+
+where $i \in [1,2]$ is the microphone index, $\Omega_i$ is the angle between wave incidence and microphone orientation axis, and $\alpha_i \in [0,1]$ is the directivity parameter of the microphone $i$, which ranges from bidirectional ($\alpha_i = 0$) to omnidirectional ($\alpha_i = 1$).
+\\
+
+For first-order differential microphones, there is a closed-form expression for the numerator and denominator of Eq.~(\ref{eq:MSCdir}):
+\begin{equation}
+	\begin{aligned}
+    &N_{12}(k r) =  \frac{\alpha_1 \alpha_2 sin(kr)}{kr} \\
+    &+ \frac{(1-\alpha_2)(1-\alpha_2)(x_1x_2+y_1y_2)}{(kr)^3}(sin(kr)-kr cos(kr)) \\
+    &+ \frac{z_1 z_2}{kr^3}[ ( (kr)^2 sin(kr) + 2kr cos(kr) )(1-\alpha_1)(1-\alpha_2) 
+    + 2 sin(kr)(1-\alpha_1)(1-\alpha_2) ] \\
+    &+ \frac{z_1}{(kr)^3}[ j(kr)^2 \alpha_2 cos(kr)(\alpha_1-1) + jkr \alpha_2 sin(kr)(1+\alpha_1) ] \\
+    &+ \frac{z_2}{(kr)^3}[ j(kr)^2 \alpha_1 cos(kr)(\alpha_2-1) + jkr \alpha_1 sin(kr)(1+\alpha_2) ],\\
+    &D_{12}(kr) =  \frac{\sqrt{3 \alpha_1^2+(1-\alpha_1)^2}\sqrt{3 \alpha_2^2+(1-\alpha_2)^2}}{3},
+    \label{eq:closedform_msc}
+    \end{aligned}
+\end{equation}
+where $x_i$, $y_i$ and $z_i$ are the cartesian coordinates of the wave incidence angle $\Omega_i$. \todo{check}.
+
+\section{Room Acoustics, Impulse Responses}
+
+\section{SOFA}
+ maybe as subsection of room acoustics
 
@@
 \section{Practical Considerations}
-
-ambisonics, normalization, etc

### 2020-06-01 Chapters/Conclusions.tex
update 20200602

@@
 
-  	\item Autoregressive B-Format Late IR Estimation \todo{ref}
+  	\item Blind reverberation time estimation from ambisonic recordings \todo{ref}
   	\begin{itemize}
-  		\item \textbf{Parameter estimation}: Novel technique for RT60 estimation from autoregressive models (subproduct of dereverberation)
-		\item \textbf{Signal enhancement}: Novel methodology to re-reverberate sound scenes (include new elements in the scene using the reverb of the recorded scene)
+  		\item \textbf{Parameter estimation}: Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models
 	\end{itemize}
@@
   	\begin{itemize}
-  		\item \textbf{Scene Description}: ovel State-of-the-Art methodology for Sound Event Localization and Detection	
+  		\item \textbf{Scene Description}: Novel state-of-the-Art methodology for Sound Event Localization and Detection	
   	\end{itemize}
@@
 
-		\item Multichannel Array Signal Processing library \todo{ref}
+		\item A Python library for Multichannel Acoustic Signal Processing \todo{ref}
 		\begin{itemize}
@@
 
-  		\item pysofaconventions \todo{ref}
+  		\item pysofaconventions, a Python API for SOFA \todo{ref}
 		\begin{itemize}

### 2020-06-01 Chapters/ReververationTimeEstimation.tex
update 20200602

\chapter{Blind reverberation time estimation}

% This is the modified v2 as after Emilia's comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%The knowledge about the acoustic properties of an enclosure results of the greatest interest for many situations within the microphone array and acoustic signal processing fields. 
Knowledge about the acoustic properties of an enclosure is a fundamental topic with many applications in the microphone array and acoustic signal processing field.
Problems such as dereverberation \cite{braun2018evaluation} or source separation \cite{gannot2017consolidated} may benefit from this information, 
%or might even require \textit{a priori} acoustic parameter estimations. 
and may require prior estimation of the related parameters.
%Reverberation time $T_{60}$ \cite{kuttruff2016room} might be one of the most widespread acoustic parameters; it represents the time required 
%%for a sound in the diffuse field to decrease its energy envelope by 60 dB. 
%for the reverberant sound field power to decay by 60 dB.
%Reverberation time can be accurately computed from the room geometry \cite{sabine1927collected} or from the Impulse Response (IR) \cite{schroeder1965new}; the problem of $T_{60}$ estimation just from observations of the reverberant signal itself is referred to as the blind reverberation time estimation, and it still remains an open research question.

The 2016 Acoustic Characterisation of Environments (ACE) Challenge \cite{eaton2016estimation} gathered dozens of methods designed for blind $T_{60}$ and Direct-to-Reverberation Ratio (DRR) estimation; nowadays, it is still considered as a state-of-the-art source for 
%method comparison, and the evaluation metrics used have become standard. 
%evaluating performance of BRTE and comparing different methods
performance evaluation and comparison among methods.

Most of the model-based $T_{60}$ estimation algorithms consider the reverberant signal envelope as an exponential decay, so that the problem is reduced to finding a signal offset and estimate the decay rate. 
%For instance, the best scoring $T_{60}$ estimation algorithm, in terms of the Pearson correlation, implements a variant of this idea on the narrowband STFT spectrum \cite{prego2012blind}; this method will be used in next sections as the baseline method. 
Moreover, in last years, data-driven models have outperformed the previous state-of-the-art results \cite{gamper2018blind, looney2020joint, bryan2020impulse}.
%Some representative examples of such approaches are based on multi-layer perceptrons \cite{xiong2018joint} or convolutional neural networks (CNN) \cite{gamper2018blind, looney2020joint, bryan2020impulse}.
A comparative review on single-channel blind $T_{60}$ estimation algorithms was recently published \cite{lollmann2019comparative}. 
%Apart from providing a complete state-of-the-art overview, this work extends the ACE Challenge evaluation to more recent methods.

However, most of the existing reverberation time estimation methods focus on the single-channel case. A representative example can be drawn from the ACE Challenge, where, despite the fact that one of the reverberant datasets was recorded with an \textit{em32 Eigenmike} spherical microphone array, none of the methods use of it for the $T_{60}$ estimation task.

%one of the data subsets was recorded with an \textit{em32 Eigenmike} spherical microphone array, featuring 32 capsules. However, only one of the participants \cite{chen2015estimation} made use of those multichannel recordings, but just for the DRR estimation problem. 
%In a slightly different context, the method proposed in \cite{falcon2019machine} estimates reverberation time from room geometry; while the groundtruth values are computed through spatial filtering of Eigenmike recordings, those are just employed for algorithm training.

On the other hand, recent years have witnessed a growing interest in immersive audio for virtual and augmented reality.
This situation has consolidated Ambisonics \cite{zotter2019ambisonics} as the \textit{de facto} standard for spatial audio. Dedicated spherical microphone arrays have reached the market in last years; their multichannel nature makes possible spatial manipulations that 
%improve traditional signal processing methods.
complement traditional signal enhancement methods.

In this chapter, we present a novel approach to the problem of multichannel blind reverberation time estimation, specifically focusing on first order ambisonic (FOA) recordings. 
%The method is based on dereverberation plus IR estimation by system identification. 
The method is based on a dereverberation stage followed by system identification.
To the best of our knowledge, the proposed algorithm is the first reverberation time estimation method specifically designed for first order ambisonic audio.

The rest of the chapter is organized as follows. Section~\ref{sec:signalmodel} introduces the nomenclature and the signal model. Sections~\ref{sec:baseline} and \ref{sec:proposed} describe the baseline and the proposed methods, respectively. The experimental setup is described in Section~\ref{sec:experimental}, and the results are discussed in Section~\ref{sec:results}. Finally, a conclusion is presented in Section~\ref{sec:conclusion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Signal Model}
\label{sec:signalmodel}

Let us consider a FOA signal $x_m(t)$, with $m = 0 \ldots, M-1$ as the 
%ACN 
channel number, and $M=4$.
Let us further assume that $x_m(t)$ represents the signal captured by an ideal spherical microphone array located in a reverberant enclosure, where a static sound source $s(t)$ is present. The ambisonic room impulse response between source and receiver is represented by $h_m(t)$. In the absence of noise, the recorded signal is the convolutive mix of the source signal $s(t)$ and the RIR, as described in Eq.~\ref{eq:convolutivemixture}:
\begin{equation}
\label{eq:signalmodel}
	x_m(t) = s(t) \ast h_m(t)
\end{equation}

It is important to remark that $T_{60}$ estimation here assumes no receiver directionality. In an ambisonic context, this corresponds to the zeroth order component. Therefore, in what follows, all methods estimating IR parameters will be applied to the zeroth order channel, $x_0(t)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Baseline method}
\label{sec:baseline}

The baseline algorithm, taken from \cite{prego2012blind}, is based on the detection of abrupt event offsets in the time-frequency domain. The subband energy decay on the transitions can be then used to compute an estimate of the full-band decay. This method performed best in the ACE Challenge regarding the Pearson correlation coefficient between estimated and true $T_{60}$\cite{eaton2016estimation}.

Let us consider the zeroth order channel of the recorded signal, $x_0(t)$, and its Short-Time Frequency Transform (STFT) counterpart $X_0(k,n)$, where $k$ and $n$ indicate frequency bin and time frame indices respectively. 
%Although the method could be potentially applied to any of the channels, in physical terms it makes more sense to work with an omnidirectional sound field representation; hence the usage of the zeroth order channel.
The \textit{subband energy} $\bar{E}(k,n)$ of the recorded signal can be expressed as:\begin{equation}
	\bar{E}(k,n) = |X_0(k,n)|^2.
\end{equation}

A \textit{Free Decay Region} (FDR) is defined as a group of consecutive bins within the same subband which exhibit a monotonically decreasing energy. 
A FDR search is performed on the subband energy spectrogram $\bar{E}(k,n)$: for each band, the algorithm tries to find at least one FDR, iterartively reducing the FDR length if no candidates are found.

The next step is the estimation of the reverberation time, which is performed using a subband equivalent of Schroeder's method \cite{schroeder1965new}.
The \textit{Subband Energy Decay Function} (SEDF) associated with a given FDR is computed as:

\begin{equation}
	\bar{c}(k,n) = 10 \log_{10} \frac{\sum_{\nu=n}^{L_c-1} \bar{E}(k,\nu)} {\sum_{\nu=0}^{L_c-1} \bar{E}(k,\nu)} \text{dB},
\end{equation}
where $n = 0 \ldots, L_c-1$ spans the length of the FDR. A linear regression is then performed on each SEDF curve: $T_{60}$ is computed as the time required by the resulting line to reach the $-60 \text{ dB}$ reference.

This procedure yields a $T_{60}$ estimate per FDR. In order to obtain a global estimate, the algorithm proposes a two-step statistical filtering. First, it obtains a narrowband estimate as the median of all estimates within each subband. Then, the resulting broadband value $\bar{T}_{60}$ is computed as the median of all subband estimates.
The last step of the method is the expansion of the resulting dynamic range by a linear mapping. This procedure is required because of the compression introduced by the median operator. The final value $T_{60}$ is thus a linear mapping of $\bar{T}_{60}$, where the parameters $\alpha$ and $\beta$ might be obtained by linear regression on a training stage:
\begin{equation}
	T_{60} = \alpha \bar{T}_{60} + \beta
\label{eq:expansion}
	\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed method}
\label{sec:proposed}

We propose a novel method for reverberation time estimation, based on two steps: signal dereverberation, and system identification. The main idea consist in obtaining an estimate of the dereverberated signal, which is later used for estimating the multichannel IR given the recorded reverberant signal. The reverberation time can be thus computed by the decay slope of the estimated IR.

\subsection{Dereverberation}

Let us consider now the \textit{CTF} model (Eq.~\ref{eq:convolutivetransferfunction}) version of the proposed signal model:
\begin{equation}
\label{eq:ctf}
	X_m(k, n) = \sum_{l=0}^{L_h-1} H_m(k, l) S(k, n-l), 
\end{equation} 
where the multichannel filter $H_m(k, l)$ of length $L_h$ contains the \textit{CTF coefficients} between the source and the microphones.

It is possible to split the former expression into two consecutive elements, which would conceptually match the early and late parts of the room's impulse response:
\begin{equation}
\label{eq:model}
\begin{aligned}
	&X_m(k, n) = D_m(k, n) + R_m(k,n) = \\
	=\sum_{l=0}^{\tau-1} H_m(&k, l) S(k, n-l) + \sum_{l=\tau}^{L_h-1} H_m(k, l) S(k, n-l),
\end{aligned}
\end{equation} 
%\begin{subequations}
%\begin{equation}
%\label{eq:model}
%	X_m(k, n) = D_m(k, n) + R_m(k, n),
%\end{equation} 
%\begin{equation}
%	D_m(k, n) = \sum_{l=0}^{\tau-1} H_m(k, l) S(k, n-l),
%\end{equation} 
%\begin{equation}
%\begin{aligned}
%	R_m(k_n) = \sum_{l=\tau}^{L_h-1} H_m(k, l) S(k, n-l),
%\end{aligned}
%\end{equation} 
%\end{subequations}
where the parameter $\tau$ represents the \textit{mixing time}, which states the transition time between early reflections and late reverberation. 
In other words, the captured signal is split between a \textit{direct} part $D_m(k, n)$, containing the direct path and the early reflections, and a \textit{reverberant} part $R_m(k, n)$, which mainly contains the diffuse part of the reverberation.

Assuming a Multichannel Auto-Regressive (MAR) model, $R_m(k, n)$ can be expressed as a multichannel Infinite Impulse Response (IIR) filter applied to the recorded signal:
\begin{equation} 
\label{eq:mar}
	R_m(k, n) = \sum_{i=1}^{M} \sum_{l=0}^{L_g-1} X_i(k,n-\tau-l) G_{mi}(k, l),
\end{equation}
where the coefficients $G_{mi}(k,l) \in \mathbb{C}$  model the relation between channels $m$ and $i$, and have a length of $L_g$ frames.

By grouping all time frames $n = 1 \ldots, N-1$, it is possible to express Eq.~\ref{eq:mar} in vector notation:
\begin{subequations} 
\begin{equation} 
	\bm{R}_m(k) =  \tilde{\bm{X}}_{\tau}(k) \bm{G}_{m}(k),
\end{equation}
\begin{equation} 
	\tilde{\bm{X}}_{\tau}(k) = [ \tilde{\bm{X}}_{\tau,1}(k), \ldots, \tilde{\bm{X}}_{\tau,M}(k)],
\end{equation}
\end{subequations}
where $\tilde{\bm{X}}_{\tau,m}(k)$ is a $N \times L_g$ matrix, and $\bm{R}_m(k)$ and $\bm{G}_{m}(k)$ are column vectors with lengths $N$ and $L_gM$, respectively.

Finally, the expression can be further simplified by omitting the frequency dependence, and by expressing the channels as columns in the vector notation. Substituting this expression in Eq.~\ref{eq:model} leads to the MAR equation:
\begin{equation} 
\label{eq:D}
	\bm{D} = \bm{X} - \tilde{\bm{X}}_{\tau} \bm{G}.
\end{equation}

Here, the dereverberation problem consists in the estimation of the MIMO filter $\bm{G}$, so that the \textit{clean} signal $\bm{D}$ (containing both direct path and early reflections) can be computed.

The algorithm proposed here is based on the method described in \cite{jukic2015group}. In this case, the dereverberation problem is tackled as an optimization problem, considering that the spectrograms of the reverberant signal are less sparse than those of the corresponding \textit{clean}, and ensuring that the inter-channel signal properties are mantained.
%(hence the reference to the \textit{group} sparsity). 
Although the presented method is applied on the whole signal in \textit{batch} mode, alternative \textit{online} methods could be also used, e.g. \cite{braun2016online}.
%Given that this is a \textit{batch} processing algorithm, the whole audio clip is processed at once. Alternative solutions using online MAR models could be also used if required \cite{braun2016online}.

By using \textit{iteratively reweighted least squares} (IRSL) \cite{chartrand2008iteratively}, it can be shown that an iterative solution for the estimation of $\bm{G}$ at the iteration $(i)$ is given by the following expression:

\begin{equation}
\begin{aligned}
	\label{eq:G}
	\bm{G}^{(i)} = ( \tilde{\bm{X}}_{\tau}^H \bm{W}^{(i)} \tilde{\bm{X}}_{\tau} )^{-1} \tilde{\bm{X}}_{\tau}^H \bm{W}^{(i)} \bm{X}, 
\end{aligned}
\end{equation}
 where $\bm{W}^{(i)}$ is a $N \times N$ diagonal matrix whose diagonal values, $w_n^{(i)}$, can be updated as:
\begin{equation}
\begin{aligned}
\label{eq:w}
	w_n^{(i)} = ( \bm{d}_n^{H (i-1)} \bm{\Phi}^{-1 (i-1)} \bm{d}_n^{(i-1)}  )^{ \frac{p-2}{2} } + \epsilon.
\end{aligned}
\end{equation}

In turn, $\bm{d}_n$ represents the rows of $\bm{D}$ arranged as column vectors of length $M$, $\bm{\Phi}$ is the $M \times M$ Spatial Covariance Matrix (SCM) of $\bm{D}$, $\epsilon$ is an arbitrary small positive value, and $p \leq 1$. The computation and update of the SCM matrix is given by:
\begin{equation}
\begin{aligned}
\label{eq:phi}
	\bm{\Phi}^{(i)} = \frac{1}{N} \bm{D}^{T(i)} \bm{W}^{(i)} \bm{D}^{*(i)}.
\end{aligned}
\end{equation}

%It is interesting to notice that, in the original proposal, the SCM matrix is assigned by default to the identity matrix, and its actual estimation is regarded as optional. However, it is known that the SCM matrix is only equal to the identity matrix in the spherical harmonic domain under isotropic noise \cite{epain2016spherical}. Furthermore, the method has also reported a slightly better performance with the SCM estimation.

To conclude the dereverberation method, Eqs.~\ref{eq:D}, \ref{eq:G}, \ref{eq:w} and \ref{eq:phi} can be applied iteratively, starting by updating Eq.~\ref{eq:w}, until convergence is reached: 
\begin{equation}
 \| \bm{D}^{(i)} - \bm{D}^{(i-1)} \|_F / \| \bm{D}^{(i)} \|_F < \eta, 
\end{equation}
where $\eta$ is an arbitrary small positive value, or alternatively until the maximum number of iterations $i_{max}$ is exceeded. For the initialization, the following values are proposed: $\bm{D} = \bm{X}$ and $\bm{\Phi} = \bm{I}_M$ (the identity matrix of size $M \times M$).\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System Identification}

%The output of the dereverberation step is the multichannel signal $D_m$, which ideally contains the direct plus early reflection components of the source. 
%Therefore, given the reverberant signal $X_m$ and the dereverberated signal $D_m$, an estimate of the late room impulse response might be derived by using System Identification (SID) through the pseudo-inverse method. 
%As in Section~\ref{sec:baseline}, we are primarily interested on the response of the omnidirectional channel; for that reason, the filter estimation is performed with the zeroth order components of both recorded and dereverberated signals: 
%%\begin{equation}
%%	\hat{\bm{H}}_0 = ( \pmb{D}_0^T \pmb{D}_0 )^{-1} \pmb{D}_0^T \bm{X}_0
%%\end{equation}
%%\begin{equation}
%%	\hat{H}_0(k,n) = ( D_0(k,n)^T D_0(k,n) )^{-1} D_0(k,n)^T X_0(k,n)
%%\end{equation}
%\begin{equation}
%	\hat{H}_0 = ( D_0^T D_0 )^{-1} D_0^T X_0
%\end{equation}

The output of the dereverberation step is the multichannel signal $D_m$, which ideally contains the direct plus early reflection components of the source. Therefore, given the reverberant signal $X_m$ and the dereverberated signal $D_m$, an estimate of the late room impulse response might be derived by identifying the filter connecting the two. 
As stated in Section~\ref{sec:signalmodel}, we are primarily interested on the response of the omnidirectional channel; for that reason, the filter estimation is performed with the zeroth order components of both recorded  and dereverberated signals. We perform system identification directly in the STFT through a linear fit between input and output independently for every frequency bin:
\begin{equation}
   \hat{H}_0(k) = \frac{\mathbf{d}^{\mathrm{H}}_0(k) \mathbf{x}^{}_0(k) } {\mathbf{d}^{\mathrm{H}}_0(k) \mathbf{d}^{}_0(k)},
\end{equation}
where $\mathbf{d}_0, \mathbf{x}_0$ are $N\times 1$ length vectors. To avoid complex cross-band modeling of the system response, we use a long STFT window, assumed longer than the twice the length of the IR so that a reduction of the CTF to a Multiplicative Transfer Function (MTF) holds \cite{avargel2007multiplicative}.

As a last sep, the estimated time-frequency filter $\hat{H}_0(k,n)$ is transformed into the time domain filter $\hat{h}(t)$. The $T_{60}$ is then computed by linear fitting of the Schroeder integral 
in the $[-5, -15] $ dB range ($T_{10}$ estimation method), after filtering $\hat{h}(t)$ with an octave-band filter centered at 1 kHz.

%It is important to notice that the first order components of the ambisonic sound field are not used for the system identification step, in a similar manner as described in Section~\ref{sec:baseline}. However, they do play an important role on the dereverberation step; as shown in \cite{jukic2015group}, the number of channels used for the MIMO IIR filter is directly proportional to the dereverberation performance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental setup}
\label{sec:experimental}

\vspace{-2mm}
\subsection{Dataset}

The proposed method is evaluated using two different reverberant datasets, containing recordings of \textit{speech} and \textit{drums} respectively. 
In order to have full control over the reverberation conditions in the experimental setup, the audio clips under consideration have been rendered by the convolutive mixture of clean monophonic recordings with FOA IRs.

The \textit{speech} dataset is composed of the LibriSpeech \cite{panayotov2015librispeech} \textit{test-clean} audio samples longer than 25 s, making a total of 30 audio clips. It contains English language sentences by male and female speakers, often with a small level of background noise. We have used only a 20 s long excerpt of each clip, preceded by an initial offset of 5 s. 
The \textit{drums} dataset is the \textit{test} subset of the isolated drum recordings from the DSD100 dataset \cite{SiSEC16}. It contains 50 different audio clips, covering a wide range of music and mixing styles. The same audio lengths and offsets as in the previous case are applied.

The IRs are FOA room impulse responses simulated by the image method with the \textit{Multichannel Acoustic Signal Processing} library \cite{masp}. There are 9 different IRs of 1 s, with random $T_{60}$ values in the range between 0.4 s and 1.1 s approximately, estimated by the $T_{10}$ method at the 1 kHz band. The angular position of the sources is randomized for each IR, while the receiver position is fixed at the room center, which has a size of $10.2 \times 7.1 \times 3.2$ m. The source distance is set to half the \textit{critical distance}, thus providing positive DRRs.

The combination of the dry audio clips with the IRs yields a total of 270 and 450 audio clips for the \textit{speech} and \textit{drums} datasets, respectively, after removing the audio clips which mostly contain silence. Those datasets will be referred in the following as the \textit{evaluation} datasets.

Finally, the baseline method requires a previous \textit{fitting} step for the computation of the mapping parameters $\alpha$ and $\beta$ from Eq.~\ref{eq:expansion}. The procedure has been performed as follows.
For the \textit{speech} dataset, we selected again the subset of audio clips longer than 25 s, but in this case on the \textit{dev-clean} dataset, which yields a total of 20 audio clips. 
For the \textit{drums} dataset, we used the 50 clips of the \textit{development} subset.
The generation of the convolutive mixes has followed the same procedure as in the previous case. We will refer to the resulting datasets as the \textit{development} datasets.

\subsection{Setup}
\label{sec:setup}

The sampling frequency for all methods is 8 kHz.
For the baseline system, the window size is 1024 samples long, with an overlap of 256 samples. The FDR length is set to 500 ms, which has been reported as the ideal theoretical minimum \cite{prego2012blind}; it corresponds to a FDR length of $L_c = 15$ samples. 
At any frequency band, the value of $L_c$ is iteratively decreased if no FDR is found, until a minimum value of 3 samples (96 ms). If still no FDR is found, the sound clip is discarded.

In order to compute $\alpha$ and $\beta$, we run the baseline method on both \textit{development} datasets. For each IR, the mean and standard deviation of the results are computed across all sound clips. Then, these values are used for a \textit{weighted least squares} linear regression against the true $T_{60}$ values.
The results are shown in Table~\ref{tab:fitting}, where $\sigma$ represents the joint standard deviation of $\alpha$ and $\beta$ after the linear regression; 
the resulting values are in the same range as the values reported in \cite{prego2012blind}.

\begin{table}[t]
\caption{Baseline system: linear regression parameters}
\begin{center}
\begin{tabular}{cccc}
\toprule
Dataset & $\alpha$  & $\beta$  & $\sigma$ \\
\midrule
Speech & 6.6619 & -1.4517 & 0.2131  \\
Drums  & 8.2421 & -2.1939 & 1.0055  \\
\bottomrule
\end{tabular}
\label{tab:fitting}
\end{center}
\end{table}

In the dereverberation stage, the STFT uses a small window size of 128 samples, with 64 samples overlap. The value of $p$ is set to 0.25, given the good results reported in \cite{jukic2015group}. Other parameter values are $\tau=2$, $i_{max} = 10$, $\eta=10^{-4}$ and  $\epsilon=10^{-4}$. After an exploratory search, the length of the IIR filter $L_g = 20$ has been chosen as a compromise between method performance and computation time. 
We have observed a tendency towards poor dereverberation and non-convergence of the IRSL when using small values of $L_g$ and short audios.

For the SID, the recorded and dereverberated signals are reshaped into much larger STFTs, with a window size of 8 s and a hop size of 0.5 s. The predicted filter size is 1 s.

For both \textit{evaluation} datasets, the two presented methods are employed; we will refer to them as \textit{Baseline} and \textit{MAR+SID}. Furthermore, with the aim of evaluating the performance of the SID method in an isolated manner, we have included a third method, \textit{Oracle SID}. As its name suggests, it performs the System Identification step using the true anechoic signal.
%, as an ideal case of perfect dereverberation.

\subsection{Evaluation metrics}
%Regarding method evaluation, 
We have considered the three metrics from the ACE Challenge \cite{eaton2016estimation}, all of them based on the differece between estimated and true values: the \textit{bias}, or mean error; the Mean Squared Error (\textit{MSE}); and the Pearson correlation coefficient. The evaluation has been performed after discarding the outliers, defined as the reverberation time estimates greater than 1.5 s.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

\begin{table}[t]
\caption{Experiment results}
\begin{center}
\begin{tabular}{ccccc}
\toprule
 & \multicolumn{2}{c}{speech} & \multicolumn{2}{c}{drums} \\
%\cline{2-5} 
Metric & \textit{Baseline}&\textit{MAR+SID}&\textit{Baseline}&\textit{MAR+SID}\\
\midrule
Bias  & -0.0599    & \textbf{0.0305}   & \textbf{0.1521}     & 0.2568   \\
MSE   & 0.6366     & \textbf{0.0594}   & \textbf{13.9376}    & 16.5261  \\
$\rho$ & 0.8212    &\textbf{ 0.9848}   & 0.3705    			 & \textbf{0.7552}  \\ 
\hline
\end{tabular}
\label{tab:results}
\end{center}
\end{table}

Figure~\ref{fig:results1} shows the experiment result specified for all audio clips individually. Each boxplot represents the statistics of the mean estimation error (\textit{bias}) for a single audio clip subject to all 9 different IRs. The results are organized by method (rows) and dataset (columns). 
Figure~\ref{fig:results2} aggregates all experiment results into the same plot, showing the statistical distribution of the \textit{bias} per method and dataset. In this case, the \textit{Oracle SID} results are omitted for clarity.
The evaluation metrics for all methods are shown in Table~\ref{tab:results}.

According to the results, the proposed method clearly outperforms the baseline in the \textit{speech} dataset by a tenfold MSE improvement. For the \textit{drums} dataset, our method only outperforms the baseline regarding correlation. 
Nevertheless, an inspection of the statistical distribution of mean estimation errors in Figure~\ref{fig:results2} brings in an interesting observation: 
%the results of the presented method are much more concentrated around zero. In other words, 
the variability of the results given by our method is substantially smaller than the results of the baseline system. This behaviour is consistent across datasets: the mean error distributions with the \textit{speech} dataset are approximately five times narrower than with the \textit{drums} dataset, regardless of the method. 
%The wide distribution of results with the baseline method might be a side effect of the dynamic range expansion of Eq.~\ref{eq:expansion}.

Moreover, all methods behave significantly better on the \textit{speech} dataset. The main reason might be the heterogeneity of the \textit{drums} dataset dataset with respect to dynamic range or timbre, and the potential application of audio effects of any kind.
Furthermore, some audio clips of the \textit{drums} dataset contain sounds with a high degree of self-similarity, such as cymbal rolls or exaggerated reverbs; these characteristics would explain the outliers on the proposed method results.
It is also interesting to notice the robustness of the proposed method against noise, present in the \textit{speech} dataset. 
%Although the signal model does not contemplate noise explicitly, such robustness is consistent with the behaviour described in \cite{jukic2015group}.
Such robustness is consistent with the behavior reported in \cite{jukic2015group}.

The performance of the \textit{ORACLE SID} method is close to ideal. The \textit{bias} is in all cases under 0.05 s (excepting a \textit{drums} clip containing mostly silence).
This result validates the system identification, and allows, in practical terms, a direct evaluation of the proposed method against the groundtruth values.

The results obtained in our analysis are very similar to the results reported in recent deep-learning state-of-the-art proposals, e.g. \cite{gamper2018blind}. 
Since all those methods perform single-channel estimation, and our method requieres FOA recordings, the results are not directly comparable. However, given the similar results obtained with the same evaluation metrics, it might be anticipated that out method may perform as well as other recent data-driven algorithms.
%Although a direct comparison is not possible due to the difference on the datasets under consideration, it might be anticipated that out method may perform as well as other recent data-driven algorithms.

%\begin{figure*}[htbp]
%	\begin{minipage}[b]{1.0\linewidth}
%		\centerline{\includegraphics[width=\textwidth]{Figures/ReverberationTimeEstimation/96.png}}
%		\centerline{(a) Estimation error computed for each audio clip.}\medskip
%	\end{minipage}
%	\begin{minipage}[b]{1.0\linewidth}
%		\centerline{\includegraphics[width=\textwidth]{Figures/ReverberationTimeEstimation/94.png}}
%		\centerline{(b) Total estimation error across audio clips and acoustic conditions. Top: boxplot. Bottom: histogram and density plot.}\medskip
%	\end{minipage}
%	\caption{Experiment results for \textit{speech} (left column) and \textit{drums} (right column) datasets.}
%	\label{fig:results}
%\end{figure*}

\begin{figure*}[htbp]
	\centerline{\includegraphics[width=1.3\textwidth]{Figures/ReverberationTimeEstimation/96.png}}
	\caption{Experiment results for \textit{speech} (left column) and \textit{drums} (right column) datasets. Estimation error computed for each audio clip.}
	\label{fig:results1}
\end{figure*}
\begin{figure*}[htbp]
	\centerline{\includegraphics[width=1.2\textwidth]{Figures/ReverberationTimeEstimation/94.png}}
	\caption{Experiment results for \textit{speech} (left column) and \textit{drums} (right column) datasets. Total estimation error across audio clips and acoustic conditions. Top: boxplot. Bottom: histogram and density plot}
	\label{fig:results2}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}

We have presented in this work a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings. Our method is based on a first dereverberation step,  performed by a multichannel autoregressive model of the late reverberation. The resulting dry signal is then used to estimate the impulse response decay by means of system identification. 
The performance of the method is evaluated in a simulated experimental environment with two different reverberant datasets, and compared against a state-of-the-art method. 
Results show that our method outperforms the baseline method in a majority of evaluation metrics and conditions, and consistently provides results with less variability than the baseline method. 
In future work, we plan to extend the experimental setup by using recorded IRs. Furthermore, the proposed method could be extended to the case of moving sources by using an \textit{online} autoregressive model. 
%Finally, an evaluation of the method in terms of the number of channels or the ambisonic order remains to be done.
Finally, an extension of the method for higher ambisonic orders remains to be done.

### 2020-06-01 Chapters/ScientificBackground.tex
update 20200602

@@
-\chapter{Scientific Background}
+\chapter{Background}
 
-\section{Reference system and conventions}
+\section{Conventions}
+\subsection{Reference system}
 
@@
 For the sake of compactness, a point in the unit sphere will be often represented by $\Omega = (\vartheta, \varphi)$.
+
+\subsection{Nomenclature}
+
+Through the document, we refer to time-domain signals with lowercase, e.g. $x(t)$, with $t$ as the time index.
+
+Time-domain signals transformed by the Short-Time Fourier Transform (STFT) are represented with uppercase, e.g. $X(k,n)$, where $k$ is the frequency bin index, and $n$ the time frame index.
+
+Multichannel signals are in general denoted by a subscript variable index, usually with the letter $m$; for example, $x_m(t)$ or $X_m(k,n)$.
+Signals with an integer subscript index, such as $x_0(t)$, represent a specific channel of the corresponding multichannel signal.
+
+In the context of ambisonic, subscripts and superscripts are used in signal names with a specific meaning; check Section~\ref{sec:ambisonics} for a detailed explanation.
+
+Vector notation is represented with boldface characters, e.g. $\bm{X}(k,n)$.  When used, the way to construct the vectors will be specified.
+
+\todo{revise all vector notations}
 
@@
 \section{Ambisonics}
+\label{sec:ambisonics}
 
@@
 \begin{equation}
-	\pmb{Y}_n^m(\Omega_s) = [Y_0^0(\Omega_s), Y_1^{-1}(\Omega_s),  \ldots ,  Y_N^N(\Omega_s)]
+	Y_n^m(\Omega_s) = [Y_0^0(\Omega_s), Y_1^{-1}(\Omega_s),  \ldots ,  Y_N^N(\Omega_s)]
 	\label{eq:sphericalharmonicvector}
@@
 \begin{equation}
-	\pmb{S}_n^m = \pmb{Y}_n^m(\Omega_s) S
+	S_n^m = Y_n^m(\Omega_s) S
 \end{equation}
@@
 
-\todo{change L for other name}
-Let us consider a loudspeaker located at the angular position $\Omega_\ell$. The signal feed $L$ is \textit{decoded} from the ambisonic signal as:
+Let us consider a loudspeaker located at the angular position $\Omega_p$. The signal feed $P$ is \textit{decoded} from the ambisonic signal as:
 \begin{equation}
-	L = \sum_{n=0}^{N} \sum_{m=-n}^{n} Y_n^m(\Omega_s) Y_n^m(\Omega_\ell)  S \alpha_n 
+	P = \sum_{n=0}^{N} \sum_{m=-n}^{n} Y_n^m(\Omega_s) S Y_n^m(\Omega_\ell)  \alpha_n 
 	\label{eq:decoding}
@@
 where $\alpha_n$ is a weighting factor which accounts for the beam directivity.
-Some of the weightings are widely used for its specific properties, such as \textit{max-rE} or \textit{in-phase} -- the reader is referred to [\todo{find ref.
-daniel?
-zotter?}] for more information.
+Some of the weightings are widely used for its specific properties, such as \textit{max-rE} or \textit{in-phase}.
+Table~\ref{tab:alpha} shows the values of $\alpha_n$ for the different weightings, and Figure~\ref{fig:alpha} plots their first-order directive patterns.
 
-The decoding equation \ref{eq:decoding} can be written in vector form as:
+\begin{table}[t]
+\caption{Ambisonic decoding: standard values of $alpha_n$ weightings. Extracted from \cite{daniel2000representation}.}
+\begin{center}
+\begin{tabular}{cccccc}
+\toprule
+Decoding & $m$ & \multicolumn{4}{c}{$n$} \\ 
+&  & 0 & 1 & 2 & 3 \\
+\midrule
+\textit{basic} & 0 & 1  \\
+	& 1 & 1 & 1  \\
+ 	& 2 & 1 & 1 & 1 \\
+ 	& 3 & 1 & 1 & 1 & 1 \\
+\midrule
+\textit{max-rE} & 0 & 0.577  \\
+	& 1 & 0.775 & 0.4  \\
+ 	& 2 & 0.861 & 0.612 & 0.305 \\
+ 	& 3 & 0.906 & 0.732 & 0.501 & 0.246 \\
+\midrule
+\textit{in-phase} & 0 & 0.333  \\
+	& 1 & 0.5 & 0.1  \\
+ 	& 2 & 0.6 & 0.2 & 0.029 \\
+ 	& 3 & 0.667 & 0.286 & 0.071 & 0.008 \\
+\bottomrule
+\end{tabular}
+\label{tab:alpha}
+\end{center}
+\end{table}
+
+\begin{figure*}[htbp]
+	\begin{center}
+	\begin{minipage}[b]{0.9\linewidth}
+		\centerline{\includegraphics[width=\textwidth]{Figures/Introduction/basic.png}}
+	\end{minipage}
+	\begin{minipage}[b]{0.9\linewidth}
+		\centerline{\includegraphics[width=\textwidth]{Figures/Introduction/maxre.png}}
+	\end{minipage}
+		\begin{minipage}[c]{0.9\linewidth}
+		\centerline{\includegraphics[width=\textwidth]{Figures/Introduction/inphase.png}}
+	\end{minipage}
+	\caption{Directive patterns of first-order ambisonic decoding.}
+	\label{fig:alpha}
+	\end{center}
+\end{figure*}
+
+The decoding equation \ref{eq:decoding} can be written in matrix form as:
 \begin{equation}
-	L =  \pmb{S}_n^m {\pmb{Y}_n^{m}(\Omega_\ell)}^T \alpha_n
+	P = S_n^m {Y_n^m (\Omega_p)}^T \alpha_n
 \end{equation}
@@
 where the superscript $T$ represents the matrix transposition. 
-This equation can be extended to the regular case of decoding to a loudspeaker array, comprised of $L$ loudspeakers located at the positions $\pmb{\Omega}_l = [\Omega_{\ell_1}, \ldots, \Omega_{\ell_L}]$ .
-In such case, the loudspeaker feed vector $\pmb{L}$ can be written as:
+This equation can be extended to the usual case of decoding to a loudspeaker array, comprised of $L$ loudspeakers located at the positions $\Omega_L = [\Omega_{p_1}, \ldots, \Omega_{p_L}]$ .
+In such case, the loudspeaker feed vector $P_L$ can be written as:
 \begin{equation}
-	\pmb{L} = \pmb{S}_n^m \pmb{D},
+	P_L = S_n^m D,
+\label{eq:decodingequation}
 \end{equation}
@@
 \begin{equation}
-	\pmb{D} = \text{diag}(\alpha_n) [{\pmb{Y}_n^{m}(\Omega_{\ell_1})}^T, \ldots, {\pmb{Y}_n^{m}(\Omega_{\ell_L})}^T]
+	D = \text{diag}(\alpha_n) [{Y_n^{m}(\Omega_{p_1})}^T, \ldots, {Y_n^{m}(\Omega_{p_L})}^T]
 \end{equation}
@@
 is a $M \times L$ matrix known as the \textit{decoding matrix}, and $\text{diag}(\alpha_n)$ is a diagonal matrix of size $M$ containing the values of $\alpha_n$ along the main diagonal. 
-$\pmb{D}$ is frequency-independent and depends solely on the loudspeaker array geometry.
-The reader is referred to [zotter?
-daniel?
-idhoa?] for more information about the vast field of study of ambisonic decoding.
+Although the matrix $D$ is frequency-independent and depends solely on the loudspeaker array geometry, in practical scenarios it is usual to include frequency-dependent weightings, $\alpha_n(k)$, to improve the broadband sound field reconstruction \cite{daniel2000representation}.
 
-\todo{figure from daniel? to sum up the subsection). Explain the allrad method and stuff because it will appear later on}.
+Furthermore, sound field reconstruction with Eq.~\ref{eq:decodingequation} is only possible when the loudspeakers are evenly located on the 3D space; in other words, the speaker layout must take the form of one of the five \textit{Platonic solids}: tetrahedron, cube, octahedron, dodecahedron or icosahedron.
+Provided that this condition is usually difficult to fulfil in real scenarios, there are several methods which allow ambisonic decoding for such \textit{irregular} layouts.
+One of the most commonly used is the AllRAD method \todo{cite zotter}.
+AllRAD proposes a two step decoding: first, the ambisonic signal is decoded to a nearly-uniform layout of virtual speakers.
+Then, the signals of the virtual speakers are further distributed into the real speakers by the \textit{Vector-Based Amplitude Panning (VBAP)} method \todo{cite pulkki}.
+
+% The reader is referred to [zotter? daniel? idhoa?] for more information about the vast field of study of ambisonic decoding.
 
@@
 \begin{equation}
-	{\pmb{Y}_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
+	{Y_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
 \end{equation}
@@
 \\
+
+\todo{check here bold vector notation}
 
@@
 
-\section{SOFA}
+ir, partes del ir
+
+%%%%%%%%%%%%%% cogido del paper de MMSP
+Reverberation time $T_{60}$ \cite{kuttruff2016room} might be one of the most widespread acoustic parameters; it represents the time required 
+%for a sound in the diffuse field to decrease its energy envelope by 60 dB. 
+for the reverberant sound field power to decay by 60 dB.
+Reverberation time can be accurately computed from the room geometry \cite{sabine1927collected} or from the Impulse Response (IR) \cite{schroeder1965new}; the problem of $T_{60}$ estimation just from observations of the reverberant signal itself is referred to as the blind reverberation time estimation, and it still remains an open research question.
+
+% TODO
+reverberation time, sabine, schroeder
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{SOFA}
  maybe as subsection of room acoustics
@@
 
+\section{Signal Model}
+
+Let us consider a sound source represented by the signal $s(t)$, located in a given acoustic enclosure.
+Assuming that the system is Linear and Time-Invariant (LTI), it can be represented by its room impulse response $h(t)$.
+The resulting reverberant signal $x(t)$ can be therefore described as the \textit{convolutive mixture} of the source and the RIR:
+\begin{equation}
+	x(t) = s(t) \ast h(t).
+\label{eq:convolutivemixture}
+\end{equation}
+
+When dealing with multichannel room impulse responses, as it is the case in ambisonics, the multichannel reverberant signal $x_m(t)$ is obtained by the convolutive mixture of each RIR channel independently:
+\begin{equation}
+	x_m(t) = s(t) \ast h_m(t).
+\label{eq:convolutivemixturemultichannel}
+\end{equation}
+
+The time domain convolution operation, under certain assumptions, is equivalent to the multiplication in frequency domain. By doing so, Eq.~\ref{eq:convolutivemixture} can be expressed as:
+\begin{equation}
+	X(k,n) = S(k,n) H(k,n).
+\label{eq:multiplicativetransferfunction}
+\end{equation}
+
+Eq.~\ref{eq:multiplicativetransferfunction}, also known as the \textit{Multiplicative Transfer Function (MTF) model} is only valid when the length of the filter $h(t)$ is smaller than the length of the analysis window used in the STFT.
+
+On the contrary, when the filter $h(t)$ spans across several analysis windows, the resulting model is referred to as the \textit{Convolutive Transfer Function (CTF) model}:
+\begin{equation}
+	X(k, n) = \sum_{l=0}^{L_h-1} H(k, l) S(k, n-l), 
+\label{eq:convolutivetransferfunction}
+\end{equation} 
+where $L_h$ is the length of the filter $H(k, n)$ in time frames.
+
+\todo{isotropic noise}
+
 \section{Practical Considerations}
+% TODO I don't remember what is this

### 2020-06-01 manuscript.tex
updated
update 20200602

@@
 \usepackage{booktabs}
+%\usepackage{hyperref}
+%\hypersetup{pdftex,colorlinks=true,allcolors=blue}
+%\usepackage{hypcap}
 
@@
 \usepackage{subcaption}
+
+%\usepackage{hyperref}
 
@@
 \department{of Information and Communication Technologies}
-\supervisor{Dr. Emilia Gómez Gutiérrez\\Dr. Adan Garriga Torres}
+\supervisor{Dr. Adan Garriga Torres\\Dr. Emilia Gómez Gutiérrez}
 
@@
 \input{Chapters/CoherenceEstimation}
-\input{Chapters/AutoregressiveReverbEstimation}
+\input{Chapters/ReververationTimeEstimation}
 \input{Chapters/SELD}
@@
 %ADD BIBLIOGRAPHY
-\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib}
+\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib}
 

### 2020-06-03 Chapters/ReververationTimeEstimation.tex
wip

@@
 
-Let us consider a FOA signal $x_m(t)$, with $m = 0 \ldots, M-1$ as the 
-%ACN 
-channel number, and $M=4$.
-Let us further assume that $x_m(t)$ represents the signal captured by an ideal spherical microphone array located in a reverberant enclosure, where a static sound source $s(t)$ is present.
-The ambisonic room impulse response between source and receiver is represented by $h_m(t)$.
-In the absence of noise, the recorded signal is the convolutive mix of the source signal $s(t)$ and the RIR, as described in Eq.~\ref{eq:convolutivemixture}:
+Let us consider a FOA signal $x_n^m(t)$, with $M=4$ as the number of channels.
+Let us further assume the convolutive mixture signal model described in Eq.~\ref{eq:convolutivemixturemultichannel}, where the reverberant signal $x_n^m(t)$ represents the signal captured by an ideal spherical microphone array located in a reverberant enclosure.
+Let $s(t)$ denote the signal of the only sound source present in the scene, and $h_n^m(t)$ denote the ambisonic RIR modelling the acoustic enclosure:
 \begin{equation}
-\label{eq:signalmodel}
-	x_m(t) = s(t) \ast h_m(t)
+	x_n^m(t) = s(t) \ast h_n^m(t)
 \end{equation}
@@
 
-Let us consider the zeroth order channel of the recorded signal, $x_0(t)$, and its Short-Time Frequency Transform (STFT) counterpart $X_0(k,n)$, where $k$ and $n$ indicate frequency bin and time frame indices respectively.
+Let us consider the zeroth order channel of the recorded signal, $x_0(t)$, and its Short-Time Frequency Transform (STFT) counterpart $X_0(k,n)$.
 %Although the method could be potentially applied to any of the channels, in physical terms it makes more sense to work with an omnidirectional sound field representation; hence the usage of the zeroth order channel.
@@
 
-It is possible to split the former expression into two consecutive elements, which would conceptually match the early and late parts of the room's impulse response:
+Considering the room impulse response model of Eq.~\ref{eq:directreverberant}, 
+it is possible to sequentially split the former expression in the following way:
 \begin{equation}
@@
 where the parameter $\tau$ represents the \textit{mixing time}, which states the transition time between early reflections and late reverberation. 
-In other words, the captured signal is split between a \textit{direct} part $D_m(k, n)$, containing the direct path and the early reflections, and a \textit{reverberant} part $R_m(k, n)$, which mainly contains the diffuse part of the reverberation.
+In other words, the captured signal is divided between a \textit{direct} part $D_m(k, n)$, containing the direct path and the early reflections, and a \textit{reverberant} part $R_m(k, n)$, which mainly contains the diffuse part of the reverberation.
 
@@
 In this case, the \textit{Oracle SID} results are omitted for clarity.
-The evaluation metrics for all methods are shown in Table~\ref{tab:results}.
+The evaluation metrics for all methods are shown in Table~\ref{tab:results}.\\
 
@@
 This behaviour is consistent across datasets: the mean error distributions with the \textit{speech} dataset are approximately five times narrower than with the \textit{drums} dataset, regardless of the method.
+\\
 %The wide distribution of results with the baseline method might be a side effect of the dynamic range expansion of Eq.~\ref{eq:expansion}.
@@
 %Although the signal model does not contemplate noise explicitly, such robustness is consistent with the behaviour described in \cite{jukic2015group}.
-Such robustness is consistent with the behavior reported in \cite{jukic2015group}.
+Such robustness is consistent with the behavior reported in \cite{jukic2015group}.\\
 

### 2020-06-03 Chapters/SELD.tex
wip

@@
 
+\todo{redo figures in this section}
+
 \begin{figure}[h]
@@
 
-The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency (TF) spatial audio analysis.
-Let us consider a first-order ($L=1$) ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
-\begin{equation}
-    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
-\end{equation}
-From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
-\begin{equation}
-    \begin{aligned}
-    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
-    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
-    \end{aligned}
-\end{equation}
-where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.
+The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency spatial audio analysis.
 
-It is desirable to identify the TF regions of $\pmb{\Omega}(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.
+Let us consider a \textit{N3D}-normalized first-order ambisonic signal $x_n^m(t)$. 
+Using the DirAC parametric model, it is possible to obtain instantaneous time-frequency estimates of the Direction of Arrival $\Omega(k,n)$ (Eq.~\ref{eq:doa}), energy density $E(k,n)$ (Eq.~\ref{eq:energydensity}) and the diffuseness $\Psi(k,n)$ (Eq.~\ref{eq:psidefinition}).\\
 
-The first mask is the \textit{energy density mask}, which is used as an activity detector. The energy density $E(k,n)$ is defined as in \cite{stanzial1996reactive} :
-\begin{equation}
-    E(k,n) = \frac{|B_w(k,n)|^2
-    + || [ B_x(k,n), B_y(k,n), B_z(k,n)] ||^2}{2Z_0c},
-\end{equation}
+%Let us consider a first order ambisonic signal $$
+%Let us consider a first order ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
+%\begin{equation}
+%    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
+%\end{equation}
+%From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
+%\begin{equation}
+%    \begin{aligned}
+%    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
+%    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
+%    \end{aligned}
+%\end{equation}
+%where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.
 
-with $c$ being the sound speed. A gaussian adaptive thresholding algorithm is then applied to $E(k,n)$, which selects TF bins with local maximum energy density, as expected from direct sounds.
+It is desirable to identify the TF regions of $\Omega(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.
 
-The \textit{diffuseness mask} selects the TF bins with high energy propagation. Diffuseness $\Psi(k,n)$ is defined in \cite{merimaa2005spatial} as:
-\begin{equation}
-    \Psi(k,n) = 1 - ||\LL \pmb{I}(k,n) \RR|| / (c \LL E(k,n) \RR),
-    \label{psi}
-\end{equation}
-where $\LL \cdot \RR$ represents the temporal expected value.
+The first mask is the \textit{energy density mask}, which is used as an activity detector.
+A gaussian adaptive thresholding algorithm is applied to $E(k,n)$.
+This procedure selects TF bins with local maximum energy density, as expected from the direct path of the sources.
+
+The \textit{diffuseness mask} selects the TF bins with high energy propagation.
+Bins with a low diffuseness value represent a TF region where one sound source is predominant.
+A fixed threshold value $\Psi_{max}$ is used for the masking.
 
 The third mask is the \textit{DOA variance mask}.
-It tries to select TF regions with small standard deviation\footnote{In this work, all statistical operators for angular position refer to the $2\pi$-\textit{periodic} operator for azimuth, and the standard operator for elevation.} with respect to their neighbor bins---a characteristic of sound fields with low diffuseness \cite{pulkki2018parametric}.
+It tries to select TF regions with small standard deviation with respect to their neighbor bins---a characteristic of sound fields with low diffuseness \cite{pulkki2018parametric}.\\
 
-The three masks are then applied to the DOA estimation, obtaining the TF-filtered DOAs $\bar{\pmb{\Omega}}(k,n)$.
+The three masks are then applied to the DOA estimation, obtaining the TF-filtered instanteneous DOAs $\bar{\Omega}(k,n)$. 
 Finally, a median filter is applied, with the aim of improving DOA estimation consistency and removing spurious TF bins.
-The median filter is applied in a TF bin belonging to $\bar{\pmb{\Omega}}(k,n)$ only if the number of TF bins belonging to $\bar{\pmb{\Omega}}(k,n)$  in its vicinity is greater than a given threshold $B_{min}$.
-The resulting filtered DOA estimation is referred to as $\breve{\pmb{\Omega}}(k,n)$.
+The median filter is applied in a TF bin belonging to $\bar{\Omega}(k,n)$ only if the number of TF bins belonging to $\bar{\\Omega}(k,n)$ in its vicinity is greater than a given threshold $B_{min}$.
+The resulting filtered DOA estimation is referred to as $\breve{\Omega}(k,n)$.
 
@@
 
-The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\pmb{\Omega}}(k,n)$ to a set of events, each one having a specific onset, offset  and location.
+The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\Omega}(k,n)$ to a set of events, each one having a specific onset, offset and location.
+\\
 
-First, DOA estimates are resampled into \textit{frames} of the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
-An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\pmb{\Omega}}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is greater than a threshold $K_{min}$.
+First, DOA estimates are resampled into \textit{frames} with the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
+An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\Omega}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is higher than a given threshold $K_{min}$.\\
 
-Next, the standard deviation in azimuth ($\sigma_{\varphi}$) and elevation ($\sigma_{\theta} $) of the frame-based DOA estimates $\breve{\pmb{\Omega}}(k,m)$ are compared to a threshold value ($\sigma_{max}$), and the result is used to estimate the frame-based event overlapping amount $o(m)$ :
+Next, the standard deviation in azimuth $\sigma_{\varphi}$ and elevation $\sigma_{\theta}$ of the frame-based DOA estimates $\breve{\Omega}(k,m)$ are compared to a threshold value $\sigma_{max}$, and the result is used to estimate the frame-based event overlapping amount $o(m)$ :
 
@@
 \end{equation}
-The clustered values $\pmb{\Omega}_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\pmb{\Omega}}(k,m)$, using a modified version of K-Means which minimizes the central angle distance.
-Notice that, for $o(m)=1$, the operation is equivalent to the median.
+
+The clustered values $\Omega_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\Omega}(k,m)$, using a modified version of K-Means which minimizes the central angle distance.
+Notice that, for $o(m)=1$, the operation is equivalent to the median.\\
 
 The following step is the grouping of clustered DOA values into events.
-Let us define $\pmb{\Omega}_S(m)$ as the frame-wise DOA estimations belonging to the event $S$.
-A given clustered DOA estimation $\pmb{\Omega}_{\text{cluster}}(m)$ belongs to the event $S$ if the following criteria are met:
+Let us define $\Omega_S(m)$ as the frame-wise DOA estimations belonging to the event $S$.
+A given clustered DOA estimation $\Omega_{\text{cluster}}(m)$ belongs to the event $S$ if the following criteria are met:
 \begin{itemize}
-    \item The central angle between  $\pmb{\Omega}_{\text{cluster}}(m)$ and the median of $\pmb{\Omega}_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
-    \item The frame distance between M and the closest frame of $\pmb{\Omega}_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
+    \item The central angle between  $\Omega_{\text{cluster}}(m)$ and the median of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
+    \item The frame distance between $m$ and the closest frame of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
 \end{itemize}
 
-The resulting DOAs $\pmb{\Omega}_S(m)$ are subject to a postprocessing step with the purpose of 
+The resulting DOAs $\Omega_S(m)$ are subject to a postprocessing step with the purpose of 
 delaying event onsets in frames where $o(m) > 2$, and discarding events shorter than a given minimum length.
-Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\pmb{\Lambda}_S = (\pmb{\Omega}_S, \text{onset}_S, \text{offset}_S)$.
+
+Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\Lambda_S = (\Omega_S, \text{onset}_S, \text{offset}_S)$.
 
@@
 The last step performed in the front-end is the input signal segmentation.
-The spatial and temporal information provided by the annotations $\pmb{\Lambda}_S$ are used to produce monophonic signal estimations of the events, $\tilde{b}_S(t)$, as the signals captured by a virtual hypercardioid:
+The spatial and temporal information provided by the annotations $\Lambda_S$ are used to produce monophonic signal estimations of the events, $\tilde{x}_S(t)$, as the signals captured by a virtual first-order cardioid (Eq.~\ref{eq:decodingequation}):
 \begin{equation}
-\tilde{b}_S(t) = \pmb{Y}(\pmb{\Omega}_S) \pmb{b}^\intercal(t),
+\tilde{x}_S(t) = x_n^m(t) Y_n^m(\Omega_S) \alpha_n,
 \end{equation}
-where $ \pmb{Y}(\pmb{\Omega}_S) = [Y_w(\pmb{\Omega}_S), Y_x(\pmb{\Omega}_S), Y_y(\pmb{\Omega}_S), Y_z(\pmb{\Omega}_S)]$ is the set of real-valued spherical harmonics up to order $L=1$ evaluated at $\pmb{\Omega}_S$.
+with $\alpha_n = [1, 1, 1, 1]$ using the \textit{basic} decoding weights.
 
@@
 %intro - problem formulation==============
-The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{b}_S(t)$.
+The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{s}_S(t)$.
 Then, the back-end classifies the resulting signals as belonging to one of a target set of 11 classes.
 Therefore, the multi-task nature of the front-end allows us to define the back-end classification task as a simple multi-class problem, even though the original SELD task is multi-label.
-It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\pmb{\Lambda}_S$ are perfectly estimated.
+It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\Lambda_S$ are perfectly estimated.
 
@@
 First, the incoming signal is transformed into the log-mel spectrogram and split into TF patches.
-Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $k \in \left\lbrace 1 ...
-K \right\rbrace$, with $K=11$.
-Then, we use a single model based on a Convolutional Recurrent Neural Network (CRNN), fed by the TF patches, and outputting probabilities for event classes $k \in \left\lbrace 1 ...
-K \right\rbrace$, with $K=11$.
+Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $c \in \left\lbrace 1 ...C \right\rbrace$, with $C=11$.
 Predictions are done at the event-level (not at the frame level), since the temporal activities have been already determined by the front-end.
@@
   \centering
-  \centerline{\includegraphics[width=\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
+  \centerline{\includegraphics[width=0.9\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
   \vspace{-2mm}
@@
 Details about the recording format and dataset specifications can be found in \cite{Adavanne2019_DCASE}.
-The dataset features a vocabulary of 11 classes encompassing human sounds and sound events typically found in indoor office environments.
+The dataset features a vocabulary of $C=11$ classes encompassing human sounds and sound events typically found in indoor office environments.
 The dataset is split into a development and evaluation sets. 
-The development set consists of a four fold cross-validation setup.
+The development set consists of a four fold cross-validation setup.\\
 
-The SELD task is evaluated with individual metrics for SED (F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments) and DOA estimation (DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise) \cite{Adavanne2018_JSTSP}.
+The SELD task is evaluated with individual metrics for the SED and DOA problems:
+\begin{itemize}
+	\item SED: F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments.
+	\item DOA: DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise.
+\end{itemize}
 The \textit{SELD score} is an averaged summary of the system performance.
+A more detailed description of the evaluation metrics can be found in \cite{Adavanne2018_JSTSP}.\\
 
@@
 We conducted a set of preliminary experiments with different types of networks including a VGG-like net, a less deep CNN \cite{Fonseca2019learning}, a Mobilenetv1 \cite{howard2017mobilenets} and a CRNN \cite{cakir2017convolutional}.
-The latter was found to stand out, and we explore certain facets of the CRNN architecture and the learning pipeline.
+The latter was found to stand out, and we explore certain facets of the CRNN architecture and the learning pipeline.\\
 
@@
 $T=50$ is the top performing value, roughly coinciding with the median event duration.
-In turn, more than 64 bands provide inconsistent improvements, at the cost of increasing the number of network weights.
+In turn, more than 64 bands provide inconsistent improvements, at the cost of increasing the number of network weights.\\
 
@@
 The main takeaways are: \textit{i)} squared 3x3 filters provide better results than larger filters, \textit{ii)} dropout of 0.5 is critical for overfitting mitigation, \textit{iii)} more than one recurrent layer does not yield improvements, while slowing down training, and \textit{iv)} surprisingly, slightly better performance is attained without BN nor pre-activation \cite{fonseca2018simple}.
+\\
 
@@
 Earlystopping was adopted with a patience of 15 epochs, monitoring validation accuracy.
-
-Prediction for every event was obtained by computing predictions at the patch level, and aggregating them with the geometric mean to produce a clip-level prediction.
+Prediction for every event was obtained by computing predictions at the patch level, and aggregating them with the geometric mean to produce a clip-level prediction. \\
 
@@
 in our method, \textit{DOA} error is reduced by a factor of 2.6, but \textit{FR}  is $\sim10$ points worst.
-In terms of \textit{SELD score}, our method performs slightly worse than the baseline in evaluation mode, while marginally outperforming it in development mode.
+In terms of \textit{SELD score}, our method performs slightly worse than the baseline in evaluation mode, while marginally outperforming it in development mode. \\m,
 
@@
 Figure \ref{fig:5a} provides further evidence for this argument: Challenge methods using some kind of parametric preprocessing (\textit{GCC-PHAT} with the microphone dataset, and \textit{Intensity Vector-Based} in ambisonics) obtained in average better \textit{DOA} error results.
+\\
 
@@
 Another option would be the usage of more sophisticated source counting methods \cite{he2010detecting, Stefanakis2017}.
+\\
 
@@
 This result suggests that the low \textit{FR} given by the front-end has a severe impact on the back-end  performance.
-Yet, the proposed system reaches similar performance to the baseline system in terms of SED metrics.
+Yet, the proposed system reaches similar performance to the baseline system in terms of SED metrics. \\
 

### 2020-06-03 Chapters/ScientificBackground.tex
wip

@@
 	\begin{aligned}
-		x = & cos \varphi sin \vartheta\\
-		y = & sin \varphi sin \vartheta\\
-		z = & cos \vartheta\\
+		x = & \cos{\varphi} \sin{\vartheta}\\
+		y = & \sin{\varphi} \sin{\vartheta}\\
+		z = & \cos{\vartheta}\\
 	\end{aligned}
@@
 
-For the sake of compactness, a point in the unit sphere will be often represented by $\Omega = (\vartheta, \varphi)$.
+For the sake of compactness, a point in the unit sphere will be often represented by $\Omega = (\vartheta, \varphi)$.\\
+
+Given the periodic nature of the azimuth angle, the descriptive statistic operations applied to $\vartheta$ will refer to the $2\pi$-\textit{periodic} version or the operator; this situation does not affect the inclination/elevation coordinate.
 
@@
 Let us consider a sound field composed of a point sound source $S$ located in far-field at the angular position $\Omega_s$.
-The sound pressure at the coordinate origin can be expressed in terms of the spherical harmonic expansion of order $N$ as: \todo{check equation, find references, how to explain the domain?
+The sound pressure at the coordinate origin $P$ can be expressed in terms of the spherical harmonic expansion of order $N$ as: \todo{check equation, find references, how to explain the domain?
 extend also to multiple sources by superposition}
@@
 Furthermore, the process of multiplying the signal $S$ by the ambisonic coefficients is known in the literature as the \textit{ambisonic encoding}.
-The resulting signal vector is usually referred to as the \textit{ambisonic} (or \textit{B-Format}) signal $\pmb{S}_n^m$:
+The resulting signal vector is usually referred to as the \textit{ambisonic} (or \textit{B-Format}) signal $S_n^m$:
 \begin{equation}
@@
 
-Let us consider a loudspeaker located at the angular position $\Omega_p$. The signal feed $P$ is \textit{decoded} from the ambisonic signal as:
+Let us consider a loudspeaker located at the angular position $\Omega_p$. In accordance with Eq.~\ref{eq:encoding}, the signal feed $P$ is \textit{decoded} from the ambisonic signal as:
 \begin{equation}
@@
 where $\alpha_n$ is a weighting factor which accounts for the beam directivity.
-Some of the weightings are widely used for its specific properties, such as \textit{max-rE} or \textit{in-phase}.
-Table~\ref{tab:alpha} shows the values of $\alpha_n$ for the different weightings, and Figure~\ref{fig:alpha} plots their first-order directive patterns.
+There are several standard weightings used for different purposes; their values are shown in Table~\ref{tab:alpha}, and the first-order directive patterns are plotted in Figure~\ref{fig:alpha}.
 
@@
 \toprule
-Decoding & $m$ & \multicolumn{4}{c}{$n$} \\ 
+Decoding & $N$ & \multicolumn{4}{c}{$n$} \\ 
 &  & 0 & 1 & 2 & 3 \\
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
 	\begin{center}
@@
 	\end{center}
-\end{figure*}
+\end{figure}
 
@@
 	P = S_n^m {Y_n^m (\Omega_p)}^T \alpha_n
+\label{eq:decodingequation}
 \end{equation}
@@
 	P_L = S_n^m D,
-\label{eq:decodingequation}
+\label{eq:decodingequation2}
 \end{equation}
@@
 
-Furthermore, sound field reconstruction with Eq.~\ref{eq:decodingequation} is only possible when the loudspeakers are evenly located on the 3D space; in other words, the speaker layout must take the form of one of the five \textit{Platonic solids}: tetrahedron, cube, octahedron, dodecahedron or icosahedron.
+Furthermore, sound field reconstruction with Eq.~\ref{eq:decodingequation2} is only possible when the loudspeakers are evenly located on the 3D space; in other words, the speaker layout must take the form of one of the five \textit{Platonic solids}: tetrahedron, cube, octahedron, dodecahedron or icosahedron.
 Provided that this condition is usually difficult to fulfil in real scenarios, there are several methods which allow ambisonic decoding for such \textit{irregular} layouts.
@@
 Historically, first-order ambisonic audio has followed what it might be called \textit{traditional B-Format} channel ordering [ambisonics in multichannel broadcasting and video]. 
-By following this scheme, the four channels are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
+By following this scheme, the four channels of a FOA signal $S_n^m$ are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
 \begin{equation}
-	{Y_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
+	{S_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
+\label{eq:fumaordering}
 \end{equation}
@@
 \begin{aligned}
-	&W = Y_0^0(\Omega) \\
-	&X = Y_1^1(\Omega) \\
-	&Y = Y_1^{-1}(\Omega) \\
-	&Z = Y_1^0(\Omega)
+	&W = S_0^0(\Omega) \\
+	&X = S_1^1(\Omega) \\
+	&Y = S_1^{-1}(\Omega) \\
+	&Z = S_1^0(\Omega)
 \end{aligned}	
@@
 Anyhow, the \textit{classical B-Format} channel naming and ordering is still widely used when referring to first-order ambisonics. 
-In what follows, we will use indistinctly both \textit{classical} and \textit{ACN} conventions.
-\todo{is that true?}
+%In what follows, we will use indistinctly both \textit{classical} and \textit{ACN} conventions. \todo{is that true?}
 
@@
 
-DirAC (acronym for \textit{Directional Audio Coding}) is a perceptually motivated time-frequency domain method, based on the assumption that any sound field may be reproduced with high perceptual quality by considering two parameters: the sound field diffuseness and the most prominent sound direction of arrivals (DOAs) [cite parametric book].
+DirAC (acronym for \textit{Directional Audio Coding}) is a perceptually motivated time-frequency (TF) domain method, based on the assumption that any sound field may be reproduced with high perceptual quality by considering two parameters: the sound field diffuseness and the most prominent sound \textit{Direction-of-arrival} (DOA) [cite parametric book].
 \\
 
-\todo{check here bold vector notation}
-
-Let us consider a \textit{N3D}-normalized first-order ambisonic signal in time-frequency domain $\pmb{B}_n^m(k, n)$, begin $k$ the frequency index and $n$ the time index.
-For the sake of clarity, we will use \textit{FuMa} channel notation and ordering:
+Let us consider a \textit{N3D}-normalized first-order ambisonic signal in time-frequency domain, $S_n^m(k, n)$. 
+For the sake of clarity, we will use in this section \textit{FuMa} channel notation and ordering (Eq.~\ref{eq:fumaordering}):
 \begin{equation}
-	\pmb{B}_n^m(k, n) = [W(k, n), X(k, n), Y(k, n), Z(k, n)]
+	S_n^m(k, n) = [W(k, n), X(k, n), Y(k, n), Z(k, n)]
 \end{equation}
@@
 	\Omega(k,n) = \angle(-\pmb{I}(k,n)),
+\label{eq:doa}
 \end{equation}
@@
 	\end{aligned}
+	\label{eq:energydensity}
 \end{equation}\\
@@
 
-In any case, in what follows, the term \textit{diffuseness} and the symbol $\Psi$ will refer by default to Eq.~\ref{eq:psidefinition}.  \\
+In any case, in what follows, the term \textit{diffuseness} and the symbol $\Psi$ will refer by default to Eq.~\ref{eq:psidefinition}.
 
@@
 \end{equation}
+
+In conclusion, Figure~\ref{fig:dirac} plots the spectrograms of the DOA $\Omega(k,n)$ and diffuseness $\Psi(k,n)$ of a FOA recording, which consists of a sound source located at the front, plus a moderate amount of reverberation and background noise.
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=\textwidth]{Figures/Introduction/spectrograms.png}
+	\caption{Parametric time-frequency spatial audio analysis of a first order ambisonic recording.
+From top to bottom: 1.) Magnitude spectrogram of the omnidirectional channel.
+2.) and 3.) Azimuth and elevation of the estimated instantaneous narrowband DOAs $\Omega(k,n)$.
+4.) Instantaneous narrowband diffuseness $\Psi(k,n)$.}
+	\label{fig:dirac}
+	\end{center}
+\end{figure}
 
@@
 
-\section{Room Acoustics, Impulse Responses}
+\section{Reverberation}
 
-ir, partes del ir
+In the context of room acoustics, reverberation refers to \textit{"the energy of a sound source that reaches a listener indirectly, by reflecting from surfaces within the surrounding space occupied by the sound source and the listener"} \cite{begault20003}.
+Conversely, in anechoic or free-field conditions, where reverberation is not present, only the direct path of the sound source exists.
+Assuming linearity and time-invariance, room reverberation can be fully characterised by its impulse response (IR). \\
 
-%%%%%%%%%%%%%% cogido del paper de MMSP
-Reverberation time $T_{60}$ \cite{kuttruff2016room} might be one of the most widespread acoustic parameters; it represents the time required 
-%for a sound in the diffuse field to decrease its energy envelope by 60 dB. 
-for the reverberant sound field power to decay by 60 dB.
-Reverberation time can be accurately computed from the room geometry \cite{sabine1927collected} or from the Impulse Response (IR) \cite{schroeder1965new}; the problem of $T_{60}$ estimation just from observations of the reverberant signal itself is referred to as the blind reverberation time estimation, and it still remains an open research question.
+Reverberation models often consider two differentiated parts of the reverberant tail, based on both physical and perceptual characteristics: the \textit{early reflections} and the \textit{late reverberation}.
+Early reflections, as the name suggests, refers to the individual sound paths arriving to the listener after a few reflections on the room surfaces, which cause some degree of attenuation.
+Early reflections typically arrive with a time difference between 1 and 80 ms after the direct path \cite{begault20003}.
+The term late reverberation encompasses all sound paths arriving to the listener after many reflections.
+Since the temporal density of such reflections increases with time, late reverberation is often modelled in statistical terms.
+An schematic representation of a room impulse response (RIR) is shown in Figure~\ref{fig:rir}.\\
 
-% TODO
-reverberation time, sabine, schroeder
-%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\textwidth]{Figures/Introduction/Acoustic_room_impulse_response.jpg}
+	\caption{Room impulse response model, from \cite{murphy2017acoustic}.}
+	\label{fig:rir}
+	\end{center}
+\end{figure}
 
-\subsection{SOFA}
- maybe as subsection of room acoustics
+By following this model, a RIR $h(t)$ can be described as a sequential combination of responses:
+\begin{equation}
+	h(t) = h_D(t) + h_R(t),
+\label{eq:directreverberant}
+\end{equation}
+where $h_D(t)$ and $h_R(t)$ represent the \textit{direct} (direct path plus early reflections) and \textit{reverberant} (late reverberation) components of the RIR. respectively.\\
 
-%
-%
-%
-%sound field capturing
-%
-%spherical harmonics
-%
-%decoding/beamforming
-%
-%\subsection{normalization, channel ordering}
-%
-%\section{Parametric spatial analysis}
-%
-%sirr, dirac
-%
-%sound field pressure and velocity approximation
-%
-%intensity vector
-%direction of arrival
-%energy density
-%diffuseness (limits)
+The room impulse response is a function of both the source and the receiver locations.
+Different levels, delays and directions of direct path and early reflections can obtained from measurements in the same room.
+However, it is generally assumed that the late reverberation is fixed for a given room, regardless of source/receiver positions.
+\\
 
-%alternative formulations
+Room reverberation plays an important role in psychoacoustics.
+While early reflections are usually perceived together with the direct path as a single auditory event, due to the \textit{precedence effect} \cite{haas1972influence}, late reverberation has often an influence on the received signal.
+In the specific case of speech, late reverberation is associated with a loss of intelligibility \cite{braun2018speech}.
+In the context of spatial perception, it has been shown that early reflections help the localization and externalization of sources [\todo{Improving Externalization in Ambisonic Binaural Decoding}], while the late reverberation is associated with a spaciousness perception of the room \cite{begault20003}.
+\\
+
+There are a number of measurable parameters which help to characterise room acoustics. 
+Perhaps one of the most widespread is the \textit{reverberation time} $T_{60}$ \cite{kuttruff2016room}. It represents the time required for the reverberant sound field power to decay by 60 dB.
+Reverberation time can be accurately computed from the room geometry \cite{sabine1927collected} or from the IR \cite{schroeder1965new}.
+
+In the latter case, the $T_{60}$ value is usually estimated from the \textit{Energy Decay Curve} (EDC), which is defined as:
+
+\begin{equation}
+	\text{EDC}(t) = 10 \log_{10} \sum_{t'=t}^{\infty} h^2(t),
+\end{equation}
+
+where $h(t)$ represents the room impulse response. The values are normalized such that the maximum peak of the curve corresponds to 0 dB.
+
+The EDC is usually modelled as a straight line in logarithmic scale.
+Therefore, the $T_{60}$ estimation is performed by estimating the slope of a straight line between two reference levels on the EDC time series.
+Some of the most used reference levels receive specific names: \textit{Early Decay Time} (EDT), $T_{60}$, and reverberation times $T_{10}$, $T_{20}$ and $T_{30}$.
+Table\ref{tab:reverberationtimes} shows their correspondent reference levels, where the maximum energy peak is normalized to 0 dB.
+An schematic representation of the reference levels is depicted in Figure~\ref{fig:reverberationtimes}.
+
+An alternative parameter is the \textit{decay rate} $\alpha_{60}$, which is related to reverberation time $T_{60}$ as:
+\begin{equation}
+	\alpha_{60} = \frac{3 \ln{(10)}} {T_{60}} (\text{dB/s}).
+\end{equation}
+
+The decay rate is thus the slope of the EDC curve, in logarithmic scale, expressed in dB per second.
+
+To conclude, it is important to notice that reverberation time is frequency-dependent.
+Accordingly, it is usual to report it for octave or third-octave bands, or alternatively to provide its value at a specific frequency.
+\\
+
+\begin{table}[t]
+\caption{Reverberation time computation: usual reference levels}
+\begin{center}
+\begin{tabular}{ccccc}
+\toprule
+   & $\text{EDT}$ & $T_{10}$ & $T_{20}$ & $T_{30}$ \\
+\midrule
+$L_{max} (dB)$ & 0 & -5 & -5 & -5  \\
+$L_{min} (dB)$ & -10 & -15 & -25 & -35 \\
+\bottomrule
+\end{tabular}
+\label{tab:reverberationtimes}
+\end{center}
+\end{table}
+
+\begin{figure}[htbp]
+	\begin{center}
+	\includegraphics[width=\textwidth]{Figures/Introduction/edt.png}
+	\caption{Room impulse response model, adapted from \url{http://www.bnoack.com/}.}
+	\label{fig:reverberationtimes}
+	\end{center}
+\end{figure}
+
+The \textit{Direct to Reverberant Ratio} (DRR) is another relevant acoustic parameter.
+DDR represents the ratio between direct and reverberant parts of the RIR, as defined in Eq.~\ref{eq:directreverberant}:
+\begin{equation}
+	DRR = 10 \log_{10} \frac{ \sum_{t=1}^{L_D} h^2_D(t) }{ \sum_{t=1}^{L_R} h^2_R(t)},
+\end{equation}
+with $L_D$ and $L_R$ as the length of the direct $h_D(t)$ and reverberant $h_R(t)$ filters, respectively.  
+At a psychoacoustic level, the direct to reverberant ratio is one of the main cues for distance perception \cite{begault20003}.
+
+Since the direct path and early reflections (but not the late reverberation) depend on the relative position between source and receiver , the filter $h_D(t)$ and therefore the DRR are as well location-dependent.
+For a given room, the source-receiver distance that produces a DRR of 0 dB is known as the \textit{critical distance}.
+
+\subsection{SOFA Conventions}
+
+The availability of recorded room impulse responses is of great importance to many acoustic signal processing problems. 
+As we have seen, many different RIRs can be obtained from the same room, by just varying the position of the source and the receiver; when the number of source and receiver positions increases, the total amount of measurements increases geometrically.
+Besides that, the actual format and organisation of the produced data (not only the RIR themselves, but also the source/position annotations) can be arbitrarily different when produced by different groups of people.
+
+In order to overcome potential interoperatibility and reusability issues, the \textit{Spatially Oriented Format for Acoustics} (SOFA) convention \cite{majdak2013spatially}, also known as the \textit{AES-69} standard \cite{majdak2015aes69}, proposes a unified file format for the storage of IR-related data.
+Despite that SOFA was initially created with a focus on \textit{Head-Related Impulse Response} (HRIR) data, its structure is very convenient to any kind of multi-location impulse responses, including ambisonics.
 
@@
 
-Let us consider a sound source represented by the signal $s(t)$, located in a given acoustic enclosure.
-Assuming that the system is Linear and Time-Invariant (LTI), it can be represented by its room impulse response $h(t)$.
+Let us consider a sound source represented by the signal $s(t)$, located in a given acoustic enclosure characterised by its room impulse response $h(t)$.
 The resulting reverberant signal $x(t)$ can be therefore described as the \textit{convolutive mixture} of the source and the RIR:

### 2020-06-03 manuscript.tex
wip

@@
-\documentclass[12pt, b5paper,twoside]{tesi_upf}
+\documentclass[12pt, b5paper,twoside]{tesi_upf_new}
+%\documentclass[12pt, b5paper,twoside]{book}
 
@@
 \usepackage{booktabs}
-%\usepackage{hyperref}
-%\hypersetup{pdftex,colorlinks=true,allcolors=blue}
-%\usepackage{hypcap}
+
+% COMMENT THESE FOR TESI_UPF
+\usepackage{hyperref}
+\hypersetup{pdftex,colorlinks=true,allcolors=blue}
+\usepackage{hypcap}
 
@@
 %ADD YOUR DATA
-\title{Spherical harmonic domain methods for 3D audio production enhancement \todo{(check capitalization)}}
-\subtitle{\todo{subtitle Required?}}
+\title{Ambisonic domain methods for immersive audio production enhancement}
+%\subtitle{\todo{subtitle Required?}}
 \author{Author: Andrés Pérez López}
+
+% TODO UNCOMMENT FOLOWING
 \thyear{2020}
@@
 
-\noindent Write here your dedication \todo{dedication}
+\noindent To my cat
 
@@
 
-%ABSTRACT IN TWO LEGUAGES.
+%ABSTRACT IN TWO LEGUAGES. 150 words max!!!
 \selectlanguage{english}
 \section*{\Large \sffamily Abstract}
-This is the abstract of the thesis in English.  Please, use less
-than 150 words. \todo{abstract in english}
+
+Recent years have witnessed a rise in immersive multimedia content, as a result of the developments in Virtual and Augmented Reality, where spatial audio plays a crucial role for the creation of a realistic experience.
+Ambisonic microphones allow the recording of sound scenes while preserving their spatial characteristics.
+For that reason, ambisonic has turned into the \textit{de facto} standard for immersive audio transmission.
+Yet, few production tools exploit all the information available in such media.
+In the present thesis we raise the research question of obtaining relevant information from ambisonic recordings, in a way that is meaningful for producers.
+In order to answer the question, we have developed several methods and tools applied to different levels of the immersive sound production workflow; with each of those elements representing a novel contribution to the state-of-the-art in their respective fields.
 
@@
 \section*{\Large \sffamily  Resum}
-Vet aquí el resum de la tesi en català.  Si us plau, utilitzeu
-menys de 150 paraules. \todo{abstract in catalan}
+
+En els últims anys hem assistit a l'auge del contingut multimèdia immersiu, afavorit pels avanços en Realitat Virtual i Augmentada.
+En aquest mitjà, el so espacial juga un paper molt important per a la creació d'una sensació de realisme.
+Els micròfons ambisonics permeten gravar escenes sonores preservant les seues característiques espacials.
+A causa d'això, ambisonics s'ha convertit en l'estàndard de facto per a la transmissió d'àudio immersiu.
+No obstant això, encara son poques les eines de producció que permeten explotar tota la informació disponible en aquest mitjà.
+En la present tesi plantegem la pregunta de com obtindre informació rellevant de gravacions ambisonics de manera automàtica, d'una forma que siga significativa per als productors.
+Per a això, hem desenvolupat una sèrie de mètodes i eines que s'apliquen a diferents nivells de la cadena de producció de so immersiu; cadascun d'aquests elements representa una contribució nova a l'estat de l'art en aquest camp.
+
+%En los últimos años hemos asistido al auge del contenido multimedia immersivo, auspiciado por los avances en Realidad Virtual y Aumentada.
+En este medio, el sonido espacial juega un papel muy importante para la creación de una sensación de realismo.
+%
+%Los micrófonos ambisonics permiten grabar escenas sonoras preservando sus características espaciales.
+%Debido a esto, ambisonics se ha convertido en el estándar de facto para la transmisión de audio immersivo.
+Sin embargo, todavía hay pocas herramientas de producción que permitan explotar toda la información disponible en este medio.
+%
+%En esta tesis planteamos la pregunta de cómo obtener información relevante de grabaciones ambisonics de manera automática, de una forma significativa para los productores.
+%
+%Para ello, hemos desarrollado una serie de métodos y herramientas que se aplican a diferentes niveles de la cadena de producción de sonido immersivo; cada uno de estos elementos representa una contribución novedosa al estado del arte en ese campo.
+
 \vspace*{\fill}
@@
 %ADD BIBLIOGRAPHY
-\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib}
+\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib, Bibliography/Introduction.bib}
 

### 2020-06-05 Chapters/CoherenceEstimation.tex
update 20200602
wip

@@
 \chapter{Coherence Estimation}
+\label{chap:coherence}
 
@@
 A number of practical applications benefit of the knowledge about the diffuseness of a sound field, including speech enhancement and dereverberation \cite{p_habets_dual-microphone_2006}, noise suppression \cite{ito_designing_2010}, source separation \cite{duong_under-determined_2009} or background estimation \cite{stefanakis_foreground_2015}.
-In the field of spatial audio, diffuseness estimation is often used for parametrization \cite{pulkki_directional_2006, politis_compass_2018}, Direction-of-Arrival estimation \cite{thiergart_localization_2009} or source separation \cite{motlicek_real-time_2013}.
+In the field of spatial audio, diffuseness estimation is often used for parametrization \cite{pulkki2006directional, politis_compass_2018}, Direction-of-Arrival estimation \cite{thiergart_localization_2009} or source separation \cite{motlicek_real-time_2013}.
 
-\todo{In this paper}, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
+\todo{In this chapter}, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
 The motivation for this work is, first, that tetrahedral arrays are a well known type of microphone arrays, which have today become popular for applications related to Virtual and Augmented Reality. 
@@
 Another interesting observation comes from the frequency response of the curves.
-For all values of $r$, the coherence of the compensated \textit{B-Format} signal (with $\Gamma_m(kR)$) is roughly flat up to around 7 kHz, which approximately corresponds to the operational spatial frequency range of the microphone \cite{gerzon_design_1975}.
+For all values of $r$, the coherence of the compensated \textit{B-Format} signal (with $\Gamma_m(kR)$) is roughly flat up to around 7 kHz, which approximately corresponds to the operational spatial frequency range of the microphone \cite{gerzon1975design}.
 Above this value, the coherence response looses the flatness due to spatial aliasing (Eq.~\ref{eq:falias}).

### 2020-06-05 Chapters/Introduction.tex
update
update 20200602
wip

@@
 
-\section{Motivation}
+\section{Problem description and motivation}
 
@@
 
-Research question: 	How can we exploit the characteristics of ambisonic recordings in order to manipulate them more adequately?
+Research question: 	How can we exploit the characteristics of ambisonic recordings in order to extract signal-dependent, meaningful information from them?
 
 \section{Context}
+\label{sec:context}
 
-Different levels of application/contribution:
+Different levels of applications:
 
@@
 \begin{figure}[hbt]
-  \includegraphics[width=\textwidth]{Figures/context_scheme.png}
+  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1.png}
   \caption{\todo{todo caption}}
+  \label{fig:scheme1}
 \end{figure}
 
-\todo{introduce rest of chapters?}
+\section{Outline and main contributions}
+
+\subsection{Thesis structure}
+
+The present Thesis is organised as follows.
+
+\textbf{Chapter~\ref{chap:scientific}} introduces the basic concepts that will be developed throughout the Thesis, including spherical harmonics and ambisonics, coherence estimation, parametric analysis or room acoustics.
+The Chapter also defines the signal models and the mathematical terminology.
+
+Chapters~\ref{chap:rt60}, \ref{chap:coherence} and \ref{chap:seld2019} develop the most significant academic contributions of the Thesis. 
+\textbf{Chapter~\ref{chap:rt60}} presents a novel method for blind reverberation time in ambisonic recordings.
+To the best of our knowledge, this is the first method proposal specifically focusing on that problem.
+The method is based on a Multichannel Auto-Regressive model of the late reverberation, which allows for an effective dereverberation of the ambisonic sound scene, and enables computation of the reverberation time from an estimation of the room impulse response.
+The evaluation metrics show a method performance similar to other state-of-the-art methods.
+\textbf{Chapter~\ref{chap:coherence}} analyses the response of tetrahedral microphone arrays, which are the simplest and most common form of ambisonic microphones, under spherically isotropic sound field.
+The analysis is performed using both simulated and recorded diffuse field, and the results quantify the differences between ideal and real values under a variety of conditions and estimators.
+In \textbf{Chapter~\ref{chap:seld2019}}, a complete system for Sound Event Localization and Detection of ambisonic sound scenes is described.
+The algorithm comprises two different parts.
+First, a parametric analysis is performed on the ambisonic signal.
+The analysis yields spatial localization and temporal activities of the sound events present in the scene.
+Then, each of those events is assigned to a class label by means of a deep-learning classifier.
+The method is able to perform in a similar way to the baseline system, while greatly improving its localization capabilities.
+
+Finally, \textbf{Chapter~\ref{chap:data}} presents some libraries and software utilities developed throughout the Thesis.
+All the code has been publicly released under open source licenses.
+The libraries include utilities for the creation of datasets, the storage and exchange of impulse response files in a standard way, and the implementation of convenience tools for acoustic and microphone array signal processing analysis.
+Although the libraries do not directly involve any scientific contribution, they can be a great help for scientific and innovative purposes; given the industial nature of the candidate's doctoral program, we have considered relevant to include them in the present Thesis.
+\todo{check}
+
+In order to place the different chapters within the problem context described in Section~\ref{sec:context}, we add to Figure~\ref{fig:scheme1} the Chapter numbers with our contributions, obtaining Figure~\ref{fig:scheme1_numbers}.
+
+\begin{figure}[hbt]
+  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
+  \caption{\todo{todo caption}}
+  \label{fig:scheme1_numbers}
+\end{figure}
+
+%\newpage
+\subsection{List of Publications}
+
+In the following list we show the main scientific contributions of the Thesis, organised by Chapters:
+
+\begin{itemize}
+
+	\item Chapter 3:\\
+	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
+	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
+	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}
+
+	\item Chapter 4:\\
+	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
+	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
+	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
+
+	\item Chapter 5:\\
+	\textbf{A hybrid parametric-deep learning approach for sound event localization and detection.}
+	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
+	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+
+	\item Chapter 6:\\
+	\textbf{Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes}.\\
+	\underline{A. P\'erez-L\'opez}.
+	In \textit{16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
+
+	\textbf{Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics.}
+	\underline{A. P\'erez-L\'opez} and J. De Muynke.
+	In \textit{Audio Engineering Society Convention 144. Audio Engineering Society, 2018.}
+
+	\textbf{pysofaconventions, a Python API for SOFA.}
+	\underline{A. P\'erez-L\'opez}.
+	In \textit{Audio Engineering Society Convention 148. Audio Engineering Society, 2020.}
+
+	\textbf{A Python library for Multichannel Acoustic Signal Processing.}
+	\underline{A. P\'erez-L\'opez} and A. Politis.
+	In \textit{Audio Engineering Society Convention 148. Audio Engineering Society, 2020.}
+
+\end{itemize}
+
+Moreover, as a result of the development of the Thesis, the following open-source libraries have been implemented and released.
+All of them are available through the author's GitHub page, \url{https://github.com/andresperezlopez}: \todo{describir cada uno?
+con respecto a paper o chapter?}
+
+\begin{itemize}
+
+	\item rt60 estimation: \todo{todo}
+
+	\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3}
+
+	\item \href{https://github.com/andresperezlopez/ambiscaper}{ambiscaper}
+
+	\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions}
+
+	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}
+
+\end{itemize}

### 2020-06-05 Chapters/ReververationTimeEstimation.tex
wip

@@
 \chapter{Blind reverberation time estimation}
+\label{chap:rt60}
 
@@
 
-The IRs are FOA room impulse responses simulated by the image method with the \textit{Multichannel Acoustic Signal Processing} library \cite{masp}.
+The IRs are FOA room impulse responses simulated by the image method with the \textit{Multichannel Acoustic Signal Processing} library \todo{[ref to chapter?]}.
 There are 9 different IRs of 1 s, with random $T_{60}$ values in the range between 0.4 s and 1.1 s approximately, estimated by the $T_{10}$ method at the 1 kHz band.

### 2020-06-05 Chapters/ScientificBackground.tex
wip

@@
-\chapter{Background}
+\chapter{Scientific Background}
+\label{chap:scientific}
 
@@
 \subsection{Nomenclature}
-
-Through the document, we refer to time-domain signals with lowercase, e.g. $x(t)$, with $t$ as the time index.
+\todo{is this name correct? }
+Throughout the Thesis, we refer to time-domain signals with lowercase, e.g. $x(t)$, with $t$ as the time index.
 
@@
 
-\todo{revise all vector notations}
-
 \section{Spherical Harmonics}
@@
 Spherical harmonics are continuous functions defined on the sphere surface.
-Due to their mathematical properties, any spherical function can be decomposed as a combination of spherical harmonics, in what is known as the \textit{Spherical Harmonics Expansion} [Jarrett book, page 17].
+Due to their mathematical properties, any spherical function can be decomposed as a combination of spherical harmonics, in what is known as the \textit{Spherical Harmonics Expansion} \cite{jarrett2017theory}.
 
@@
 Many different spherical harmonic definitions exist in the literature, with minor variations among them.
-In the following, we will use the real-valued, fully normalized spherical harmonics as defined by [Zotter]:
+In the following, we will use the real-valued, fully normalized spherical harmonics as defined by \cite{zotter2019ambisonics}:
 
@@
 Let us consider a sound field captured with a spherical microphone array, which contains $Q$ capsules distributed around a spherical surface of radius $R$ at the positions $\Omega_q, 1 \leq q \leq Q$.
-The captured frequency-domain signals $X_q(k)$ can be represented as the spherical harmonic domain signals $X_n^m(k)$ through the spherical harmonic transform of order $n$ and degree $m$ (Moreau et al., 2006):
+The captured frequency-domain signals $X_q(k)$ can be represented as the spherical harmonic domain signals $X_n^m(k)$ through the spherical harmonic transform of order $n$ and degree $m$ \cite{moreau20063d}:
 
@@
 
-where the term $\Gamma_n(kR)$ models the radial transfer function, and depends on the microphone geometry[SOME REFS: moreau, rafaely].
-There are several possible sampling schemes of capsules along the sphere, each one having different properties; the reader is redirected to [rafaely] for a deeper insight.
+where the term $\Gamma_n(kR)$ models the radial transfer function, and depends on the microphone geometry.
+There are several possible sampling schemes of capsules along the sphere, each one having different properties; the reader is referred to  \cite{rafaely2004analysis} for a deeper insight.
 \\
 
-By using this model, the maximum spherical harmonic order $N$ that can be retrieved with negligible spatial aliasing depends on the number of microphone capsules [moreau]:
+By using this model, the maximum spherical harmonic order $N$ that can be retrieved with negligible spatial aliasing depends on the number of microphone capsules \cite{moreau20063d}:
 \begin{equation}
@@
 Furthermore, the sphere radius $R$ has also an effect on the operational bandwidth of the microphone.
-More precisely, for a given spherical harmonic order $n$, the maximum aliasing-free operational frequency is given by [moreau, rafaely]:
+More precisely, for a given spherical harmonic order $n$, the maximum aliasing-free operational frequency is given by:
 \begin{equation}
@@
 
-Ambisonics is a spatial sound recording and playback technology initially developed during the 1970's, and further expanded into its modern formulation around the 2000s [ZOTTER, page 53].  
+Ambisonics is a spatial sound recording and playback technology initially developed during the 1970s \cite{gerzon1973periphony}, and further expanded into its modern formulation around the 2000s \cite{daniel2000representation}.
 Ambisonics is based on the idea of decomposing a sound field into its spherical harmonic representation.
 
-Originally, the decomposition was limited to first-order spherical harmonics, mainly due to practical limitations [CITE GERZON], as the so-called \textit{First Order Ambisonics} (FOA).
-The technique was later formalized for arbitrary spherical harmonic orders, known as \textit{Higher Order Ambisonics} or HOA [CITE DANIEL].
-In general, with the term Ambisonics we will be referring to the latter definition.\\
+Originally, the decomposition was limited to first-order spherical harmonics, as the so-called \textit{First Order Ambisonics} (FOA); mainly because of practical limitations.
+The technique was later formalized for arbitrary spherical harmonic orders, known as \textit{Higher Order Ambisonics} (HOA).
+In general, with the term \textit{ambisonics} we will be referring to the latter definition.\\
 
@@
 
-Historically, the name \textit{B-Format} was used as an opposite of \textit{A-Format}, which describes the signals recorded by a tetrahedral microphone array [cite gerzon?].
-The tetrahedron is the simplest and most common form of spherical microphone arrays (\textit{ambisonic microphones}) with uniform capsule distribution.
+Historically, the name \textit{B-Format} was used as an opposite of \textit{A-Format}, which describes the signals recorded by a tetrahedral microphone array \cite{gerzon1975design}.
+The tetrahedron is the simplest and most common form of spherical microphone arrays (indistinctly referred to as ambisonic microphones) with uniform capsule distribution.
 Again, the term \textit{A-Format} is also currently employed for referring to the signals recorded by any spherical microphone array, regardless of the number or arrangement of capsules.
@@
 Likewise, the process of signal conversion from the spatial domain (microphone capsules) to the spherical harmonic domain (ambisonic signals), as in Eq.~\ref{eq:a2b}, is known as \textit{A-B conversion}.
-A number of different approaches have been developed for this process, and the interested reader is referred to [\todo{find ref}] for more information.
+A number of different approaches have been developed for this process, and the interested reader is referred to \cite{moreau20063d} for more information.
 
@@
 Conversely, the sound field reconstruction is performed by the \textit{ambisonic decoding} operation.
-This process is equivalent to weight-and-sum beamforming in the spherical harmonic domain, and it is sometimes also referred to as the \textit{virtual microphone} technique [ambisonic book].
+This process is equivalent to weight-and-sum beamforming in the spherical harmonic domain, and it is sometimes also referred to as the \textit{virtual microphone} technique \cite{zotter2019ambisonics}.
 
@@
 \begin{table}[t]
-\caption{Ambisonic decoding: standard values of $alpha_n$ weightings. Extracted from \cite{daniel2000representation}.}
+\caption{Ambisonic decoding: standard values of $alpha_n$ weightings. Adapted from \cite{daniel2000representation}.}
 \begin{center}
@@
 For instance, \textit{N3D} is the most mathematically straightforward, and spherical harmonics defined in that way can be directly used for both encoding and decoding (as in Eqs~\ref{eq:encoding} and Eq.~\ref{eq:decoding}) -- however, from a sound engineer point of view, other normalization schemes with maximum values below the unity might be preferred, such as \textit{SN3D}.
-Besides this, \textit{FuMa} has been historically the default normalization \todo{[gerzon?]}, while the more modern \textit{N3D} and \textit{SN3D} were popularized after Daniel \todo{[daniel]}.
+Besides this, \textit{FuMa} has been historically the default normalization \cite{gerzon1985ambisonics}, while the more modern \textit{N3D} and \textit{SN3D} were popularized after \cite{daniel2000representation}.
 \\
@@
   \includegraphics[width=\textwidth]{Figures/ScientificBackground/normalization.png}
-  \caption{Maximum value of each ambisonic channel up to order 5, for all different normalization schemes. Image from T. Carpentier [\todo{cite}].}
+  \caption{Maximum value of each ambisonic channel up to order 5, for all different normalization schemes. Image from \cite{carpentier2017ambisonic}.}
   \label{fig:normalization}
@@
 
-As a summary, Figure~\ref{fig:normalization} displaysthe different normalization schemes. The reader is referred to \todo{[thibaut]} for an extensive review on the topic.
+As a summary, Figure~\ref{fig:normalization} displays the different normalization schemes. The reader is referred to \cite{carpentier2017ambisonic} for an extensive review on the topic.
 
@@
 
-Historically, first-order ambisonic audio has followed what it might be called \textit{traditional B-Format} channel ordering [ambisonics in multichannel broadcasting and video]. 
-By following this scheme, the four channels of a FOA signal $S_n^m$ are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
+Historically, first-order ambisonic audio has followed what it might be called \textit{traditional B-Format} channel ordering \cite{gerzon1985ambisonics}. 
+By this scheme, the four channels of a FOA signal $S_n^m$ are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
 \begin{equation}
@@
 \end{equation}
-with:
+where:
 \begin{equation}
@@
 
-This nomenclature was extended to second and third order by Malham, and is currently known as the \textit{Furse-Malham} or \textit{FuMa} channel ordering.
-The channel names use all english alphabet letters from K to Z in third order and, although there would be enough letters to go up to fourth order, the unconvenience of the system was evident [Higher order Ambisonic systems].
+This nomenclature was extended to second and third order, and is currently known as the \textit{Furse-Malham} or \textit{FuMa} channel ordering.
+The channel names use all english alphabet letters from K to Z in third order and, although there would be enough letters to go up to fourth order, the unconvenience of the system was clear \cite{malham2003higher}.
 Figure~\ref{fig:normalization} shows the equivalence between \textit{FuMa} (\textit{"letter code"}) and \textit{ACN} channel names.\\
@@
 Trough parametric analysis, sound fields may be described in terms of a small amount of sound sources and associate parameters.
-Such representation might reduce to a great extent the complexity of processing methods [book jarrett].
+Such representation might reduce to a great extent the complexity of processing methods \cite{jarrett2017theory}.
 
-One of the most successful sound field parametric models is DirAC[pulkki2007 dirac].
-Originally conceived as a method for impulse response processing and spatial sound reproduction [merima2005], it has been widely used in many different audio-related problems [\todo{find some references!}].
+One of the most successful sound field parametric models is DirAC \cite{Pulkki07}, which was originally conceived as a method for impulse response processing and spatial sound reproduction \cite{merimaa2005spatial}.
 
-DirAC (acronym for \textit{Directional Audio Coding}) is a perceptually motivated time-frequency (TF) domain method, based on the assumption that any sound field may be reproduced with high perceptual quality by considering two parameters: the sound field diffuseness and the most prominent sound \textit{Direction-of-arrival} (DOA) [cite parametric book].
+DirAC (acronym for \textit{Directional Audio Coding}) is a perceptually motivated time-frequency (TF) domain method, based on the assumption that any sound field may be reproduced with high perceptual quality by considering two parameters: the sound field diffuseness and the most prominent sound \textit{Direction-of-Arrival} (DOA) \cite{pulkki2018parametric}.
 \\
 
-Let us consider a \textit{N3D}-normalized first-order ambisonic signal in time-frequency domain, $S_n^m(k, n)$. 
+Let us consider a \textit{SN3D}-normalized first-order ambisonic signal in time-frequency domain, $S_n^m(k, n)$. 
 For the sake of clarity, we will use in this section \textit{FuMa} channel notation and ordering (Eq.~\ref{eq:fumaordering}):
@@
 
-The \textit{active intensity} $\pmb{I}(k,n)$, defined as the amount of transmitted acoustic energy, can be expressed in terms of sound pressure and velocity [fahy2002]:
+The \textit{active intensity} $\pmb{I}(k,n)$, defined as the amount of transmitted acoustic energy, can be expressed in terms of sound pressure and velocity \cite{fahy1990sound}:
 \begin{equation}
@@
 
-Another useful parameter is the \textit{energy density} $E(k,n)$ [stazial 1996]:
+Another useful parameter is the \textit{energy density} $E(k,n)$ \cite{stanzial1996reactive}:
 \begin{equation}
 	\begin{aligned}
-		E(k,n) &= \frac{|P(k,n)|^2 + \|\pmb{U}(k,n)\|^2}{2\rho_0 c^2} \\ 
-		&=  \frac{|W(k,n)|^2 + \| [X(k,n), Y(k,n), Z(k,n)] \|^2}{2\rho_0 c^2}.
+		E(k,n) &= \frac{1}{2\rho_0 c^2} |P(k,n)|^2 + \frac{1}{2} {\|\pmb{U}(k,n)\|^2} \\
+		&=  \frac{1}{2\rho_0 c^2} \Big(|W(k,n)|^2 + \| [X(k,n), Y(k,n), Z(k,n)] \|^2\Big).
 	\end{aligned}
@@
 
-Finally, the \textit{diffuseness} $\Psi(k,n)$ can be computed from the sound intensity and  energy density [merimaa2005]:
+Finally, the \textit{diffuseness} $\Psi(k,n)$ can be computed from the sound intensity and  energy density \cite{merimaa2005spatial}:
 \begin{equation}
@@
 
- where the symbols $\langle \cdot \rangle$ represent the expectation operator, which is usually implemented as time-domain averaging. \todo{check equation}
+ where the symbols $\langle \cdot \rangle$ represent the expectation operator, which is usually implemented as time-domain averaging.
 
 Even though Eq.~\ref{eq:psidefinition} (known as \textit{DirAC's diffuseness}) is one of the most common ambisonic diffuseness estimators, several alternative formulations exist. 
-Other diffuseness estimation procedures include the \textit{coefficient of variation method} [ahonen, pulkki, del galdo] and the more recent \textit{COMEDIE} estimator [epain, craig].
-
+Other diffuseness estimation procedures include the \textit{coefficient of variation method} \cite{ahonen2009diffuseness} and the more recent \textit{COMEDIE} estimator \cite{epain2016spherical}. 
 In any case, in what follows, the term \textit{diffuseness} and the symbol $\Psi$ will refer by default to Eq.~\ref{eq:psidefinition}.
@@
 In the specific case of speech, late reverberation is associated with a loss of intelligibility \cite{braun2018speech}.
-In the context of spatial perception, it has been shown that early reflections help the localization and externalization of sources [\todo{Improving Externalization in Ambisonic Binaural Decoding}], while the late reverberation is associated with a spaciousness perception of the room \cite{begault20003}.
+In the context of spatial perception, it has been shown that early reflections help the localization and externalization of sources \cite{rudrich2019improving}, while the late reverberation is associated with a spaciousness perception of the room \cite{begault20003}.
 \\
@@
 
-\section{Signal Model}
+\section{Signal Models}
 
@@
 
-\section{Practical Considerations}
+%\section{Practical Considerations}
 % TODO I don't remember what is this

### 2020-06-05 manuscript.tex
wip
modified title and abstract

@@
 
-%\usepackage{hyperref}
-
 %WITHOUT HEADINGS: NO MODIFICATION
@@
 %ADD YOUR DATA
-\title{Ambisonic domain methods for immersive audio production enhancement}
-%\subtitle{\todo{subtitle Required?}}
+%\title{Ambisonic domain methods for immersive audio production enhancement}
+\title{Parametric analysis of ambisonic audio}
+\subtitle{Contributions to methods, applications and data generation}
 \author{Author: Andrés Pérez López}
@@
 Yet, few production tools exploit all the information available in such media.
-In the present thesis we raise the research question of obtaining relevant information from ambisonic recordings, in a way that is meaningful for producers.
-In order to answer the question, we have developed several methods and tools applied to different levels of the immersive sound production workflow; with each of those elements representing a novel contribution to the state-of-the-art in their respective fields.
+%In the present thesis we raise the research question of obtaining relevant information from ambisonic recordings, in a way that is meaningful for producers.
+In the present thesis we raise the research question of obtaining meaningful information from ambisonic recordings, based on the parametric analysis of the sound field.
+In order to answer the question we have developed different state-of-the-art methods and tools, which cover a broad range of applications within the parametric audio analysis scope.
+
+% Each of these elements represent a novel contribution to the state-of-the-art in their respective fields.
+%In order to answer the question, we have developed several methods and tools applied to different levels of the immersive sound production workflow; with each of those elements representing a novel contribution to the state-of-the-art in their respective fields.
 
@@
 No obstant això, encara son poques les eines de producció que permeten explotar tota la informació disponible en aquest mitjà.
-En la present tesi plantegem la pregunta de com obtindre informació rellevant de gravacions ambisonics de manera automàtica, d'una forma que siga significativa per als productors.
-Per a això, hem desenvolupat una sèrie de mètodes i eines que s'apliquen a diferents nivells de la cadena de producció de so immersiu; cadascun d'aquests elements representa una contribució nova a l'estat de l'art en aquest camp.
+En la present tesi plantegem la pregunta de com obtenir informació rellevant de gravacions ambisonics de manera automàtica, basant-nos en l'an\`alisi param\`etric del camp ac\'ustic. 
+%d'una forma que siga significativa per als productors.
+Per a això, hem desenvolupat una sèrie de mètodes i eines a nivell de l'estat de l'art, 
+que cobreixen un ampli espectre d'aplicaciones a l'\`ambit de l'an\`alisi param\`etric del so. 
+%que s'apliquen a diferents nivells de la cadena de producció de so immersiu; cadascun d'aquests elements representa una contribució nova a l'estat de l'art en aquest camp.
 
@@
 \input{Chapters/ScientificBackground.tex}
+\input{Chapters/ReververationTimeEstimation}
 \input{Chapters/CoherenceEstimation}
-\input{Chapters/ReververationTimeEstimation}
 \input{Chapters/SELD}

### 2020-06-11 Chapters/Introduction.tex
restructured Introduction, written Motivation

@@
 
-\section{Problem description and motivation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Motivation}
 
-\subsection{3D arrays}
+Ambisonics is a spatial audio theory based on the directional decomposition of the sound field.
+Conceived in its primal form during the 70s \cite{gerzon1973periphony}, it was not until the 21st century, with a modern mathematical formulation \cite{daniel2000representation} and much more computational power available, that it definitely drew the attention of the research community.
 
-Sound propagates in 3D: need for 3D mic arrays to capture spatial properties
+%\begin{figure}[t]
+%  \includegraphics[width=\textwidth]{Figures/Introduction/hardware.jpg}
+%    \caption{An ambisonic hardware processor, manufactured by Audio \& Design Recording Ltd. Source: \href{https://commons.wikimedia.org/w/index.php?curid=11813129}{Relen, Wikimedia Commons}}
+%  \label{fig:hardware}
+%\end{figure}
 
-\subsection{spherical microphone arrays}
+Nevertheless, the greatest contributor to the current interest in ambisonics has been the rise of Virtual Reality (VR) in recent years.
+Although VR focuses primarily on visual cues, the immersive experience might be greatly enhanced by spatial audio \todo{CITE REQUIRED, begault?}.
+In this context, Ambisonics has been rapidly adopted as \textit{de facto} standard for spatial audio transmission, supposedly due to a variety of factors:
 
-\begin{itemize}
-  \item even distribution of capsules
-  \item mathematical convenience: spherical harmonics
-\end{itemize}
+\begin{description}
 
-\subsection{ambisonics}
+  \item [Layout independence] As opposed to other audio spatialization techniques that rely on specific playback layouts, ambisonics makes use of an intermediate sound field representation, known as \textit{B-Format} (or just ambisonic audio).
+This representation, often referred to as \textit{scene-based}, can be then further processed to match any playback configuration.
 
-advantages on the vr/ar context
-\begin{itemize}
-  \item device independent
-  \item intermediate storage format
-  \item signal-independent transformations are easy
-  \item de-facto standard for vr
-\end{itemize}
+  \item [Recording device independence] Regardless of the specific characteristics of an ambisonic microphone, the recorded signal is usually converted into B-Format, which effectively represents the standard exchange format.
 
-\subsection{Current limitations of vr/ar production}
+  \item [Ease of manipulations] Signal-independent transformations of the ambisonic stream, and specifically rotations, are computationally inexpensive.
 
-\section{Goals}
+  \item [Binaural transformation] Spatial audio in VR is mostly consumed as binaural; methods for ambisonic to binaural conversion have been known for a long time \todo{cite?}.
+Furthermore, VR headsets can easily provide head rotation information, which can be used in combination with scene rotations to provide head-locked audio, which greatly improves immersiveness \todo{cite required!}.
+This is a key feature of ambisonics when compared to static binaural recordings.
 
-Research question: 	How can we exploit the characteristics of ambisonic recordings in order to extract signal-dependent, meaningful information from them?
+\end{description}
 
-\section{Context}
-\label{sec:context}
+Coming back to the Virtual Reality popularity, the following three events might be representative of the growth that it experienced in the second half of the 2010s: 
+(i) the billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}; 
+(ii) the Time cover page devoted to VR: \textit{""The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
+and (ii) M.
+Zuckerberg's invited talk at the \textit{Samsung Unpacked} event within the World Mobile Congress 2016 in Barcelona: \textit{"VR is the next platform where anyone can experience anything they want"} \cite{bbcnews}.
 
-Different levels of applications:
+\begin{figure}[t!]
+  \includegraphics[width=\textwidth]{Figures/Introduction/num_mics_ticks.png}
+  \caption{Number of ambisonic microphones released in last years (from \cite{List_of_Ambisonic_hardware}).
+From left to right, the vertical lines correspond to (i) Oculus acquisition, (ii) Time cover page on VR, (iii) M.
+Zuckerberg's speech in MWC, and (iv) Jaunt announcement of shift towards AR.}
+  \label{fig:nummics}
+\end{figure}
+
+\begin{table}[t!]
+\centering
+\caption{List of ambisonic microphones released in recent years (from \cite{List_of_Ambisonic_hardware}).}
+\begin{tabular}{cccc}
+  \toprule
+Manufacturer & Model   & Year & Order \\
+\midrule
+MH Acoustics & EigenMike   & 2013                     & 4                         \\
+Brahma       & (Brahma)                                                   & 2014                     & 1                         \\
+Sennheiser   & Ambeo                                                      & 2016                     & 1                         \\
+Twirling     & 720 VR    & 2016                     & 1                         \\
+Zoom         & H2n                                                        & 2016                     & 1                         \\
+Zylia        & ZM-1                                                       & 2017                     & 3                         \\
+Twirling     & 720 Lite    & 2017                     & 1                         \\
+Ricoh        &  TA-1        & 2017                     & 1                         \\
+Nevaton      & Nevaton VR                                                 & 2017                     & 1                         \\
+Rode         &  Rode NT-SF1 & 2018                     & 1                         \\
+CoreSound    &  OctoMic     & 2018                     & 2                         \\
+Zoom         & H3-VR                                                      & 2018                     & 1                         \\
+Brahma       & Brahma 8                                                   & 2019                     & 2                         \\
+Voyage Audio & Spatial Mic  & 2019    & 2  \\
+\bottomrule
+\end{tabular}
+\label{tab:nummics}
+\end{table}
+
+Given this context, many microphone manufacturers and audio-related companies have followed the industry trend in the search for new markets.
+Only in the interval 2016-2019, 12 different ambisonic microphones have reached the market (Figure~\ref{fig:nummics}) --- a greater amount than all previous existing ambisonic microphones together.
+A comprehensive list of recent ambisonic microphone releases is shown in \ref{tab:nummics}.
+
+At the present time, however, the high expectations in VR have significantly lowered, as shown in Figure~\ref{fig:vrfunding}.
+This is due to a variety of reasons, including lack of interesting content and the high production cost of the headsets \cite{fortune}.
+The change of focus of Jaunt (formerly one of the biggest VR film production companies) towards Augmented Reality (AR), as of October 2018, might be a paradigmatic example of this tendency \cite{theverge}.
+
+\begin{figure}[t!]
+  \includegraphics[width=\textwidth]{Figures/Introduction/vr_funding.png}
+  \caption{Venture investments in VR in the period 2014-2018. Adapted from \cite{fortune}.}
+  \label{fig:vrfunding}
+\end{figure}
+
+In any case, the current high availability and affordability of ambisonic microphones brings new challenges from the signal processing perspective.
+More specifically, ambisonic microphones conform a subset of near-coincident spherical microphone arrays, a category which possesses some specific characteristics.
+
+Although the VR momentum has also reached spherical microphone array processing, many challenges remain still opened, and the number of research works specifically focusing on such geometry are still low.
+Besides that, the growing interest in AR posses new problems related to acoustical signal processing.
+But since ambisonics is still the standard choice for immersive audio, existing solutions might be successfully adapted.
+
+Lastly, the advance in signal processing methods for ambisonics can give rise to applications that enhance the work of immersive audio producers, providing meaningful information about the recorded scenes and automating some of the repetitive tasks, thus allowing a more flexible and creative workflow.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Problem Description}
+
+\begin{figure}[hbt]
+  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1.png}
+  \caption{\todo{todo caption}}
+  \label{fig:scheme1}
+\end{figure}
+
+EXPLAIN PARTS OF FIGURE:
+
+Different levels of applications in multichannel acoustic signal processing (inspired by \cite{jarrett2017theory})
 
@@
 
-\begin{figure}[hbt]
-  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1.png}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Scientific Objectives}
+
+preguntas cientificas concretas, coincidiendo con contribuciones
+
+(base para la conclusion)
+
+%\section{Goals}
+%
+%Research question: How can we exploit the characteristics of ambisonic recordings in order to extract signal-dependent, meaningful information from them?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Outline}
+
+\begin{figure}[t]
+  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
   \caption{\todo{todo caption}}
-  \label{fig:scheme1}
+  \label{fig:scheme1_numbers}
 \end{figure}
-
-\section{Outline and main contributions}
-
-\subsection{Thesis structure}
 
@@
 
-\begin{figure}[hbt]
-  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
-  \caption{\todo{todo caption}}
-  \label{fig:scheme1_numbers}
-\end{figure}
+\section{Context}
+\label{sec:context}
 
-%\newpage
-\subsection{List of Publications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Contributions}
 
@@
 	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+
+	\todo{DCASE2020}
 

### 2020-06-11 manuscript.tex
restructured Introduction, written Motivation

--- 2020-06-01 manuscript.tex
@@
 \usepackage{subcaption}
-
-%\usepackage{hyperref}
 
@@
 %ADD YOUR DATA
-\title{Spherical harmonic domain methods for 3D audio production enhancement \todo{(check capitalization)}}
-\subtitle{\todo{subtitle Required?}}
+%\title{Ambisonic domain methods for immersive audio production enhancement}
+\title{Parametric analysis of ambisonic audio}
+\subtitle{Contributions to methods, applications and data generation}
 \author{Author: Andrés Pérez López}
+
+% TODO UNCOMMENT FOLOWING
 \thyear{2020}
@@
 
-%ABSTRACT IN TWO LEGUAGES.
+%ABSTRACT IN TWO LEGUAGES. 150 words max!!!
 \selectlanguage{english}
 \section*{\Large \sffamily Abstract}
-This is the abstract of the thesis in English.  Please, use less
-than 150 words. \todo{abstract in english}
+
+%Recent years have witnessed a rise in immersive multimedia content, as a result of the developments in Virtual and Augmented Reality, where spatial audio plays a crucial role for the creation of a realistic experience.
+%Ambisonic microphones allow the recording of sound scenes while preserving their spatial characteristics.
+For that reason, ambisonic has turned into the \textit{de facto} standard for immersive audio transmission.
+Yet, few production tools exploit all the information available in such media.
+%In the present thesis we raise the research question of obtaining meaningful information from ambisonic recordings, based on the parametric analysis of the sound field.
+%In order to answer the question we have developed different state-of-the-art methods, which cover a broad range of applications within the parametric audio analysis scope. 
+%Specifically, we present novel contributions in the scope of blind reverberation time estimation, diffuseness estimation, and sound event localization and detection.
+Furthermore, several software tools developed for dataset generation and storage are also presented.
+
+Due to the recent advances in virtual and augmented reality, ambisonics has emerged as the \textit{de facto} standard for immersive audio.
+Ambisonic audio can be captured using spherical microphone arrays, which are becoming increasingly popular as well.
+Yet, most methods for acoustic and microphone array signal processing are not specifically tailored for spherical geometries.
+Therefore, there is still room for improvements in the field of automatic analysis and description of ambisonic recordings. 
+In the present thesis, we tackle this problem using methods based on the parametric analysis of the sound field. 
+Specifically, we present novel contributions in the scope of blind reverberation time estimation, diffuseness estimation, and sound event localization and detection. 
+Furthermore, several software tools developed for ambisonic dataset generation and management are also presented.
 
@@
 \section*{\Large \sffamily  Resum}
-Vet aquí el resum de la tesi en català.  Si us plau, utilitzeu
-menys de 150 paraules. \todo{abstract in catalan}
+
+Ambisonics ha esdevingut l'estàndard d'àudio immersiu als últims anys, afavorit pels avanços en realitat virtual i augmentada.
+L'àudio ambisonic es pot obtindre mitjançant \textit{arrays} de micròfons esfèrics, que també són cada vegada més populars.
+Tot i això, la majoria de mètodes acústics i de processament de senyal basats en \textit{arrays} de micròfons no estan adaptats al cas específic de geometries esfèriques.
+Per tant, encara hi ha moltes possibilitats de millora en l'àmbit d'anàlisi automàtic i descripció de gravacions ambisonic.
+En la present tesi plantegem aquest problema basant-nos en l'anàlisi paramètric del camp acústic.
+Més concretament, presentem contribucions originals en les àrees d'estimació de reverberació, estimació de difusió acústica, i detecció i localització d'esdeveniments sonors.
+Tanmateix, presentem diverses eines de programari desenvolupades per la generació i manteniment de bases de dades d'àudio ambisonic.
+
+%Els micròfons ambisonics permeten gravar escenes sonores preservant les seues característiques espacials.
+A causa d'això, ambisonics s'ha convertit en l'estàndard de facto per a la transmissió d'àudio immersiu.
+No obstant això, encara son poques les eines de producció que permeten explotar tota la informació disponible en aquest mitjà.
+%En la present tesi plantegem la pregunta de com obtenir informació rellevant de gravacions ambisonics de manera automàtica, basant-nos en l'an\`alisi param\`etric del camp ac\'ustic. 
+%%d'una forma que siga significativa per als productors.
+%Per a això, hem desenvolupat una sèrie de mètodes i eines a nivell de l'estat de l'art, 
+%que cobreixen un ampli espectre d'aplicaciones a l'\`ambit de l'an\`alisi param\`etric del so. 
+%%que s'apliquen a diferents nivells de la cadena de producció de so immersiu; cadascun d'aquests elements representa una contribució nova a l'estat de l'art en aquest camp.
+%
+%%En los últimos años hemos asistido al auge del contenido multimedia immersivo, auspiciado por los avances en Realidad Virtual y Aumentada.
+En este medio, el sonido espacial juega un papel muy importante para la creación de una sensación de realismo.
+%%
+%%Los micrófonos ambisonics permiten grabar escenas sonoras preservando sus características espaciales.
+%%Debido a esto, ambisonics se ha convertido en el estándar de facto para la transmisión de audio immersivo.
+Sin embargo, todavía hay pocas herramientas de producción que permitan explotar toda la información disponible en este medio.
+%%
+%%En esta tesis planteamos la pregunta de cómo obtener información relevante de grabaciones ambisonics de manera automática, de una forma significativa para los productores.
+%%
+%%Para ello, hemos desarrollado una serie de métodos y herramientas que se aplican a diferentes niveles de la cadena de producción de sonido immersivo; cada uno de estos elementos representa una contribución novedosa al estado del arte en ese campo.
+
 \vspace*{\fill}
@@
 %ADD BIBLIOGRAPHY
-\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib}
+\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib, Bibliography/Introduction.bib, Bibliography/ScientificBackground.bib}
 

### 2020-06-26 Chapters/DataGeneration.tex
wip
update 20200626

--- 2020-01-23 Chapters/DataGeneration.tex
@@
 \chapter{Data generation and storage}
+\label{chap:data}
 
-Explain about mono files plus ambisonics IRs.
+\section{Introduction}
+\label{sec:intro_data}
 
-\section{Recorded IRs}
+This chapter gathers several proposals related with the creation, storage and transmission of ambisonic data for research purposes.
+The main objective of the contributions described here is the support for the generation of parametrizable ambisonic datasets, using both synthetic and recorded materials, and specifically emphasizing the usage of Room Impulse Responses.\\
 
-AmbisonicsDRIR: explain problematics, standard, etc.
-also: pysofaconventions
+Most of the contributions listed here (and also most of the code developed for this thesis) have been implemented in Python.
+Indeed, Python has recently become one of the most used programming languages worldwide \cite{theoverflow, PYPL, TIOBE}; as shown also in Figure~\ref{fig:popularity}.\\
 
-\section{Simulated IRs}
+\begin{figure}
+\label{fig_architecture}
+  \centering
+    \includegraphics[width=\textwidth]{Figures/DataGeneration/projections-1-1400x1200.png}
+    \caption{2018 projections of future internet traffic for major programming languages. Adapted from \cite{theoverflow}.}
+    \label{fig:popularity}
+\end{figure}
 
-explain different methods and libraries.
-tell about masp.
+One of the reasons behind this tendency shift is the popularity of the language among machine learning and data science communities, fields where Python holds the first place by usage \cite{githubblog}.
+Since data-driven paradigms currently conform the state-of-the-art of many applied sciences, including audio signal processing, the availability of convenience Python packages and libraries is therefore of the highest interest to the research community.
 
-\section{High-level scene description}
+It is important to remark the predominant position that Matlab has always had regarding scientific computing.
+Indeed, it is still the tool of choice for many researchers, and the availability of libraries is accordingly very high.
+But the aforementioned tendency shift towards Python causes, as a side effect, the lack of many tools developed in Matlab by the research community.
 
-ambiscaper
+Although Matlab code can be called and executed from Python, in practice this approach is suboptimal under several criteria.
+A better solution in the long run is the effective port of the code towards native Python code.
+Some of the libraries presented in this Chapter are partially or totally  motivated by this scenario.
+
+\section{MASP: a Python library for multichannel acoustic signal processing}
+
+\subsection{Description}
+
+The Multichannel Acoustic Signal Processing (\textit{MASP}) is a Python library consisting of a collection of methods related with acoustics and microphone array processing.
+The library is mostly a transcoding from several Matlab libraries by A. Politis\cite{politis2016microphone, github_politis}. 
+It can be conveniently installed using \textit{pip}.
+
+\textit{MASP} implements a variety of methods for the simulation and analysis of reverberant acoustic scenes, with emphasis on microphone arrays with spherical geometries.
+More specifically, MASP is structured in submodules, with the following structure :
+
+\begin{description}
+
+	\item [Array Response Simulator] Simulation of
+spherical microphones:
+	\begin{itemize}
+		\item Rigid/open configurations.
+		\item Scattering simulation.
+		\item Arbitrary capsule distances, positions and directivities.
+	\end{itemize}
+
+	\item [Shoebox Room Model] Fast implementation of the Image Source Method \cite{imagemethod}:
+	\begin{itemize}
+		\item Convex 3D rooms.
+		\item Arbitrary number of sources and receivers, with arbitrary positions, orientations and directivities.
+		\item ISM expansion limited by order or time.
+		\item Frequency-dependent wall absorption.
+		\item RIR with spherical harmonic expansion.
+	\end{itemize}
+
+	\item [Spherical Array Processing] Transformation and analysis of signals measured with a spherical microphone array:
+	\begin{itemize}
+		\item A2B conversion with theoretical or measured filters.
+		\item Signal-independent beamforming.
+		\item Signal-dependent and adaptive beamforming.
+		\item Direction of Arrival estimation.
+		\item Diffuseness estimation.
+	\end{itemize}
+
+	\item [Spherical Harmonic Transform] Mathematical convenience tools.
+\end{description}
+
+The library implements a Unit Testing system, which numerically assesses the validity of the methods. 
+More specifically, each function test calls the equivalent Matlab code under the hood. The numeric result is then sent back to Python, where
+it is evaluated against the own result.\\
+
+Two example applications of the library are shown in Figures~\ref{fig:array_response} and \ref{fig:sht_filters}.
+In Figure~\ref{fig:array_response} , obtained with the Array Response Simulator package, the frequency response of a spherical microphone array to an impinging plane-wave with varying incidence angle is shown.
+The microphone array consists of a 2nd order supercardioid and a 3rd order hypercardioid, located at opposite directions of an open sphere, and both of them facing to the front direction.
+\\
+
+\begin{figure}[h!]
+\label{fig_architecture}
+  \centering
+    \includegraphics[width=\textwidth]{Figures/DataGeneration/array_response.png}
+    \caption{Frequency response of an arbitrary spherical array.}
+    \label{fig:array_response}
+\end{figure}
+
+One of the features of the Spherical Array Processing package is shown in Figure~\ref{fig:sht_filters}.
+The plot shows the evaluation of radial filters $\Gamma_n(kR)$ for an arbitrary spherical array, generated by inverting the theoretical response of the array \cite{Bertet2006}.
+The evaluation is performed following the metrics presented in the same paper, which compare spatial correlation, level difference and maximum amplification with respect to the ideal case. \\
+
+\begin{figure}[h!]
+\label{fig_architecture}
+  \centering
+    \includegraphics[width=1\textwidth]{Figures/DataGeneration/sht_filters.png}
+    \caption{Evaluation of radial filters for an arbitrary spherical microphone array.}
+    \label{fig:sht_filters}
+\end{figure}
+
+\subsection{Related software}
+
+There exists another recent Python library which covers a similar scope: \textit{pyroomacoustics} \cite{scheibler2018pyroomacoustics}.
+This framework provides an object-oriented interface with two main application scopes: allow RIR simulation of complex rooms based on the image source method, and provide a reference implementation of standard microphone array processing algorithms.
+
+Although some of the features are common to both libraries, there is a significant difference regarding their target usage.
+While \textit{MASP} primarily focuses on spherical geometries, \textit{pyroomacoustics} is more concerned about arbitrary room geometries and computational performance.
+Therefore, both libraries might be considered as complementary to some extent. 
+A comparative list of their features is shown in Table~\ref{tab:masp_features}.
+
+\begin{table}[th!]
+\centering
+\caption{Features of \textit{MASP} compared to \textit{pyroomacoustics}.}
+
+\begin{tabular}{cccc}
+  \toprule
+Package & Feature &  MASP & PRA \\
+\midrule
+Shoebox & Convex 3D room         & \checkmark	& \checkmark \\
+Room    & Non-convex 3D room     & -    			& \checkmark   \\
+Model   & Arbitrary \#sources    & \checkmark   & \checkmark      \\
+ & Arbitrary \#receivers, arrays & \checkmark   & \checkmark   \\
+& ISM by max\_order                                 & \checkmark                              & \checkmark                              \\
+& ISM by max\_time                                  & \checkmark                              & -                              \\
+& Wall absorption                                   & \checkmark                              & \checkmark                              \\
+& Frequency-dependent absorption                    & \checkmark                              & -                              \\
+& Plot methods                                      & -                              & \checkmark                              \\
+& RIR rendering                                     & \checkmark                              & \checkmark                              \\
+& Audio simulation                                  & \checkmark                              & \checkmark                              \\
+& Acoustic descriptor estimation                    & \checkmark                              & -                              \\
+& Microphone orientation                            & \checkmark                              & -                              \\
+& Custom microphone directivity                     & \checkmark                              & -                              \\
+& RIR Spherical Harmonic Expansion                  & \checkmark                              & -                              \\
+
+\midrule
+Array   	& Rigid spherical arrays   & \checkmark                              & -     \\
+Simulator 	& Arbitrary capsule geometries & \checkmark   & -  \\
+           	& Recorded array IRs  & \checkmark                              & - \\
+\midrule
+Spherical & A2B conversion   & \checkmark      & -       \\
+Processing  & Beamforming          		 & \checkmark      & \checkmark                              \\
+ & Plane-wave decomposition         		 & \checkmark      & -                              \\
+ & Nullformer                            & \checkmark      & -                              \\
+& Adaptive Beamforming                   & \checkmark      & -                              \\
+ & Adaptive Filtering             		 & -               & \checkmark                              \\
+& DoA Estimation                         & \checkmark      & \checkmark                              \\
+ & Diffuseness Estimation             	 & \checkmark      & -                              \\
+ & Diffuse-field coherence               & \checkmark      & -                              \\
+ & Blind Source Separation               & -               & \checkmark          \\                   
+\bottomrule
+
+\end{tabular}
+\label{tab:masp_features}
+\end{table}
+
+\section{SOFA}
+
+\subsection{Problem statement}
+
+The availability of recorded room impulse responses is of great importance to many acoustic signal processing problems. 
+Many different RIRs can be obtained from the same room, just by varying the position of the source and the receiver; when the number of source and receiver positions increases, the total amount of measurements increases geometrically.
+Besides that, the final format and organisation of the produced data (not only the RIR themselves, but also the source/position annotations) can be arbitrarily different when produced by different groups of people.
+
+In order to overcome potential interoperatibility and reusability issues, the \textit{Spatially Oriented Format for Acoustics} (SOFA) convention \cite{majdak2013spatially}, also known as the \textit{AES-69} standard \cite{majdak2015aes69}, proposes a unified file format for the storage of IR-related data.
+Despite that SOFA was initially created with an emphasis on \textit{Head-Related Impulse Response} (HRIR) data, the framework that SOFA provides can be potentially applied to a variety of recording procedures and audio-related data.
+Such variety is associated with the concept of \textit{conventions}: a specific data structure designed to hold a concrete type of data or measurement.
+Some examples of widespread conventions might be \textit{SimpleFreeFieldHRIR} (for anechoic binaural measurements), \textit{SimpleHeadphoneIR} (intended for storing headphone impulse responses), or \textit{MultiSpeakerBRIR} (for binaural RIRs measured from loudspeaker arrays), to name a few of them.
+
+\subsection{Ambisonics Directional Room Impulse Response as a SOFA convention}
+
+Given the intrinsic spatial characterization capabilities of ambisonics, Gerzon proposed the technique as a potentially successful candidate format for acoustical heritage preservation, as early as 1975 \cite{gerzon1975recording}.
+
+The increase in popularity of ambisonics since the beginning of the present century has turned this idea into reality; OpenAIRlib, a freely accesible dataset that gathers dozens of RIRs, might be a good example of it \cite{openair}.
+
+In any case, the usage of recorded ambisonic RIRs is not limited to the field of acoustic heritage.
+Among others, the availability of such recordings has powered works in a variety of works, auralization \cite{postma2016virtual}, room acoustics analysis \cite{embrechts2015measurement,clapp2011investigations} and modelling \cite{romblom2017diffuse}, spatial audio synthesis \cite{coleman2017object} or source separation \cite{baque2016separation}.
+\\
+
+In general, all publicly available ambisonic RIR measurements share some common approaches for describing and organizing the recorded data. 
+For instance, recordings from different rooms are usually stored as separated folders. 
+Each combination of emitter and receiver positions is often saved as an individual file, and the different spherical harmonics match the audio channels.
+Moreover, it is also usual to provide a \textit{metadata} file, describing the different emitter and receiver positions, and potentially some information about the measurement setup, methodology, etc.
+Such files might be formatted as plain text or delimiter-separated files.
+
+Despite the common approach, it can be easily foreseen that each database generated by a different individual or institution might potentially have a different naming convention, folder structure, file format, and so on.
+This is exactly the same situation that motivated the development of the SOFA conventions.
+
+On the other hand, the SOFA specification defines some criteria that must be fulfilled in order to propose a new convention \cite{sofaconventions}. These criteria are:
+\begin{enumerate}
+    \item Data must exist.
+    \item Data can not be described by existing SOFA conventions.
+    \item Relevant information about the data must be available.
+\end{enumerate}
+
+Given that the described situation meets all requirements, the \textit{Ambisonics Directional Room Impulse Response} (AmbisonicsDRIR) convention has been therefore proposed as a new member of SOFA.\\
+
+The technical specifications of the proposed convention in its current state (version 0.2) are available online \cite{ambisonicsdrir}. 
+Despite a timid adoption of the convention, the AmbisonicsDRIR proposal has arisen interest in the community.
+At the moment of writing, the \textit{Standardisation Committee on AES-69 Standard} is discussing potential modifications to the SOFA file format required for the adoption of several new features; and the support for spherical harmonic-based measurements is among the topics for deliberation.
+
+\subsection{pysofaconventions}
+
+The situation described in Section~\label{sec:intro_data}, regarding the availability of acoustic signal processing libraries in the Python programming language, can be easily extended to the case of SOFA APIs.
+
+The library \textit{pysofaconventions} has been created with the aim to provide an alternative to the existing Matlab/Octave and C/C++ implementations. 
+For ease of installation, it is integrated in the standard python package manager, \textit{pip}.
+
+The current software version is 0.1.5. The library structure is inspired by the C++ implementation \cite{api_cpp}. 
+It features all functionalities described by SOFA version 1.0, plus the proposed AmbisonicsDRIR convention. 
+The implementation is based on extensive error checking, to ensure code consistency.
+
+\textit{pysofaconventions} has gained a moderate amount of attraction.
+Among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}.
+
+\section{Ambiscaper}
+
+%
+%Explain about mono files plus ambisonics IRs.
+%
+%
+%\section{Recorded IRs}
+%
+%
+%Impulse Responses (IRs) measurements constitute a compact way of representing the acoustic propertiesof a linear time-invariant system.
+When such measurements are performed in a specific room or enclosure, the so-called Room Impulse Responses (RIRs) are able to capture the intrinsic reverberation and acoustic characteristics of the enclosure, for which several methods have been developed these past years \cite{stan2002comparison}.
+Furthermore, it is possible to account for different emitter/receiver positions in the measurement, usually performing the measurement with a microphone array.
+In that case, this kind of measurements is referred to as Directional Room Impulse Responses (DRIRs) \cite{embrechts2005computation}.
+DRIRs have a wide range of applications: auralization \cite{embrechts2005computation}, room acoustics analysis \cite{embrechts2015measurement,clapp2011investigations} and modelling \cite{romblom2017diffuse}, spatial audio synthesis \cite{coleman2017object}, source separation and dereverberation \cite{baque2016separation}, acoustic heritage preservation \cite{gerzon1975recording,murphy2005multi}, etc.
+%
+%
+%\subsection{Ambisonics Recording}
+%
+%It is also possible to capture Ambisonics audio scenes by using specific recording devices.
+Spherical microphone arrays (also known as \textit{Ambisonics microphones}) are a type of microphone arrays in which the capsules are located around a spherical surface, presenting rotational symmetry.
+Such geometric arrangement allows the recorded signals to be transformed to the Ambisonics domain, by means of projection into the Spherical Harmonics basis, and further equalization with radial filters \cite{Bertet2006}.
+%
+%This process is known in the audio production domain as \textit{A-B Conversion}.
+All major spherical microphone array manufacturers provide tools for achieving this transformation \cite{soundfield,ambeo,zylia,tetra,em32}.
+%
+%
+%\subsection{Ambisonics DRIRs}
+%
+%The intrinsic spatial capabilities of Ambisonics microphones might be applied to DRIR recording, as originally proposed by Gerzon in the context of acoustical heritage preservation \cite{gerzon1975recording}.
+In recent years, several datasets of Ambisonics DRIRs have been publicly released, such as the OpenAIR database \cite{murphy2010openair} or the set of  measurements performed in the scope of the S3A project \cite{coleman2015s3a,openair}.
+%
+%\subsection{The SOFA Conventions}
+%
+%In general, the Ambisonics DRIR databases show a common approach for describing the measurements: given a specific room, usually represented as a folder, IR data consist of several multichannel audio files, with one audio channel per spherical harmonic, and one file per emitter/receiver combination.
+Furthermore, it is also usual to provide a \textit{metadata} file, describing the different emitter and receiver positions, and eventually some information about the measurement setup, methodology, etc.
+Such files might be formatted as plain text or delimiter-separated value files.
+%
+%Despite the common approach, it can be foreseen that each database generated by a different individual or institution might potentially have a different naming convention, folder structure, file format, and so on.
+This situation hinders data manipulation and exchange, and forces users to write ad-hoc parsers and algorithms for each specific database.
+%
+%The \textit{Spatially Oriented Format for Acoustics} (SOFA) is a file format designed for a consistent, standardized storage and manipulation of IR data \cite{majdak2015aes69}.
+The need for such standard arose from dealing with different databases of \textit{Head-Related Transfer Functions} (HRTFs), in a similar manner as the one mentioned for Ambisonics DRIRs.
+%
+%There are several SOFA conventions, each one addressing a particular type of IR measurement.
+In the case of Ambisonics DRIRs, given the data representation in existing databases, one could outline the following specificities:
+% 
+%\begin{itemize}
+%    \item Presence of Ambisonics-related information (Ambisonics order, channel ordering and normalization)
+%    \item Audio stored in the Ambisonics (spherical harmonics) domain  
+%    \item Data structure support for different combinations of source and receiver positions
+%\end{itemize}
+%
+%
+%However, none of the existing SOFA conventions meets those requisites.
+The potentially first candidate by name, \textit{SingleRoomDRIR}, is limited to one source position per file.
+On the other hand, \textit{MultiSpeakerBRIR} allows for multiple sound sources, but restricts the number of receivers (microphone capsules) to two, as expected in a binaural recording.
+\textit{GeneralFIRE} is intended for "data which are too general to store in more specific conventions" \cite{generalfire}.
+%
+%
+%\subsection{Convention Proposal}
+%
+%\subsubsection{Considerations}
+%
+%The SOFA specification defines some criteria that must be fulfilled in order to propose a new convention \cite{sofaconventions}. These criteria are:
+%\begin{enumerate}
+%    \item Data must exist.
+%    \item Data can not be described by existing SOFA conventions.
+%    \item Relevant information about the data must be available.
+%\end{enumerate}
+%
+%
+%As we already mentioned, existing databases of Ambisonics DRIRs can be found in OpenAIR and S3A databases (criterium 1).
+Criterium 2 has been discussed in the previous paragraph.
+All available Ambisonics DRIR datasets are accompanied by explanations, pictures, diagrams and related information (criterium 3).
+%
+%\subsubsection{Specifications}
+%
+%We therefore propose a new SOFA convention, \textit{AmbisonicsDRIR}, designed for DRIRs measured with spherical microphone arrays, and presented in the Ambisonics domain.
+In other words, we propose to store the instantaneous $B_{mn}$ values for a given Ambisonics order $L$, provided that $S$ is a unit impulse (Eq.
+\ref{eq1}).
+%
+%The \textit{Listener} - as defined by the SOFA specification - is embodied by the spherical microphone array.
+The different \textit{Receivers} are the different Ambisonics Components $B_{mn}$, so that their number is fixed provided $L$ (more precisely, by following the relationship $R = (L+1)^2$), and their positions are not applicable.
+Furthermore, there might be many different \textit{Sources}, all of them omnidirectional and consisting of one only \textit{Emitter}.
+In that sense, \textit{M} represents the number of different \textit{Source} positions.
+%
+%
+%The proposed convention is based on \textit{GeneralFIRE}, with the following modifications:
+%\begin{itemize}
+%
+%    \item Mandatory field \textit{GLOBAL:AmbisonicsOrder} (type \textit{double}, dimension \textit{I}, default \textit{1}). Indicates the order of the Spherical Harmonic expansion.
+%    
+%    \item Mandatory field \textit{DATA:ChannelOrdering} (type \textit{attribute}, default \textit{acn}).
+Describes the ordering of the different Ambisonics Channels.
+Must be one of: \textit{acn} or \textit{fuma}.
+%    
+%    \item Mandatory field \textit{DATAChannelNormalization} (type \textit{attribute}, default \textit{sn3d}).
+Describes the Ambisonics normalization convention used in the data.
+Must be one of: \textit{sn3d}, \textit{n3d}, \textit{fuma} or \textit{maxn}.
+%    
+%    \item Fields \textit{ReceiverPosition}, \textit{ReceiverPosition:Type} and \textit{ReceiverPosition:Units} are not needed, since it is assumed that \textit{Receivers} represent the Ambisonics channels.
+Furthermore, the values of \textit{R} (number of receivers) are defined by $R = (L+1)^2$.
+Accordingly, the values of \textit{GLOBAL:AmbisonicsOrder} and \textit{R} are valid only if they follow the given equation.
+%    
+%    \item Field \textit{DATA:Delay} is not mandatory.
+%    
+%    \item Relevant information about the microphone model and brand, Ambisonics encoding methodology, software used, etc, is recommended to be added into the field \textit{GLOBAL:Comment}.
+%    
+%\end{itemize}
+%
+%\subsection{Results}  
+%
+%As a preliminary result, we have extended the current SOFA C++  \cite{sofacpp} and Matlab \cite{sofamo} APIs to be fully compatible with the AmbisonicsDRIR convention. \todo{pysofaconventions}
+%
+%Furthermore, as a use-case, a selection of existing Ambisonics DRIRs have been transcoded to the proposed convention: \textit{Emmanuel Main Church} from S3A database \cite{coleman2015s3a}, and \textit{Heslington Church} and \textit{York Guildhall Council Chamber} from the OpenAIRlib \cite{openair}.
+Figure \ref{guildhall} shows a schematic diagram of the different \textit{Emitter/Receiver} positions at the \textit{York Guildhall Council Chamber} recordings.
+The data, as well as the tools used to perform the conversion, are available online under an open source license \cite{ambisonicsdrirexamples}.
+%
+%Finally, we must remark that the \textit{AmbisonicsDRIR} SOFA convention proposal is currently under discussion, and the described specifications are subject to change with upcoming versions. 
+%
+%% TODO: include some examples!!! about data dimensions etc
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{Figures/DataGeneration/guildhall}
+%	\caption{Source and Listener position diagram for the "York Guildhall Council Chamber" Ambisonics DRIR (original diagram attribution to \cite{guildhall} }
+%	\label{guildhall}
+%\end{figure}
+%
+%\subsection{Summary} 
+%
+%This document addresses the lack of compatibility among different databases of Ambisonics DRIRs.
+The present proposal consists in defining a new SOFA convention, specifically designed for Ambisonics DRIRs.
+That way we contribute to improve the ease of data manipulation and database interoperability.
+Software implementations and tools for automatic conversion are provided.
+%
+%
+%
+%\section{Simulated IRs}
+%
+%explain different methods and libraries.
+%tell about masp.
+%
+%
+%
+%
+%
+%
+%% ==================================================================
+%\section{High-level scene description}
+%
+%\subsection{Motivation}
+%
+%\todo{introduce the topic}
+%
+%\todo{ARRANGE this introduction in a meaningful way}\\
+%The work by Moore \cite{Moore2015} was the first to combine Intensity Vector statistics with DPD-test preprocessing.
+A similar approach is found on the \textit{Single Source Zones (SSZs)} algorithm, first presented by Pavlidi \cite{Pavlidi2015}.
+Other recent proposals, as for example the work by He \cite{He2017}, consider the local DOA variance as an estimator of reliability.
+%
+%Source Localization based on IV statistics has been also successfully used as a preprocessing step for Source Separation tasks.
+The proposal of G\"unel \cite{Gunel2008}, which uses beamforming over the DOA estimation, was a pioneer work on this scope.
+Other similar approaches, which rely on DOA-based Time-Frequency (TF) masking, can be found on the works by Shujau \cite{Shujau2011}, Riaz \cite{Riaz2015} or Chen \cite{Chen2015}.
+%
+%
+%
+%\subsection{Evaluation data}
+%\label{subsec:evaluationdata}
+%
+%Audio data used for evaluation in the works presented in Section \ref{subsec:algorithms} can be classified in  three main categories:
+%
+%
+%\begin{itemize}
+%
+%    \item \textit{IR simulations:} Impulse Responses simulated by numerical methods, using software tools such as \textit{SMIR Generator} \cite{smir}.
+IR simulation provides the most flexible approach for evaluation, since it allows custom design and parametric analysis (algorithm evaluation as a function of reverberation time, source(s) distance(s), etc).
+However, the resulting audio data realism is restricted by the simulation algorithm, which is usually limited to empty \textit{shoebox} rooms.
+%    
+%    \item \textit{IR recordings:} Impulse Responses (IRs) of specific rooms recorded with an Ambisonics microphone, as first presented by Gerzon in the context of acoustic heritage preservation \cite{gerzon1975recording}.
+IR recordings do not provide the flexibility of IR simulations, but allow realistic captures of the room's acoustic properties, while maintaining the potential of using different source contents and positions.
+OpenAIRlib \cite{murphy2010openair} is the reference public repository for IRs, even including a dedicated Ambisonics IR section.
+%
+%    \item \textit{Audio recordings:} a sound scene recorded in a specific room, using an Ambisonics microphone and real sound sources.
+This approach  represents the  most  accurate  option  for algorithm evaluation in a real scenario.
+However, its cost and lack of flexibility makes this approach only attractive for last stages of algorithm  evaluation.
+%
+%\end{itemize}
+%
+%In the case of IR simulations and recordings, the Ambisonics IRs are convolved with anechoic monophonic recordings to produce the actual evaluation data.
+The anechoic recordings are usually taken from standard audio processing databases, such as TIMIT \cite{timit}.
+\\
+%
+%Table \ref{table:evaluationdata} summarizes the evaluation data types used in each work review in Section \ref{subsec:algorithms}, along with the audio content type and origin database (if any).    
+%
+%
+%%% DOA ESTIMATION
+%\begin{table}[htpb]
+%    \centering
+%    \footnotesize
+%
+%	\begin{tabular}{ p{2cm} p{2.8cm} p{2.5cm} p{2.3cm} }
+%    \toprule
+%    % \cmidrule(r){1-2}
+%    
+%    % \centering 
+%    \textbf{Article} &
+%    % \textbf{\textit{$L$}} &
+%    \textbf{Evaluation Data Type} &
+%    \textbf{Audio Content Type} &
+%    \textbf{Database}\\
+%    \midrule
+%    
+%    Thiergart \cite{thiergart_localization_2009}
+%    % & 1 h
+%    & Audio recording 
+%    & Speech
+%    & -\\
+%    
+%    Tervo \cite{Tervo2009}
+%    % & 1 h
+%    & Audio recording
+%    & Noise, music
+%    & -\\
+%    
+%    Jarret \cite{Jarrett2010} 
+%    % & 4 
+%    & IR simulation, Audio recording
+%    & Noise
+%    & -\\
+%    
+%    Nadiri \cite{Nadiri2014}   
+%    % & 4 
+%    & IR simulation, Audio recording
+%    & Speech
+%    & TIMIT \\
+%    
+%    Moore \cite{Moore2015}   
+%    % & 4 
+%    & IR simulation 
+%    & Speech
+%    & APLAWD \\
+%    
+%    Pavlidi \cite{Pavlidi2015}
+%    % & 4
+%    & IR simulation 
+%    & Noise, speech
+%    & -\\
+%    
+%    He \cite{He2017}   
+%    % & 1 h
+%    & IR simulation, Audio recording
+%    & Speech
+%    & TIMIT \\
+%
+%    
+%    % Ding \cite{Ding2017}   
+%    % % & 1 h
+%    % & IR simulation \& Audio recording \\
+%    
+%    \midrule
+%    
+%    Gunel \cite{Gunel2008}
+%    & IR recording
+%    & Speech, music
+%    & Music for Archimedes \\
+%    
+%    Shujau \cite{Shujau2011} 
+%    & Audio recording
+%    & Speech
+%    & TIMIT \\
+%    
+%    Riaz \cite{Riaz2015}
+%    & IR recording
+%    & Speech, music
+%    & Music for Archimedes \\
+%    
+%    Chen \cite{Chen2015}
+%    & IR simulation, IR recording
+%    & Speech
+%    & TIMIT \\
+%
+%    \bottomrule
+%    \end{tabular}
+%    \caption{Summary of audio data used across Ambisonics-based Source Localization (above) and Source Separation (below)  algorithm proposals. \todo{update table? fix references}}
+%    \label{table:evaluationdata}
+%\end{table}
+%
+%
+%
+%Although algorithm evaluation in real scenarios provides the most realistic approach to the problem, this methodology has not been widely adopted due to its cost and lack of flexibility.
+As an example, none of the works using real recordings in Table \ref{table:evaluationdata} performs the evaluation with more than one  recording in each case.
+In contrast, when using IR-based  scenes, the number of audios evaluated are usually one or two magnitude orders greater.
+%
+%However, it is important to notice the lack of public availability of evaluation data.
+None of the analyzed articles provide a way to access neither the used audio dataset, nor the groundtruth (position annotations in the case of Sound Localization, and original sound sources for Sound Separation).
+Only in the case of simulated IRs it is possible to partially replicate the experimental setup, since some of the parameters used in the simulation software are usually provided.
+Furthermore, the process of creation of the custom datasets (selecting an anechoic audio dataset, convolving with custom IRs, or performing real recordings) seems to be performed \textit{ad-hoc} on each article.
+%
+%Taking into account the flexibility offered by IR-based scenes, it would be desirable to have a tool for automatic generation of reverberant Ambisonics scenes (and their associated groundtruth) for analysis purposes.
+Such tool would help the scientific community in several ways: reducing the amount of time dedicated to build custom datasets, reusing publicly available resources and recordings, and enhancing experiment reproducibility by making easier the exchange of datasets.
+Furthermore, the capacity of producing a big number of diverse audio scenes for analysis will help the algorithm design and early testing stages,and specially the training stage of machine learning-based algorithms.
+%
+%
+%\subsection{AMBISCAPER}
+%\label{sec:ambiscaper}
+%
+%\todo{this has been downgraded to subsection. maybe place it at the introduction or something to find a more meaningful explanation}
+%
+%AmbiScaper \cite{ambiscaper} is a tool designed to provide a flexible way of creating complex Ambisonics sound scenes and their associated groundtruth, to be used in the context of Source Localization and Source Separation algorithms.
+%AmbiScaper offers a high level control of the sound scene parameters, and provides an easy way of creating large datasets with custom characteristics.
+%AmbiScaper is based on Scaper, a framework designed to generate ground truth information to train Sound Event Detection models \cite{Salamon2017}.
+%
+%\subsection{Sound scene description}
+%\label{subsec:description}
+%
+%One of the main features of AmbiScaper is that all parameters of the sound scene can be specified in a non-deterministic way.
+In that sense, the parameters for each \textit{event} (sound source) are actually generated through a two-step process.
+First, in the \textit{Event Specification}, all parameters related to an event are defined in terms of statistical distributions.
+During the \textit{Event Instanciation}, the actual values for each parameter are then sampled from the statistical distributions.
+%This two-step process allows the user to describe abstract \textit{templates} of sound scenes, rather than manually assigning values to parameters.
+Therefore, a single \textit{event specification} might produce potentially infinite different sound scenes.
+%
+%
+%
+%\subsection{Architecture}
+%\label{subsec:architecture}
+%
+%\begin{figure}
+%\label{fig_architecture}
+%  \centering
+%    \includegraphics[width=\textwidth]{Figures/DataGeneration/figure_architecture_V2.png}
+%    \caption{AmbiScaper architecture.}
+%\end{figure}
+%
+%%%%%%%%%
+%In order to generate a sound scene, AmbiScaper requires three different inputs: the original \textit{mono signals}, which will be the basis for the scene's audio content, an optional \textit{Ambisonics IR}, and the \textit{event specification}.
+%
+%The process of dataset creation starts with the \textit{event instanciation}, as described in Section \ref{subsec:description}.
+Once all values are sampled, three different types of output are generated: the \textit{Ambisonics scene}, the \textit{instanciated mono signals} (the original mono signals after data augmentation and duration changes), and the \textit{annotations}, in the form of a JAMS file \cite{humphrey2014jams}, containing all information about the scene specification and the instanciated values.
+%
+%AmbiScaper's architecture is depicted in Figure \ref{fig_architecture}.
+%
+%
+%\subsection{Reverberation}
+%
+%When no reverberation is specified, AmbiScaper can generate anechoic sound scenes, which can be useful for baseline performance evaluation.
+In this case, there is no upper limit on the Ambisonics order of the rendered scene.
+Furthermore, the anoechic case allows for the specification of source \textit{spread} or apparent size through Ambisonics order downgrade \cite{carpentier2017ambisonic}.
+%
+%AmbiScaper supports the usage of recorded Ambisonics IRs, although it is currently limited to IRs from the S3A database \cite{coleman2015s3a}.
+The development of a standardized file format for Ambisonics IRs, which is being discussed at the moment of writing, will provide the flexibility to work with arbitrary Ambisonics IRs.
+%
+%Lastly, AmbiScaper features the possibility of using simulated Ambisonics IRs, through a wrapper to \textit{SMIR Generator}.
+In this case, the reverberation model specifications might be defined as well in statistical terms, and the generated IRs are stored for evaluation purposes.
+A working copy of Matlab is required to run this option.
+%
+%When a reverberant sound scene is created, the specific Ambisonics IRs used on the scene are also provided as an output.
+Other research problems such as dereverberation or room reflection modelling might therefore benefit from such data.
+%
+%
+%
+%
+%\subsection{AmbiScaper and experiment reproducibility}
+%
+%As mentioned in Section \ref{subsec:evaluationdata}, there is a generalized lack of publicly available datasets for Source Localization and Source Separation in the Ambisonics domain.
+Even when using general purpose audio/speech datasets (such as TIMIT), the actual reverberant Ambisonics evaluation data is usually not available.
+%In that sense, the compatibility of AmbiScaper with public Ambisonics IR databases is a key aspect for reproducibility, since it allows for the reutilization of acoustical measurements in a systematic way.
+%
+%Furthermore, the output of the AmbiScaper dataset generation process is not limited to the actual dataset.
+In fact, as explained in Section \ref{subsec:architecture}, the resulting annotation file does not only contain the \textit{instanciation} (the actual values of each parameter in the sound scene), but also the \textit{specification} (the statistical distributions from which the instanciated values are sampled).
+In the scope of experiment reproducibility, the exchange of \textit{specification files} instead of actual audio files greatly reduces the storage capacitiy and bandwith required to transfer big databases.
+%
+%AmbiScaper is implemented in the form of a Python package, publicly available through the \textit{Python Package Index} repository. 
+%AmbiScaper is free software under the GPL license, easing the software adoption and the potential engagement of the scientific community with the development. 
+%
+%
+%\subsection{Sample Dataset}
+%\label{ssec:dataset}
+%
+%As an example of the potential capabilities of AmbiScaper,  we have created and published a dataset for the evaluation of source localization algorithms \footnote{https://zenodo.org/record/1186907 \todo{link}}.
+%
+%The dataset contains 300 first order Ambisonics sound scenes, each one containing from one to three static sound sources (be they simultaneous or not), with different gains and SNRs, and placed at random positions around the sphere.
+Each sound scene has a duration between 1 and 2 seconds.
+%
+%The sources are randomly chosen from a subset of the Anechoic OpenAirlib database, which mostly contains recordings from baroque musical instruments.
+%The AudioBooth IR database from S3A has been used for all scenes.
+%It features SoundField recordings from an acoustically treated room with a geodesic dome structure, to which 17 speakers have been attached.
+%
+%
+%
+%\subsection{CONCLUSIONS AND FUTURE WORK}
+%\label{sec:conclusions}
+%
+%AmbiScaper is a tool designed for easy dataset creation and exchange, in the context of reverberant Ambisonics Source Localization and Source Separation.
+It responds to the lack of public datasets for algorithm design, evaluation and reproducibility.
+An example dataset generated with AmbiScaper, and its analysis with state-of-the-art Source Localization algorithms are also provided.
+%Further studies in Ambisonics IR position interpolation would allow creating non-static audio scenes, which might be an interesting feature. 
+%Another feature in progress is the support for a standardized Ambisonics IR format \cite{perez2018ambisonics}, which will ease data reusability and provide the potential to cover a big variety of acoustic scenarios.
+%
+%
+%
+%
+%
+%
+%

### 2020-06-26 Chapters/Introduction.tex
update 20200626

@@
 Nevertheless, the greatest contributor to the current interest in ambisonics has been the rise of Virtual Reality (VR) in recent years.
-Although VR focuses primarily on visual cues, the immersive experience might be greatly enhanced by spatial audio \todo{CITE REQUIRED, begault?}.
+Although VR focuses primarily on visual cues, the immersive experience might be greatly enhanced by spatial audio \cite{begault20003}	.
 In this context, Ambisonics has been rapidly adopted as \textit{de facto} standard for spatial audio transmission, supposedly due to a variety of factors:
@@
 
-  \item [Binaural transformation] Spatial audio in VR is mostly consumed as binaural; methods for ambisonic to binaural conversion have been known for a long time \todo{cite?}.
-Furthermore, VR headsets can easily provide head rotation information, which can be used in combination with scene rotations to provide head-locked audio, which greatly improves immersiveness \todo{cite required!}.
+  \item [Binaural transformation] Spatial audio in VR is mostly consumed as binaural; methods for ambisonic to binaural conversion have been known for a long time \cite{noisternig20033d}.
+Furthermore, VR headsets can easily provide head rotation information, which can be used in combination with scene rotations to provide head-locked audio, which greatly improves localization accuracy and inmersiveness \cite{begault20003}.
 This is a key feature of ambisonics when compared to static binaural recordings.
@@
 (i) the billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}; 
-(ii) the Time cover page devoted to VR: \textit{""The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
+(ii) the Time cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
 and (ii) M.
@@
 
-Although the VR momentum has also reached spherical microphone array processing, many challenges remain still opened, and the number of research works specifically focusing on such geometry are still low.
+Although the VR momentum has also reached spherical microphone array processing, many challenges remain still open, and the number of research works specifically focusing on such geometry are low yet.
 Besides that, the growing interest in AR posses new problems related to acoustical signal processing.
@@
 \section{Problem Description}
+\label{sec:problemdescription}
 
-\begin{figure}[hbt]
+The scientific context of the work developed in the present dissertation is depicted in Figure~\ref{fig:scheme1}, which has been inspired by \cite{jarrett2017theory}. 
+As we can observe, there are two main topics related with B-Format audio: \textit{generation} and \textit{analysis}, with the signal flow going from the former to the latter.
+Although the conceptual approach of the scheme might be very similar for any type of audio, the specific spatial information conveyed by the ambisonic signal places an emphasis on the informed analysis and description of the sound scene.
+\\
+
+\begin{figure}[th]
   \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1.png}
-  \caption{\todo{todo caption}}
+  \caption{General scheme of the B-Format audio generation and analysis framework.  Solid lines represent audio signals, while outlined arrows refer to non-audio information.}
   \label{fig:scheme1}
@@
 
-EXPLAIN PARTS OF FIGURE:
+The problem of B-Format generation is mostly related with dataset generation, which is a common issue for many audio signal processing problems. 
+In our case, we consider two different approaches to the data generation problem:
+\begin{description}
+	\item[Recording] Using spherical microphone arrays for the recording of ambisonic material.
+	\item[Simulation] Using numerical methods for the simulation of acoustic scenes.
+\end{description}
+While recordings are by definition more similar to real scenarios, they are expensive to perform, and can only provide a limited set of parameter possibilities.
+Simulations, on the other hand, have the potential to cover any desired condition.
+Therefore, it is of our interest to consider the strengths of both signal generation paths.\\
 
-Different levels of applications in multichannel acoustic signal processing (inspired by \cite{jarrett2017theory})
+B-Format analysis has been divided into three different categories:
+\begin{description}
+	\item[Acoustic Parameter Estimation] Low-level analysis of the sound field, which yields quantitative information about different acoustic parameters used to model the acoustic scene.
+	\item[Signal Enhancement] Modification of the sound field in order to obtain a set of output signals with desired attributes. 
+	\item[Scene Description] Textual representation of different high-level characteristics of the sound field.
+\end{description}
 
-\begin{itemize}
-	\item Acoustic Parameter Estimation (low level, audio2data)
-	\begin{itemize}
-		\item Direction of Arrival estimation
-		\item Coherence analysis
-		\item Acoustic description (RT60, etc)
-		\item Source counting
-	\end{itemize}
+The knowledge about the acoustic parameters of a sound field can be considered either as a goal by itself, or alternatively as a preprocessing step which complements the other analysis categories.
+Examples of typical estimated acoustic parameters are the \textit{Direction-of-Arrival} (DOA), the sound field diffuseness, or the reverberation time of the enclosure \cite{jarrett2017theory}.
 
-	\item Signal Enhancement  (high level, audio2audio)
-	\begin{itemize}
-		\item Source Separation
-		\item Dereverberation / denoising
-		\item IR estimation
-	\end{itemize}
+There are a variety of well known signal enhancement problems, including dereverberation \cite{braun2018speech}, source separation \cite{gannot2017consolidated}, or foreground-background segmentation \cite{stefanakis_foreground_2015}.
+As it has been shown, many of them benefit (or even depend) from the knowledge derived by the acoustic parameter estimation step.
 
-	\item Scene Description (high level, audio2data)
-	\begin{itemize}
-		\item Event Detection
-		\item Acoustic Scene Classification
-	\end{itemize}
-\end{itemize}
+Finally, under the scene description typology we can find a set of applications that provide abstract representations of the sound scene under analysis.
+Most of the recent research performed in this scope is grouped around the Detection and Classification of Acoustic Scenes and Events (DCASE) community\footnote{\url{http://dcase.community/community_info}}.
+Examples of the problems under consideration are acoustic scene classification \cite{Mesaros2018_DCASE}, sound event localization and detection \cite{Adavanne2018_JSTSP} or audio captioning \cite{drossos2017automated}.
 
@@
 
-preguntas cientificas concretas, coincidiendo con contribuciones
+The list that follows concentrates the main scientific objectives to be developed on this thesis:
 
-(base para la conclusion)
-
+\begin{itemize}
+	\item To develop methods for the characterisation of acoustic parameters from recordings originated from ambisonic microphones. 
+	\item To propose methodologies for sound event localization and detection in ambisonic domain which are grounded on spatial parametric analysis. 
+	\item To contribute in the generation and storage of ambisonic sound scenes, for their usage in controlled experimental environments. 
+\end{itemize}
 %\section{Goals}
@@
   \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
-  \caption{\todo{todo caption}}
+  \caption{General scheme of the B-Format audio generation and analysis framework, including the thesis contributions in form of Chapter numbers.}
   \label{fig:scheme1_numbers}
@@
 
-The present Thesis is organised as follows.
+The present dissertation is organised as follows.
 
-\textbf{Chapter~\ref{chap:scientific}} introduces the basic concepts that will be developed throughout the Thesis, including spherical harmonics and ambisonics, coherence estimation, parametric analysis or room acoustics.
+\textbf{Chapter~\ref{chap:scientific}} introduces the basic concepts that will be developed throughout this thesis, including spherical harmonics and ambisonics, coherence estimation, parametric analysis or room acoustics.
 The Chapter also defines the signal models and the mathematical terminology.
 
-Chapters~\ref{chap:rt60}, \ref{chap:coherence} and \ref{chap:seld2019} develop the most significant academic contributions of the Thesis. 
+Chapters~\ref{chap:rt60}, \ref{chap:coherence} and \ref{chap:seld2019} develop the most significant academic contributions of this thesis. 
 \textbf{Chapter~\ref{chap:rt60}} presents a novel method for blind reverberation time in ambisonic recordings.
@@
 
-Finally, \textbf{Chapter~\ref{chap:data}} presents some libraries and software utilities developed throughout the Thesis.
+Finally, \textbf{Chapter~\ref{chap:data}} presents some libraries and software utilities developed throughout this thesis.
 All the code has been publicly released under open source licenses.
 The libraries include utilities for the creation of datasets, the storage and exchange of impulse response files in a standard way, and the implementation of convenience tools for acoustic and microphone array signal processing analysis.
-Although the libraries do not directly involve any scientific contribution, they can be a great help for scientific and innovative purposes; given the industial nature of the candidate's doctoral program, we have considered relevant to include them in the present Thesis.
-\todo{check}
+Although the libraries do not directly involve any scientific contribution, they can be a great help for scientific and innovative purposes; given the industrial nature of this thesis, we have considered relevant to include them in the present dissertation.\\
 
-In order to place the different chapters within the problem context described in Section~\ref{sec:context}, we add to Figure~\ref{fig:scheme1} the Chapter numbers with our contributions, obtaining Figure~\ref{fig:scheme1_numbers}.
-
-\section{Context}
-\label{sec:context}
+In order to provide a schematic representation of the thesis structure and scope, Figure~\ref{fig:scheme1_numbers} features the problem structure stated in Figure~\ref{fig:scheme1}, with the addition of the Chapter numbers with the contributions of this thesis.
 
@@
 
-In the following list we show the main scientific contributions of the Thesis, organised by Chapters:
+In the following list we show the main scientific contributions of this dissertation, organised by Chapters:
 
@@
 	\item Chapter 5:\\
-	\textbf{A hybrid parametric-deep learning approach for sound event localization and detection.}
+
+	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
+	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
+
+	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
 	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
@@
 
-	\todo{DCASE2020}
-
 	\item Chapter 6:\\
-	\textbf{Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes}.\\
+	\textbf{"Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes".}\\
 	\underline{A. P\'erez-L\'opez}.
@@
 
-	\textbf{Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics.}
+	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
 	\underline{A. P\'erez-L\'opez} and J. De Muynke.
@@
 
-	\textbf{pysofaconventions, a Python API for SOFA.}
+	\textbf{"pysofaconventions, a Python API for SOFA".}
 	\underline{A. P\'erez-L\'opez}.
@@
 
-	\textbf{A Python library for Multichannel Acoustic Signal Processing.}
+	\textbf{"A Python library for Multichannel Acoustic Signal Processing".}
 	\underline{A. P\'erez-L\'opez} and A. Politis.
@@
 
-Moreover, as a result of the development of the Thesis, the following open-source libraries have been implemented and released.
-All of them are available through the author's GitHub page, \url{https://github.com/andresperezlopez}: \todo{describir cada uno?
-con respecto a paper o chapter?}
+Moreover, although not strictly aligned with the research direction of this thesis, the author has collaborated in the following publications:
 
 \begin{itemize}
+	\item \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
+	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
 
-	\item rt60 estimation: \todo{todo}
+\end{itemize}
+
+Finally, as a result of the development of this thesis, the following open-source libraries have been implemented and released.
+All of them are available through the author's GitHub page\footnote{\url{https://github.com/andresperezlopez}}:
+\begin{itemize}
+
+	\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation}
+
+	\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020}
 

### 2020-06-26 Chapters/SELD.tex
wip
update 20200626

@@
 \chapter{Sound Event Localization and Detection}
+\label{chap:seld2019}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 \label{sec:intro}
+%
 
-Sound Event Localization and Detection (SELD) refers to the problem of identifying, for each individual  event present in a sound field, the temporal activity, spatial location, and sound class to which it belongs.
+Sound Event Localization and Detection (SELD) refers to the problem of identifying, for each individual event present in a sound field, the \textbf{spatial location} $\Omega$, \textbf{temporal activity} $\Upsilon$, and \textbf{sound class} $\kappa$ to which it belongs.
+
+The organization of a dedicated SELD task within the IEEE AASP Challenge on Detection and Classification of Acoustic Scenes and Events (DCASE) 2019 can be considered as a milestone for the development of the SELD research problem.
+Indeed, a large number of novel methodologies were developed for the Challenge, most of them based on Convolutional Recurrent Neural Networks (CRNN).
+The performance of the baseline method, a CRNN that performed jointly the localization and classification tasks \cite{Adavanne2018_JSTSP}, was vastly exceeded by a variety of deep-learning based algorithms \cite{kapka2019sound, Cao2019, grondin2019sound}.
+Some of these improvements have been included in the baseline system for the SELD Challenge of DCASE 2020.
+
+Despite the predominant trend towards high-complexity deep-learning architectures, some recent works have been able to match or even improve CRNN-based methods with regard to localization, by using parametric analysis of the ambisonic sound field \cite{perez2019hybrid, nguyen2020sequence}.
+Apart from the benefit derived by their simplicity, these approaches are able to resolve the case of overlapping events of the same class, a situation difficult to disambiguate for CRNN-based methods \cite{politis2020dataset}.
+
+The present work continues the exploration of possibilities of parametric SELD methods, focusing on a low-complexity architecture that makes use of traditional, feature-based machine learning techniques.
+The method has been developed in the context of the SELD task within DCASE 2020 Challenge, and therefore utilizes the proposed dataset, baseline system and evaluation metrics.
+
+Finally, it is important to remark that the method described here is the continuation of an algorithm presented at the DCASE 2019 Challenge \cite{perez2019hybrid}; both algorithms share a common structure and a similar approach to the SELD problem.
+However, the current proposal tries to solve some of the problems identified on our early approach, mainly related with a low frame recall derived from a naive approach to event segmentation.
+Moreover, the current method also presents a complexity reduction, regarding the single-source
+DOA estimation and the classifier.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{SYSTEM DESCRIPTION}
+\label{sec:methodology}
+
+The proposed method, referred to as \textit{PAPAFIL}, can be summed up in four steps:
+
+\begin{enumerate}
+    \item Estimate single-source time-frequency bins.
+    \item Use a particle tracking system to estimate event trajectories and activation times from single-source bins.
+    \item Perform spatio-temporal filtering on the input signal.
+    \item Assign a class label to the estimated event.
+    % \item The estimated event is assigned a class label by a classifier.
+\end{enumerate}
+
+A scheme of the method is shown in Fig.~\ref{fig:scheme}.
+
+\begin{figure}[th!]
+  \centering
+  \centerline{\includegraphics[width=0.75\textwidth]{Figures/SELD/ARCH.png}}
+  \caption{Architecture of the proposed methodology.}
+  \label{fig:scheme}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Single-source estimation}
+
+The first step is the transformation of the B-Format input signal $x_n^m(t)$ using the Short-Time Fourier Transform (STFT) into the time-frequency (TF) signal $X_n^m(k,n)$.
+
+In the resulting spectrogram, the frequencies above a given limit $f_{max}$ are discarded; this procedure speeds up the method while maintaining the directional information, given that the microphone geometry aliases spatial measurements above approx.
+5 kHz \cite{Bertet2006}.
+
+Assuming that the sources are sparse in time-frequency, it could be possible to identify TF bins which contain a significant energetic contribution from only one source.
+% i.e., without significant cross-talk from other sources or background noise. 
+These bins could be then used to produce accurate DOA estimates. The effectiveness of this approach has already been demonstrated \cite{tho2014robust, nguyen2020sequence}.
+
+Single-source TF bins are computed from the DirAC parametric analysis.
+A variety of alternative subspace methods are known \cite{epain2016spherical, madmoni2018direction}; however, those methods require local estimation of eigenvalues, which is a computationally expensive procedure.
+% This is the main reason for the choice of DirAC-based analysis in this work.
+
+A TF bin is counted as single-source if its diffuseness $\Psi(k,n)$ is lower than a threshold $\Psi_{max}$. Diffuseness is computed here using Eq.~\ref{eq:psidefinition}. 
+ Finally, the DOA $\Omega(k,n)$ of the TF bins passing the aforementioned single-source test is computed as the angle of the active intensity vector by Eq.~\ref{eq:doa}.
+ To illustrate the process, an example of the method output is plotted in Fig.~\ref{fig:plots} (top).
+% As an example, Fig.~\ref{fig:plots} (top) plots the estimated DOAs of single-source TF bins.
+
+\begin{figure}[th!]
+  \centering
+  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/plots.png}}
+  \caption{Estimation of localization and temporal activation.
+Top: azimuth spectrogram after diffuseness mask; color indicates estimated position of a TF bin passing the single-source test.
+Bottom: input/output of the particle tracking; the crosses represent the measurement space, and the continuous lines are the resulting events.}
+  \label{fig:plots}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Particle tracking}
+
+Once a set of reliable TF DOA estimates is obtained, the next step is the generalization of the individual measurements into trajectories and temporal activations.
+% Given that the instantaneous number of sources is unknown, it is necessary to use a multiple target tracker. 
+In our case, we opted for the Rao-Blackwellized Monte-Carlo Data Association (RBMCDA) algorithm \cite{sarkka2004rao}, which decomposes the multiple target tracking problem in two: it solves first the data association problem, and then performs the single target tracking individually.
+This method has been recently used in the context of sound event localization and tracking with successful results \cite{Adavanne2018_JSTSP, adavanne2019localization}; the code used for our implementation has been adapted from the same authors\footnote{\url{https://github.com/sharathadavanne/multiple-target-tracking}}.
+
+The system takes as the input the set of TF DOA values passing the single source test, and produces spatio-temporal event trajectories, considering an event as an entity with contiguous temporal activation and continuous spatial position.
+More specifically, for each time frame of the DOA masked spectrogram, the median\footnote{Circular median in the case of azimuth.} of all narrowband DOA estimates is computed.
+The resulting value is added to the measurement space of the tracker if the number of single-source frequency bins for that frame exceeds a minimum $K_{min}$.
+
+The performance of the RBMCDA algorithm is controlled by several parameters.
+Some of the most relevant include the angular velocity prior $v$, the standard deviation $\sigma_{\nu}$ and the spectral density $s_{\nu}$ of the measurement noise, the prior probabilities of birth $p_{birth}$ and noise percentage $p_{\nu}$, and the number of Monte-Carlo particles $N$.
+Position-related parameters are adjusted with respect to their ranges, so that azimuth-related magnitudes double elevation values.
+
+The procedure is followed by a numerical post-processing step, which includes data interpolation, resampling (if needed), and removal of elements shorter than $T_{min}$.
+Finally, the system provides a list of $J$ events, each one having an instantaneous position $\Omega_j(t)$ and a temporal activation $\Upsilon_j$. 
+An example of the system inputs and outputs is depicted in Fig.~\ref{fig:plots} (bottom).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Signal filter}
+
+The information provided by the particle tracking system is used to spatially filter the input signal.
+This can provide an enhanced monophonic estimate of an event $\tilde{s}_j(t)$ with reduced influence of simultaneous events.
+The process is performed by steering a virtual first-order cardioid in the direction of interest, using Eq.~\ref{eq:decoding}.
+%\begin{equation}
+%	\tilde{s}_j(t) = \sum_{m=0}^{M-1} x_m(t) Y_m(\Omega_j) \alpha_n
+%	\label{eq:decoding}
+%\end{equation}
+%where $\bm{Y}(\Omega_j) = [Y_0(\Omega_j), \dots, Y_3(\Omega_j)]^\intercal$ are the real-valued spherical harmonics up to first order evaluated in the direction $\Omega_j$ \cite{daniel2000representation}, and the column vector $\alpha_n$ controls the beam pattern directivity.
+ The result of this process is a monophonic estimate for each event, $\tilde{s}_j(t)$, temporally delimited by $\Upsilon_j$. 
+ As a last step, each estimate is amplitude peak-normalized, in order to minimize potential amplitude variability due to arbitrary configurations of the scene.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Event classification}
+As a final step, a class label is assigned to each estimated event $\tilde{s}_j(t)$ using a single-class classifier.
+Since the objective is to keep complexity low and make results interpretable, a machine learning algorithm is used instead of deep learning frameworks.
+The main advantages of this choice are: (i) low number of parameters; (ii) low train and predict computational time, easing reproducibility; and (iii) relative importance of the features in the output can be interpreted, which is not possible with deep learning approaches.
+
+Gradient Boosting Machine (GBM, Fig.~\ref{fig:gbm}) has been selected as the classification algorithm since it is a powerful yet simple technique for predictive modeling. 
+In essence, the algorithm is aimed to minimize the loss of the objective function by adding many weak learners.
+These learners are typically simple decision trees and their parameters are tuned using gradient descent techniques.
+GBM implementation makes use of the \textit{scikit-learn} library \cite{scikit-learn}.
+% Specifically, XGBoost framework is implemented for training process due to its proven performance in a wide range of classification problems \cite{xgboost}.
+
+\begin{figure}[t]
+  \centering
+  \centerline{\includegraphics[width=0.75\columnwidth]{Figures/SELD/gbm.png}}
+  \caption{Gradient boosting machine learning process. Adding weak estimators allows reducing overall error in the predictions.}
+  \label{fig:gbm}
+\end{figure}
+
+Sound features are obtained using extractors from \textit{Essentia}, an open-source library for audio analysis \cite{bogdanov2013essentia}. 
+Given the heterogeneous nature of the sound classes, a mixture of spectral, temporal and harmonic features are used, as shown in Table~\ref{tab:features}. 
+Features are be computed either frame-based or on the whole event; in the former case, the classifier is fed with their temporal first-order statistics. 
+% This is the situation for all \textit{Low-level}, and some \textit{SFX} features.
+% Conversely, event-level features are passed directly to the classifier.
+
+\begin{table}[t!]
+\caption{Acoustic features used for classification, grouped by type.}
+\begin{footnotesize}
+\begin{center}
+\begin{tabular}{ccc}
+\toprule
+Type & Features & Number \\
+\midrule
+\textit{Low-level} & Mel bands & 24       \\
+& MFCC & 13 \\
+& Spectral Features & 26 \\
+% & Pitch Salience &1 \\
+\midrule
+% SFX & Total and perceived&2 \\
+% & sound duration & \\
+% & Descriptors based on pitch & 4 \\
+% & and harmonics estimation & \\
+% & Sound envelope descriptors & 11\\
+% & Pitch envelope descriptors & 4\\
+\textit{SFX }& Duration &2 \\
+& Harmonic & 4 \\
+& Sound envelope & 11\\
+& Pitch envelope & 4\\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{footnotesize}
+\label{tab:features}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{EXPERIMENTS}
+\label{sec:experiments}
+
+\subsection{Dataset and baseline system}
+
+The dataset used is the FOA subset of the development set of the \textit{TAU-NIGENS Spatial Sound Events 2020} \cite{politis2020dataset}, which features 600 different B-Format clips of 60 seconds long each.
+Each clip contains multiple sound events, which belong to one of the fourteen sound classes from the NIGENS database \cite{trowitzsch2019nigens}.
+Events are also located at a potentially time-varying positions, and the maximum instantaneous overlapping of sources allowed is limited to two.
+Fifteen different Room Impulse Responses (RIR) are used for scene reverberation, covering a vast range of acoustic conditions.
+Furthermore, the audio clips contain a moderate amount of recorded background sounds.
+
+The baseline method is based on the recently proposed SELDnet architecture  \cite{Adavanne2018_JSTSP}, which features a Convolutional Recurrent Neural Network (CRNN) that solves both localization and classification problems jointly.
+Additionally, the baseline implementation has been improved with several changes inspired by one of the best performing methods in DCASE 2019 Task 3 Challenge \cite{Cao2019}.
+
+\subsection{Experimental setup}
+
+In order to explore the performance of the system, two different approaches have been undertaken regarding the creation of the training dataset for the monophonic single-class classifier. 
+The first approach, referred to as \textit{PAPAFIL1}, collects all event localization, temporal activation and class information by parsing the annotation files.
+Conversely, the second approach, called \textit{PAPAFIL2}, uses the proposed parametric particle filter to estimate localizations and activations, and the class label is assigned to each event by a custom association algorithm based on spatio-temporal distance.
+In both cases, the input signal is filtered with the obtained information in order to conform the monophonic event estimates.
+
+Therefore, the difference between training datasets is noticeable: while the training events in \textit{PAPAFIL1} are more accurately determined than in \textit{PAPAFIL2}, the differences with respect to the prediction scenario are much bigger in the former case.
+% Consequently, a slightly better performance of the second method might be expected, provided that the parametric particle filter performance has some degree of robustness and accuracy.
+The number of individual events for each of the approaches is plotted in Fig.~\ref{fig:num_events}.
+Approximately half of the classes have similar number of instances in both datasets.
+However, the other half presents noticeable differences, which might be explained by the different criteria applied for the consideration of event temporal activations: the groundtruth seems to follow a frame-based activity detection approach, while the output of the proposed method tends to consider events as time-continuous manifestations, influenced by the particle filter.
+
+This situation leads to two different \textit{oracle} systems (referred to by appending \textit{-O} in the method name), which represent the best performance theoretically achievable for the corresponding method.
+
+\begin{figure}[t]
+  \centering
+  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/numclasses2_tight.png}}
+  \caption{Number of occurrences of each event class in the training set, for both proposed methods. }
+  \label{fig:num_events}
+\end{figure}
+
+The accurate information of the \textit{PAPAFIL1} training set suggests a need for data augmentation; 
+in contrast, the training material used in \textit{PAPAFIL2} is already provided by a certain extent of variability.
+This situation motivates the implementation of data augmentation methods in the \textit{PAPAFIL1} training set.
+Specifically, several standard data augmentation techniques are implemented: pitch shifting, time shifting, time stretching and white noise addition.
+Furthermore, given the observed high influence of reverberation in the system performance, a reverberant data augmentation technique based on synthetic RIRs has been considered. 
+Ten different single-channel RIRs, with reverberation times between 0.3 and 1.1 seconds, have been synthetically created using the \textit{masp} library \cite{perez2020python}.
+During training, each event estimate is convolved with one of the RIRs, randomly chosen.
+RIR augmentation has recently been shown very effective for blind reverberation time estimation \cite{bryan2020impulse} but, to the best of the authors' knowledge, this is the first application in SELD.
+
+% The presented scheme leads to two different \textit{oracle} systems (named by an \textit{-O} appendix in the method name), which represent the best performance theoretically achievable for the corresponding method.
+
+% In this way, it is expected that \textit{PAPAFIL1-O} obtains very high localization scores overall, while the performance of \textit{PAPAFIL2-O} can be much similar to the non-oracle case.
+
+% It is important to notice that the spatial filtering is performed with a first-order cardioid, which provides a broad directive pattern.
+Accordingly, in the case of overlapping events, there will be always signal cross-talk, even when using the groundtruth annotations.
+The usage of higher ambisonic orders could easily mitigate this effect.
+
+Table~\ref{tab:parameter} shows a comprehensive list of the parameters used throughout the different steps of the proposed method.
+All values are equal for both presented approaches, except for the number of Monte-Carlo particles $N$.
+The values for Single-Source Estimation and Particle Filtering parameters have been iteratively refined by manual tuning and inspection, departing from standard values. 
+The beamforming weights $\alpha_m$ correspond to the \textit{maximum directivity beamformer}, which minimizes the energy contributions from directions other than the lookup direction \cite{rafaely2015fundamentals}.
+In the spatial audio field, such property is also known as the \textit{max-rE} decoder \cite{daniel2000representation}.
+Regarding event classification, a cross-validation scheme has been implemented for tuning GBM hyperparameters. 
+% In addition, unlike other methods such as CRNN, GBM allows to analyze the relative importance of each feature in the classifier.
+As it can be seen in Figure \ref{fig:importance}, the most representative variable is sfx.duration, being also important lowlevel.mfcc and spectral variables.
+
+% \begin{figure}[t]
+%   \centering
+%   \centerline{\includegraphics[width=1\columnwidth]{Figure_2.png}}
+%   \caption{Most representative features in event classifier }
+%   \label{fig:importance}
+% \end{figure}
+
+\begin{table}[th!]
+\begin{footnotesize}
+\caption{(Hyper-)parameter values.}
+    \begin{center}
+\begin{tabular}{cccc}
+\toprule
+Step & Parameter & Value & Unit \\
+\midrule
+Single-Source &sample rate     & 24    &  kHz\\
+% Estimation&window type     & hann  &  \\
+&window size      & 2400 &  samples\\
+&window overlap  & 50  &  \% \\
+&$f_{max}$   &  6     &  kHz\\
+&$N_\Psi$  & 2     &  frames\\
+&$\Psi_{max}$    & 0.1  &  \\
+
+\midrule
+Particle & $v$ & 2     &  $^\circ$/frame\\
+Filtering & $\sigma_{\nu}$   & 5     &  \\
+& $s_{\nu}$ & 20    &  \\
+& $p_{birth}$     & 0.25  &  \\
+& $p_{\nu}$    & 0.25  &  \\
+& $N$  & 100 / 30    &  \\
+&$K_{min}$ & 10     &  bins/frame\\ 
+&$T_{min}$  & 10     &  frames\\
+
+\midrule
+Signal & $\alpha_0$     &  0.775 & \\
+Filter & $\alpha_1$     &   3 * 0.4 & \\
+
+\midrule
+Event & number of estimators   & 1300       & trees  \\
+Classification & loss  &  \textit{mlogloss}    &\\
+& learning rate & 0.05& \\
+& max depth & 4 & \\
+& min samples leaf & 10 & samples \\
+\bottomrule
+\label{tab:parameter}
+\end{tabular}
+ \end{center}
+\end{footnotesize}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation metrics}
+
+The system is evaluated according to the joint metrics proposed in the Challenge \cite{Mesaros_2019_WASPAA}. 
+The metrics evaluate jointly the localization and the classification, and are divided into two types: location-aware classification, and classification-aware localization. 
+There are two classification metrics: Error Rate ($\text{ER}_{20}$) and F-Score ($\text{F}_{20}$).
+As the name suggests, the metrics are conditioned to a minimum localization performance, which is set to 20$^{\circ}$ in this case.
+Localization metrics are also two-fold: Localization Error ($\text{LE}_{CD}$) and Localization Recall ($\text{LR}_{CD}$); 
+as their name suggests, the metrics are class-dependent, and thus are conditioned to a correct classification.
+Finally, the $\text{SELD}$ score is an average of the four other metrics, used to conveniently sum up the results.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{RESULTS}
+\label{sec:results}
+
+Table~\ref{tab:results} summarizes the results of the experiments using the recommended data split: training with folds 3 to 6, validation with fold 2 and testing with fold 1. 
+% This structure has been promoted by the Challenge organization as a fair way for method comparison. 
+% Furthermore, the usage of an unseen split 
+% Furthermore, it is expected to provide similar results to the evaluation set, given that part of the data remains unseen at training.
+Results are reported for three different systems: the baseline and the two proposed methods \textit{PAPAFIL1} and \textit{PAPAFIL2}.
+The results of their respective oracle results, \textit{PAPAFIL1-O} and \textit{PAPAFIL2-O}, are also provided.
+
+Both proposed approaches outperform the baseline system in three out of the four evaluation metrics ($\text{ER}_{20}$, $\text{F}_{20}$ and $\text{LE}_{CD}$). 
+Although the results obtained by both of them are similar, \textit{PAPAFIL2} obtains better classification scores ($\text{ER}_{20}$ and $\text{F}_{20}$), and \textit{PAPAFIL1} performs subtly better regarding localization error ($\text{LE}_{CD}$).
+However, the localization recall results ($\text{LR}_{CD}$) are slightly worst than the baseline in both cases.
+This fact does not prevent the proposed methods to have a $\text{SELD}$ score better than the baseline: 0.41 (\textit{PAPAFIL1}) and 0.38 (\textit{PAPAFIL2}), against 0.47 (\textit{BASELINE}).
+
+The results obtained by the oracle methods are within the expected ranges.
+\textit{PAPAFIL1-O} performs almost perfectly regarding $\text{LE}_{CD}$, but the classification errors influence the $\text{LR}_{CD}$ result.
+In turn, \textit{PAPAFIL2-O} performs better than \textit{PAPAFIL1-O} regarding all metrics, excepting $\text{LE}_{CD}$; this improvement is specially noticeable in $\text{LR}_{CD}$, with a performance difference of about 15\%.
+The good results obtained by \textit{PAPAFIL2-O} validate the proposed particle filtering approach, an leave space for improvements that might be given by a better understanding and fine tuning of the model.
+
+The performance of the proposed methods deteriorates noticeably with overlapping sounds.
+% Although the performance is good in general terms with both \textit{PAPAFIL} approaches, the results deteriorate noticeably with overlapping sounds.
+A closer inspection reveals that, in many occasions, the TF bins passing the single-source test mostly belong to one out of two simultaneous sources. 
+It is a known issue that performance of DirAC diffuseness is reduced when two sources are present \cite{epain2016spherical}; similar problems have been reported in \cite{adavanne2019localization}, where an instantaneous source number estimator is used in combination with the particle filter.
+As in that case, the results suggest the need for more sophisticated source detection and counting methods.
+
+\begin{table}[th!]
+\begin{footnotesize}
+\caption{Evaluation results on the development set.}
+    \begin{center}
+    \begin{tabular}{cccccc}
+    \toprule
+    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
+    \midrule
+    \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & \textbf{60.7} \% & 0.47 \\
+    \textit{PAPAFIL1} & 0.60 & 49.8 \% & \textbf{13.4}$^{\circ}$ & 54.4 \% &0.41\\
+
+    \textit{PAPAFIL2} & \textbf{0.57} & \textbf{54.0} \% &13.8$^{\circ}$ & 59.7 \% & \textbf{ 0.38} \\
+
+    \midrule
+    % this is oracle beam
+    \textit{PAPAFIL1-O} & 0.37 & 67.0 \% & 2.0$^{\circ}$ & 68.6 \% & 0.26 \\
+    \textit{PAPAFIL2-O} & 0.32 & 79.6 \% & 8.5$^{\circ}$ & 82.4\% & 0.19 \\
+
+    % \midrule
+    % \textit{PAPAFIL1} &  0.57 & 55.6 \% & 15.6$^{\circ}$ & 66.7 \% & 0.36 \\
+    % % PAPAFIL_1 & 0.47  & 64.1 \% & 14.1$^{\circ}$ & 70.1 \% & 0.30 \\ ?????????
+    % % this is oracle beam
+    % \textit{PAPAFIL1-O} & 0.08 & 93.7 \% &  0.2$^{\circ}$ & 94.0 \% & 0.05 \\
+
+    %  \textit{PAPAFIL2} & \textbf{ 0.44} & \textbf{68.0} \% & \textbf{13.3}$^{\circ}$ & \textbf{79.6} \% & \textbf{0.26} \\
+
+    % \textit{PAPAFIL2-O} & 0.41 & 71.1 \% & 12.3$^{\circ}$ & 82.0\% &  0.24 \\
+    \bottomrule
+    \end{tabular}
+    \end{center}
+    \label{tab:results}
+\end{footnotesize}
+\end{table}
+
+\begin{figure}[t]
+  \centering
+  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/importance_new_tight.png}}
+  \caption{Most representative features in event classifier.}
+  \label{fig:importance}
+\end{figure}
+
+To conclude the analysis, Fig.~\ref{fig:importance} shows the relative importance of the fifteen most relevant acoustic features for the \textit{PAPAFIL2} classifier model.
+Event duration is clearly the feature with the highest importance, and effective duration (duration of the signal discarding silence) also appears in the eighth position.
+This fact can help to explain the better performance of \textit{PAPAFIL2} over \textit{PAPAFIL1}: the temporal activities of the events in training and prediction are much more similar to each other in the former method, as a consequence of the training set generation approach.
+Furthermore, it is interesting to notice the high relevance of low-level features, and specifically several MFCC combinations (eight of the fifteen reported features) and  various extractors related to the spectral structure.
+The absence of pitch, harmonic and envelope features in the list represents a significant finding as well.
+
+\todo{TODO: DISCUSS RESULTS OF EVALUATION SET!!}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+\label{sec:results}
+
+% We have presented a novel method for Sound Event Localization and Detection (SELD), based on parametric particle filtering and gradient boosting single-class event classification of audio features.
+Results show that the proposed method outperforms the baseline method, a state-of-the-art Convolutional Recurrent Neural Network (CRNN).
+Specifically, the proposed method is able to improve the baseline SELD score by almost ten points, while also increasing the evaluation score in three out of the four proposed metrics, by means of a low complexity machine learning architecture.
+
+We present a novel low-complexity method for Sound Event Localization and Detection of First Order Ambisonic signals, based on four steps: estimation of single-source spectrogram regions by parametric analysis; computation of event trajectories and activations by means of a particle tracker; spatio-temporal filtering of the input signal; and single-class monophonic event classification by Gradient Boosting.
+Results show that the proposed method outperforms the baseline method, a state-of-the-art Convolutional Recurrent Neural Network.
+Specifically, our method is able to improve the baseline SELD score by almost ten points, while increasing the scores in three out of the four metrics under consideration.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section{Introduction}
+%\label{sec:intro}
+%
+%Sound Event Localization and Detection (SELD) refers to the problem of identifying, for each individual  event present in a sound field, the temporal activity, spatial location, and sound class to which it belongs.
 SELD is a current research topic which deals with microphone array processing and sound classification, with potential applications in the fields of signal enhancement, autonomous navigation, acoustic scene description or surveillance, among others.
-
-SELD arises from the combination of two different problems: Sound Event Detection (SED) and Direction of Arrival (DOA) estimation.
+%
+%SELD arises from the combination of two different problems: Sound Event Detection (SED) and Direction of Arrival (DOA) estimation.
 The number of works in the literature which jointly address SED and DOA problems is relatively small.
 It is possible to classify them by the type of microphone arrays used: distributed \cite{grobler2017sound, butko2011two, chakraborty2014sound} or near-coincident \cite{hirvonen2015classification, lopatka2016detection, Adavanne2018_JSTSP}.
-As mentioned in \cite{Adavanne2018_JSTSP}, the usage of near-coincident circular/spherical arrays enables the representation of the sound field in the spatial domain, using the spherical harmonic decomposition, also known as Ambisonics \cite{gerzon1973periphony, daniel2000representation}.
+%As mentioned in \cite{Adavanne2018_JSTSP}, the usage of near-coincident circular/spherical arrays enables the representation of the sound field in the spatial domain, using the spherical harmonic decomposition, also known as Ambisonics \cite{gerzon1973periphony, daniel2000representation}.
 Such spatial representation allows a flexible, device-independent comparison between methods.
 Furthermore, the number of commercially available ambisonic microphones has increased in recent years due to their suitability for immersive multimedia applications.
-Taking advantage of the compact spatial representation provided by the spherical harmonic decomposition, several methods for parametric analysis of the sound field in the ambisonic domain have been proposed  \cite{pulkki2006directional, berge2010high, Politis2018, pulkki2018parametric}.
-These methods ease sound field segmentation into direct and diffuse components, and further localization of the direct sounds.
-The advent of deep learning techniques for DOA estimation has also improved the results of traditional methods \cite{Adavanne2018_JSTSP}.
+%Taking advantage of the compact spatial representation provided by the spherical harmonic decomposition, several methods for parametric analysis of the sound field in the ambisonic domain have been proposed  \cite{pulkki2006directional, berge2010high, Politis2018, pulkki2018parametric}.
+%These methods ease sound field segmentation into direct and diffuse components, and further localization of the direct sounds.
+%The advent of deep learning techniques for DOA estimation has also improved the results of traditional methods \cite{Adavanne2018_JSTSP}.
 However, none of the deep learning-based DOA estimation methods explicitly exploits the spatial parametric analysis.
 This situation is further extended to the SELD problem, with the exception of \cite{lopatka2016detection}, where DOAs are estimated from the \textit{active intensity vector}  \cite{pulkki2006directional}.
-
-The motivation for the proposed methodology is two-fold.
-First, we would like to check whether the usage of spatial parametric analysis in the ambisonic domain can improve the performance of SELD algorithms.
-Second, temporal information derived by the parametric analysis could be further exploited to estimate event onsets and offsets, thus lightening the event classifier complexity; such reduction might positively impact algorithm's performance.
-
-In what follows, we present the methodology and the architecture of the proposed system (Section~\ref{sec:method}).
+%
+%The motivation for the proposed methodology is two-fold.
+%First, we would like to check whether the usage of spatial parametric analysis in the ambisonic domain can improve the performance of SELD algorithms.
+%Second, temporal information derived by the parametric analysis could be further exploited to estimate event onsets and offsets, thus lightening the event classifier complexity; such reduction might positively impact algorithm's performance.
+%
+%In what follows, we present the methodology and the architecture of the proposed system (Section~\ref{sec:method}).
 Then, we describe the design choices and the experimental setup (Section~\ref{sec:expe}), and discuss the results in the context of DCASE2019 Challenge - Task 3 (Section~\ref{sec:results}).
 A summary is presented in Section~\ref{sec:conclusion}.
-In order to support open access and reproducibility, all code is freely available at \cite{code}.
-\todo{code ref}.
-
-\section{Method}
-\label{sec:method}
-
-The proposed method presents a solution for the SELD problem splitting the task into four different problems: \textit{DOA  estimation}, \textit{association}, \textit{beamforming} and \textit{classification}, which will be described in the following subsections.
-The former three systems follow a heuristic approach---in what follows, they will be jointly referred to as the \textit{parametric front-end}.
+%%In order to support open access and reproducibility, all code is freely available at \cite{code}. \todo{code ref}.
+%
+%
+%\section{Method}
+%\label{sec:method}
+%
+%The proposed method presents a solution for the SELD problem splitting the task into four different problems: \textit{DOA  estimation}, \textit{association}, \textit{beamforming} and \textit{classification}, which will be described in the following subsections.
+%The former three systems follow a heuristic approach---in what follows, they will be jointly referred to as the \textit{parametric front-end}.
 Conversely, the \textit{classification} system is data-driven, and will be referred to as the \textit{deep learning back-end}.
 The method architecture is depicted in Figure~\ref{fig:diagram_all}.
-
-\todo{redo figures in this section}
-
-\begin{figure}[h]
-	\centering
-    \includegraphics[width=\columnwidth]{Figures//SELD/dcase_challenge_tech_report_all2.pdf}
-    \caption{System architecture.}
-    \label{fig:diagram_all}
-
-\end{figure}
-
-\subsection{DOA estimation}
-\label{ssec:doa_estimation}
-
-\begin{figure*}[h]
-
-    \includegraphics[width=\textwidth]{Figures/SELD/DOA1.pdf}
-    \caption{DOA estimation architecture.}
-    \label{fig:doa}
-
-\end{figure*}
-
-The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency spatial audio analysis.
-
-Let us consider a \textit{N3D}-normalized first-order ambisonic signal $x_n^m(t)$. 
-Using the DirAC parametric model, it is possible to obtain instantaneous time-frequency estimates of the Direction of Arrival $\Omega(k,n)$ (Eq.~\ref{eq:doa}), energy density $E(k,n)$ (Eq.~\ref{eq:energydensity}) and the diffuseness $\Psi(k,n)$ (Eq.~\ref{eq:psidefinition}).\\
-
-%Let us consider a first order ambisonic signal $$
-%Let us consider a first order ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
-%\begin{equation}
-%    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
-%\end{equation}
-%From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
-%\begin{equation}
-%    \begin{aligned}
-%    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
-%    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
-%    \end{aligned}
-%\end{equation}
-%where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.
-
-It is desirable to identify the TF regions of $\Omega(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.
-
-The first mask is the \textit{energy density mask}, which is used as an activity detector.
+%
+%\todo{redo figures in this section}
+%
+%\begin{figure}[h]
+%	\centering
+%    \includegraphics[width=\columnwidth]{Figures//SELD/dcase_challenge_tech_report_all2.pdf}
+%    \caption{System architecture.}
+%    \label{fig:diagram_all}
+%
+%\end{figure}
+%
+%\subsection{DOA estimation}
+%\label{ssec:doa_estimation}
+%
+%\begin{figure*}[h]
+%
+%    \includegraphics[width=\textwidth]{Figures/SELD/DOA1.pdf}
+%    \caption{DOA estimation architecture.}
+%    \label{fig:doa}
+%
+%\end{figure*}
+%
+%
+%
+%The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency spatial audio analysis.
+%
+%Let us consider a \textit{N3D}-normalized first-order ambisonic signal $x_n^m(t)$. 
+%Using the DirAC parametric model, it is possible to obtain instantaneous time-frequency estimates of the Direction of Arrival $\Omega(k,n)$ (Eq.~\ref{eq:doa}), energy density $E(k,n)$ (Eq.~\ref{eq:energydensity}) and the diffuseness $\Psi(k,n)$ (Eq.~\ref{eq:psidefinition}).\\
+%
+%
+%
+%%Let us consider a first order ambisonic signal $$
+%%Let us consider a first order ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
+%%\begin{equation}
+%%    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
+%%\end{equation}
+%%From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
+%%\begin{equation}
+%%    \begin{aligned}
+%%    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
+%%    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
+%%    \end{aligned}
+%%\end{equation}
+%%where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.
+%
+%It is desirable to identify the TF regions of $\Omega(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.
+%
+%The first mask is the \textit{energy density mask}, which is used as an activity detector.
 A gaussian adaptive thresholding algorithm is applied to $E(k,n)$.
 This procedure selects TF bins with local maximum energy density, as expected from the direct path of the sources.
-
-The \textit{diffuseness mask} selects the TF bins with high energy propagation.
+%
+%The \textit{diffuseness mask} selects the TF bins with high energy propagation.
 Bins with a low diffuseness value represent a TF region where one sound source is predominant.
 A fixed threshold value $\Psi_{max}$ is used for the masking.
-
-The third mask is the \textit{DOA variance mask}.
+%
+%The third mask is the \textit{DOA variance mask}.
 It tries to select TF regions with small standard deviation with respect to their neighbor bins---a characteristic of sound fields with low diffuseness \cite{pulkki2018parametric}.\\
-
-The three masks are then applied to the DOA estimation, obtaining the TF-filtered instanteneous DOAs $\bar{\Omega}(k,n)$. 
-Finally, a median filter is applied, with the aim of improving DOA estimation consistency and removing spurious TF bins.
-The median filter is applied in a TF bin belonging to $\bar{\Omega}(k,n)$ only if the number of TF bins belonging to $\bar{\\Omega}(k,n)$ in its vicinity is greater than a given threshold $B_{min}$.
-The resulting filtered DOA estimation is referred to as $\breve{\Omega}(k,n)$.
-
-\subsection{Association}
-\label{ssec:association}
-
-\begin{figure*}[h]
-    \includegraphics[width=\textwidth]{Figures/SELD//ASS.pdf}
-    \caption{Association architecture.}
-    \label{fig:association}
-
-\end{figure*}
-
-The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\Omega}(k,n)$ to a set of events, each one having a specific onset, offset and location.
+%
+%The three masks are then applied to the DOA estimation, obtaining the TF-filtered instanteneous DOAs $\bar{\Omega}(k,n)$. 
+%Finally, a median filter is applied, with the aim of improving DOA estimation consistency and removing spurious TF bins.
+%The median filter is applied in a TF bin belonging to $\bar{\Omega}(k,n)$ only if the number of TF bins belonging to $\bar{\\Omega}(k,n)$ in its vicinity is greater than a given threshold $B_{min}$.
+%The resulting filtered DOA estimation is referred to as $\breve{\Omega}(k,n)$.
+%
+%
+%\subsection{Association}
+%\label{ssec:association}
+%
+%\begin{figure*}[h]
+%    \includegraphics[width=\textwidth]{Figures/SELD//ASS.pdf}
+%    \caption{Association architecture.}
+%    \label{fig:association}
+%
+%\end{figure*}
+%
+%The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\Omega}(k,n)$ to a set of events, each one having a specific onset, offset and location.
 \\
-
-First, DOA estimates are resampled into \textit{frames} with the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
-An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\Omega}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is higher than a given threshold $K_{min}$.\\
-
-Next, the standard deviation in azimuth $\sigma_{\varphi}$ and elevation $\sigma_{\theta}$ of the frame-based DOA estimates $\breve{\Omega}(k,m)$ are compared to a threshold value $\sigma_{max}$, and the result is used to estimate the frame-based event overlapping amount $o(m)$ :
-
-\begin{equation}
-\begin{aligned}
-	o(m) =  \begin{cases}
-    		1, &\text{if } \sigma_{\varphi}/2 + \sigma_{\theta} < \sigma_{max},\\
-    		2, &\text{otherwise}.
-    \end{cases}    		
-\end{aligned}
-\end{equation}
-
-The clustered values $\Omega_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\Omega}(k,m)$, using a modified version of K-Means which minimizes the central angle distance.
+%
+%First, DOA estimates are resampled into \textit{frames} with the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
+%An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\Omega}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is higher than a given threshold $K_{min}$.\\
+%
+%Next, the standard deviation in azimuth $\sigma_{\varphi}$ and elevation $\sigma_{\theta}$ of the frame-based DOA estimates $\breve{\Omega}(k,m)$ are compared to a threshold value $\sigma_{max}$, and the result is used to estimate the frame-based event overlapping amount $o(m)$ :
+%
+%\begin{equation}
+%\begin{aligned}
+%	o(m) =  \begin{cases}
+%    		1, &\text{if } \sigma_{\varphi}/2 + \sigma_{\theta} < \sigma_{max},\\
+%    		2, &\text{otherwise}.
+%    \end{cases}    		
+%\end{aligned}
+%\end{equation}
+%
+%The clustered values $\Omega_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\Omega}(k,m)$, using a modified version of K-Means which minimizes the central angle distance.
 Notice that, for $o(m)=1$, the operation is equivalent to the median.\\
-
-The following step is the grouping of clustered DOA values into events.
-Let us define $\Omega_S(m)$ as the frame-wise DOA estimations belonging to the event $S$.
+%
+%The following step is the grouping of clustered DOA values into events.
+%Let us define $\Omega_S(m)$ as the frame-wise DOA estimations belonging to the event $S$.
 A given clustered DOA estimation $\Omega_{\text{cluster}}(m)$ belongs to the event $S$ if the following criteria are met:
-\begin{itemize}
-    \item The central angle between  $\Omega_{\text{cluster}}(m)$ and the median of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
-    \item The frame distance between $m$ and the closest frame of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
-\end{itemize}
-
-The resulting DOAs $\Omega_S(m)$ are subject to a postprocessing step with the purpose of 
-delaying event onsets in frames where $o(m) > 2$, and discarding events shorter than a given minimum length.
-
-Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\Lambda_S = (\Omega_S, \text{onset}_S, \text{offset}_S)$.
-
-\subsection{Beamforming}
-\label{ssec:Beamforming}
-
-The last step performed in the front-end is the input signal segmentation.
+%\begin{itemize}
+%    \item The central angle between  $\Omega_{\text{cluster}}(m)$ and the median of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
+%    \item The frame distance between $m$ and the closest frame of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
+%\end{itemize}
+%
+%The resulting DOAs $\Omega_S(m)$ are subject to a postprocessing step with the purpose of 
+%delaying event onsets in frames where $o(m) > 2$, and discarding events shorter than a given minimum length. 
+%
+%Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\Lambda_S = (\Omega_S, \text{onset}_S, \text{offset}_S)$.
+%
+%\subsection{Beamforming}
+%\label{ssec:Beamforming}
+%
+%The last step performed in the front-end is the input signal segmentation.
 The spatial and temporal information provided by the annotations $\Lambda_S$ are used to produce monophonic signal estimations of the events, $\tilde{x}_S(t)$, as the signals captured by a virtual first-order cardioid (Eq.~\ref{eq:decodingequation}):
-\begin{equation}
-\tilde{x}_S(t) = x_n^m(t) Y_n^m(\Omega_S) \alpha_n,
-\end{equation}
-with $\alpha_n = [1, 1, 1, 1]$ using the \textit{basic} decoding weights.
-
-\subsection{Deep learning classification back-end}
-\label{ssec:backend_method}
-
-%intro - problem formulation==============
-The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{s}_S(t)$.
-Then, the back-end classifies the resulting signals as belonging to one of a target set of 11 classes.
-Therefore, the multi-task nature of the front-end allows us to define the back-end classification task as a simple multi-class problem, even though the original SELD task is multi-label.
-It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\Lambda_S$ are perfectly estimated.
-
-%2 stages: context
-The classification method is divided into two stages.
-First, the incoming signal is transformed into the log-mel spectrogram and split into TF patches.
-Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $c \in \left\lbrace 1 ...C \right\rbrace$, with $C=11$.
-Predictions are done at the event-level (not at the frame level), since the temporal activities have been already determined by the front-end.
-Only one label is predicted for each incoming event, such that there is no binarization stage needed as in a standard SED task.
-
-%archi=============
-The proposed CRNN is depicted in Figure \ref{fig:archi}.
-It presents three convolutional \textit{blocks} to extract local features from the input representation.
-Each convolutional block consists of one convolutional layer, after which the resulting feature maps are passed through a ReLU non-linearity \cite{nair2010rectified}.
-This is followed by a max-pooling operation to downsample the feature maps and add invariance along the frequency dimension.
-The target classes vary to a large extent in terms of their temporal dynamics, with some of them being rather impulsive (e.g., \textit{Door slam}), while others being more stationary (e.g., \textit{Phone ringing}).
-Therefore, after stacking the feature maps resulting from the convolutional blocks, this representation is fed into one bidirectional recurrent layer in order to model discriminative temporal structures.
-Specifically, 64 nodes of gated recurrent units (GRU) are used with \textit{tanh} activations.
-The recurrent layer is followed by a Fully Connected (FC) layer, and finally a 11-way softmax classifier layer produces the event-level probabilities.
-Dropout is applied extensively.
-The loss function used is categorical cross-entropy.
-The model has $\sim$175k weights.
-
-\begin{figure}[ht]
-  \vspace{-3mm}
-  \centering
-  \centerline{\includegraphics[width=0.9\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
-  \vspace{-2mm}
-  \caption{Back-end architecture.}
-  \label{fig:archi}
-    \vspace{-5mm}
-\end{figure}
-
-%===============================================================================
-
-\section{Experiments}
-\label{sec:expe}
-
-\subsection{Dataset, evaluation metrics and baseline system}
-\label{ssec:dataset}
-We use the TAU Spatial Sound Events 2019 - Ambisonic, which provides first-order ambisonic recordings.
-Details about the recording format and dataset specifications can be found in \cite{Adavanne2019_DCASE}.
-The dataset features a vocabulary of $C=11$ classes encompassing human sounds and sound events typically found in indoor office environments.
-The dataset is split into a development and evaluation sets. 
-The development set consists of a four fold cross-validation setup.\\
-
-The SELD task is evaluated with individual metrics for the SED and DOA problems:
-\begin{itemize}
-	\item SED: F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments.
-	\item DOA: DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise.
-\end{itemize}
-The \textit{SELD score} is an averaged summary of the system performance.
-A more detailed description of the evaluation metrics can be found in \cite{Adavanne2018_JSTSP}.\\
-
-The baseline system features a CRNN that jointly performs DOA and SED through multi-task learning \cite{Adavanne2018_JSTSP}.
-Baseline results are shown in Table \ref{tab:results_real}.
-
-\subsection{Parametric front-end}
-\label{ssec:param_frontend}
-
-Based on the method's exploratory analysis, we propose the following set of parameter values, which are shown in Table~\ref{tab:params_front}.
-
-\begin{table}[!htbp]
-
-\centering
-\caption{Parameter values for the selected configuration. Top: \textit{DOA analysis} parameters. Bottom: \textit{Association} parameters.}
-\begin{tabular}{cccc}
-\toprule
-Parameter & Unit & Value\\
-\midrule
-sampling rate & Hz  & 48000\\
-STFT window size & sample  & 256\\
-STFT window overlap & sample & 128\\
-STFT window type & -  &  Hann\\
-analysis frequency range & Hz & [0,8000]\\
-time average vicinity radius $r$ & bin& 10 \\
-diffuseness mask threshold $\Psi_{max}$ & - & 0.5 \\
-energy density filter length & bin & 11\\
-std mask vicinity radius & bin &  2\\
-std mask normalized threshold & - & 0.15\\
-median filter minimum ratio $B_{min}$ & - & 0.5 \\
-median filter vicinity radius (k,n) & bin & (20, 20) \\
-\midrule
-frame size $h$ & s  & 0.02\\
-resampling minimum valid bins $K_{min}$ & bin & 1 \\
-overlapping std threshold $\sigma_{max}$ & degree & 10 \\
-grouping maximum angle $d_{max}^\text{ANGLE}$ & degree & 20\\
-grouping maximum distance $d_{max}^\text{FRAME}$ & frame & 20\\
-event minimum length & frame & 8\\
-\bottomrule
-\end{tabular}
-\label{tab:params_front}
-\end{table}
-
-\subsection{Deep learning classification back-end}
-\label{ssec:backend_experiments}
-
-We use the provided four fold cross-validation setup.
-Training and validation stages use the outcome of an \textit{ideal} front-end, where the groundtruth DOA estimation and activation times are used to feed the beamformer for time-space segmentation.
-Conversely, we test the trained models with the signals coming from the \textit{complete} front-end described in Section \ref{sec:method}.
-
-We conducted a set of preliminary experiments with different types of networks including a VGG-like net, a less deep CNN \cite{Fonseca2019learning}, a Mobilenetv1 \cite{howard2017mobilenets} and a CRNN \cite{cakir2017convolutional}.
-The latter was found to stand out, and we explore certain facets of the CRNN architecture and the learning pipeline.\\
-
-Sound events in the dataset last from $\sim$ 0.2 to 3.3 s.
-First, clips shorter than 2s are replicated to meet this length.
-Then, we compute TF patches of log-mel spectrograms of $T=50$ frames (1 s) and $F=64$ bands.
-The values come from the exploration of $T \in \left\lbrace 25,50,75,100\right\rbrace$ and $F \in \left\lbrace 40,64,96,128\right\rbrace$.
-$T=50$ is the top performing value, roughly coinciding with the median event duration.
+%\begin{equation}
+%\tilde{x}_S(t) = x_n^m(t) Y_n^m(\Omega_S) \alpha_n,
+%\end{equation}
+%with $\alpha_n = [1, 1, 1, 1]$ using the \textit{basic} decoding weights.
+%
+%
+%
+%\subsection{Deep learning classification back-end}
+%\label{ssec:backend_method}
+%
+%%intro - problem formulation==============
+%The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{s}_S(t)$.
+%Then, the back-end classifies the resulting signals as belonging to one of a target set of 11 classes.
+%Therefore, the multi-task nature of the front-end allows us to define the back-end classification task as a simple multi-class problem, even though the original SELD task is multi-label.
+%It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\Lambda_S$ are perfectly estimated.
+%
+%%2 stages: context
+%The classification method is divided into two stages.
+%First, the incoming signal is transformed into the log-mel spectrogram and split into TF patches.
+%Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $c \in \left\lbrace 1 ...C \right\rbrace$, with $C=11$.
+%Predictions are done at the event-level (not at the frame level), since the temporal activities have been already determined by the front-end.
+%Only one label is predicted for each incoming event, such that there is no binarization stage needed as in a standard SED task.
+%
+%%archi=============
+%The proposed CRNN is depicted in Figure \ref{fig:archi}.
+%It presents three convolutional \textit{blocks} to extract local features from the input representation.
+%Each convolutional block consists of one convolutional layer, after which the resulting feature maps are passed through a ReLU non-linearity \cite{nair2010rectified}.
+%This is followed by a max-pooling operation to downsample the feature maps and add invariance along the frequency dimension.
+%The target classes vary to a large extent in terms of their temporal dynamics, with some of them being rather impulsive (e.g., \textit{Door slam}), while others being more stationary (e.g., \textit{Phone ringing}).
+%Therefore, after stacking the feature maps resulting from the convolutional blocks, this representation is fed into one bidirectional recurrent layer in order to model discriminative temporal structures.
+%Specifically, 64 nodes of gated recurrent units (GRU) are used with \textit{tanh} activations.
+%The recurrent layer is followed by a Fully Connected (FC) layer, and finally a 11-way softmax classifier layer produces the event-level probabilities.
+%Dropout is applied extensively.
+%The loss function used is categorical cross-entropy.
+%The model has $\sim$175k weights.
+%
+%\begin{figure}[ht]
+%  \vspace{-3mm}
+%  \centering
+%  \centerline{\includegraphics[width=0.9\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
+%  \vspace{-2mm}
+%  \caption{Back-end architecture.}
+%  \label{fig:archi}
+%    \vspace{-5mm}
+%\end{figure}
+%
+%
+%
+%
+%%===============================================================================
+%
+%\section{Experiments}
+%\label{sec:expe}
+%
+%\subsection{Dataset, evaluation metrics and baseline system}
+%\label{ssec:dataset}
+%We use the TAU Spatial Sound Events 2019 - Ambisonic, which provides first-order ambisonic recordings.
+%Details about the recording format and dataset specifications can be found in \cite{Adavanne2019_DCASE}.
+%The dataset features a vocabulary of $C=11$ classes encompassing human sounds and sound events typically found in indoor office environments.
+%The dataset is split into a development and evaluation sets. 
+%The development set consists of a four fold cross-validation setup.\\
+%
+%The SELD task is evaluated with individual metrics for the SED and DOA problems:
+%\begin{itemize}
+%	\item SED: F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments.
+%	\item DOA: DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise.
+%\end{itemize}
+%The \textit{SELD score} is an averaged summary of the system performance.
+%A more detailed description of the evaluation metrics can be found in \cite{Adavanne2018_JSTSP}.\\
+%
+%The baseline system features a CRNN that jointly performs DOA and SED through multi-task learning \cite{Adavanne2018_JSTSP}.
+%Baseline results are shown in Table \ref{tab:results_real}.
+%
+%
+%\subsection{Parametric front-end}
+%\label{ssec:param_frontend}
+%
+%Based on the method's exploratory analysis, we propose the following set of parameter values, which are shown in Table~\ref{tab:params_front}.
+%
+%
+%\begin{table}[!htbp]
+%
+%\centering
+%\caption{Parameter values for the selected configuration. Top: \textit{DOA analysis} parameters. Bottom: \textit{Association} parameters.}
+%\begin{tabular}{cccc}
+%\toprule
+%Parameter & Unit & Value\\
+%\midrule
+%sampling rate & Hz  & 48000\\
+%STFT window size & sample  & 256\\
+%STFT window overlap & sample & 128\\
+%STFT window type & -  &  Hann\\
+%analysis frequency range & Hz & [0,8000]\\
+%time average vicinity radius $r$ & bin& 10 \\
+%diffuseness mask threshold $\Psi_{max}$ & - & 0.5 \\
+%energy density filter length & bin & 11\\
+%std mask vicinity radius & bin &  2\\
+%std mask normalized threshold & - & 0.15\\
+%median filter minimum ratio $B_{min}$ & - & 0.5 \\
+%median filter vicinity radius (k,n) & bin & (20, 20) \\
+%\midrule
+%frame size $h$ & s  & 0.02\\
+%resampling minimum valid bins $K_{min}$ & bin & 1 \\
+%overlapping std threshold $\sigma_{max}$ & degree & 10 \\
+%grouping maximum angle $d_{max}^\text{ANGLE}$ & degree & 20\\
+%grouping maximum distance $d_{max}^\text{FRAME}$ & frame & 20\\
+%event minimum length & frame & 8\\
+%\bottomrule
+%\end{tabular}
+%\label{tab:params_front}
+%\end{table}
+%
+%
+%\subsection{Deep learning classification back-end}
+%\label{ssec:backend_experiments}
+%
+%We use the provided four fold cross-validation setup.
+%Training and validation stages use the outcome of an \textit{ideal} front-end, where the groundtruth DOA estimation and activation times are used to feed the beamformer for time-space segmentation.
+%Conversely, we test the trained models with the signals coming from the \textit{complete} front-end described in Section \ref{sec:method}.
+%
+%We conducted a set of preliminary experiments with different types of networks including a VGG-like net, a less deep CNN \cite{Fonseca2019learning}, a Mobilenetv1 \cite{howard2017mobilenets} and a CRNN \cite{cakir2017convolutional}.
+%The latter was found to stand out, and we explore certain facets of the CRNN architecture and the learning pipeline.\\
+%
+%Sound events in the dataset last from $\sim$ 0.2 to 3.3 s.
+%First, clips shorter than 2s are replicated to meet this length.
+%Then, we compute TF patches of log-mel spectrograms of $T=50$ frames (1 s) and $F=64$ bands.
+%The values come from the exploration of $T \in \left\lbrace 25,50,75,100\right\rbrace$ and $F \in \left\lbrace 40,64,96,128\right\rbrace$.
+%$T=50$ is the top performing value, roughly coinciding with the median event duration.
 In turn, more than 64 bands provide inconsistent improvements, at the cost of increasing the number of network weights.\\
-
-Regarding the network structure, several variants of the CRNN architecture were explored until reaching the network of Figure \ref{fig:archi}.
-This included a small grid search over number of CNN filters, CNN filter size and shape, number of GRU units, number of FC units, dropout \cite{srivastava2014dropout}, learning rate, and the usage of Batch Normalization (BN) \cite{ioffe2015batch}.
-Network extensions (involving more weights) were considered only if providing major improvements, as a measure against overfitting.
-The main takeaways are: \textit{i)} squared 3x3 filters provide better results than larger filters, \textit{ii)} dropout of 0.5 is critical for overfitting mitigation, \textit{iii)} more than one recurrent layer does not yield improvements, while slowing down training, and \textit{iv)} surprisingly, slightly better performance is attained without BN nor pre-activation \cite{fonseca2018simple}.
+%
+%Regarding the network structure, several variants of the CRNN architecture were explored until reaching the network of Figure \ref{fig:archi}.
+%This included a small grid search over number of CNN filters, CNN filter size and shape, number of GRU units, number of FC units, dropout \cite{srivastava2014dropout}, learning rate, and the usage of Batch Normalization (BN) \cite{ioffe2015batch}.
+%Network extensions (involving more weights) were considered only if providing major improvements, as a measure against overfitting.
+%The main takeaways are: \textit{i)} squared 3x3 filters provide better results than larger filters, \textit{ii)} dropout of 0.5 is critical for overfitting mitigation, \textit{iii)} more than one recurrent layer does not yield improvements, while slowing down training, and \textit{iv)} surprisingly, slightly better performance is attained without BN nor pre-activation \cite{fonseca2018simple}.
 \\
-
-For all experiments, the batch size was 100 and Adam optimizer was used \cite{kingma2014adam} with initial learning rate of 0.001, halved each time the validation accuracy plateaus for 5 epochs.
-Earlystopping was adopted with a patience of 15 epochs, monitoring validation accuracy.
-Prediction for every event was obtained by computing predictions at the patch level, and aggregating them with the geometric mean to produce a clip-level prediction. \\
-
-Finally, we apply \textit{mixup} \cite{zhang2017mixup} as data augmentation technique.
-Mixup consists in creating virtual training examples through linear interpolations in the feature space, assuming that they correspond to linear interpolations in the label space.
-Essentially, virtual TF patches are created on the fly as convex combinations of the input training patches, with a hyper-parameter $\alpha$ controlling the interpolation strength.
-Mixup has been proven successful for sound event classification, even in adverse conditions of corrupted labels \cite{Fonseca2019model}.
-It seems appropriate for this task since the front-end outcome can present leakage due to overlapping sources, effectively mixing two sources while only one training label is available, which can be understood as a form of label noise \cite{Fonseca2019learning}.
-Experiments revealed that mixup with $\alpha=0.1$ boosted testing accuracy in $\sim1.5$\%.
-
-%==============================================
-
-\section{Results and Discussion}
-\label{sec:results}
-\begin{table}[!htbp]
-\centering
-\caption{Results for development (top) and evaluation (bottom) sets.}
-
-\begin{tabular}{cccccc}
-\toprule
-Method & \textit{ER} & \textit{F} & \textit{DOA} & \textit{FR} & \textbf{\textit{SELD}}\\
-\midrule
-Baseline & 0.34 & 79.9\% & 28.5$^{\circ}$  & 85.4\% & 0.2113\\
-Proposed & 0.32 & 79.7\% & 9.1$^{\circ}$   & 76.4\% & \textbf{0.2026}\\
-Ideal front-end & 0.08 & 93.2\% & $\sim0^{\circ}$   &$ \sim100$\% & 0.0379\\
-\midrule
-Baseline & 0.28 & 85.4\% & 24.6$^{\circ}$  & 85.7\% &  0.1764\\
-Proposed & 0.29 & 82.1\% & 9.3$^{\circ}$   & 75.8\% & \textbf{ 0.1907}\\
-\bottomrule
-\end{tabular}
-\label{tab:results_real}
-\end{table}
-
-Table~\ref{tab:results_real} shows the results of the proposed method for both development and evaluation sets, compared to the baseline.
-Focusing on evaluation results, our method and the baseline obtain similar performance in SED (\textit{ER} and \textit{F}). However, there is a clear difference in the DOA metrics:
-in our method, \textit{DOA} error is reduced by a factor of 2.6, but \textit{FR}  is $\sim10$ points worst.
-In terms of \textit{SELD score}, our method performs slightly worse than the baseline in evaluation mode, while marginally outperforming it in development mode. \\m,
-
-The most relevant observation is the great improvement in \textit{DOA} error.
+%
+%For all experiments, the batch size was 100 and Adam optimizer was used \cite{kingma2014adam} with initial learning rate of 0.001, halved each time the validation accuracy plateaus for 5 epochs.
+%Earlystopping was adopted with a patience of 15 epochs, monitoring validation accuracy.
+%Prediction for every event was obtained by computing predictions at the patch level, and aggregating them with the geometric mean to produce a clip-level prediction. \\
+%
+%
+%Finally, we apply \textit{mixup} \cite{zhang2017mixup} as data augmentation technique.
+%Mixup consists in creating virtual training examples through linear interpolations in the feature space, assuming that they correspond to linear interpolations in the label space.
+%Essentially, virtual TF patches are created on the fly as convex combinations of the input training patches, with a hyper-parameter $\alpha$ controlling the interpolation strength.
+%Mixup has been proven successful for sound event classification, even in adverse conditions of corrupted labels \cite{Fonseca2019model}.
+%It seems appropriate for this task since the front-end outcome can present leakage due to overlapping sources, effectively mixing two sources while only one training label is available, which can be understood as a form of label noise \cite{Fonseca2019learning}.
+%Experiments revealed that mixup with $\alpha=0.1$ boosted testing accuracy in $\sim1.5$\%. 
+%
+%
+%
+%
+%
+%
+%%==============================================
+%
+%\section{Results and Discussion}
+%\label{sec:results}
+%\begin{table}[!htbp]
+%\centering
+%\caption{Results for development (top) and evaluation (bottom) sets.}
+%
+%\begin{tabular}{cccccc}
+%\toprule
+%Method & \textit{ER} & \textit{F} & \textit{DOA} & \textit{FR} & \textbf{\textit{SELD}}\\
+%\midrule
+%Baseline & 0.34 & 79.9\% & 28.5$^{\circ}$  & 85.4\% & 0.2113\\
+%Proposed & 0.32 & 79.7\% & 9.1$^{\circ}$   & 76.4\% & \textbf{0.2026}\\
+%Ideal front-end & 0.08 & 93.2\% & $\sim0^{\circ}$   &$ \sim100$\% & 0.0379\\
+%\midrule
+%Baseline & 0.28 & 85.4\% & 24.6$^{\circ}$  & 85.7\% &  0.1764\\
+%Proposed & 0.29 & 82.1\% & 9.3$^{\circ}$   & 75.8\% & \textbf{ 0.1907}\\
+%\bottomrule
+%\end{tabular}
+%\label{tab:results_real}
+%\end{table}
+%
+%
+%Table~\ref{tab:results_real} shows the results of the proposed method for both development and evaluation sets, compared to the baseline.
+%Focusing on evaluation results, our method and the baseline obtain similar performance in SED (\textit{ER} and \textit{F}). However, there is a clear difference in the DOA metrics:
+%in our method, \textit{DOA} error is reduced by a factor of 2.6, but \textit{FR}  is $\sim10$ points worst.
+%In terms of \textit{SELD score}, our method performs slightly worse than the baseline in evaluation mode, while marginally outperforming it in development mode. \\m,              
+%
+%The most relevant observation is the great improvement in \textit{DOA} error.
 Results suggest that using spatial audio parametric analysis as a preprocessing step can help to substantially improve localization.
-Figure \ref{fig:5a} provides further evidence for this argument: Challenge methods using some kind of parametric preprocessing (\textit{GCC-PHAT} with the microphone dataset, and \textit{Intensity Vector-Based} in ambisonics) obtained in average better \textit{DOA} error results.
+%Figure \ref{fig:5a} provides further evidence for this argument: Challenge methods using some kind of parametric preprocessing (\textit{GCC-PHAT} with the microphone dataset, and \textit{Intensity Vector-Based} in ambisonics) obtained in average better \textit{DOA} error results.
 \\
-
-\begin{figure}
-  \begin{subfigure}[b]{\textwidth}
-    \includegraphics[width=\textwidth]{Figures/SELD//doaerror2.jpg}
-    \caption{\textit{DOA error} across submissions.
+%
+%\begin{figure}
+%  \begin{subfigure}[b]{\textwidth}
+%    \includegraphics[width=\textwidth]{Figures/SELD//doaerror2.jpg}
+%    \caption{\textit{DOA error} across submissions.
 Hatched bars denote methods using parametric preprocessing.
 Horizontal lines depict average DOA error accross different subsets: all methods (solid), parametric methods (dashed), non-parametric methods (dotted).}
-    \label{fig:5a}
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \includegraphics[width=\textwidth]{Figures/SELD//complexity_vs_seld2.jpg}
-    \caption{\textit{SELD score} versus complexity.}
-    \label{fig:5b}
-  \end{subfigure}
-\caption{DCASE2019 Challenge Task 3 results, evaluation set.}
-\end{figure}
-
-Conversely, the front-end fails regarding \textit{FR}.
+%    \label{fig:5a}
+%  \end{subfigure}
+%  \begin{subfigure}[b]{\textwidth}
+%    \includegraphics[width=\textwidth]{Figures/SELD//complexity_vs_seld2.jpg}
+%    \caption{\textit{SELD score} versus complexity.}
+%    \label{fig:5b}
+%  \end{subfigure}
+%\caption{DCASE2019 Challenge Task 3 results, evaluation set.}
+%\end{figure}
+%
+%Conversely, the front-end fails regarding \textit{FR}.
 This is probably due to the complexity added by the association step \cite{Adavanne2018_JSTSP}, and its lack of robustness under highly reverberant scenarios.
@@
 \\
-
-In order to gain a better insight of the classification back-end performance, Table~\ref{tab:results_real} shows the method results when the testing clips are obtained by feeding  the beamformer  with groundtruth annotations (\textit{ideal} front-end).
-In this ideal scenario of DOA performance, the SED metrics show a significant boost. 
-This result suggests that the low \textit{FR} given by the front-end has a severe impact on the back-end  performance.
-Yet, the proposed system reaches similar performance to the baseline system in terms of SED metrics. \\
-
-Finally, we would like to discuss algorithm complexity among Challenge methods. 
-As depicted in Figure \ref{fig:5b}, there is a general trend towards architectures with very high number of weights, as a consequence of the usage of ensembles and large capacity networks.
-Specifically, 66\% of submitted methods employ 1M weights or more, 30\% employ 10M or more, and 15\% employ 100M or more.
+%
+%In order to gain a better insight of the classification back-end performance, Table~\ref{tab:results_real} shows the method results when the testing clips are obtained by feeding  the beamformer  with groundtruth annotations (\textit{ideal} front-end).
+%In this ideal scenario of DOA performance, the SED metrics show a significant boost. 
+%This result suggests that the low \textit{FR} given by the front-end has a severe impact on the back-end  performance.
+%Yet, the proposed system reaches similar performance to the baseline system in terms of SED metrics. \\
+%
+%Finally, we would like to discuss algorithm complexity among Challenge methods. 
+%As depicted in Figure \ref{fig:5b}, there is a general trend towards architectures with very high number of weights, as a consequence of the usage of ensembles and large capacity networks.
+%Specifically, 66\% of submitted methods employ 1M weights or more, 30\% employ 10M or more, and 15\% employ 100M or more.
 Such complexities are several orders of magnitude greater than the baseline (150k weights) or the proposed method ($\sim$175k weights).
-In this context, our method represents a low-complexity solution to the SELD problem, featuring a number of parameters and a performance comparable to the baseline method.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-We present a novel approach for the SELD task.
+%In this context, our method represents a low-complexity solution to the SELD problem, featuring a number of parameters and a performance comparable to the baseline method.
+%
+%
+%\section{Conclusion}
+%\label{sec:conclusion}
+%
+%We present a novel approach for the SELD task.
 Our method relies on spatial parametric analysis for the computation of event DOAs, onsets and offsets.
 This information is used to filter the input signals in time and space, and the resulting event estimations are fed into a CRNN which predicts the class to which the events belong; the classification problem is thereby handled from a simple multi-class perspective.
-The proposed method is able to obtain an overall performance comparable to the baseline system. 
-The localization accuracy achieved by our method greatly improves the baseline performance, suggesting that spatial parametric analysis might enhance performance of SELD algorithms.
+%The proposed method is able to obtain an overall performance comparable to the baseline system. 
+%The localization accuracy achieved by our method greatly improves the baseline performance, suggesting that spatial parametric analysis might enhance performance of SELD algorithms.
 Moreover, detection and classification performance in our method suffers from a low Frame Recall; improving this metric could lead to promising SELD scores.
+%
+%
+%

### 2020-06-26 Chapters/ScientificBackground.tex
update 20200626

@@
 
-where the term $\Gamma_n(kR)$ models the radial transfer function, and depends on the microphone geometry.
-There are several possible sampling schemes of capsules along the sphere, each one having different properties; the reader is referred to  \cite{rafaely2004analysis} for a deeper insight.
-\\
+where the term $\Gamma_n(kR)$ models the radial transfer function, and depends on a number of factors, being the \textit{sphere configuration} one of the most prominent considerations. 
+Sphere configuration, in its basic form, refers to the physical properties of the baffle where the capsules are mounted, and it can be either \textit{open} or \textit{rigid}. 
+While open configuration is the simplest solution, it might present numerical problems in the form of zeros in its frequency response.
+Conversely, a rigid baffle interferes with the sound field and might create undesired interferences, but it improves the numerical condition from the open case.
+Fig.~\ref{fig:magnitude_kr} shows the simulated magnitude response of $\Gamma_n(kR)$  for a spherical array considering both configurations.
+The reader is referred to \cite{moreau20063d} and \cite{rafaely2004analysis}  for a deeper insight into the topic of spherical microphone array design.
 
-By using this model, the maximum spherical harmonic order $N$ that can be retrieved with negligible spatial aliasing depends on the number of microphone capsules \cite{moreau20063d}:
+%\begin{figure}[h!]
+%  \includegraphics[width=\textwidth]{Figures/ScientificBackground/magnitude_kr.png}  
+%  \caption{Magnitude of $\Gamma_n(kR)$ for different ambisonic orders, in the case of (a) rigid sphere, and (b) open sphere configurations. Adapted from \cite{rafaely2004analysis}. }
+%  \label{fig:magnitude_kr}
+%\end{figure}
+%
+
+\begin{figure}[hbt]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{Figures/ScientificBackground/magnitude_kr.png}
+  \caption{Magnitude of $\Gamma_n(kR)$ for different ambisonic orders, in the case of (a) rigid sphere, and (b) open sphere configurations. Adapted from \cite{rafaely2004analysis}. }
+	\label{fig:magnitude_kr}
+	\end{center}
+\end{figure}
+
+%and depends on the microphone geometry, the directivity of the capsules and the physical properties of the baffle where the capsules are mounted.
+%
+% There are several possible sampling schemes of capsules along the sphere, each one having different properties; the reader is referred to  \cite{rafaely2004analysis} for a deeper insight. \\
+
+By using the model from Eq.~\ref{eq:a2b}, the maximum spherical harmonic order $N$ that can be retrieved with negligible spatial aliasing depends on the number of microphone capsules \cite{moreau20063d}:
 \begin{equation}
@@
 Furthermore, the sphere radius $R$ has also an effect on the operational bandwidth of the microphone.
-More precisely, for a given spherical harmonic order $n$, the maximum aliasing-free operational frequency is given by:
+According to \cite{moreau20063d}, the maximum aliasing-free operational frequency of a spherical microphone array is given by:
 \begin{equation}
-	f_{max} = \frac{n c} {2 \pi R},
+	f_{max} = \frac{c} {2 R \gamma},
 	\label{eq:falias}
@@
 
-with $c$ being the sound speed.
-
-\todo{moreau has c/2Rgamma}.
+with $c$ being the sound speed, and $\gamma$ the maximum aperture angle between two capsules. 
+It is important to notice the existence of a practical minimum frequency of the spherical microphone array, given by the low magnitude in low frequencies of high ambisonic order components, as shown in Fig.~\ref{fig:magnitude_kr}.
 
@@
 
-\subsection{SOFA Conventions}
+%\subsection{SOFA Conventions}
+%
+%The availability of recorded room impulse responses is of great importance to many acoustic signal processing problems. 
+%As we have seen, many different RIRs can be obtained from the same room, by just varying the position of the source and the receiver; when the number of source and receiver positions increases, the total amount of measurements increases geometrically.
+%Besides that, the actual format and organisation of the produced data (not only the RIR themselves, but also the source/position annotations) can be arbitrarily different when produced by different groups of people.
+%
+%In order to overcome potential interoperatibility and reusability issues, the \textit{Spatially Oriented Format for Acoustics} (SOFA) convention \cite{majdak2013spatially}, also known as the \textit{AES-69} standard \cite{majdak2015aes69}, proposes a unified file format for the storage of IR-related data.
+%Despite that SOFA was initially created with a focus on \textit{Head-Related Impulse Response} (HRIR) data, its structure is very convenient to any kind of multi-location impulse responses, including ambisonics.
 
-The availability of recorded room impulse responses is of great importance to many acoustic signal processing problems. 
-As we have seen, many different RIRs can be obtained from the same room, by just varying the position of the source and the receiver; when the number of source and receiver positions increases, the total amount of measurements increases geometrically.
-Besides that, the actual format and organisation of the produced data (not only the RIR themselves, but also the source/position annotations) can be arbitrarily different when produced by different groups of people.
-
-In order to overcome potential interoperatibility and reusability issues, the \textit{Spatially Oriented Format for Acoustics} (SOFA) convention \cite{majdak2013spatially}, also known as the \textit{AES-69} standard \cite{majdak2015aes69}, proposes a unified file format for the storage of IR-related data.
-Despite that SOFA was initially created with a focus on \textit{Head-Related Impulse Response} (HRIR) data, its structure is very convenient to any kind of multi-location impulse responses, including ambisonics.
+\todo{RIR SIMULATION: IMAGE METHOD}
 

### 2020-06-26 manuscript.tex
update 20200626

@@
 Ambisonic audio can be captured using spherical microphone arrays, which are becoming increasingly popular as well.
-Yet, most methods for acoustic and microphone array signal processing are not specifically tailored for spherical geometries.
+Yet, many methods for acoustic and microphone array signal processing are not specifically tailored for spherical geometries.
 Therefore, there is still room for improvements in the field of automatic analysis and description of ambisonic recordings. 
@@
 %ADD BIBLIOGRAPHY
-\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib, Bibliography/Introduction.bib, Bibliography/ScientificBackground.bib}
+\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, 
+	Bibliography/Data.bib, 
+	Bibliography/AmbisonicsDRIR.bib, 
+	Bibliography/ReverberationTimeEstimation.bib,
+	Bibliography/Introduction.bib, 
+	Bibliography/ScientificBackground.bib}
 

### 2020-06-27 Chapters/DataGeneration.tex
finished data chapter

--- 2020-01-23 Chapters/DataGeneration.tex
@@
 \chapter{Data generation and storage}
+\label{chap:data}
 
-Explain about mono files plus ambisonics IRs.
+\section{Introduction}
+\label{sec:intro_data}
 
-\section{Recorded IRs}
+This chapter gathers several proposals related with the creation, storage and transmission of ambisonic data for research purposes.
+The main objective of the contributions described here is the support for the generation of parametrizable ambisonic datasets, using both synthetic and recorded materials, and specifically emphasizing the usage of Room Impulse Responses.\\
 
-AmbisonicsDRIR: explain problematics, standard, etc.
-also: pysofaconventions
+Most of the contributions listed here (and also most of the code developed for this thesis) have been implemented in Python.
+Indeed, Python has recently become one of the most used programming languages worldwide \cite{theoverflow, PYPL, TIOBE}; as shown also in Figure~\ref{fig:popularity}.\\
 
-\section{Simulated IRs}
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{Figures/DataGeneration/projections-1-1400x1200.png}
+    \caption{2018 projections of future internet traffic for major programming languages. Adapted from \cite{theoverflow}.}
+    \label{fig:popularity}
+\end{figure}
 
-explain different methods and libraries.
-tell about masp.
+One of the reasons behind this tendency shift is the popularity of the language among machine learning and data science communities, fields where Python holds the first place by usage \cite{githubblog}.
+Since data-driven paradigms currently conform the state-of-the-art of many applied sciences, including audio signal processing, the availability of convenience Python packages and libraries is therefore of the highest interest to the research community.
 
-\section{High-level scene description}
+It is important to remark the predominant position that Matlab has always had regarding scientific computing.
+Indeed, it is still the tool of choice for many researchers, and the availability of libraries is accordingly very high.
+But the aforementioned tendency shift towards Python causes, as a side effect, the lack of many tools developed in Matlab by the research community.
 
-ambiscaper
+Although Matlab code can be called and executed from Python, in practice this approach is suboptimal under several criteria.
+A better solution in the long run is the effective port of the code towards native Python code.
+Some of the libraries presented in this Chapter are partially or totally  motivated by this scenario.
+
+\section{MASP: a Python library for multichannel acoustic signal processing}
+
+\subsection{Description}
+
+The Multichannel Acoustic Signal Processing (\textit{MASP}) is a Python library consisting of a collection of methods related with acoustics and microphone array processing.
+The library is mostly a transcoding from several Matlab libraries by A. Politis\cite{politis2016microphone, github_politis}. 
+It can be conveniently installed using \textit{pip}.
+
+\textit{MASP} implements a variety of methods for the simulation and analysis of reverberant acoustic scenes, with emphasis on microphone arrays with spherical geometries.
+More specifically, MASP is structured in submodules, with the following structure :
+
+\begin{description}
+
+	\item [Array Response Simulator] Simulation of
+spherical microphones:
+	\begin{itemize}
+		\item Rigid/open configurations.
+		\item Scattering simulation.
+		\item Arbitrary capsule distances, positions and directivities.
+	\end{itemize}
+
+	\item [Shoebox Room Model] Fast implementation of the Image Source Method \cite{imagemethod}:
+	\begin{itemize}
+		\item Convex 3D rooms.
+		\item Arbitrary number of sources and receivers, with arbitrary positions, orientations and directivities.
+		\item ISM expansion limited by order or time.
+		\item Frequency-dependent wall absorption.
+		\item RIR with spherical harmonic expansion.
+	\end{itemize}
+
+	\item [Spherical Array Processing] Transformation and analysis of signals measured with a spherical microphone array:
+	\begin{itemize}
+		\item A2B conversion with theoretical or measured filters.
+		\item Signal-independent beamforming.
+		\item Signal-dependent and adaptive beamforming.
+		\item Direction of Arrival estimation.
+		\item Diffuseness estimation.
+	\end{itemize}
+
+	\item [Spherical Harmonic Transform] Mathematical convenience tools.
+\end{description}
+
+The library implements a Unit Testing system, which numerically assesses the validity of the methods. 
+More specifically, each function test calls the equivalent Matlab code under the hood. The numeric result is then sent back to Python, where
+it is evaluated against the own result.\\
+
+Two example applications of the library are shown in Figures~\ref{fig:array_response} and \ref{fig:sht_filters}.
+In Figure~\ref{fig:array_response} , obtained with the Array Response Simulator package, the frequency response of a spherical microphone array to an impinging plane-wave with varying incidence angle is shown.
+The microphone array consists of a 2nd order supercardioid and a 3rd order hypercardioid, located at opposite directions of an open sphere, and both of them facing to the front direction.
+\\
+
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=\textwidth]{Figures/DataGeneration/array_response.png}
+    \caption{Frequency response of an arbitrary spherical array.}
+    \label{fig:array_response}
+\end{figure}
+
+One of the features of the Spherical Array Processing package is shown in Figure~\ref{fig:sht_filters}.
+The plot shows the evaluation of radial filters $\Gamma_n(kR)$ for an arbitrary spherical array, generated by inverting the theoretical response of the array \cite{Bertet2006}.
+The evaluation is performed following the metrics presented in the same paper, which compare spatial correlation, level difference and maximum amplification with respect to the ideal case. \\
+
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=1\textwidth]{Figures/DataGeneration/sht_filters.png}
+    \caption{Evaluation of radial filters for an arbitrary spherical microphone array.}
+    \label{fig:sht_filters}
+\end{figure}
+
+\subsection{Related software}
+
+There exists another recent Python library which covers a similar scope: \textit{pyroomacoustics} \cite{scheibler2018pyroomacoustics}.
+This framework provides an object-oriented interface with two main application scopes: allow RIR simulation of complex rooms based on the image source method, and provide a reference implementation of standard microphone array processing algorithms.
+
+Although some of the features are common to both libraries, there is a significant difference regarding their target usage.
+While \textit{MASP} primarily focuses on spherical geometries, \textit{pyroomacoustics} is more concerned about arbitrary room geometries and computational performance.
+Therefore, both libraries might be considered as complementary to some extent. 
+A comparative list of their features is shown in Table~\ref{tab:masp_features}.
+
+\begin{table}[th!]
+\centering
+\caption{Features of \textit{MASP} compared to \textit{pyroomacoustics}.}
+
+\begin{tabular}{cccc}
+  \toprule
+Package & Feature &  MASP & PRA \\
+\midrule
+Shoebox & Convex 3D room         & \checkmark	& \checkmark \\
+Room    & Non-convex 3D room     & -    			& \checkmark   \\
+Model   & Arbitrary \#sources    & \checkmark   & \checkmark      \\
+ & Arbitrary \#receivers, arrays & \checkmark   & \checkmark   \\
+& ISM by max\_order                                 & \checkmark                              & \checkmark                              \\
+& ISM by max\_time                                  & \checkmark                              & -                              \\
+& Wall absorption                                   & \checkmark                              & \checkmark                              \\
+& Frequency-dependent absorption                    & \checkmark                              & -                              \\
+& Plot methods                                      & -                              & \checkmark                              \\
+& RIR rendering                                     & \checkmark                              & \checkmark                              \\
+& Audio simulation                                  & \checkmark                              & \checkmark                              \\
+& Acoustic descriptor estimation                    & \checkmark                              & -                              \\
+& Microphone orientation                            & \checkmark                              & -                              \\
+& Custom microphone directivity                     & \checkmark                              & -                              \\
+& RIR Spherical Harmonic Expansion                  & \checkmark                              & -                              \\
+
+\midrule
+Array   	& Rigid spherical arrays   & \checkmark                              & -     \\
+Simulator 	& Arbitrary capsule geometries & \checkmark   & -  \\
+           	& Recorded array IRs  & \checkmark                              & - \\
+\midrule
+Spherical & A2B conversion   & \checkmark      & -       \\
+Processing  & Beamforming          		 & \checkmark      & \checkmark                              \\
+ & Plane-wave decomposition         		 & \checkmark      & -                              \\
+ & Nullformer                            & \checkmark      & -                              \\
+& Adaptive Beamforming                   & \checkmark      & -                              \\
+ & Adaptive Filtering             		 & -               & \checkmark                              \\
+& DoA Estimation                         & \checkmark      & \checkmark                              \\
+ & Diffuseness Estimation             	 & \checkmark      & -                              \\
+ & Diffuse-field coherence               & \checkmark      & -                              \\
+ & Blind Source Separation               & -               & \checkmark          \\                   
+\bottomrule
+
+\end{tabular}
+\label{tab:masp_features}
+\end{table}
+
+\section{SOFA}
+
+\subsection{Problem statement}
+
+The availability of recorded room impulse responses is of great importance to many acoustic signal processing problems. 
+Many different RIRs can be obtained from the same room, just by varying the position of the source and the receiver; when the number of source and receiver positions increases, the total amount of measurements increases geometrically.
+Besides that, the final format and organisation of the produced data (not only the RIR themselves, but also the source/position annotations) can be arbitrarily different when produced by different groups of people.
+
+In order to overcome potential interoperatibility and reusability issues, the \textit{Spatially Oriented Format for Acoustics} (SOFA) convention \cite{majdak2013spatially}, also known as the \textit{AES-69} standard \cite{majdak2015aes69}, proposes a unified file format for the storage of IR-related data.
+Despite that SOFA was initially created with an emphasis on \textit{Head-Related Impulse Response} (HRIR) data, the framework that SOFA provides can be potentially applied to a variety of recording procedures and audio-related data.
+Such variety is associated with the concept of \textit{conventions}: a specific data structure designed to hold a concrete type of data or measurement.
+Some examples of widespread conventions might be \textit{SimpleFreeFieldHRIR} (for anechoic binaural measurements), \textit{SimpleHeadphoneIR} (intended for storing headphone impulse responses), or \textit{MultiSpeakerBRIR} (for binaural RIRs measured from loudspeaker arrays), to name a few of them.
+
+\subsection{Ambisonics Directional Room Impulse Response as a SOFA convention}
+\label{subsec:adrir}
+
+Given the intrinsic spatial characterization capabilities of ambisonics, Gerzon proposed the technique as a potentially successful candidate format for acoustical heritage preservation, as early as 1975 \cite{gerzon1975recording}.
+
+The increase in popularity of ambisonics since the beginning of the present century has turned this idea into reality; OpenAIRlib, a freely accesible dataset that gathers dozens of RIRs, might be a good example of it \cite{murphy2010openair, openair}.
+
+In any case, the usage of recorded ambisonic RIRs is not limited to the field of acoustic heritage.
+Among others, the availability of such recordings has powered works in a variety of works, auralization \cite{postma2016virtual}, room acoustics analysis \cite{embrechts2015measurement,clapp2011investigations} and modelling \cite{romblom2017diffuse}, spatial audio synthesis \cite{coleman2017object} or source separation \cite{baque2016separation}.
+\\
+
+In general, all publicly available ambisonic RIR measurements share some common approaches for describing and organizing the recorded data. 
+For instance, recordings from different rooms are usually stored as separated folders. 
+Each combination of emitter and receiver positions is often saved as an individual file, and the different spherical harmonics match the audio channels.
+Moreover, it is also usual to provide a \textit{metadata} file, describing the different emitter and receiver positions, and potentially some information about the measurement setup, methodology, etc.
+Such files might be formatted as plain text or delimiter-separated files.
+
+Despite the common approach, it can be easily foreseen that each database generated by a different individual or institution might potentially have a different naming convention, folder structure, file format, and so on.
+This is exactly the same situation that motivated the development of the SOFA conventions.
+
+On the other hand, the SOFA specification defines some criteria that must be fulfilled in order to propose a new convention \cite{sofaconventions}. These criteria are:
+\begin{enumerate}
+    \item Data must exist.
+    \item Data can not be described by existing SOFA conventions.
+    \item Relevant information about the data must be available.
+\end{enumerate}
+
+Given that the described situation meets all requirements, the \textit{Ambisonics Directional Room Impulse Response} (AmbisonicsDRIR) convention has been therefore proposed as a new member of SOFA.\\
+
+The technical specifications of the proposed convention in its current state (version 0.2) are available online \cite{ambisonicsdrir}. 
+Despite a timid adoption of the convention, the AmbisonicsDRIR proposal has arisen interest in the community.
+At the moment of writing, the \textit{Standardisation Committee on AES-69 Standard} is discussing potential modifications to the SOFA file format required for the adoption of several new features; and the support for spherical harmonic-based measurements is among the topics for deliberation.
+
+\subsection{Pysofaconventions}
+
+The situation described in Section~\label{sec:intro_data}, regarding the availability of acoustic signal processing libraries in the Python programming language, can be easily extended to the case of SOFA APIs.
+
+The library \textit{pysofaconventions} has been created with the aim to provide an alternative to the existing Matlab/Octave and C/C++ implementations. 
+For ease of installation, it is integrated in the standard python package manager, \textit{pip}.
+
+The current software version is 0.1.5. The library structure is inspired by the C++ implementation \cite{api_cpp}. 
+It features all functionalities described by SOFA version 1.0, plus the proposed AmbisonicsDRIR convention. 
+The implementation is based on extensive error checking, to ensure code consistency.
+
+\textit{pysofaconventions} has gained a moderate amount of attraction.
+Among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}.
+Furthermore, given the open nature of the project, several external researchers have contributed with bug fixes or new feature implementations.
+
+\section{Ambiscaper}
+
+%==================================================================
+
+\subsection{Motivation}
+
+The availability of data is a fundamental requirement for research. 
+More specifically, in the scope of the B-Format data analysis in which this thesis focuses, audio is frequently generated on an IR-based manner, employing both acoustic simulations and recordings. 
+%As already explained, the actual ambisonic audio under analysis is obtained from the convolution of the RIRs with any desired signal.
+ In that way, the analysis can focus on different signal types, such as speech or music, usually employing dedicated external datasets.
+
+An alternative procedure for ambisonic data generation is the actual recording of sound scenes with spherical microphone arrays. 
+Although this procedure might yield the most realistic sound field representations, the high cost, lack of scene control, and technical difficulty to obtain reliable groundtruth annotations lead to a limited usage of the technique, often reserved for real-life algorithm validation.
+%in two different ways: IR simulation and IR recording.
+As already explained, the actual ambisonic audio under analysis is obtained from the convolution of the RIRs with any desired signal.
+In that way, the analysis can focus on different signal types, such as speech or music, usually employing dedicated external datasets.
+
+%We may also consider a third type of B-Format audio generation, which is the actual recording of a sound scene.
+Although this procedure might yield the most realistic representation of the sound field, the high cost and lack of scene control leads to a limited usage of the technique, often reserved for real-life algorithm validation.
+
+It is of interest to have an insight of the diverse methods used in the literature for ambisonic sound generation.
+Table \ref{table:evaluationdata} summarizes the information gathered from works on the scope of sound source localization and separation, which have been published until
+2018\footnote{The original research to which this section refers was conducted in 2018, for the publication of \cite{perez2018ambiscaper}.
+Since then, the outbreak of localization-related challenges, such as LOCATA 2018 and specially DCASE 2019 and 2020, and the consolidation of data-based methods, has largely contributed to the homogenization of datasets \cite{evers2020locata, adavanne2019multi, politis2020dataset}.
+Still, most assumptions and results of our analysis continue to hold at present.}.
+ The table displays, for each article, how the evaluation data was generated, and which was the type of audio content considered.
+
+%% DOA ESTIMATION
+\begin{table}[htpb]
+    \centering
+    \footnotesize
+
+%	\begin{tabular}{ p{3cm} p{2.8cm} p{2.5cm} p{2.3cm} }
+	\begin{tabular}{ ccc }
+    \toprule
+    % \cmidrule(r){1-2}
+
+    % \centering 
+    Article &
+    % \textbf{\textit{$L$}} &
+    Data Generation Method &
+    Audio Content\\
+    \midrule
+
+    \cite{thiergart_localization_2009}
+    % & 1 h
+    & Audio recording 
+    & Speech\\
+
+    \cite{Tervo2009}
+    % & 1 h
+    & Audio recording
+    & Noise, music\\
+
+    \cite{Jarrett2010} 
+    % & 4 
+    & IR simulation, audio recording
+    & Noise\\
+
+    \cite{Nadiri2014}   
+    % & 4 
+    & IR simulation, audio recording
+    & Speech\\
+
+    \cite{Moore2015}   
+    % & 4 
+    & IR simulation 
+    & Speech\\
+
+    \cite{Pavlidi2015}
+    % & 4
+    & IR simulation 
+    & Noise, speech\\
+
+    \cite{He2017}   
+    % & 1 h
+    & IR simulation, audio recording
+    & Speech\\
+
+    % Ding \cite{Ding2017}   
+    % % & 1 h
+    % & IR simulation \& Audio recording \\
+
+    \midrule
+
+    \cite{Gunel2008}
+    & IR recording
+    & Speech, music\\
+
+    \cite{Shujau2011} 
+    & Audio recording
+    & Speech\\
+
+    \cite{Riaz2015}
+    & IR recording
+    & Speech, music\\
+
+    \cite{Chen2015}
+    & IR simulation, IR recording
+    & Speech\\
+
+    \bottomrule
+    \end{tabular}
+    \caption{Summary of audio data used across Ambisonics-based Source Localization (above) and Source Separation (below) methods.}
+    \label{table:evaluationdata}
+\end{table}
+
+The statistics of the global usage of each generation method show that there is not a clear tendency towards any method.
+Nevertheless, it is also noticeable that the methods with evaluations performed on real recordings make use of only one one sound scene in each case.
+In contrast, when using IR-based scenes, the number of audios evaluated are usually one or two magnitude orders greater.
+
+In addition, it is important to notice the lack of availability of the generated data.
+None of the analyzed articles provide a way to access neither the used audio dataset, nor the groundtruth (position annotations in the case of localization, and original sound sources for sound separation).
+Only when simulated IRs are used, it is possible to partially replicate the experimental setup --- the parameters used in the simulation software are usually provided. 
+Furthermore, the process of dataset creation seems to be performed \textit{ad-hoc} in each case.
+
+Taking into account the flexibility offered by IR-based scenes, it would be desirable to have a tool for automatic generation of ambisonics scenes, and their associated annotations, for analysis purposes.
+A tool with such characteristics would help the scientific community in several ways: (i) reducing the amount of time dedicated to build custom datasets, (ii) reusing publicly available resources and recordings, and (iii) enhancing experiment reproducibility by making easier the exchange of datasets.
+Moreover, the capacity of batch processing of large ambisonic collections might also contribute to the development of data-driven approaches.
+
+\subsection{Implementation}
+\label{sec:ambiscaper}
+
+\textit{AmbiScaper} is a tool designed to provide a flexible way of creating complex ambisonics sound scenes and their associated groundtruth annotations, to be used in the context of source localization and separation algorithms.
+\textit{AmbiScaper} offers a high level control of the sound scene parameters, and provides a simple interface for the creation of large datasets with custom characteristics.
+\textit{AmbiScaper} is based on \textit{Scaper}, a framework designed to generate annotations for training Sound Event Detection models \cite{Salamon2017}.\\
+
+One of the main features of \textit{AmbiScaper} is that all parameters of the sound scene can be specified in a non-deterministic way.
+In that sense, the parameters for each \textit{event} (sound source) are actually generated through a two-step process.
+First, in the \textit{Event Specification}, all parameters related to an event are defined in terms of statistical distributions.
+During the \textit{Event Instantiation}, the actual values for each parameter are then sampled from the statistical distributions.
+This two-step process allows the user to describe abstract \textit{templates} of sound scenes, rather than manually assigning values to parameters.
+Therefore, a single \textit{event specification} might produce potentially infinite different sound scenes.
+\\
+
+%%%%%%%%
+In order to generate a sound scene, \textit{AmbiScaper} requires three different inputs: the original \textit{mono signals}, which will provide the actual audio content; an optional ambisonic RIR; and an \textit{event specification}.
+
+The process of dataset creation starts with the \textit{event instantiation}, as described above.
+Once all values are sampled, three different types of output are generated: the \textit{ambisonic scene}; the \textit{instantiated mono signals} (the original mono signals after data augmentation); and the \textit{annotations}, in the form of a \textit{jams} file \cite{humphrey2014jams}, containing all information about the instantiated values.
+
+The complete architecture of \textit{AmbiScaper} architecture is depicted in Figure~\ref{fig:architecture}.\\
+
+\begin{figure}
+  \centering
+	\includegraphics[width=\textwidth]{Figures/DataGeneration/figure_architecture_V2.png}
+    \caption{\textit{AmbiScaper} architecture.}
+	\label{fig:architecture}
+\end{figure}
+
+When no reverberation is specified, \textit{AmbiScaper} can generate anechoic sound scenes, using the theoretical expression of the ambisonic encoding as in Eq.~\ref{eq:encoding}.
+In this case, there is no upper limit on the Ambisonics order of the rendered scene. 
+Furthermore, the anechoic case allows the modification of the directivity of the source(s) through ambisonic order downgrade \cite{carpentier2017ambisonic}.
+
+\textit{AmbiScaper} partially supports RIR-based scene creation.
+It features a limited set of recorded ambisonic IRs from the S3A database \cite{coleman2015s3a}; compatibility with the AmbisonicsDRIR (Subsection \ref{subsec:adrir}) could be easily included.
+On the other hand, the usage of simulated IRs is provided through a wrapper to the Matlab library \textit{SMIR Generator} \cite{smir}.
+In this case, the reverberation model specifications are defined as well in statistical terms, and the generated RIRs are also stored for evaluation purposes.
+
+%\textit{AmbiScaper} features the possibility of using simulated Ambisonics IRs, through a wrapper to \textit{SMIR Generator}.
+In this case, the reverberation model specifications might be defined as well in statistical terms, and the generated IRs are stored for evaluation purposes.
+A working copy of Matlab is required to run this option.
+
+%\textit{AmbiScaper} supports the usage of recorded Ambisonics IRs, although it is currently limited to IRs from the S3A database \cite{coleman2015s3a}.
+The development of a standardized file format for Ambisonics IRs, which is being discussed at the moment of writing, will provide the flexibility to work with arbitrary Ambisonics IRs.
+%
+%Lastly, AmbiScaper features the possibility of using simulated Ambisonics IRs, through a wrapper to \textit{SMIR Generator}.
+In this case, the reverberation model specifications might be defined as well in statistical terms, and the generated IRs are stored for evaluation purposes.
+A working copy of Matlab is required to run this option.
+
+%When a reverberant sound scene is created, the specific Ambisonics IRs used on the scene are also provided as an output.
+Other research problems such as dereverberation or room reflection modelling might therefore benefit from such data.
+%
+
+\subsection{Experiment reproducibility}
+
+As already mentioned, there is a generalized lack of publicly available datasets of ambisonic reverberant sound scenes. 
+Even when using general purpose audio/speech datasets, the actual evaluation data is usually not available.
+In that sense, the potential compatibility of \textit{AmbiScaper} with public ambisonic RIR databases is a key aspect for reproducibility, since it would allow the systematic reutilization of acoustical measurements in the analysis context.
+
+Furthermore, the output of the \textit{AmbiScaper} dataset generation process is not limited to the actual dataset. 
+In fact, the resulting annotation file does not only contain the \textit{instantiation} (the actual values of each parameter in the sound scene), but also the \textit{specification} (the statistical distributions from which the instantiated values are sampled).
+In the scope of experiment reproducibility, the exchange of \textit{specification files} instead of actual audio files might greatly reduce the storage capacity and bandwidth required to transfer large databases.
+
+\textit{AmbiScaper} is implemented in the form of a Python package, and is publicly available through the \textit{Python Package Index} repository  under the GPL license, thus easing the software adoption and the potential engagement of the scientific community with the development.
+
+As an example of the potential capabilities of the software, a sample dataset for the evaluation of source localization algorithms has been created \cite{ambiscaperDataset}.
+The dataset contains 300 FOA sound scenes, with a duration between 1 and 2 seconds, each one containing a number of static sound sources between one and three.
+Sound sources might have different gains, and they are located at random positions around the sphere. 
+The sources are randomly chosen from a subset of the Anechoic OpenAirlib database \cite{openair}, which mostly contains recordings from baroque musical instruments.
+Reverberation is provided by the \textit{AudioBooth} recorded RIR from the S3A collection \cite{coleman2015s3a}.\\
+
+\todo{explain current state, discussion, etc?}
+
+%\subsection{Current state}
+%
+%
+%
+%
+%AmbiScaper is a tool designed for easy dataset creation and exchange, in the context of reverberant Ambisonics Source Localization and Source Separation.
+It responds to the lack of public datasets for algorithm design, evaluation and reproducibility.
+An example dataset generated with AmbiScaper, and its analysis with state-of-the-art Source Localization algorithms are also provided.
+%Further studies in Ambisonics IR position interpolation would allow creating non-static audio scenes, which might be an interesting feature. 
+%Another feature in progress is the support for a standardized Ambisonics IR format \cite{perez2018ambisonics}, which will ease data reusability and provide the potential to cover a big variety of acoustic scenarios.
+%
+%
+%
+%
+%
+%
+%

### 2020-06-28 Chapters/CoherenceEstimation.tex
on work

@@
 
-\todo{In this chapter}, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
+In this Chapter, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
 The motivation for this work is, first, that tetrahedral arrays are a well known type of microphone arrays, which have today become popular for applications related to Virtual and Augmented Reality. 
@@
 Spherical isotropic noise has been generated following the \textit{geometrical method} \cite{habets_generating_2007, habets_comments_2010}, using $I = 1024$ plane waves.
-The resulting \textit{A-Format} signals correspond to a virtual tetrahedral microphone array mimicking the Ambeo\footnote{Sennheiser Ambeo VR Mic.
-\todo{https://en-us.sennheiser.com/microphone-3d-audio-ambeo-vr-mic}}
+The resulting \textit{A-Format} signals correspond to a virtual tetrahedral microphone array mimicking the Ambeo\footnote{Sennheiser Ambeo VR Mic \cite{sennheiser}.}
 characteristics ($R=0.015$ meter, $\alpha=0.5$). 
@@
 
-\begin{figure}
-	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure1}
-    \caption{\label{fig:Fig1}\textit{A-Format} coherence between microphone signals.
-Left: $\text{MSC}$ as a function of the frequency of theoretical, simulated and recorded \textit{((BLD,BRU)},  $N=5, I=64$) signals.
-Right: mean error $\bar{\varepsilon}$ of the recorded signals' $MSC$ \textit{(BLD,BRU)} compared to the simulated values, for all values of $N$ and $I$.\todo{REDO FIGURE WITH N AND I}}
-\end{figure}
-
 The coherence of the generated \textit{A-Format} signals is exemplified in Fig.~\ref{fig:Fig1} (left), which shows the $MSC$ between the capsule pair (\textit{BLD,BRU}) for the theoretical, simulated and recorded cases.
@@
 
+\begin{figure}[th]
+	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Untitled1}
+    \caption{\label{fig:Fig1}\textit{A-Format} coherence between microphone signals.
+Left: $\text{MSC}$ as a function of the frequency of theoretical, simulated and recorded \textit{((BLD,BRU)},  $N=5, I=64$) signals.
+Right: mean error $\bar{\varepsilon}$ of the recorded signals' $MSC$ \textit{(BLD,BRU)} compared to the simulated values, for all values of $N$ and $I$.}
+\end{figure}
+
 \subsection{B-Format}
+
+In order to evaluate the dependency of the \textit{B-Format coherence} $\Delta$ on the number of time frames used for averaging, the following procedure is presented.
+The simulated \textit{A-Format} sound field has been transformed into the spherical harmonic domain, with and without the application of radial filters $\Gamma_n(kR)$ (Eq.~\ref{eq:a2b}).
+Then, $\Delta$ has been computed with Eq.~(\ref{eq:delta}) for exponentially growing values of $r$ between 1 (8 ms) and 2048 (10.92 s), where $r$ is the vicinity radius used for time averaging, and the number of time windows is given by $T = 2r+1$.
+The time-frequency representation is derived by applying the STFT with the same window parameters as in Subsection \ref{subsec:results_aformat}.\\
+
 \begin{figure}
@@
 \end{figure}
-
-In order to evaluate the dependency of the \textit{B-Format coherence} $\Delta$ on the number of time frames used for averaging, the following procedure is presented.
-The simulated \textit{A-Format} sound field has been transformed into the spherical harmonic domain, with and without the application of radial filters $\Gamma_n(kR)$ (Eq.~\ref{eq:a2b}).
-Then, $\Delta$ has been computed with Eq.~(\ref{eq:delta}) for exponentially growing values of $r$ between 1 (8 ms) and 2048 (10.92 s), where $r$ is the vicinity radius used for time averaging, and the number of time windows is given by $T = 2r+1$.
-The time-frequency representation is derived by applying the STFT with the same window parameters as in Subsection \ref{subsec:results_aformat}.\\
 
@@
 
-As a remark, the comparison between Figs.~\ref{fig:Fig1} and \ref{fig:Fig2} provides evidence that the application of the spherical harmonic transform might be able to yield more accurate diffuseness estimations, due to a better signal conditioning \cite{epain_spherical_2016}.\\
+As a remark, the comparison between Figs.~\ref{fig:Fig1} and \ref{fig:Fig2} provides evidence that the application of the spherical harmonic transform might be able to yield more accurate diffuseness estimations, due to a better signal conditioning \cite{epain2016spherical}.\\
 
 \begin{figure}
-	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Figure3}
+	\includegraphics[width=\textwidth]{Figures/CoherenceEstimation/Untitled2}
 	\caption{\label{fig:Fig3}\textit{B-Format coherence} between microphone signals.
 Left: $\Delta$ of simulated and recorded ($N=5, I=64$) signals.
-Right: $\bar{\varepsilon}$ of the recorded signals coherence across all values of $N$ and $I$.\todo{REDO FIGURE} }
+Right: $\bar{\varepsilon}$ of the recorded signals coherence across all values of $N$ and $I$.}
 \end{figure}

### 2020-06-28 Chapters/ReververationTimeEstimation.tex
on work

@@
 
-The IRs are FOA room impulse responses simulated by the image method with the \textit{Multichannel Acoustic Signal Processing} library \todo{[ref to chapter?]}.
+The IRs are FOA room impulse responses simulated by the image method with the \textit{Multichannel Acoustic Signal Processing} library (Section~\ref{sec:masp}).
 There are 9 different IRs of 1 s, with random $T_{60}$ values in the range between 0.4 s and 1.1 s approximately, estimated by the $T_{10}$ method at the 1 kHz band.

### 2020-06-28 Chapters/ScientificBackground.tex
on work

@@
 \subsection{Nomenclature}
-\todo{is this name correct? }
+
 Throughout the Thesis, we refer to time-domain signals with lowercase, e.g. $x(t)$, with $t$ as the time index.
@@
 Provided that this condition is usually difficult to fulfil in real scenarios, there are several methods which allow ambisonic decoding for such \textit{irregular} layouts.
-One of the most commonly used is the AllRAD method \todo{cite zotter}.
+One of the most commonly used is the AllRAD method \cite{zotter2012all}.
 AllRAD proposes a two step decoding: first, the ambisonic signal is decoded to a nearly-uniform layout of virtual speakers.
-Then, the signals of the virtual speakers are further distributed into the real speakers by the \textit{Vector-Based Amplitude Panning (VBAP)} method \todo{cite pulkki}.
+Then, the signals of the virtual speakers are further distributed into the real speakers by the \textit{Vector-Based Amplitude Panning (VBAP)} method \cite{pulkki1997virtual}.
 

### 2020-06-29 Chapters/Conclusions.tex
wip
beta1

@@
 
-\begin{itemize}
-  \item Academic Contributions
+The scientific objectives of this thesis, as they were described in Section \ref{sec:objectives}, are:
 
 \begin{enumerate}
-  	\item Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone \todo{ref}
-  	\begin{itemize}
-  		\item \textbf{Parameter estimation}: Contribution to the characterization of coherence with tetrahedral microphones (the most common spherical arrangement)
-	\end{itemize}
-
-  	\item Blind reverberation time estimation from ambisonic recordings \todo{ref}
-  	\begin{itemize}
-  		\item \textbf{Parameter estimation}: Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models
-	\end{itemize}
-
-  	\item A hybrid parametric-deep learning approach for sound event localization and detection \todo{ref}
-  	\begin{itemize}
-  		\item \textbf{Scene Description}: Novel state-of-the-Art methodology for Sound Event Localization and Detection	
-  	\end{itemize}
-
+	\item The development of methods for the characterization of acoustic parameters.
+	\item The research on parametric-based methodologies for sound event localization and detection.
+	\item The contribution in the generation and storage of annotated ambisonic datasets.
 \end{enumerate}
 
-  \item Software Contributions
+In the following list we provide the main contributions of the present thesis, both in the academic and software scopes.
+Contrubitions are organized by the Chapter where they appear; the index of the respective contribution is also indicated:
 
-	\begin{enumerate}
+  \begin{itemize}
 
-  		\item Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes \todo{ref}
-  		\begin{itemize}
-  			\item \textbf{Data Generation}: Novel tool for reverberant ambisonic dataset generation
-		\end{itemize}
+  	 \item Blind reverberation time estimation (Chapter~\ref{chap:rt60}):
+  	\begin{description}
+  		\item [1. Parameter estimation] Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models.
+	\end{description}
 
-  		\item  Ambisonics Directional Room Impulse Response as a New Convention of the Spatially Oriented Format for Acoustics \todo{ref}
-		\begin{itemize}
-  			\item \textbf{Recorded IRs}: File standard/convention proposal for storage of recorded ambisonic IRs
-		\end{itemize}
+  	\item Coherence estimation (Chapter~\ref{chap:coherence}):
+  	\begin{description}
+  		\item [1. Parameter estimation] Contribution to the characterization of coherence with tetrahedral microphones (the most common spherical arrangement).
+	\end{description}
 
-		\item A Python library for Multichannel Acoustic Signal Processing \todo{ref}
-		\begin{itemize}
-  			\item \textbf{Simulated IRs}: Library for acoustic simulation (IR generation, microphone array simulation, etc)
-		\end{itemize}
+  	\item Sound Event Localization and Detection (Chapter~\ref{chap:seld2019}):
+  	\begin{description}
+  		\item [2. Scene Description]: Novel state-of-the-Art methodology for Sound Event Localization and Detection	.
+  	\end{description}
 
-  		\item pysofaconventions, a Python API for SOFA \todo{ref}
-		\begin{itemize}
-  			\item \textbf{Recorded IRs}: implementation of SOFA for python 
-		\end{itemize}
+  	\item Data generation and storage (Chapter~\ref{chap:data}):
+  		\begin{description}
+  			\item [3. Data Generation]: Library for acoustic simulation and spherical microphone array processing.
+  			\item [3. Data Generation]: Proposal and implementation of a file convention for the storage of recorded ambisonic RIRs.
+  			\item [3. Data Generation]: Novel tool for high-level description and generation of of ambisonic datasets.
 
-	\end{enumerate}
+ 		\end{description}
 
@@
 
-\section{List of related publications}
+Figure~\ref{fig:scheme1_numbers2}, which is a copy of Figure~\ref{fig:scheme1_numbers}, 
+ has been included here again in order to help the contextualization of the contributions.
+
+  \begin{figure}[h!]
+  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
+  \caption{General scheme of the B-Format audio generation and analysis framework, including the thesis contributions in form of Chapter numbers.}
+  \label{fig:scheme1_numbers2}
+\end{figure}
+
+\section{Conclusion and Future Work}
+
+This thesis has tackled several research problems associated with the analysis of ambisonic recordings, making use primarily of the parametric sound field modelling.
+Chapter~\ref{chap:rt60} presents a novel methodology for the blind estimation of reverberation time from ambisonic audio.
+The method is based on two main steps: first, dereverberation using a multichannel auto-recursive model (MAR), and second, estimation of the  filter from reverberant and dereverberated signals.
+The actual reverberation time value is estimated from the energy decay of the estimated filter.
+
+The proposed system is the first attempt in the literature to address the blind reverberation time estimation problem specifically for ambisonic signals.
+Compared with a state-of-the-art monophonic estimator, our method is able to improve in all the standard evaluation metrics under consideration.
+
+Still, there is much room for improvement.
+Among others, the method could be easily extended to higher order ambisonic signals, which might still improve the reported results due to the availability of much more audio channels.
+Moreover, the usage of online MAR methods would enable the possibility of analyzing sound scenes with moving sources; the statistical time-invariance property of late reverberation supports this hypothesis.
+Finally, it is important to mention that the proposed method is resource-intensive.
+An analysis of the trade-off between computation time and evaluation performance, mostly dependent on the estimation filter length, remains to be done.
+
+Given the current interest in the field of augmented reality, new related research topics emerge.
+One of them, which has been recently baptized as \textit{acoustic matching} \cite{su2020acoustic}, deals with the analysis of acoustic properties of real enclosures, with the aim of later introduction of virtual elements whose reverberation would match real conditions.
+The application of our method to the acoustic matching problem is straightforward: given an ambisonic recording with a target reverberation, estimate its reverberation time and synthesize a reverberant tail with the target energy decay; early reflections might be generated by various methods, including physical models or perceptually motivated approaches.
+We can foresee a growing interest on the topic in the near future; our contribution might help to establish the foundation of a new family of methods. \\
+
+In Chapter~\ref{chap:coherence}, we have characterised the response of tetrahedral microphones to isotropic noise field, which is one of the most used models for diffuse sound.
+Furthermore, the capabilities of a spherical loudspeaker array with respect to the reconstruction of diffuse sound fields using ambisonics are also tested.
+Future work in this direction might include the characterisation of different spherical microphone array geometries, and different types of diffuse field models.
+Additionally, the loudspeaker array diffuse field reconstruction experiment could be easily extended.
+\\
+
+Chapter~\ref{chap:seld2019} describes an algorithm for sound event localization and detection (SELD), developed in the context of the DCASE 2020 challenge.
+The method estimates the localization and temporal activity of the sound events based on a particle filter that tracks event trajectories obtained from the parametric analysis of the ambisonic sound field.
+Each event is assigned to a sound class by a machine learning classifier that uses low- and mid- level audio features.
+Results on the cross-validation development set show a significant performance increase in most evaluation metrics, compared with a state-of-the-art deep learning baseline.
+This result shows that our approach, which is substantially different to the baseline and the majority of state-of-the-art methods, represents an alternative in specific situations with low-complexity or small database constraints.
+
+The wide scope of the SELD task allows for a wide range of research continuation possibilities.
+For instance, one of the major problems of the proposed algorithm is the inaccurate parametric estimation when two events are simultaneously active.
+Although it is a known problem in the literature, a successful solution in the given context remains still to be found.
+
+Another source of potential improvements is the refinement of the particle filter applied to this specific task.
+A deeper understanding of control theory, as well as collaborations with experts on the field, might bring a noticeable improvement on the overall system performance.
+
+The performance of the event classifier might be also further analyzed.
+Although in this case we opted for a classical feature-based machine learning approach, different methods and architectures could be compared, including more modern deep-learning approaches.\\
+
+Finally, the thesis contributions to more practical aspects are presented in Chapter~\ref{chap:data}.
+Those contributions comprise two software libraries written in Python: one of them focused on spherical microphone array and acoustic simulation, and another one implementing the SOFA standard, which has also been revised and modified for allowing the representation of ambisonic data.
+Finally, a software for the procedural creation of annotated reverberant ambisonic datasets has ben also presented.
+
+Apart from the straightforward task of software maintenance, the engagement of the research community with the usage and eventual contribution of the libraries would represent a desired situation in the near future.
+Moreover, the proposed file format convention is being currently discussed for its adoption into the SOFA standard, which can be considered an achievement of the original proposal.

### 2020-06-29 Chapters/Introduction.tex
beta1

@@
 \section{Scientific Objectives}
+\label{sec:objectives}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
 	\item To develop methods for the characterisation of acoustic parameters from recordings originated from ambisonic microphones. 
@@
 	\item To contribute in the generation and storage of ambisonic sound scenes, for their usage in controlled experimental environments. 
-\end{itemize}
-%\section{Goals}
-%
-%Research question: How can we exploit the characteristics of ambisonic recordings in order to extract signal-dependent, meaningful information from them?
+\end{enumerate}
 
@@
 Finally, as a result of the development of this thesis, the following open-source libraries have been implemented and released.
-All of them are available through the author's GitHub page\footnote{\url{https://github.com/andresperezlopez}}:
+All of them are available through the author's GitHub page \cite{andresgithub}:
 \begin{itemize}
@@
 
-	\item \href{https://github.com/andresperezlopez/ambiscaper}{ambiscaper}
+	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}
 
@@
 
-	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}
+	\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}
+
+	\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper}
 

### 2020-06-29 Chapters/SELD.tex
finished data chapter
beta1

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{SYSTEM DESCRIPTION}
+\section{System description}
 \label{sec:methodology}
@@
 
-\section{EXPERIMENTS}
+\section{Experiments}
 \label{sec:experiments}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{RESULTS}
+\section{Results}
 \label{sec:results}
@@
 This fact can help to explain the better performance of \textit{PAPAFIL2} over \textit{PAPAFIL1}: the temporal activities of the events in training and prediction are much more similar to each other in the former method, as a consequence of the training set generation approach.
+In order to provide a deeper insight on the temporal characteristics of the data under consideration, the durations of all sound events in the \textit{PAPAFIL2} training set are summarized in Figure~\ref{fig:eventdurations}.
+
+\begin{figure}[t]
+  \centering
+  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/event_durations.png}}
+  \caption{Event durations of all elements in the \textit{PAPAFIL2} training set.}
+  \label{fig:eventdurations}
+\end{figure}
+
 Furthermore, it is interesting to notice the high relevance of low-level features, and specifically several MFCC combinations (eight of the fifteen reported features) and  various extractors related to the spectral structure.

### 2020-06-29 manuscript.tex
beta1

@@
 Yet, many methods for acoustic and microphone array signal processing are not specifically tailored for spherical geometries.
-Therefore, there is still room for improvements in the field of automatic analysis and description of ambisonic recordings. 
+Therefore, there is still room for improvement in the field of automatic analysis and description of ambisonic recordings. 
 In the present thesis, we tackle this problem using methods based on the parametric analysis of the sound field. 
@@
 %PREFACE. 
-{\bf Preface \todo{is that really needed?}}
+{\bf Preface \todo{TODO}}
 
@@
 	Bibliography/Introduction.bib, 
-	Bibliography/ScientificBackground.bib}
+	Bibliography/ScientificBackground.bib,
+	Bibliography/Conclusion.bib}
 

### 2020-07-05 Chapters/Conclusions.tex
tbc

@@
 
-The scientific objectives of this thesis, as they were described in Section \ref{sec:objectives}, are:
+In this thesis we have presented our contributions to different components of an ambisonics analysis and generation framework, with a focus on reproducibility and portability to real-world scenarios.
+The main scientific objectives of this thesis, as they were described in Section \ref{sec:objectives}, are:
 

### 2020-07-05 Chapters/DataGeneration.tex
on work
tbc

@@
 \section{MASP: a Python library for multichannel acoustic signal processing}
-
+\label{sec:masp}
 \subsection{Description}
@@
 
-\textit{pysofaconventions} has gained a moderate amount of attraction.
+%\textit{pysofaconventions} has gained a moderate amount of attraction.
 Among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}.
@@
 \subsection{Experiment reproducibility}
+\label{experiment_reproducibility}
 
@@
 
+\section{Conclusion}
+
+The present Chapter has presented the main contributions of the Thesis regarding the availability of ambisonic data, specifically targeting software packages and libraries developed for the generation and storage of ambisonic sound scenes.
+
+The concept of \textit{research reproducibility} is directly related with these topics.
+According to \cite{cannam2012sound}, providing just an article as a research outcome would not be sufficient for reproducing the research in a feasible way.
+Instead, supplementary material, such as the code implementation, the dataset, and instructions on how to set up the environment, should be also conveniently provided.
+
+In this regard, the software discussed in this Chapter try to foster reproducible methods and procedures in several different ways.
+First, all code is freely available online with open source licenses. 
+Second, the contributed software is completely written in Python, which is itself an open-source project.
+Actually, all code used in this follows the same path, with with the exception of the particle tracker used in Chapter~\ref{chap:seld2019}.
+And third, the motivation of the Ambiscaper project itself is the reproducibility of experimental conditions for signal processing methods using ambisonic data, as explained in detail in Section~\ref{experiment_reproducibility}.\\
+
+Regarding the lifecycle of the libraries described in this Chapter, their status is diverse.
+Although \textit{pysofaconventions} has been recently published in an academic context (AES Convention in May 2020), it has been available online for more than a year.
+In fact, during that time it has gained a moderate amount of attraction: among others, it is being used as a dependency for the \textit{Real-Time Spherical Microphone Renderer (ReTiSAR) for binaural reproduction}, developed by Chalmers University in collaboration with Facebook Reality Labs \cite{helmholz2019real}.
+Furthermore, given the open nature of the project, several external researchers have contributed with bug fixes or new feature implementations.
+
+The \textit{masp} library has been also presented in the same context.
+However, in this case, the library has not been available beforehand, so the interest generated in the community has been moderate.
+
+For its part, \textit{Ambiscaper} has not been able to attract the community in a significant way.
+This might be due to a number of reasons, including not enough dissemination among potentially interested communities, or a lack of a paradigmatic use-case situation.
+On the contrary, the library that inspired this work, \textit{Scaper}, has been used for the creation of the dataset in Task 4 of DCASE 2019 and 2020 (sound event detection in domestic environments).
+This situation generates a feedback loop for the maintenance and improvement of the software, motivated by requests from the community.
+Given this situation, the future of \textit{Ambiscaper} is linked to the evolution of the research topics and tools in the communities which miggt be potentially interested in using it.
+
 %\subsection{Current state}

### 2020-07-05 Chapters/Introduction.tex
tbc

@@
 Nevertheless, the greatest contributor to the current interest in ambisonics has been the rise of Virtual Reality (VR) in recent years.
-Although VR focuses primarily on visual cues, the immersive experience might be greatly enhanced by spatial audio \cite{begault20003}	.
+Although VR focuses primarily on visual cues, the immersive experience can be greatly enhanced by spatial audio \cite{begault20003}.
 In this context, Ambisonics has been rapidly adopted as \textit{de facto} standard for spatial audio transmission, supposedly due to a variety of factors:
@@
 
-  \item [Recording device independence] Regardless of the specific characteristics of an ambisonic microphone, the recorded signal is usually converted into B-Format, which effectively represents the standard exchange format.
+  \item [Recording device independence] Regardless of the specific characteristics of an ambisonic microphone, the recorded signal is usually converted into B-Format, which is effectively  the standard exchange format.
 
@@
 
-Coming back to the Virtual Reality popularity, the following three events might be representative of the growth that it experienced in the second half of the 2010s: 
-(i) the billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}; 
-(ii) the Time cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
-and (ii) M.
+Coming back to the issue of the popularity of Virtual Reality, we have selected three events that might epitomize the growth undergone in the second half of the 2010s:
+
+%Coming back to the Virtual Reality popularity, the following three events might be representative of the growth experienced in the second half of the 2010s: 
+\begin{enumerate}
+	\item The billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}
+	\item \textit{Time} maagazine cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
+	\item  M.
+Zuckerberg's invited talk at the \textit{Samsung Unpacked} event within the World Mobile Congress 2016 in Barcelona: \textit{"VR is the next platform where anyone can experience anything they want"} \cite{bbcnews}.
+\end{enumerate}
+
+%(i) the billionaire acquisition by Facebook of the VR headset manufacturer Oculus, in March 2014 \cite{facebookoculus}; 
+%(ii) the Time cover page devoted to VR: \textit{"The surprising joy of Virtual Reality. And why it's about to change the world"} (August 2015) \cite{time};
+%and (iii) M.
 Zuckerberg's invited talk at the \textit{Samsung Unpacked} event within the World Mobile Congress 2016 in Barcelona: \textit{"VR is the next platform where anyone can experience anything they want"} \cite{bbcnews}.
@@
   \caption{Number of ambisonic microphones released in last years (from \cite{List_of_Ambisonic_hardware}).
-From left to right, the vertical lines correspond to (i) Oculus acquisition, (ii) Time cover page on VR, (iii) M.
-Zuckerberg's speech in MWC, and (iv) Jaunt announcement of shift towards AR.}
+From left to right, the vertical lines correspond to (1) Oculus acquisition, (2) \textit{Time} cover page on VR, (3) M.
+Zuckerberg's speech in MWC, and (4) Jaunt announcement of shift towards AR.}
   \label{fig:nummics}
@@
 
-At the present time, however, the high expectations in VR have significantly lowered, as shown in Figure~\ref{fig:vrfunding}.
+At the present time, however, the high expectations put into VR have significantly lowered, as shown in Figure~\ref{fig:vrfunding}.
 This is due to a variety of reasons, including lack of interesting content and the high production cost of the headsets \cite{fortune}.
@@
 
-B-Format analysis has been divided into three different categories:
+Ambisonic signal analysis has been divided into three different categories:
 \begin{description}
+	\item[Signal Enhancement] Modification of the input signal in order to obtain one or more output signals with desired attributes.
+
+	There are a variety of well known signal enhancement problems, including dereverberation \cite{braun2018speech}, source separation \cite{gannot2017consolidated}, or foreground-background segmentation \cite{stefanakis_foreground_2015}.
+	As it has been shown, many of them benefit (or even depend) from the knowledge derived by acoustic parameter estimation methods.
+
 	\item[Acoustic Parameter Estimation] Low-level analysis of the sound field, which yields quantitative information about different acoustic parameters used to model the acoustic scene.
-	\item[Signal Enhancement] Modification of the sound field in order to obtain a set of output signals with desired attributes. 
-	\item[Scene Description] Textual representation of different high-level characteristics of the sound field.
-\end{description}
 
@@
 
-There are a variety of well known signal enhancement problems, including dereverberation \cite{braun2018speech}, source separation \cite{gannot2017consolidated}, or foreground-background segmentation \cite{stefanakis_foreground_2015}.
-As it has been shown, many of them benefit (or even depend) from the knowledge derived by the acoustic parameter estimation step.
+	\item[Scene Description] Textual representation of different high-level characteristics of the sound field.
 
-Finally, under the scene description typology we can find a set of applications that provide abstract representations of the sound scene under analysis.
-Most of the recent research performed in this scope is grouped around the Detection and Classification of Acoustic Scenes and Events (DCASE) community\footnote{\url{http://dcase.community/community_info}}.
+	Under the scene description typology we can find a set of applications that provide abstract representations of the sound scene under analysis.
+Most of the recent research performed in this scope is grouped around the Detection and Classification of Acoustic Scenes and Events (DCASE) community \cite{dcase}.
 	Examples of the problems under consideration are acoustic scene classification \cite{Mesaros2018_DCASE}, sound event localization and detection \cite{Adavanne2018_JSTSP} or audio captioning \cite{drossos2017automated}.
+\end{description}
 
@@
 	\item To propose methodologies for sound event localization and detection in ambisonic domain which are grounded on spatial parametric analysis. 
-	\item To contribute in the generation and storage of ambisonic sound scenes, for their usage in controlled experimental environments. 
+	\item To contribute to the generation and storage of ambisonic sound scenes, for their usage in controlled experimental environments. 
 \end{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Contributions}
-
-In the following list we show the main scientific contributions of this dissertation, organised by Chapters:
-
-\begin{itemize}
-
-	\item Chapter 3:\\
-	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
-	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
-	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}
-
-	\item Chapter 4:\\
-	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
-	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
-	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
-
-	\item Chapter 5:\\
-
-	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
-	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
-	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
-
-	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
-	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
-	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
-
-	\item Chapter 6:\\
-	\textbf{"Ambiscaper: A Tool for Automatic Generation and Annotation of Reverberant Ambisonics Sound Scenes".}\\
-	\underline{A. P\'erez-L\'opez}.
-	In \textit{16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
-
-	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
-	\underline{A. P\'erez-L\'opez} and J. De Muynke.
-	In \textit{Audio Engineering Society Convention 144. Audio Engineering Society, 2018.}
-
-	\textbf{"pysofaconventions, a Python API for SOFA".}
-	\underline{A. P\'erez-L\'opez}.
-	In \textit{Audio Engineering Society Convention 148. Audio Engineering Society, 2020.}
-
-	\textbf{"A Python library for Multichannel Acoustic Signal Processing".}
-	\underline{A. P\'erez-L\'opez} and A. Politis.
-	In \textit{Audio Engineering Society Convention 148. Audio Engineering Society, 2020.}
-
-\end{itemize}
-
-Moreover, although not strictly aligned with the research direction of this thesis, the author has collaborated in the following publications:
-
-\begin{itemize}
-	\item \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
-	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
-	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
-
-\end{itemize}
-
-Finally, as a result of the development of this thesis, the following open-source libraries have been implemented and released.
+%\section{Contributions}
+%
+%
+%In the following list we show the main scientific contributions of this dissertation, organised by Chapters:
+%
+%\begin{itemize}
+%
+%	\item Chapter 3:\\
+%	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
+%	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
+%	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}
+%	
+%	\item Chapter 4:\\
+%	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
+%	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
+%	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
+%
+%	\item Chapter 5:\\
+%	
+%	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
+%	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
+%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
+%	
+%	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
+%	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
+%	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+%	
+%	
+%	\item Chapter 6:\\
+%	\textbf{"Ambiscaper: A tool for automatic generation and annotation of reverberant ambisonics sound scenes".}\\
+%	\underline{A. P\'erez-L\'opez}.
+%	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
+%	
+%	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
+%	\underline{A. P\'erez-L\'opez} and J. De Muynke.
+%	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}
+%	
+%	\textbf{"pysofaconventions, a Python API for SOFA".}
+%	\underline{A. P\'erez-L\'opez}.
+%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}
+%	
+%	\textbf{"A Python library for multichannel acoustic signal processing".}
+%	\underline{A. P\'erez-L\'opez} and A. Politis.
+%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}
+%	
+%\end{itemize}
+%
+%Moreover, although not strictly aligned with the research direction of this thesis, the author has collaborated in the following publications:
+%
+%\begin{itemize}
+%	\item \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
+%	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
+%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.
+%	
+%\end{itemize}
+%
+%
+%Finally, as a result of the development of this thesis, the following open-source libraries have been implemented and released.
 All of them are available through the author's GitHub page \cite{andresgithub}:
-\begin{itemize}
-
-	\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation}
-
-	\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020}
-
-	\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3}
-
-	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}
-
-	\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions}
-
-	\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}
-
-	\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper}
-
-\end{itemize}
+%\begin{itemize}
+%
+%	\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation}
+%	
+%	\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020}
+%	
+%	\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3}
+%	
+%	\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library}
+%
+%	\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions}
+%	
+%	\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}	
+%
+%	\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper} 
+%	
+%
+%
+%\end{itemize}

### 2020-07-05 Chapters/ReververationTimeEstimation.tex
tbc

@@
 
-We have presented in this work a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings.
+I n this Chapter, we have presented  a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings.
 Our method is based on a first dereverberation step,  performed by a multichannel autoregressive model of the late reverberation.

### 2020-07-05 Chapters/SELD.tex
tbc

@@
   \centering
-  \centerline{\includegraphics[width=0.75\textwidth]{Figures/SELD/ARCH.png}}
+  \centerline{\includegraphics[width=0.55\textwidth]{Figures/SELD/ARCH.png}}
   \caption{Architecture of the proposed methodology.}
@@
 
-Table~\ref{tab:results} summarizes the results of the experiments using the recommended data split: training with folds 3 to 6, validation with fold 2 and testing with fold 1. 
-% This structure has been promoted by the Challenge organization as a fair way for method comparison. 
-% Furthermore, the usage of an unseen split 
-% Furthermore, it is expected to provide similar results to the evaluation set, given that part of the data remains unseen at training.
+\begin{table}[t]
+\begin{footnotesize}
+\caption{System evaluation. Top: results on the cross-validation development set. Bottom: results on the evaluation set.}
+    \begin{center}
+    \begin{tabular}{cccccc}
+    \toprule
+    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
+    \midrule
+    \textit{BASELINE} & 0.70   & 39.5 \% & 23.2$^{\circ}$ & \textbf{62.1} \% & 0.45 \\
+    \textit{PAPAFIL1} & 0.60 & 49.8 \% & \textbf{13.4}$^{\circ}$ & 54.4 \% &0.41\\
+    \textit{PAPAFIL2} & \textbf{0.57} & \textbf{54.0} \% &13.8$^{\circ}$ & 59.7 \% & \textbf{0.38} \\
+
+    % this is oracle beam
+    \textit{PAPAFIL1-O} & 0.37 & 67.0 \% & 2.0$^{\circ}$ & 68.6 \% & 0.26 \\
+    \textit{PAPAFIL2-O} & 0.32 & 79.6 \% & 8.5$^{\circ}$ & 82.4\% & 0.19 \\
+	\midrule
+	 \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & 60.7 \% & 0.47 \\
+    \textit{PAPAFIL1} & 0.55 & 56 \% & 12.8$^{\circ}$ & 61.1 \% &0.36\\
+    \textit{PAPAFIL2} & \textbf{0.51} & \textbf{60.1} \% &\textbf{12.4}$^{\circ}$ & \textbf{65.1} \% & \textbf{ 0.33} \\
+
+    \bottomrule
+    \end{tabular}
+    \end{center}
+    \label{tab:results}
+\end{footnotesize}
+\end{table}
+
+Table~\ref{tab:results} summarizes the results of the experiments for development and evaluation datasets.
+Results on the development set have been computed using the provided cross-validation scheme: split 1 for testing, split 2 for validation, and splits 3 to 6 for training.
+
 Results are reported for three different systems: the baseline and the two proposed methods \textit{PAPAFIL1} and \textit{PAPAFIL2}.
-The results of their respective oracle results, \textit{PAPAFIL1-O} and \textit{PAPAFIL2-O}, are also provided.
+The results of their respective oracle results, \textit{PAPAFIL1-O} and \textit{PAPAFIL2-O}, are also provided for the development set.
 
-Both proposed approaches outperform the baseline system in three out of the four evaluation metrics ($\text{ER}_{20}$, $\text{F}_{20}$ and $\text{LE}_{CD}$). 
+Regarding the development set, both proposed approaches outperform the baseline system in three out of the four evaluation metrics ($\text{ER}_{20}$, $\text{F}_{20}$ and $\text{LE}_{CD}$). 
 Although the results obtained by both of them are similar, \textit{PAPAFIL2} obtains better classification scores ($\text{ER}_{20}$ and $\text{F}_{20}$), and \textit{PAPAFIL1} performs subtly better regarding localization error ($\text{LE}_{CD}$).
@@
 In turn, \textit{PAPAFIL2-O} performs better than \textit{PAPAFIL1-O} regarding all metrics, excepting $\text{LE}_{CD}$; this improvement is specially noticeable in $\text{LR}_{CD}$, with a performance difference of about 15\%.
-The good results obtained by \textit{PAPAFIL2-O} validate the proposed particle filtering approach, an leave space for improvements that might be given by a better understanding and fine tuning of the model.
+
+The good results obtained by \textit{PAPAFIL2-O} validate the proposed particle filtering approach, and leave space for improvements that might be given by a better understanding and fine tuning of the model.
+
+The overall tendency is maintained in the evaluation set results.
+Both proposed methods outperform again the baseline, improving the development set SELD score in five points; since the baseline slightly decreases the performance for the evaluation set, the score difference with respect to the proposed methods diverges significantly.
+\textit{PAPAFIL2} is the method that clearly performs better on the evaluation set, with better results than \textit{PAPAFIL1} in all evaluation metrics. \\
 
@@
 % Although the performance is good in general terms with both \textit{PAPAFIL} approaches, the results deteriorate noticeably with overlapping sounds.
-A closer inspection reveals that, in many occasions, the TF bins passing the single-source test mostly belong to one out of two simultaneous sources. 
+A closer inspection to the development set results reveals that, in many occasions, the TF bins passing the single-source test mostly belong to one out of two simultaneous sources. 
 It is a known issue that performance of DirAC diffuseness is reduced when two sources are present \cite{epain2016spherical}; similar problems have been reported in \cite{adavanne2019localization}, where an instantaneous source number estimator is used in combination with the particle filter.
@@
 
-\begin{table}[th!]
-\begin{footnotesize}
-\caption{Evaluation results on the development set.}
-    \begin{center}
-    \begin{tabular}{cccccc}
-    \toprule
-    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
-    \midrule
-    \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & \textbf{60.7} \% & 0.47 \\
-    \textit{PAPAFIL1} & 0.60 & 49.8 \% & \textbf{13.4}$^{\circ}$ & 54.4 \% &0.41\\
-
-    \textit{PAPAFIL2} & \textbf{0.57} & \textbf{54.0} \% &13.8$^{\circ}$ & 59.7 \% & \textbf{ 0.38} \\
-
-    \midrule
-    % this is oracle beam
-    \textit{PAPAFIL1-O} & 0.37 & 67.0 \% & 2.0$^{\circ}$ & 68.6 \% & 0.26 \\
-    \textit{PAPAFIL2-O} & 0.32 & 79.6 \% & 8.5$^{\circ}$ & 82.4\% & 0.19 \\
-
+%\begin{table}[th!]
+%\begin{footnotesize}
+%\caption{Evaluation results on the cross-validation development set.}
+%    \begin{center}
+%    \begin{tabular}{cccccc}
+%    \toprule
+%    Method   & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
 %    \midrule
-    % \textit{PAPAFIL1} &  0.57 & 55.6 \% & 15.6$^{\circ}$ & 66.7 \% & 0.36 \\
-    % % PAPAFIL_1 & 0.47  & 64.1 \% & 14.1$^{\circ}$ & 70.1 \% & 0.30 \\ ?????????
-    % % this is oracle beam
-    % \textit{PAPAFIL1-O} & 0.08 & 93.7 \% &  0.2$^{\circ}$ & 94.0 \% & 0.05 \\
-
-    %  \textit{PAPAFIL2} & \textbf{ 0.44} & \textbf{68.0} \% & \textbf{13.3}$^{\circ}$ & \textbf{79.6} \% & \textbf{0.26} \\
-
-    % \textit{PAPAFIL2-O} & 0.41 & 71.1 \% & 12.3$^{\circ}$ & 82.0\% &  0.24 \\
-    \bottomrule
-    \end{tabular}
-    \end{center}
-    \label{tab:results}
-\end{footnotesize}
-\end{table}
+%    \textit{BASELINE} & 0.72   & 37.4 \% & 22.8$^{\circ}$ & \textbf{60.7} \% & 0.47 \\
+%    \textit{PAPAFIL1} & 0.55 & 56 \% & 12.8$^{\circ}$ & 61.1 \% &0.36\\
+%    \textit{PAPAFIL2} & \textbf{0.51} & \textbf{60.1} \% &\textbf{12,4}$^{\circ}$ & 65.1 \% & \textbf{ 0.33} \\
+%    \bottomrule
+%    \end{tabular}
+%    \end{center}
+%    \label{tab:results}
+%\end{footnotesize}
+%\end{table}
 
@@
 
-To conclude the analysis, Fig.~\ref{fig:importance} shows the relative importance of the fifteen most relevant acoustic features for the \textit{PAPAFIL2} classifier model.
+Fig.~\ref{fig:importance} shows the relative importance of the fifteen most relevant acoustic features for the \textit{PAPAFIL2} classifier model.
 Event duration is clearly the feature with the highest importance, and effective duration (duration of the signal discarding silence) also appears in the eighth position.
@@
 The absence of pitch, harmonic and envelope features in the list represents a significant finding as well.
+\\
 
-\todo{TODO: DISCUSS RESULTS OF EVALUATION SET!!}
+The results of the DCASE 2020 Challenge for the SELD task are shown in Table~\ref{tab:dcase2020results}.
+The method entries corresponds to the best scoring method for each team submission, respecting the naming conventions established for the Challenge.
+Accordingly, \textit{PerezLopez} refers there to the \textit{PAPAFIL2} method.
+The final rank is computed as the cumulative rank across all evaluation metrics, sorted by ascending order.
+The data has been taken from the Task results website \cite{dcasetask}, which provides complete information regarding the challenge.
+
+\begin{table}[t]
+\begin{footnotesize}
+\caption{DCASE 2020 Challenge Task 3 evaluation results}
+    \begin{center}
+
+\begin{tabular}{llccccc}
+    \toprule
+Rank       & Method  & $\text{ER}_{20}$ & $\text{F}_{20}$   & $\text{LE}_{CD}$ & $\text{LR}_{CD}$ & $\text{SELD}$ \\
+    \midrule
+\textbf{1} & Du & \textbf{0.20}  & \textbf{84.9} \%    & \textbf{6.0}$^{\circ}$   & 88.5 \%    &  \textbf{0.12 }    \\
+2          & Nguyen          & 0.23  & 82.0 \%    & 9.3$^{\circ}$   &\textbf{ 90.0} \%    &  0.14     \\
+3          & Shimada         & 0.25  & 83.2 \%    & 7.0$^{\circ}$   & 86.2 \%    &  0.15     \\
+4          & Cao             & 0.36  & 71.2 \%    & 13.3$^{\circ}$  & 81.1 \%    &  0.22     \\
+5          & Park            & 0.43  & 65.2 \%    & 16.8$^{\circ}$  & 81.9 \%    &  0.26     \\
+6          & Phan            & 0.49  & 61.7 \%    & 15.2$^{\circ}$  & 72.4 \%    &  0.30     \\
+7          & PerezLopez      & 0.51  & 60.1 \%    & 12.4$^{\circ}$  & 65.1 \%    &  0.33     \\
+8          & Sampathkumar    & 0.53  & 56.6 \%    & 14.8$^{\circ}$  & 66.5 \%    &  0.35     \\
+9          & Patel           & 0.55  & 55.5 \%    & 14.4$^{\circ}$  & 65.5 \%    &  0.38     \\
+10         & Ronchini        & 0.58  & 50.8 \%    & 16.9$^{\circ}$  & 65.5 \%    &  0.39     \\
+11         & Naranjo-Alcazar & 0.61  & 49.1 \%    & 19.5$^{\circ}$  & 67.1 \%    &  0.38     \\
+12         & Song            & 0.57  & 50.4 \%    & 20.0$^{\circ}$  & 64.3 \%    &  0.38     \\
+13         & Tian            & 0.64  & 47.6 \%    & 24.5$^{\circ}$  & 67.5 \%    &  0.40     \\
+14         & Singla          & 0.88  & 18.0 \%    & 53.4$^{\circ}$  & 66.2 \%    &  0.58     \\
+15         & Baseline        & 0.69  & 41.3 \%    & 23.1$^{\circ}$  & 62.4 \%    &  0.45     \\
+\bottomrule      
+\end{tabular}
+\end{center}
+    \label{tab:dcase2020results}
+    \end{footnotesize}
+\end{table}
+
+It is important to remark that, apart from our proposed method, all other systems rely completely on deep learning algorithms, specially making use of different configurations and combinations of CRNNs.
+The only exception to that tendency is the method presented by Nguyen and colleagues, which is an adaptation of their recently proposed methodology \cite{Nguyen2020_task3_report, nguyen2020sequence}.
+In this method, localization is also based on the identification of the single-source TF bins; instead of DirAC diffuseness, they use a subspace analysis to find the TF bins associated with nearly rank-1 spatial covariance matrices (SCM).
+Furthermore, the instantaneous DOA estimates are obtained by finding the azimuth and elevation histogram peaks, with the steering directions obtained from the covariance matrix diagonals. 
+While this method might contribute to provide a smaller localization error compared with our approach (9.3$^{\circ}$ versus 14.2$^{\circ}$ in $\text{LE}_{CD}$), it is more expense computationally, since it involves building a SCM for each TF bin, plus peak picking in two 1D histograms (one for each spherical coordinate).
+However, the $\text{LE}_{CD}$ results are also influenced by the classification accuracy; a deeper analysis would be required to perform a direct comparison of both approaches.
+
+Table~\ref{tab:dcase2020results} also highlights the good result obtained by our method regarding localization error: it is the fourth best method with respect to $\text{LE}_{CD}$. 
+Moreover, the two model-based localization methods (our system and Nguyen) rank fourth and second in the $\text{LE}_{CD}$ classification, respectively. 
+This result suggests that parametric spatial audio analysis, combined with a robust tracking system, is able to produce state-of-the-art localization results in the context of SELD. \\
+
+To conclude the analysis, Figure~\ref{fig:importance} shows the relationship between the system complexity, measured in number of parameters of the model, against $\text{SELD}$ score, for all submitted systems, denoted by their ranking position.
+The overall tendency is clear: the higher the number of parameters, the better (smaller) the $\text{SELD}$ score; the regression curve, computed from all observations, confirms such tendency. 
+It must be noticed that the presented method (\textit{7} in the Figure) lies outside the 95\% confidence interval of the regression curve (shaded area).
+In other words, its $\text{SELD}$ score is significantly better than the tendency observed from all submissions, considering the system complexity.
+Figure~\ref{fig:importance} also highlights the great difference in complexity among submissions.
+Between our method (the less complex, with 20k parameters) and the first scoring method (the most complex, with 123M parameters) there are four magnitude orders of different, constituting a remarkable difference.
+The logarithmic mean of all submitted systems' complexities is around 2.3M parameters.
+
+\begin{figure}[t]
+  \centering
+  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/seld_complexity_v4.png}}
+  \caption{DCASE 2020 Task 3 submissions: complexity versus SELD score.}
+  \label{fig:importance}
+\end{figure}
 
@@
 Specifically, our method is able to improve the baseline SELD score by almost ten points, while increasing the scores in three out of the four metrics under consideration.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section{Introduction}
-%\label{sec:intro}
-%
-%Sound Event Localization and Detection (SELD) refers to the problem of identifying, for each individual  event present in a sound field, the temporal activity, spatial location, and sound class to which it belongs.
-SELD is a current research topic which deals with microphone array processing and sound classification, with potential applications in the fields of signal enhancement, autonomous navigation, acoustic scene description or surveillance, among others.
-%
-%SELD arises from the combination of two different problems: Sound Event Detection (SED) and Direction of Arrival (DOA) estimation.
-The number of works in the literature which jointly address SED and DOA problems is relatively small.
-It is possible to classify them by the type of microphone arrays used: distributed \cite{grobler2017sound, butko2011two, chakraborty2014sound} or near-coincident \cite{hirvonen2015classification, lopatka2016detection, Adavanne2018_JSTSP}.
-%As mentioned in \cite{Adavanne2018_JSTSP}, the usage of near-coincident circular/spherical arrays enables the representation of the sound field in the spatial domain, using the spherical harmonic decomposition, also known as Ambisonics \cite{gerzon1973periphony, daniel2000representation}.
-Such spatial representation allows a flexible, device-independent comparison between methods.
-Furthermore, the number of commercially available ambisonic microphones has increased in recent years due to their suitability for immersive multimedia applications.
-%Taking advantage of the compact spatial representation provided by the spherical harmonic decomposition, several methods for parametric analysis of the sound field in the ambisonic domain have been proposed  \cite{pulkki2006directional, berge2010high, Politis2018, pulkki2018parametric}.
-%These methods ease sound field segmentation into direct and diffuse components, and further localization of the direct sounds.
-%The advent of deep learning techniques for DOA estimation has also improved the results of traditional methods \cite{Adavanne2018_JSTSP}.
-However, none of the deep learning-based DOA estimation methods explicitly exploits the spatial parametric analysis.
-This situation is further extended to the SELD problem, with the exception of \cite{lopatka2016detection}, where DOAs are estimated from the \textit{active intensity vector}  \cite{pulkki2006directional}.
-%
-%The motivation for the proposed methodology is two-fold.
-%First, we would like to check whether the usage of spatial parametric analysis in the ambisonic domain can improve the performance of SELD algorithms.
-%Second, temporal information derived by the parametric analysis could be further exploited to estimate event onsets and offsets, thus lightening the event classifier complexity; such reduction might positively impact algorithm's performance.
-%
-%In what follows, we present the methodology and the architecture of the proposed system (Section~\ref{sec:method}).
-Then, we describe the design choices and the experimental setup (Section~\ref{sec:expe}), and discuss the results in the context of DCASE2019 Challenge - Task 3 (Section~\ref{sec:results}).
-A summary is presented in Section~\ref{sec:conclusion}.
-%%In order to support open access and reproducibility, all code is freely available at \cite{code}. \todo{code ref}.
-%
-%
-%\section{Method}
-%\label{sec:method}
-%
-%The proposed method presents a solution for the SELD problem splitting the task into four different problems: \textit{DOA  estimation}, \textit{association}, \textit{beamforming} and \textit{classification}, which will be described in the following subsections.
-%The former three systems follow a heuristic approach---in what follows, they will be jointly referred to as the \textit{parametric front-end}.
-Conversely, the \textit{classification} system is data-driven, and will be referred to as the \textit{deep learning back-end}.
-The method architecture is depicted in Figure~\ref{fig:diagram_all}.
-%
-%\todo{redo figures in this section}
-%
-%\begin{figure}[h]
-%	\centering
-%    \includegraphics[width=\columnwidth]{Figures//SELD/dcase_challenge_tech_report_all2.pdf}
-%    \caption{System architecture.}
-%    \label{fig:diagram_all}
-%
-%\end{figure}
-%
-%\subsection{DOA estimation}
-%\label{ssec:doa_estimation}
-%
-%\begin{figure*}[h]
-%
-%    \includegraphics[width=\textwidth]{Figures/SELD/DOA1.pdf}
-%    \caption{DOA estimation architecture.}
-%    \label{fig:doa}
-%
-%\end{figure*}
-%
-%
-%
-%The \textit{DOA estimation} system (Figure~\ref{fig:doa}) is based on parametric time-frequency spatial audio analysis.
-%
-%Let us consider a \textit{N3D}-normalized first-order ambisonic signal $x_n^m(t)$. 
-%Using the DirAC parametric model, it is possible to obtain instantaneous time-frequency estimates of the Direction of Arrival $\Omega(k,n)$ (Eq.~\ref{eq:doa}), energy density $E(k,n)$ (Eq.~\ref{eq:energydensity}) and the diffuseness $\Psi(k,n)$ (Eq.~\ref{eq:psidefinition}).\\
-%
-%
-%
-%%Let us consider a first order ambisonic signal $$
-%%Let us consider a first order ambisonic signal vector $\pmb{b}(t)$ with N3D normalization \cite{carpentier2017normalization}:
-%%\begin{equation}
-%%    \pmb{b}(t) = [b_w(t), \sqrt{3}b_x(t),\sqrt{3}b_y(t), \sqrt{3}b_z(t)].
-%%\end{equation}
-%%From its short-time frequency domain representation $\pmb{B}(k,n)$, the instantaneous DOA at each TF bin $\pmb{\Omega}(k,n)$ can be estimated as:
-%%\begin{equation}
-%%    \begin{aligned}
-%%    \pmb{I}(k,n) = -\frac{1}{Z_0}\mathbb{R}\{& [B_x(k,n), B_y(k,n), B_z(k,n)]  B_w(k,n)^* \},\\
-%%    \pmb{\Omega}(k,n) &= [\varphi(k,n), \theta (k,n)] =\angle ( -\pmb{I}(k,n) ),
-%%    \end{aligned}
-%%\end{equation}
-%%where $\pmb{I}(k,n)$ stands for the \textit{active intensity vector} \cite{pulkki2006directional}, $Z_0$ is the characteristic impedance of the medium, $^*$ represents the complex conjugate operator, and $\angle$ is the spherical coordinates angle operator, expressed in terms of azimuth $\varphi$ and elevation $\theta$.
-%
-%It is desirable to identify the TF regions of $\Omega(k,n)$ which carry information from the sound events, and discard the rest. Three binary masks are computed with that aim.
-%
-%The first mask is the \textit{energy density mask}, which is used as an activity detector.
-A gaussian adaptive thresholding algorithm is applied to $E(k,n)$.
-This procedure selects TF bins with local maximum energy density, as expected from the direct path of the sources.
-%
-%The \textit{diffuseness mask} selects the TF bins with high energy propagation.
-Bins with a low diffuseness value represent a TF region where one sound source is predominant.
-A fixed threshold value $\Psi_{max}$ is used for the masking.
-%
-%The third mask is the \textit{DOA variance mask}.
-It tries to select TF regions with small standard deviation with respect to their neighbor bins---a characteristic of sound fields with low diffuseness \cite{pulkki2018parametric}.\\
-%
-%The three masks are then applied to the DOA estimation, obtaining the TF-filtered instanteneous DOAs $\bar{\Omega}(k,n)$. 
-%Finally, a median filter is applied, with the aim of improving DOA estimation consistency and removing spurious TF bins.
-%The median filter is applied in a TF bin belonging to $\bar{\Omega}(k,n)$ only if the number of TF bins belonging to $\bar{\\Omega}(k,n)$ in its vicinity is greater than a given threshold $B_{min}$.
-%The resulting filtered DOA estimation is referred to as $\breve{\Omega}(k,n)$.
-%
-%
-%\subsection{Association}
-%\label{ssec:association}
-%
-%\begin{figure*}[h]
-%    \includegraphics[width=\textwidth]{Figures/SELD//ASS.pdf}
-%    \caption{Association architecture.}
-%    \label{fig:association}
-%
-%\end{figure*}
-%
-%The association step (Figure~\ref{fig:association}) tackles the problem of assigning the time-frequency-space observation $\breve{\Omega}(k,n)$ to a set of events, each one having a specific onset, offset and location.
-\\
-%
-%First, DOA estimates are resampled into \textit{frames} with the task's required length ($0.02$ s). In what follows, frames will be represented by index $m$. 
-%An additional constraint is applied: for a given window $n_0$, the DOA estimates $\breve{\Omega}(k,n_0)$ are assigned to the corresponding frame $m_0$ only if the number of estimates is higher than a given threshold $K_{min}$.\\
-%
-%Next, the standard deviation in azimuth $\sigma_{\varphi}$ and elevation $\sigma_{\theta}$ of the frame-based DOA estimates $\breve{\Omega}(k,m)$ are compared to a threshold value $\sigma_{max}$, and the result is used to estimate the frame-based event overlapping amount $o(m)$ :
-%
-%\begin{equation}
-%\begin{aligned}
-%	o(m) =  \begin{cases}
-%    		1, &\text{if } \sigma_{\varphi}/2 + \sigma_{\theta} < \sigma_{max},\\
-%    		2, &\text{otherwise}.
-%    \end{cases}    		
-%\end{aligned}
-%\end{equation}
-%
-%The clustered values $\Omega_{\text{cluster}}(m)$ are then computed as the $K=o(m)$ centroids of $\breve{\Omega}(k,m)$, using a modified version of K-Means which minimizes the central angle distance.
-Notice that, for $o(m)=1$, the operation is equivalent to the median.\\
-%
-%The following step is the grouping of clustered DOA values into events.
-%Let us define $\Omega_S(m)$ as the frame-wise DOA estimations belonging to the event $S$.
-A given clustered DOA estimation $\Omega_{\text{cluster}}(m)$ belongs to the event $S$ if the following criteria are met:
-%\begin{itemize}
-%    \item The central angle between  $\Omega_{\text{cluster}}(m)$ and the median of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{ANGLE}$, and
-%    \item The frame distance between $m$ and the closest frame of $\Omega_S(m)$ is smaller than a given threshold $d_{max}^\text{FRAME}$.
-%\end{itemize}
-%
-%The resulting DOAs $\Omega_S(m)$ are subject to a postprocessing step with the purpose of 
-%delaying event onsets in frames where $o(m) > 2$, and discarding events shorter than a given minimum length. 
-%
-%Finally, the frame-based event estimations are converted into \textit{metadata annotations} in the form $\Lambda_S = (\Omega_S, \text{onset}_S, \text{offset}_S)$.
-%
-%\subsection{Beamforming}
-%\label{ssec:Beamforming}
-%
-%The last step performed in the front-end is the input signal segmentation.
-The spatial and temporal information provided by the annotations $\Lambda_S$ are used to produce monophonic signal estimations of the events, $\tilde{x}_S(t)$, as the signals captured by a virtual first-order cardioid (Eq.~\ref{eq:decodingequation}):
-%\begin{equation}
-%\tilde{x}_S(t) = x_n^m(t) Y_n^m(\Omega_S) \alpha_n,
-%\end{equation}
-%with $\alpha_n = [1, 1, 1, 1]$ using the \textit{basic} decoding weights.
-%
-%
-%
-%\subsection{Deep learning classification back-end}
-%\label{ssec:backend_method}
-%
-%%intro - problem formulation==============
-%The parametric front-end performs DOA estimation, temporal activity detection and time/space segmentation, and produces monophonic estimations of the events, $\tilde{s}_S(t)$.
-%Then, the back-end classifies the resulting signals as belonging to one of a target set of 11 classes.
-%Therefore, the multi-task nature of the front-end allows us to define the back-end classification task as a simple multi-class problem, even though the original SELD task is multi-label.
-%It must be noted, however, that due to the limited directivity of the first-order beamformer, the resulting monophonic signals can present a certain leakage from additional sound sources when two events overlap, even when the annotations $\Lambda_S$ are perfectly estimated.
-%
-%%2 stages: context
-%The classification method is divided into two stages.
-%First, the incoming signal is transformed into the log-mel spectrogram and split into TF patches.
-%Then, the TF patches are fed into a single-mode based on a Convolutional Recurrent Neural Network (CRNN), which outputs probabilities for event classes $c \in \left\lbrace 1 ...C \right\rbrace$, with $C=11$.
-%Predictions are done at the event-level (not at the frame level), since the temporal activities have been already determined by the front-end.
-%Only one label is predicted for each incoming event, such that there is no binarization stage needed as in a standard SED task.
-%
-%%archi=============
-%The proposed CRNN is depicted in Figure \ref{fig:archi}.
-%It presents three convolutional \textit{blocks} to extract local features from the input representation.
-%Each convolutional block consists of one convolutional layer, after which the resulting feature maps are passed through a ReLU non-linearity \cite{nair2010rectified}.
-%This is followed by a max-pooling operation to downsample the feature maps and add invariance along the frequency dimension.
-%The target classes vary to a large extent in terms of their temporal dynamics, with some of them being rather impulsive (e.g., \textit{Door slam}), while others being more stationary (e.g., \textit{Phone ringing}).
-%Therefore, after stacking the feature maps resulting from the convolutional blocks, this representation is fed into one bidirectional recurrent layer in order to model discriminative temporal structures.
-%Specifically, 64 nodes of gated recurrent units (GRU) are used with \textit{tanh} activations.
-%The recurrent layer is followed by a Fully Connected (FC) layer, and finally a 11-way softmax classifier layer produces the event-level probabilities.
-%Dropout is applied extensively.
-%The loss function used is categorical cross-entropy.
-%The model has $\sim$175k weights.
-%
-%\begin{figure}[ht]
-%  \vspace{-3mm}
-%  \centering
-%  \centerline{\includegraphics[width=0.9\textwidth]{Figures/SELD/DCASE19Task3_backend_archi_v3.png}}
-%  \vspace{-2mm}
-%  \caption{Back-end architecture.}
-%  \label{fig:archi}
-%    \vspace{-5mm}
-%\end{figure}
-%
-%
-%
-%
-%%===============================================================================
-%
-%\section{Experiments}
-%\label{sec:expe}
-%
-%\subsection{Dataset, evaluation metrics and baseline system}
-%\label{ssec:dataset}
-%We use the TAU Spatial Sound Events 2019 - Ambisonic, which provides first-order ambisonic recordings.
-%Details about the recording format and dataset specifications can be found in \cite{Adavanne2019_DCASE}.
-%The dataset features a vocabulary of $C=11$ classes encompassing human sounds and sound events typically found in indoor office environments.
-%The dataset is split into a development and evaluation sets. 
-%The development set consists of a four fold cross-validation setup.\\
-%
-%The SELD task is evaluated with individual metrics for the SED and DOA problems:
-%\begin{itemize}
-%	\item SED: F-score (\textit{F}) and error rate (\textit{ER}) calculated in one-second segments.
-%	\item DOA: DOA error (\textit{DOA}) and frame recall (\textit{FR}) calculated frame-wise.
-%\end{itemize}
-%The \textit{SELD score} is an averaged summary of the system performance.
-%A more detailed description of the evaluation metrics can be found in \cite{Adavanne2018_JSTSP}.\\
-%
-%The baseline system features a CRNN that jointly performs DOA and SED through multi-task learning \cite{Adavanne2018_JSTSP}.
-%Baseline results are shown in Table \ref{tab:results_real}.
-%
-%
-%\subsection{Parametric front-end}
-%\label{ssec:param_frontend}
-%
-%Based on the method's exploratory analysis, we propose the following set of parameter values, which are shown in Table~\ref{tab:params_front}.
-%
-%
-%\begin{table}[!htbp]
-%
-%\centering
-%\caption{Parameter values for the selected configuration. Top: \textit{DOA analysis} parameters. Bottom: \textit{Association} parameters.}
-%\begin{tabular}{cccc}
-%\toprule
-%Parameter & Unit & Value\\
-%\midrule
-%sampling rate & Hz  & 48000\\
-%STFT window size & sample  & 256\\
-%STFT window overlap & sample & 128\\
-%STFT window type & -  &  Hann\\
-%analysis frequency range & Hz & [0,8000]\\
-%time average vicinity radius $r$ & bin& 10 \\
-%diffuseness mask threshold $\Psi_{max}$ & - & 0.5 \\
-%energy density filter length & bin & 11\\
-%std mask vicinity radius & bin &  2\\
-%std mask normalized threshold & - & 0.15\\
-%median filter minimum ratio $B_{min}$ & - & 0.5 \\
-%median filter vicinity radius (k,n) & bin & (20, 20) \\
-%\midrule
-%frame size $h$ & s  & 0.02\\
-%resampling minimum valid bins $K_{min}$ & bin & 1 \\
-%overlapping std threshold $\sigma_{max}$ & degree & 10 \\
-%grouping maximum angle $d_{max}^\text{ANGLE}$ & degree & 20\\
-%grouping maximum distance $d_{max}^\text{FRAME}$ & frame & 20\\
-%event minimum length & frame & 8\\
-%\bottomrule
-%\end{tabular}
-%\label{tab:params_front}
-%\end{table}
-%
-%
-%\subsection{Deep learning classification back-end}
-%\label{ssec:backend_experiments}
-%
-%We use the provided four fold cross-validation setup.
-%Training and validation stages use the outcome of an \textit{ideal} front-end, where the groundtruth DOA estimation and activation times are used to feed the beamformer for time-space segmentation.
-%Conversely, we test the trained models with the signals coming from the \textit{complete} front-end described in Section \ref{sec:method}.
-%
-%We conducted a set of preliminary experiments with different types of networks including a VGG-like net, a less deep CNN \cite{Fonseca2019learning}, a Mobilenetv1 \cite{howard2017mobilenets} and a CRNN \cite{cakir2017convolutional}.
-%The latter was found to stand out, and we explore certain facets of the CRNN architecture and the learning pipeline.\\
-%
-%Sound events in the dataset last from $\sim$ 0.2 to 3.3 s.
-%First, clips shorter than 2s are replicated to meet this length.
-%Then, we compute TF patches of log-mel spectrograms of $T=50$ frames (1 s) and $F=64$ bands.
-%The values come from the exploration of $T \in \left\lbrace 25,50,75,100\right\rbrace$ and $F \in \left\lbrace 40,64,96,128\right\rbrace$.
-%$T=50$ is the top performing value, roughly coinciding with the median event duration.
-In turn, more than 64 bands provide inconsistent improvements, at the cost of increasing the number of network weights.\\
-%
-%Regarding the network structure, several variants of the CRNN architecture were explored until reaching the network of Figure \ref{fig:archi}.
-%This included a small grid search over number of CNN filters, CNN filter size and shape, number of GRU units, number of FC units, dropout \cite{srivastava2014dropout}, learning rate, and the usage of Batch Normalization (BN) \cite{ioffe2015batch}.
-%Network extensions (involving more weights) were considered only if providing major improvements, as a measure against overfitting.
-%The main takeaways are: \textit{i)} squared 3x3 filters provide better results than larger filters, \textit{ii)} dropout of 0.5 is critical for overfitting mitigation, \textit{iii)} more than one recurrent layer does not yield improvements, while slowing down training, and \textit{iv)} surprisingly, slightly better performance is attained without BN nor pre-activation \cite{fonseca2018simple}.
-\\
-%
-%For all experiments, the batch size was 100 and Adam optimizer was used \cite{kingma2014adam} with initial learning rate of 0.001, halved each time the validation accuracy plateaus for 5 epochs.
-%Earlystopping was adopted with a patience of 15 epochs, monitoring validation accuracy.
-%Prediction for every event was obtained by computing predictions at the patch level, and aggregating them with the geometric mean to produce a clip-level prediction. \\
-%
-%
-%Finally, we apply \textit{mixup} \cite{zhang2017mixup} as data augmentation technique.
-%Mixup consists in creating virtual training examples through linear interpolations in the feature space, assuming that they correspond to linear interpolations in the label space.
-%Essentially, virtual TF patches are created on the fly as convex combinations of the input training patches, with a hyper-parameter $\alpha$ controlling the interpolation strength.
-%Mixup has been proven successful for sound event classification, even in adverse conditions of corrupted labels \cite{Fonseca2019model}.
-%It seems appropriate for this task since the front-end outcome can present leakage due to overlapping sources, effectively mixing two sources while only one training label is available, which can be understood as a form of label noise \cite{Fonseca2019learning}.
-%Experiments revealed that mixup with $\alpha=0.1$ boosted testing accuracy in $\sim1.5$\%. 
-%
-%
-%
-%
-%
-%
-%%==============================================
-%
-%\section{Results and Discussion}
-%\label{sec:results}
-%\begin{table}[!htbp]
-%\centering
-%\caption{Results for development (top) and evaluation (bottom) sets.}
-%
-%\begin{tabular}{cccccc}
-%\toprule
-%Method & \textit{ER} & \textit{F} & \textit{DOA} & \textit{FR} & \textbf{\textit{SELD}}\\
-%\midrule
-%Baseline & 0.34 & 79.9\% & 28.5$^{\circ}$  & 85.4\% & 0.2113\\
-%Proposed & 0.32 & 79.7\% & 9.1$^{\circ}$   & 76.4\% & \textbf{0.2026}\\
-%Ideal front-end & 0.08 & 93.2\% & $\sim0^{\circ}$   &$ \sim100$\% & 0.0379\\
-%\midrule
-%Baseline & 0.28 & 85.4\% & 24.6$^{\circ}$  & 85.7\% &  0.1764\\
-%Proposed & 0.29 & 82.1\% & 9.3$^{\circ}$   & 75.8\% & \textbf{ 0.1907}\\
-%\bottomrule
-%\end{tabular}
-%\label{tab:results_real}
-%\end{table}
-%
-%
-%Table~\ref{tab:results_real} shows the results of the proposed method for both development and evaluation sets, compared to the baseline.
-%Focusing on evaluation results, our method and the baseline obtain similar performance in SED (\textit{ER} and \textit{F}). However, there is a clear difference in the DOA metrics:
-%in our method, \textit{DOA} error is reduced by a factor of 2.6, but \textit{FR}  is $\sim10$ points worst.
-%In terms of \textit{SELD score}, our method performs slightly worse than the baseline in evaluation mode, while marginally outperforming it in development mode. \\m,              
-%
-%The most relevant observation is the great improvement in \textit{DOA} error.
-Results suggest that using spatial audio parametric analysis as a preprocessing step can help to substantially improve localization.
-%Figure \ref{fig:5a} provides further evidence for this argument: Challenge methods using some kind of parametric preprocessing (\textit{GCC-PHAT} with the microphone dataset, and \textit{Intensity Vector-Based} in ambisonics) obtained in average better \textit{DOA} error results.
-\\
-%
-%\begin{figure}
-%  \begin{subfigure}[b]{\textwidth}
-%    \includegraphics[width=\textwidth]{Figures/SELD//doaerror2.jpg}
-%    \caption{\textit{DOA error} across submissions.
-Hatched bars denote methods using parametric preprocessing.
-Horizontal lines depict average DOA error accross different subsets: all methods (solid), parametric methods (dashed), non-parametric methods (dotted).}
-%    \label{fig:5a}
-%  \end{subfigure}
-%  \begin{subfigure}[b]{\textwidth}
-%    \includegraphics[width=\textwidth]{Figures/SELD//complexity_vs_seld2.jpg}
-%    \caption{\textit{SELD score} versus complexity.}
-%    \label{fig:5b}
-%  \end{subfigure}
-%\caption{DCASE2019 Challenge Task 3 results, evaluation set.}
-%\end{figure}
-%
-%Conversely, the front-end fails regarding \textit{FR}.
-This is probably due to the complexity added by the association step \cite{Adavanne2018_JSTSP}, and its lack of robustness under highly reverberant scenarios.
-Including spectral information at the grouping stage might help to improve \textit{FR} --- such information could be provided by the classification back-end, in a similar approach to the baseline system.
-Another option would be the usage of more sophisticated source counting methods \cite{he2010detecting, Stefanakis2017}.
-\\
-%
-%In order to gain a better insight of the classification back-end performance, Table~\ref{tab:results_real} shows the method results when the testing clips are obtained by feeding  the beamformer  with groundtruth annotations (\textit{ideal} front-end).
-%In this ideal scenario of DOA performance, the SED metrics show a significant boost. 
-%This result suggests that the low \textit{FR} given by the front-end has a severe impact on the back-end  performance.
-%Yet, the proposed system reaches similar performance to the baseline system in terms of SED metrics. \\
-%
-%Finally, we would like to discuss algorithm complexity among Challenge methods. 
-%As depicted in Figure \ref{fig:5b}, there is a general trend towards architectures with very high number of weights, as a consequence of the usage of ensembles and large capacity networks.
-%Specifically, 66\% of submitted methods employ 1M weights or more, 30\% employ 10M or more, and 15\% employ 100M or more.
-Such complexities are several orders of magnitude greater than the baseline (150k weights) or the proposed method ($\sim$175k weights).
-%In this context, our method represents a low-complexity solution to the SELD problem, featuring a number of parameters and a performance comparable to the baseline method.
-%
-%
-%\section{Conclusion}
-%\label{sec:conclusion}
-%
-%We present a novel approach for the SELD task.
-Our method relies on spatial parametric analysis for the computation of event DOAs, onsets and offsets.
-This information is used to filter the input signals in time and space, and the resulting event estimations are fed into a CRNN which predicts the class to which the events belong; the classification problem is thereby handled from a simple multi-class perspective.
-%The proposed method is able to obtain an overall performance comparable to the baseline system. 
-%The localization accuracy achieved by our method greatly improves the baseline performance, suggesting that spatial parametric analysis might enhance performance of SELD algorithms.
-Moreover, detection and classification performance in our method suffers from a low Frame Recall; improving this metric could lead to promising SELD scores.
-%
-%
-%

### 2020-07-05 Chapters/ScientificBackground.tex
beta1
tbc

@@
 The value ranges are $0 \leq \vartheta \leq \pi$ for the inclination, and $0 \leq \varphi \leq 2\pi$ for the azimuth. 
+The spherical coordinate system used in this thesis is depicted in Figure~\ref{fig:sphericalreference}.
+
+\begin{figure}[h]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{Figures/ScientificBackground/path829_2.png}
+  \caption{Spherical coordinate system used.}
+	\label{fig:sphericalreference}
+	\end{center}
+\end{figure}
 
 Table~\ref{tab:cartesian} shows the spherical coordinate values for some reference points on the unit sphere.
-Notice that the poles ($\vartheta = ±\pi$) are a special case for the spherical coordinate system -- in that case, the azimuth angle is not defined. \\
+Notice that the poles ($\vartheta = 0, \pi$) are a special case for the spherical coordinate system -- in that case, the azimuth angle is not defined. \\
 
-\begin{table}[!htbp]
+\begin{table}[h]
 \centering
@@
 
-Given the periodic nature of the azimuth angle, the descriptive statistic operations applied to $\vartheta$ will refer to the $2\pi$-\textit{periodic} version or the operator; this situation does not affect the inclination/elevation coordinate.
+Given the periodic nature of the azimuth angle, the descriptive statistic operations applied to $\varphi$ will refer to the $2\pi$-\textit{periodic} version or the operator; this situation does not affect the inclination/elevation coordinate.
 
@@
 
-Time-domain signals transformed by the Short-Time Fourier Transform (STFT) are represented with uppercase, e.g. $X(k,n)$, where $k$ is the frequency bin index, and $n$ the time frame index.
+Time-domain signals transformed by the Short-Time Fourier Transform (STFT) are represented with uppercase, e.g.
+$X(k,n)$, where $k \in [0, K-1]$ is the frequency bin index, and $n \in [0, N-1]$ the time frame index.
 
@@
 	\includegraphics[width=\textwidth]{Figures/Introduction/edt.png}
-	\caption{Room impulse response model, adapted from \url{http://www.bnoack.com/}.}
+	\caption{Room impulse response model, adapted from \cite{avinfo}.}
 	\label{fig:reverberationtimes}
@@
 
-\todo{RIR SIMULATION: IMAGE METHOD}
+\todo{RIR SIMULATION: IMAGE METHOD?}
 

### 2020-07-05 manuscript.tex
tbc

@@
 \usepackage{subcaption}
+
+\usepackage[none]{hyphenat}
+\tolerance=1
+\emergencystretch=\maxdimen
+\hyphenpenalty=10000
+\hbadness=10000
+
+\usepackage[titletoc]{appendix}
 
@@
 %%%%%% Thanks
-\noindent {\Large \sffamily Thanks} thanks to....  \todo{thanks}
+\noindent {\Large \sffamily Thanks}
+
+First of all
+
+Adan, eurecat
+
+Emilia, upf
+
+nikos, archontis, tuomas
+
+rafa
+
+pe;a de barcelona
+marta
 
@@
 Due to the recent advances in virtual and augmented reality, ambisonics has emerged as the \textit{de facto} standard for immersive audio.
-Ambisonic audio can be captured using spherical microphone arrays, which are becoming increasingly popular as well.
+Ambisonic audio can be captured using spherical microphone arrays, which are becoming increasingly popular.
 Yet, many methods for acoustic and microphone array signal processing are not specifically tailored for spherical geometries.
@@
 
-Ambisonics ha esdevingut l'estàndard d'àudio immersiu als últims anys, afavorit pels avanços en realitat virtual i augmentada.
-L'àudio ambisonic es pot obtindre mitjançant \textit{arrays} de micròfons esfèrics, que també són cada vegada més populars.
+\textit{Ambisonics} ha esdevingut l'estàndard d'àudio immersiu als últims anys, afavorit pels avançaments en realitat virtual i augmentada.
+L'àudio \textit{ambisonic} es pot obtindre mitjançant \textit{arrays} de micròfons esfèrics, que són cada vegada més populars.
 Tot i això, la majoria de mètodes acústics i de processament de senyal basats en \textit{arrays} de micròfons no estan adaptats al cas específic de geometries esfèriques.
-Per tant, encara hi ha moltes possibilitats de millora en l'àmbit d'anàlisi automàtic i descripció de gravacions ambisonic.
-En la present tesi plantegem aquest problema basant-nos en l'anàlisi paramètric del camp acústic.
+Per tant, encara hi ha moltes possibilitats de millora en l'àmbit d'anàlisi automàtica i descripció de gravacions \textit{ambisonic}.
+En la present tesi plantegem aquest problema basant-nos en l'anàlisi paramètrica del camp acústic.
 Més concretament, presentem contribucions originals en les àrees d'estimació de reverberació, estimació de difusió acústica, i detecció i localització d'esdeveniments sonors.
-Tanmateix, presentem diverses eines de programari desenvolupades per la generació i manteniment de bases de dades d'àudio ambisonic.
+Així mateix, presentem diverses eines de programari desenvolupades per la generació i manteniment de bases de dades d'àudio \textit{ambisonic}.
 
@@
 %PREFACE. 
-{\bf Preface \todo{TODO}}
+%{\bf Preface }
+\textbf{Preface} \\
+
+The present thesis has been carried out within the context of the Industrial Doctorate program from the Catalan government, as a collaboration between the Music Technology Group of the Pompeu Fabra University, and the Multimedia Unit of Eurecat, the Catalan Technology Center.
+It has been jointly supervised by Dr. Emilia Gómez and Dr. Adan Garriga.
+
+The industrial nature of the doctorate program implies a tendency towards the innovative application of the findings and conclusions generated during the development of the thesis. 
+Therefore, one of the targets of the research performed in the context of the thesis is to ensure the applicability to real-world problems, scenarios and technologies. 
+Moreover, there is also an stress on research reproducibility.
+This has not been motivated only by an academic perspective; but also by the open research and open innovation paradigms present in our today's society.\\
+
+The work presented here would not have been possible without the contributions from an excellent network of researchers.
+Apart from internal collaborations (UPF and Eurecat), two of the main contributions in the thesis followed two research visits: to the Audio Research Group from Tampere University, Finland (Chapter~\ref{chap:rt60}), and to the Signal Processing Group at the FORTH - Foundation for Research and Technology Hellas, Heraklion, Greece, (Chapter~\ref{chap:coherence}).
+
+The alphabetical list of academic collaborators is: Eduardo Fonseca, Emilia G\'omez, Rafael Ib\'añez-Usach,  Julien de Muynke, Archontis Politis, Xavier Serra and Nikolaos Stefanakis.
 
@@
 \input{Chapters/Conclusions}
+
+\begin{appendices}
+\appendixpage
+\noappendicestocpagenum
+\addappheadtotoc
+
+\chapter{List of Contributions}
+
+In the following list we show the main scientific contributions, as main author, related to this dissertation:
+
+\begin{itemize}
+
+	\item Peer-reviewed journal articles
+
+	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
+	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
+	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
+
+	\item Peer-reviewed conference articles
+
+	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
+	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
+	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}\\
+
+	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
+	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+
+	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
+	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
+	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+
+	\textbf{"Ambiscaper: A tool for automatic generation and annotation of reverberant ambisonics sound scenes".}\\
+	\underline{A. P\'erez-L\'opez}.
+	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
+
+	\item Conference engineering briefs
+
+	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
+	\underline{A. P\'erez-L\'opez} and J. De Muynke.
+	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}\\
+
+	\textbf{"pysofaconventions, a Python API for SOFA".}
+	\underline{A. P\'erez-L\'opez}.
+	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+
+	\textbf{"A Python library for multichannel acoustic signal processing".}
+	\underline{A. P\'erez-L\'opez} and A. Politis.
+	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+
+\end{itemize}
+
+\newpage
+
+Moreover, although not strictly aligned with the research direction of this thesis, the author has supervised the following publication:\\
+
+	 \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
+	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+
+\chapter{List of Software Resources}
+
+As a result of the development of this thesis, the following open-source libraries and repositories have been produced.
+All of them are freely available through the author's GitHub page \cite{andresgithub} under open-source licenses.
+
+\begin{itemize}
+
+	\item Software tools:
+
+	\begin{itemize}
+
+		\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library} 
+		Tools for acoustic simulation and spherical array processing.
+
+		\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions} Implementation of the SOFA convention in Python.
+
+		\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}	 Ambisonic SOFA Convention proposal.
+
+		\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper} Tool for automatic generation of annotated ambisonic datasets.
+
+	\end{itemize}
+	\newpage
+	\item Method implementations:
+
+	\begin{itemize}
+		\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation} Contains the code implementing the methods described in Chapter~\ref{chap:rt60}.
+
+		\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020} Full code implementing the SELD algorithm from Chapter~\ref{chap:seld2019}.
+
+		\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3} 
+		Implementation of the method submitted to DCASE2019 (not included in this thesis).
+
+	\end{itemize}
+
+\end{itemize}
+
+\end{appendices}
 

### 2020-07-05 Chapters/CoherenceEstimation.tex
beta_v2

@@
 The theoretical coherence is derived from Eq.~(\ref{eq:closedform_msc}), while simulated and recorded MSC have been computed by Welch's method, using a \textit{hanning} window of 256 samples and 1/2 overlap.
+
 The difference between theoretical and simulated coherence is negligible for practical applications.
@@
 Above this frequency, the recorded $MSC$ presents several spectral peaks, which might be partially explained by the interference of the microphone itself in the recorded sound field, and by the non-ideal directivity of the capsules.
+
 The squared error $\varepsilon(k)$ with respect to the simulated curve is shown in Fig.~\ref{fig:Fig1} (left), while Fig.~\ref{fig:Fig1} (right) represents the same error averaged over frequency $\bar{\varepsilon}$ for different spatial resolution values of the diffuse field reproduction algorithm.
@@
 \end{figure}
+
+\newpage
 
@@
 
+\newpage
 This effect can be also appreciated in Fig.~\ref{fig:Fig3} (right): the mean squared error is around two orders of magnitude higher in \textit{B-Format}.

### 2020-07-05 Chapters/Conclusions.tex
beta_v2

@@
 \begin{enumerate}
-	\item The development of methods for the characterization of acoustic parameters.
+	\item The development of methods to support and improve the characterization of acoustic parameters.
 	\item The research on parametric-based methodologies for sound event localization and detection.
@@
 
-In the following list we provide the main contributions of the present thesis, both in the academic and software scopes.
-Contrubitions are organized by the Chapter where they appear; the index of the respective contribution is also indicated:
+In what follows, we summarize the main contributions of the present thesis, both in the academic and software scopes.
 
-  \begin{itemize}
+%  \begin{itemize}
 
-  	 \item Blind reverberation time estimation (Chapter~\ref{chap:rt60}):
-  	\begin{description}
-  		\item [1. Parameter estimation] Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models.
-	\end{description}
-
-  	\item Coherence estimation (Chapter~\ref{chap:coherence}):
-  	\begin{description}
-  		\item [1. Parameter estimation] Contribution to the characterization of coherence with tetrahedral microphones (the most common spherical arrangement).
-	\end{description}
-
-  	\item Sound Event Localization and Detection (Chapter~\ref{chap:seld2019}):
-  	\begin{description}
-  		\item [2. Scene Description]: Novel state-of-the-Art methodology for Sound Event Localization and Detection	.
-  	\end{description}
-
-  	\item Data generation and storage (Chapter~\ref{chap:data}):
-  		\begin{description}
-  			\item [3. Data Generation]: Library for acoustic simulation and spherical microphone array processing.
-  			\item [3. Data Generation]: Proposal and implementation of a file convention for the storage of recorded ambisonic RIRs.
-  			\item [3. Data Generation]: Novel tool for high-level description and generation of of ambisonic datasets.
-
- 		\end{description}
-
-  \end{itemize}
-
-Figure~\ref{fig:scheme1_numbers2}, which is a copy of Figure~\ref{fig:scheme1_numbers}, 
- has been included here again in order to help the contextualization of the contributions.
-
-  \begin{figure}[h!]
-  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
-  \caption{General scheme of the B-Format audio generation and analysis framework, including the thesis contributions in form of Chapter numbers.}
-  \label{fig:scheme1_numbers2}
-\end{figure}
-
-\section{Conclusion and Future Work}
-
-This thesis has tackled several research problems associated with the analysis of ambisonic recordings, making use primarily of the parametric sound field modelling.
+%  	 \item Blind reverberation time estimation
+\newpage
+\paragraph{Blind reverberation time estimation}
 Chapter~\ref{chap:rt60} presents a novel methodology for the blind estimation of reverberation time from ambisonic audio.
@@
 The proposed system is the first attempt in the literature to address the blind reverberation time estimation problem specifically for ambisonic signals.
+Compared with a state-of-the-art monophonic estimator, our method is able to improve in all the evaluation metrics under consideration.
+
+\paragraph{Coherence estimation}
+In Chapter~\ref{chap:coherence}, we have characterised the response of tetrahedral microphones to isotropic noise field, which is one of the most used models for diffuse sound.
+Furthermore, the capabilities of a spherical loudspeaker array with respect to the reconstruction of diffuse sound fields using ambisonics are also quantitatively analyzed.
+
+\paragraph{Sound event localization and detection}
+Chapter~\ref{chap:seld2019} describes an algorithm for sound event localization and detection (SELD), developed in the context of the DCASE 2020 challenge.
+The method estimates the localization and temporal activity of the sound events based on a particle filter that tracks event trajectories obtained from the parametric analysis of the ambisonic sound field.
+Each event is assigned to a sound class by a machine learning classifier that uses low- and mid- level audio features.
+
+Results show a significant performance increase in all evaluation metrics under consideration, compared with a state-of-the-art deep learning baseline.
+This suggests that our approach, substantially different to the baseline and the majority of state-of-the-art methods, represents a feasible alternative in situations with low-complexity or small database constraints.
+
+\paragraph{Data generation and management}
+Finally, the thesis contributions to more practical aspects are presented in Chapter~\ref{chap:data}.
+Those contributions comprise two software libraries written in Python: one of them focused on spherical microphone array and acoustic simulation, and another one implementing the SOFA standard, which has also been revised and modified for allowing the representation of ambisonic data.
+Finally, a novel software tool for the procedural creation of annotated reverberant ambisonic datasets has ben also presented.
+
+%  	\begin{description}
+%  		\item [1. Parameter estimation] Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models.
+%	\end{description}
+%
+%  
+%  	\item Coherence estimation (Chapter~\ref{chap:coherence}):
+%  	\begin{description}
+%  		\item [1. Parameter estimation] Contribution to the characterization of coherence with tetrahedral microphones (the most common spherical arrangement).
+%	\end{description}
+%
+%
+%  	\item Sound Event Localization and Detection (Chapter~\ref{chap:seld2019}):
+%  	\begin{description}
+%  		\item [2. Scene Description]: Novel state-of-the-Art methodology for Sound Event Localization and Detection	.
+%  	\end{description}
+%  	
+%  	\item Data generation and storage (Chapter~\ref{chap:data}):
+%  		\begin{description}
+%  			\item [3. Data Generation]: Library for acoustic simulation and spherical microphone array processing.
+%  			\item [3. Data Generation]: Proposal and implementation of a file convention for the storage of recorded ambisonic RIRs.
+%  			\item [3. Data Generation]: Novel tool for high-level description and generation of of ambisonic datasets.
+%
+% 		\end{description}
+%
+%  \end{itemize}
+
+%Figure~\ref{fig:scheme1_numbers2}, which is a copy of Figure~\ref{fig:scheme1_numbers}, 
+% has been included here again in order to help the contextualization of the contributions. 
+%
+%  \begin{figure}[h!]
+%  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
+%  \caption{General scheme of the B-Format audio generation and analysis framework, including the thesis contributions in form of Chapter numbers.}
+%  \label{fig:scheme1_numbers2}
+%\end{figure}
+
+\section{Future Work}
+
+This thesis has tackled several research problems associated with the analysis of ambisonic recordings, making use primarily of the parametric sound field modelling analysis.
+The presented techniques improve existing state-of-the-art methodologies, or present novel approaches for known research problems, which in turn bring new research questions. \\
+
+%Chapter~\ref{chap:rt60} presents a novel methodology for the blind estimation of reverberation time from ambisonic audio.
+The method is based on two main steps: first, dereverberation using a multichannel auto-recursive model (MAR), and second, estimation of the  filter from reverberant and dereverberated signals.
+The actual reverberation time value is estimated from the energy decay of the estimated filter.
+%
+%The proposed system is the first attempt in the literature to address the blind reverberation time estimation problem specifically for ambisonic signals.
 Compared with a state-of-the-art monophonic estimator, our method is able to improve in all the standard evaluation metrics under consideration.
 
-Still, there is much room for improvement.
-Among others, the method could be easily extended to higher order ambisonic signals, which might still improve the reported results due to the availability of much more audio channels.
+A novel blind reverberation time estimation method for first-order ambisonic recordings is introduced in Chapter~\ref{chap:rt60}.
+Among others, the method could be straightforwardly extended to higher order ambisonic signals, which might still improve the reported results due to the availability of many more audio channels. 
 Moreover, the usage of online MAR methods would enable the possibility of analyzing sound scenes with moving sources; the statistical time-invariance property of late reverberation supports this hypothesis.
@@
 An analysis of the trade-off between computation time and evaluation performance, mostly dependent on the estimation filter length, remains to be done.
-
+\newpage
 Given the current interest in the field of augmented reality, new related research topics emerge.
@@
 
-In Chapter~\ref{chap:coherence}, we have characterised the response of tetrahedral microphones to isotropic noise field, which is one of the most used models for diffuse sound.
+%In Chapter~\ref{chap:coherence}, we have characterised the response of tetrahedral microphones to isotropic noise field, which is one of the most used models for diffuse sound.
 Furthermore, the capabilities of a spherical loudspeaker array with respect to the reconstruction of diffuse sound fields using ambisonics are also tested.
-Future work in this direction might include the characterisation of different spherical microphone array geometries, and different types of diffuse field models.
-Additionally, the loudspeaker array diffuse field reconstruction experiment could be easily extended.
+Regarding the diffuse field characterization performed in Chapter~\ref{chap:coherence}, an immediate extension of the work would include the study of different spherical microphone array geometries, from the ones that are commercially available.
+The usage of different models of diffuse fields, specially extending to the anisotropic case \cite{alary2019assessing}, might also constitute an interesting research continuation direction. 
+Both cases could be also applied to the experiment of diffuse sound field reconstruction using loudspeaker arrays. \\
+
+%in Future work in this direction might include the characterisation of different spherical microphone array geometries, and different types of diffuse field models.
+Additionally, the experiment on loudspeaker array diffuse field reconstruction could be easily extended, with the aim of including more .
 \\
 
-Chapter~\ref{chap:seld2019} describes an algorithm for sound event localization and detection (SELD), developed in the context of the DCASE 2020 challenge.
+%Chapter~\ref{chap:seld2019} describes an algorithm for sound event localization and detection (SELD), developed in the context of the DCASE 2020 challenge.
 The method estimates the localization and temporal activity of the sound events based on a particle filter that tracks event trajectories obtained from the parametric analysis of the ambisonic sound field.
@@
 
-The wide scope of the SELD task allows for a wide range of research continuation possibilities.
-For instance, one of the major problems of the proposed algorithm is the inaccurate parametric estimation when two events are simultaneously active.
-Although it is a known problem in the literature, a successful solution in the given context remains still to be found.
-
+The wide scope of the SELD problem, as described in Chapter~\ref{chap:seld2019}, allows for a wide range of possibilities regarding a potential follow-up of the proposed method.
+For instance, one of the major problems of our algorithm is the inaccurate parametric estimation when two events are simultaneously active.
+Although it is a known problem in the literature \cite{epain2016spherical} , a successful solution in the given context remains still to be explored.
+\newpage
 Another source of potential improvements is the refinement of the particle filter applied to this specific task.
@@
 
-Finally, the thesis contributions to more practical aspects are presented in Chapter~\ref{chap:data}.
+%Finally, the thesis contributions to more practical aspects are presented in Chapter~\ref{chap:data}.
 Those contributions comprise two software libraries written in Python: one of them focused on spherical microphone array and acoustic simulation, and another one implementing the SOFA standard, which has also been revised and modified for allowing the representation of ambisonic data.
@@
 
+Finally, we discuss briefly on the practical contributions described in  Chapter~\ref{chap:data}.
 Apart from the straightforward task of software maintenance, the engagement of the research community with the usage and eventual contribution of the libraries would represent a desired situation in the near future.
-Moreover, the proposed file format convention is being currently discussed for its adoption into the SOFA standard, which can be considered an achievement of the original proposal.
+Moreover, the proposed file format convention is being currently discussed by the \textit{Standardisation Committee on AES-69 Standard}, which can be considered a great achievement of the original proposal.
+The results of the discussion might lead to the addition of a modified version of our proposal into version 2 of the \textit{AES-69 Standard}.
+
+In the medium term, the application at commercial level of some of the methods  described in this thesis would be a highly desirable outcome of our research.
+Such application would probably imply a re-implementation at the software level, intended for compatibility with the workflows and formats used in the VR/AR content production industry.

### 2020-07-05 Chapters/DataGeneration.tex
beta_v2

@@
 This chapter gathers several proposals related with the creation, storage and transmission of ambisonic data for research purposes.
-The main objective of the contributions described here is the support for the generation of parametrizable ambisonic datasets, using both synthetic and recorded materials, and specifically emphasizing the usage of Room Impulse Responses.\\
+The main objective of the contributions described here is the support for the generation of parametrizable ambisonic datasets, using both synthetic and recorded materials, and specifically emphasizing the usage of Room Impulse Responses.
 
 Most of the contributions listed here (and also most of the code developed for this thesis) have been implemented in Python.
-Indeed, Python has recently become one of the most used programming languages worldwide \cite{theoverflow, PYPL, TIOBE}; as shown also in Figure~\ref{fig:popularity}.\\
+Indeed, Python has recently become one of the most used programming languages worldwide \cite{theoverflow, PYPL, TIOBE}; as shown also in Figure~\ref{fig:popularity}.
 
@@
 More specifically, each function test calls the equivalent Matlab code under the hood. The numeric result is then sent back to Python, where
-it is evaluated against the own result.\\
+it is evaluated against the own result.
 
-Two example applications of the library are shown in Figures~\ref{fig:array_response} and \ref{fig:sht_filters}.
-In Figure~\ref{fig:array_response} , obtained with the Array Response Simulator package, the frequency response of a spherical microphone array to an impinging plane-wave with varying incidence angle is shown.
-The microphone array consists of a 2nd order supercardioid and a 3rd order hypercardioid, located at opposite directions of an open sphere, and both of them facing to the front direction.
+%Two example applications of the library are shown in Figures~\ref{fig:array_response} and \ref{fig:sht_filters}.
+In Figure~\ref{fig:array_response}, obtained with the Array Response Simulator package, the frequency response of a spherical microphone array to a plane-wave with varying incidence angle is shown.
+The array consists of a 2nd order supercardioid and a 3rd order hypercardioid, located at opposite directions of an open sphere, both of them facing front.
 \\
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
@@
 
-\begin{figure}[h!]
+\begin{figure}[th!]
   \centering
@@
 This framework provides an object-oriented interface with two main application scopes: allow RIR simulation of complex rooms based on the image source method, and provide a reference implementation of standard microphone array processing algorithms.
+\\
 
@@
 
-\begin{table}[th!]
+\begin{table}[ht!]
 \centering
@@
 Among others, the availability of such recordings has powered works in a variety of works, auralization \cite{postma2016virtual}, room acoustics analysis \cite{embrechts2015measurement,clapp2011investigations} and modelling \cite{romblom2017diffuse}, spatial audio synthesis \cite{coleman2017object} or source separation \cite{baque2016separation}.
-\\
 
@@
 Despite the common approach, it can be easily foreseen that each database generated by a different individual or institution might potentially have a different naming convention, folder structure, file format, and so on.
-This is exactly the same situation that motivated the development of the SOFA conventions.
+This is exactly the same situation that motivated the development of the SOFA conventions. \\
 
@@
 
-Given that the described situation meets all requirements, the \textit{Ambisonics Directional Room Impulse Response} (AmbisonicsDRIR) convention has been therefore proposed as a new member of SOFA.\\
+Given that the described situation meets all requirements, the \textit{Ambisonics Directional Room Impulse Response} (AmbisonicsDRIR) convention has been therefore proposed as SOFA convention.\\
 
 The technical specifications of the proposed convention in its current state (version 0.2) are available online \cite{ambisonicsdrir}. 
-Despite a timid adoption of the convention, the AmbisonicsDRIR proposal has arisen interest in the community.
+%Despite a timid adoption of the convention, the AmbisonicsDRIR proposal has arisen interest in the community.
 At the moment of writing, the \textit{Standardisation Committee on AES-69 Standard} is discussing potential modifications to the SOFA file format required for the adoption of several new features; and the support for spherical harmonic-based measurements is among the topics for deliberation.
@@
 %% DOA ESTIMATION
-\begin{table}[htpb]
+\begin{table}[ht!]
     \centering
@@
 Only when simulated IRs are used, it is possible to partially replicate the experimental setup --- the parameters used in the simulation software are usually provided. 
-Furthermore, the process of dataset creation seems to be performed \textit{ad-hoc} in each case.
+Furthermore, the process of dataset creation seems to be performed \textit{ad-hoc} in each case. \\
 
@@
 %%%%%%%%
+\newpage
 In order to generate a sound scene, \textit{AmbiScaper} requires three different inputs: the original \textit{mono signals}, which will provide the actual audio content; an optional ambisonic RIR; and an \textit{event specification}.
@@
 
+\newpage
 When no reverberation is specified, \textit{AmbiScaper} can generate anechoic sound scenes, using the theoretical expression of the ambisonic encoding as in Eq.~\ref{eq:encoding}.
@@
 The sources are randomly chosen from a subset of the Anechoic OpenAirlib database \cite{openair}, which mostly contains recordings from baroque musical instruments.
-Reverberation is provided by the \textit{AudioBooth} recorded RIR from the S3A collection \cite{coleman2015s3a}.\\
-
-\todo{explain current state, discussion, etc?}
+Reverberation makes use of the \textit{AudioBooth} RIR from S3A database \cite{coleman2015s3a}.\\
 

### 2020-07-05 Chapters/Introduction.tex
beta_v2

@@
 \begin{figure}[t!]
-  \includegraphics[width=\textwidth]{Figures/Introduction/num_mics_ticks.png}
+\begin{center}
+  \includegraphics[width=0.8\textwidth]{Figures/Introduction/num_mics_ticks.png}
   \caption{Number of ambisonic microphones released in last years (from \cite{List_of_Ambisonic_hardware}).
@@
   \label{fig:nummics}
+\end{center}
 \end{figure}
@@
 \begin{figure}[t!]
-  \includegraphics[width=\textwidth]{Figures/Introduction/vr_funding.png}
+\begin{center}
+  \includegraphics[width=0.8\textwidth]{Figures/Introduction/vr_funding.png}
   \caption{Venture investments in VR in the period 2014-2018. Adapted from \cite{fortune}.}
   \label{fig:vrfunding}
+\end{center}
 \end{figure}
@@
 
-The scientific context of the work developed in the present dissertation is depicted in Figure~\ref{fig:scheme1}, which has been inspired by \cite{jarrett2017theory}. 
+The scientific context of the work developed in this thesis is shown in Figure~\ref{fig:scheme1}, which has been inspired by \cite{jarrett2017theory}. 
 As we can observe, there are two main topics related with B-Format audio: \textit{generation} and \textit{analysis}, with the signal flow going from the former to the latter.
-Although the conceptual approach of the scheme might be very similar for any type of audio, the specific spatial information conveyed by the ambisonic signal places an emphasis on the informed analysis and description of the sound scene.
+Although the conceptual approach of the scheme might be very similar for any type of audio, the spatial information conveyed by the ambisonic signal places an emphasis on the informed analysis and description of the sound scene.
 \\
 
-\begin{figure}[th]
-  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1.png}
+\begin{figure}[t!]
+\begin{center}
+  \includegraphics[width=0.8\textwidth]{Figures/Introduction/SCHEME1.png}
   \caption{General scheme of the B-Format audio generation and analysis framework.  Solid lines represent audio signals, while outlined arrows refer to non-audio information.}
   \label{fig:scheme1}
+\end{center}
 \end{figure}
@@
 
-Ambisonic signal analysis has been divided into three different categories:
+Ambisonic signal analysis has been divided in three categories:
 \begin{description}
@@
 \begin{figure}[t]
-  \includegraphics[width=\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
+\begin{center}
+  \includegraphics[width=0.8\textwidth]{Figures/Introduction/SCHEME1_NUMBERS.png}
   \caption{General scheme of the B-Format audio generation and analysis framework, including the thesis contributions in form of Chapter numbers.}
   \label{fig:scheme1_numbers}
+\end{center}
 \end{figure}
@@
 The evaluation metrics show a method performance similar to other state-of-the-art methods.
+
 \textbf{Chapter~\ref{chap:coherence}} analyses the response of tetrahedral microphone arrays, which are the simplest and most common form of ambisonic microphones, under spherically isotropic sound field.
 The analysis is performed using both simulated and recorded diffuse field, and the results quantify the differences between ideal and real values under a variety of conditions and estimators.
+
 In \textbf{Chapter~\ref{chap:seld2019}}, a complete system for Sound Event Localization and Detection of ambisonic sound scenes is described.
@@
 The method is able to perform in a similar way to the baseline system, while greatly improving its localization capabilities.
+\newpage
 

### 2020-07-05 Chapters/ReververationTimeEstimation.tex
beta_v2

--- 2020-06-28 Chapters/ReververationTimeEstimation.tex
@@
 The proposed method is evaluated using two different reverberant datasets, containing recordings of \textit{speech} and \textit{drums} respectively. 
-In order to have full control over the reverberation conditions in the experimental setup, the audio clips under consideration have been rendered by the convolutive mixture of clean monophonic recordings with FOA IRs.
+In order to have full control over the reverberation conditions in the experimental setup, the audio clips under consideration have been rendered by the convolutive mixture of clean monophonic recordings with FOA IRs.\\
 
@@
 Those datasets will be referred in the following as the \textit{evaluation} datasets.
+\\
 
@@
 
+\newpage
 \subsection{Setup}
@@
 In the dereverberation stage, the STFT uses a small window size of 128 samples, with 64 samples overlap.
-The value of $p$ is set to 0.25, given the good results reported in \cite{jukic2015group}.
+The value of $p$ is 0.25, given the good results reported in \cite{jukic2015group}.
 Other parameter values are $\tau=2$, $i_{max} = 10$, $\eta=10^{-4}$ and  $\epsilon=10^{-4}$.
-After an exploratory search, the length of the IIR filter $L_g = 20$ has been chosen as a compromise between method performance and computation time.
-We have observed a tendency towards poor dereverberation and non-convergence of the IRSL when using small values of $L_g$ and short audios.
+After an exploratory search, the length of the IIR filter $L_g = 20$ has been chosen as a compromise between performance and computation time.
+We have observed a tendency towards poor dereverberation and non-convergence of the IRSL when using small values of $L_g$.
 
@@
 Moreover, all methods behave significantly better on the \textit{speech} dataset.
-The main reason might be the heterogeneity of the \textit{drums} dataset dataset with respect to dynamic range or timbre, and the potential application of audio effects of any kind.
+The main reason might be the heterogeneity of the \textit{drums} dataset with respect to dynamic range or timbre, and the potential application of audio effects of any kind.
 Furthermore, some audio clips of the \textit{drums} dataset contain sounds with a high degree of self-similarity, such as cymbal rolls or exaggerated reverbs; these characteristics would explain the outliers on the proposed method results.
-It is also interesting to notice the robustness of the proposed method against noise, present in the \textit{speech} dataset. 
+It is also interesting to notice the robustness of our method against noise, present in the \textit{speech} dataset; 
 %Although the signal model does not contemplate noise explicitly, such robustness is consistent with the behaviour described in \cite{jukic2015group}.
-Such robustness is consistent with the behavior reported in \cite{jukic2015group}.\\
+such robustness is consistent with the behavior reported in \cite{jukic2015group}.\\
 
@@
 
-We have presented in this work a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings.
+In this Chapter, we have presented  a novel method for blind reverberation time estimation for multichannel audio, with the aim of applying it to the context of ambisonic recordings.
 Our method is based on a first dereverberation step,  performed by a multichannel autoregressive model of the late reverberation.
@@
 Results show that our method outperforms the baseline method in a majority of evaluation metrics and conditions, and consistently provides results with less variability than the baseline method. 
-In future work, we plan to extend the experimental setup by using recorded IRs.
+%In future work, we plan to extend the experimental setup by using recorded IRs.
 Furthermore, the proposed method could be extended to the case of moving sources by using an \textit{online} autoregressive model.
-%Finally, an evaluation of the method in terms of the number of channels or the ambisonic order remains to be done.
-Finally, an extension of the method for higher ambisonic orders remains to be done.
+%%Finally, an evaluation of the method in terms of the number of channels or the ambisonic order remains to be done.
+%Finally, an extension of the method for higher ambisonic orders remains to be done.

### 2020-07-05 Chapters/SELD.tex
beta_v2

@@
 The performance of the baseline method, a CRNN that performed jointly the localization and classification tasks \cite{Adavanne2018_JSTSP}, was vastly exceeded by a variety of deep-learning based algorithms \cite{kapka2019sound, Cao2019, grondin2019sound}.
-Some of these improvements have been included in the baseline system for the SELD Challenge of DCASE 2020.
+Some of these improvements have been included in the baseline system for the SELD Challenge of DCASE 2020. \\
 
 Despite the predominant trend towards high-complexity deep-learning architectures, some recent works have been able to match or even improve CRNN-based methods with regard to localization, by using parametric analysis of the ambisonic sound field \cite{perez2019hybrid, nguyen2020sequence}.
-Apart from the benefit derived by their simplicity, these approaches are able to resolve the case of overlapping events of the same class, a situation difficult to disambiguate for CRNN-based methods \cite{politis2020dataset}.
+Apart from the benefit derived by their simplicity, these approaches are able to resolve the case of overlapping events of the same class, a situation difficult to disambiguate for CRNN-based methods \cite{politis2020dataset}.\\
 
 The present work continues the exploration of possibilities of parametric SELD methods, focusing on a low-complexity architecture that makes use of traditional, feature-based machine learning techniques.
-The method has been developed in the context of the SELD task within DCASE 2020 Challenge, and therefore utilizes the proposed dataset, baseline system and evaluation metrics.
+The method has been developed in the context of the SELD task within DCASE 2020 Challenge, and therefore utilizes the proposed dataset, baseline system and evaluation metrics.\\
 
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In the resulting spectrogram, the frequencies above a given limit $f_{max}$ are discarded; this procedure speeds up the method while maintaining the directional information, given that the microphone geometry aliases spatial measurements above approx.
+In the resulting spectrogram, the frequencies above a given limit $f_{max}$ are discarded; this procedure speeds up the method while maintaining the directional information, given that the microphone geometry produces spatial aliasing above approx.
 5 kHz \cite{Bertet2006}.
@@
 Single-source TF bins are computed from the DirAC parametric analysis.
-A variety of alternative subspace methods are known \cite{epain2016spherical, madmoni2018direction}; however, those methods require local estimation of eigenvalues, which is a computationally expensive procedure.
-% This is the main reason for the choice of DirAC-based analysis in this work.
+A variety of alternative subspace methods are known \cite{epain2016spherical, madmoni2018direction}; however, those methods require local estimation of eigenvalues through the Spatial Covariance Matrix (SCM), which is a computationally expensive procedure;
+this is the main reason for the choice of DirAC-based analysis in this work.
 
@@
 The system takes as the input the set of TF DOA values passing the single source test, and produces spatio-temporal event trajectories, considering an event as an entity with contiguous temporal activation and continuous spatial position.
-More specifically, for each time frame of the DOA masked spectrogram, the median\footnote{Circular median in the case of azimuth.} of all narrowband DOA estimates is computed.
+More specifically, for each time frame, the median\footnote{Circular median in the case of azimuth.} of all narrowband masked DOA estimates is computed.
 The resulting value is added to the measurement space of the tracker if the number of single-source frequency bins for that frame exceeds a minimum $K_{min}$.
@@
 The performance of the RBMCDA algorithm is controlled by several parameters.
-Some of the most relevant include the angular velocity prior $v$, the standard deviation $\sigma_{\nu}$ and the spectral density $s_{\nu}$ of the measurement noise, the prior probabilities of birth $p_{birth}$ and noise percentage $p_{\nu}$, and the number of Monte-Carlo particles $N$.
+Some of the most relevant ones are the angular velocity prior $v$, the standard deviation $\sigma_{\nu}$ and the spectral density $s_{\nu}$ of the measurement noise, the prior probabilities of birth $p_{birth}$ and noise percentage $p_{\nu}$, and the number of Monte-Carlo particles $N$.
 Position-related parameters are adjusted with respect to their ranges, so that azimuth-related magnitudes double elevation values.
@@
 It is a known issue that performance of DirAC diffuseness is reduced when two sources are present \cite{epain2016spherical}; similar problems have been reported in \cite{adavanne2019localization}, where an instantaneous source number estimator is used in combination with the particle filter.
-As in that case, the results suggest the need for more sophisticated source detection and counting methods.
+As in that case, the results suggest the need for more sophisticated source detection and counting methods. \\
 
@@
 In order to provide a deeper insight on the temporal characteristics of the data under consideration, the durations of all sound events in the \textit{PAPAFIL2} training set are summarized in Figure~\ref{fig:eventdurations}.
+\\
 
@@
 
-Furthermore, it is interesting to notice the high relevance of low-level features, and specifically several MFCC combinations (eight of the fifteen reported features) and  various extractors related to the spectral structure.
-The absence of pitch, harmonic and envelope features in the list represents a significant finding as well.
+Furthermore, it is interesting to notice the high relevance of low-level features; specifically, several MFCC combinations (eight of the fifteen reported features) and various extractors related to the spectral structure.
+The absence of pitch, harmonic and envelope features in the list represents also a significant finding.
 \\
@@
 It is important to remark that, apart from our proposed method, all other systems rely completely on deep learning algorithms, specially making use of different configurations and combinations of CRNNs.
+
 The only exception to that tendency is the method presented by Nguyen and colleagues, which is an adaptation of their recently proposed methodology \cite{Nguyen2020_task3_report, nguyen2020sequence}.
-In this method, localization is also based on the identification of the single-source TF bins; instead of DirAC diffuseness, they use a subspace analysis to find the TF bins associated with nearly rank-1 spatial covariance matrices (SCM).
-Furthermore, the instantaneous DOA estimates are obtained by finding the azimuth and elevation histogram peaks, with the steering directions obtained from the covariance matrix diagonals. 
+In this method, localization is also based on the identification of the single-source TF bins; instead of DirAC diffuseness, they use a subspace analysis to find the TF bins associated with nearly rank-1 spatial covariance matrices.
+Furthermore, the instantaneous DOA estimates are obtained by finding the azimuth and elevation histogram peaks, with the steering directions obtained from the covariance matrix diagonals. \\
+
 While this method might contribute to provide a smaller localization error compared with our approach (9.3$^{\circ}$ versus 14.2$^{\circ}$ in $\text{LE}_{CD}$), it is more expense computationally, since it involves building a SCM for each TF bin, plus peak picking in two 1D histograms (one for each spherical coordinate).
@@
   \centering
-  \centerline{\includegraphics[width=\columnwidth]{Figures/SELD/seld_complexity_v4.png}}
+  \centerline{\includegraphics[width=0.8\columnwidth]{Figures/SELD/seld_complexity_v4.png}}
   \caption{DCASE 2020 Task 3 submissions: complexity versus SELD score.}

### 2020-07-05 Chapters/ScientificBackground.tex
beta_v2

@@
 
-\begin{table}[h]
+\begin{table}[hb!]
 \centering
@@
 \end{table}
+
+\newpage
 
@@
 
+\newpage
+
 One of the properties of the spherical harmonics is orthonormality on the sphere surface:
@@
 \end{equation}
-\todo{is that actually valid? or Y should be the complex-valued spherical harmonics?}
+%\todo{is that actually valid? or Y should be the complex-valued spherical harmonics?}
 
@@
 
-\begin{figure}[hbt]
+\begin{figure}[ht]
 \begin{center}
@@
 
+\newpage
 By using the model from Eq.~\ref{eq:a2b}, the maximum spherical harmonic order $N$ that can be retrieved with negligible spatial aliasing depends on the number of microphone capsules \cite{moreau20063d}:
@@
 
+\newpage
 Furthermore, the sphere radius $R$ has also an effect on the operational bandwidth of the microphone.
@@
 Ambisonics is a spatial sound recording and playback technology initially developed during the 1970s \cite{gerzon1973periphony}, and further expanded into its modern formulation around the 2000s \cite{daniel2000representation}.
-Ambisonics is based on the idea of decomposing a sound field into its spherical harmonic representation.
+Ambisonics is based on the idea of decomposing a sound field into its spherical harmonic representation. \\
 
@@
 
+\newpage
 \subsubsection{Ambisonic encoding}
@@
 Let us consider a sound field composed of a point sound source $S$ located in far-field at the angular position $\Omega_s$.
-The sound pressure at the coordinate origin $P$ can be expressed in terms of the spherical harmonic expansion of order $N$ as: \todo{check equation, find references, how to explain the domain?
-extend also to multiple sources by superposition}
+The sound pressure at the coordinate origin $P$ can be expressed in terms of the spherical harmonic expansion of order $N$ as:
+%\todo{check equation, find references, how to explain the domain? extend also to multiple sources by superposition}
 \begin{equation}
@@
 
-Although the term \textit{B-Format} was initially introduced as an alternative name for first-order ambisonic signals [gerzon, tesis de daniel], it is nowadays common to use it as a synonim of ambisonic signals, without any order restriction.
+Note that, because of the superposition principle, a sound field composed of several different point sources can be broken down to the addition of the individual contributions.
+
+Although the term \textit{B-Format} was initially introduced as an alternative name for first-order ambisonic signals \cite{daniel2000representation}, it is nowadays common to use it as a synonim of ambisonic signals, without any order restriction.
 We will use the latter acception in what follows.
@@
 The second one is the \textit{recording} with a spherical microphone array, followed by the aforementioned domain conversion.
-\\
 
@@
 Due to the orthonormal property showed in Eq.~\ref{eq:orthonormality}, they follow the \textit{fully 3d normalized} or \textit{N3D} channel normalization convention.
-\todo{what about the 1/sqrt(4pi)???}.
+%\todo{what about the 1/sqrt(4pi)???}.
 
-Alternatively, the \textit{Schmidt 3d semi-normalized} or \textit{SN3D} [daniel] convention is also of widespread usage.
+Alternatively, the \textit{Schmidt 3d semi-normalized} or \textit{SN3D} \cite{daniel2000representation} convention is also of widespread usage.
 The conversion between \textit{N3D} and \textit{SN3D} is driven by the following expression:
@@
 For instance, \textit{N3D} is the most mathematically straightforward, and spherical harmonics defined in that way can be directly used for both encoding and decoding (as in Eqs~\ref{eq:encoding} and Eq.~\ref{eq:decoding}) -- however, from a sound engineer point of view, other normalization schemes with maximum values below the unity might be preferred, such as \textit{SN3D}.
-Besides this, \textit{FuMa} has been historically the default normalization \cite{gerzon1985ambisonics}, while the more modern \textit{N3D} and \textit{SN3D} were popularized after \cite{daniel2000representation}.
+Besides this, \textit{FuMa} has been historically the default normalization \cite{gerzon1985ambisonics}, while the more modern \textit{N3D} and \textit{SN3D} were popularized after J.
+Daniel's work \cite{daniel2000representation}.
 \\
 
-\begin{figure}[hbt!]
-  \includegraphics[width=\textwidth]{Figures/ScientificBackground/normalization.png}
+\begin{figure}[ht!]
+  \includegraphics[width=1.1\textwidth]{Figures/ScientificBackground/normalization.png}
   \caption{Maximum value of each ambisonic channel up to order 5, for all different normalization schemes. Image from \cite{carpentier2017ambisonic}.}
@@
 
+\newpage
 In practice, there exist two main combinations of channel normalization and ordering schemes:
@@
 
+\newpage
 Let us consider a \textit{SN3D}-normalized first-order ambisonic signal in time-frequency domain, $S_n^m(k, n)$. 
@@
 \end{equation}
-where $\rho_0$ is the mean density of the medium, and $c$ is the speed of sound.\\
+where $\rho_0$ is the mean medium density, and $c$ is the sound speed.
 
@@
 \end{equation}
-where $^*$ represents the complex conjugate operator. \\
+where $^*$ represents the complex conjugate operator.
 
@@
 	\begin{center}
-	\includegraphics[width=\textwidth]{Figures/Introduction/spectrograms.png}
+	\includegraphics[width=1.1\textwidth]{Figures/Introduction/spectrograms.png}
 	\caption{Parametric time-frequency spatial audio analysis of a first order ambisonic recording.
@@
 \section{Spatial Coherence Analysis}
-\todo{put this chapter in context or something}\\
+%\todo{put this chapter in context or something}\\
 
@@
 \textit{Magnitude Squared Coherence} (MSC) \cite{elko_spatial_2001} between two frequency-domain signals $S_1$ and $S_2$, as a function of the
-\textit{wavenumber} $k$ and the microphone distance $r$:
+\textit{wavenumber} $k$ and the capsule distance $r$:
 \begin{equation}
@@
 
+\newpage
 For first-order differential microphones, there is a closed-form expression for the numerator and denominator of Eq.~(\ref{eq:MSCdir}):
@@
 \end{equation}
-where $x_i$, $y_i$ and $z_i$ are the cartesian coordinates of the wave incidence angle $\Omega_i$. \todo{check}.
+where $x_i$, $y_i$ and $z_i$ are the cartesian coordinates of the wave incidence angle $\Omega_i = (\varphi_i, \vartheta_i)$.
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[t!]
 	\begin{center}
@@
 \end{equation}
-where $h_D(t)$ and $h_R(t)$ represent the \textit{direct} (direct path plus early reflections) and \textit{reverberant} (late reverberation) components of the RIR. respectively.\\
+where $h_D(t)$ and $h_R(t)$ represent the \textit{direct} (direct path plus early reflections) and \textit{reverberant} (late reverberation) components of the RIR. respectively.
 
@@
 However, it is generally assumed that the late reverberation is fixed for a given room, regardless of source/receiver positions.
-\\
 
@@
 In the context of spatial perception, it has been shown that early reflections help the localization and externalization of sources \cite{rudrich2019improving}, while the late reverberation is associated with a spaciousness perception of the room \cite{begault20003}.
-\\
 
@@
 
-\todo{RIR SIMULATION: IMAGE METHOD?}
+%\todo{RIR SIMULATION: IMAGE METHOD?}
 
@@
 
+\newpage
 When dealing with multichannel room impulse responses, as it is the case in ambisonics, the multichannel reverberant signal $x_m(t)$ is obtained by the convolutive mixture of each RIR channel independently:
@@
 
-\todo{isotropic noise}
+%\todo{isotropic noise}
 

### 2020-07-05 manuscript.tex
beta_v2

@@
 
+max
+
 rafa
@@
 %{\bf Preface }
-\textbf{Preface} \\
+\section*{\Large \sffamily  Preface}
 
@@
 The industrial nature of the doctorate program implies a tendency towards the innovative application of the findings and conclusions generated during the development of the thesis. 
-Therefore, one of the targets of the research performed in the context of the thesis is to ensure the applicability to real-world problems, scenarios and technologies. 
-Moreover, there is also an stress on research reproducibility.
+Therefore, one of the targets of the research performed in the context of this thesis is the stress on real-world problems, scenarios and technologies, with the aim of developing methods that could be easily adapted into applications with a high level of technology readiness.
+Moreover, there has been also an stress on research reproducibility.
 This has not been motivated only by an academic perspective; but also by the open research and open innovation paradigms present in our today's society.\\
@@
 The work presented here would not have been possible without the contributions from an excellent network of researchers.
-Apart from internal collaborations (UPF and Eurecat), two of the main contributions in the thesis followed two research visits: to the Audio Research Group from Tampere University, Finland (Chapter~\ref{chap:rt60}), and to the Signal Processing Group at the FORTH - Foundation for Research and Technology Hellas, Heraklion, Greece, (Chapter~\ref{chap:coherence}).
+Apart from internal collaborations (UPF and Eurecat), two of the main contributions in the thesis followed two research visits to internationally recognized institutions: the Audio Research Group from Tampere University, Finland (Chapter~\ref{chap:rt60}), and the Signal Processing Group at the FORTH - Foundation for Research and Technology Hellas, Heraklion, Greece, (Chapter~\ref{chap:coherence}).
 
-The alphabetical list of academic collaborators is: Eduardo Fonseca, Emilia G\'omez, Rafael Ib\'añez-Usach,  Julien de Muynke, Archontis Politis, Xavier Serra and Nikolaos Stefanakis.
+The alphabetical list of academic collaborators is: Eduardo Fonseca, Emilia G\'omez, Rafael Ib\'añez-Usach,  Julien de Muynke, Archontis Politis, Xavier Serra, and Nikolaos Stefanakis.
 

### 2020-07-05 Chapters/Conclusions.tex
beta_v3

@@
 Such application would probably imply a re-implementation at the software level, intended for compatibility with the workflows and formats used in the VR/AR content production industry.
+
+\newpage
+\section{List of Contributions}
+
+In the following list we show the main scientific contributions, as main author, related to this dissertation:
+
+\begin{itemize}
+
+	\item Peer-reviewed journal articles
+
+	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
+	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
+	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
+
+	\item Peer-reviewed conference articles
+
+	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
+	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
+	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}\\
+
+	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
+	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+
+	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
+	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
+	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+
+	\textbf{"Ambiscaper: A tool for automatic generation and annotation of reverberant ambisonics sound scenes".}\\
+	\underline{A. P\'erez-L\'opez}.
+	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
+
+	\item Conference engineering briefs
+
+	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
+	\underline{A. P\'erez-L\'opez} and J. De Muynke.
+	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}\\
+
+	\textbf{"pysofaconventions, a Python API for SOFA".}
+	\underline{A. P\'erez-L\'opez}.
+	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+
+	\textbf{"A Python library for multichannel acoustic signal processing".}
+	\underline{A. P\'erez-L\'opez} and A. Politis.
+	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+
+\end{itemize}
+
+Moreover, although not strictly aligned with the research direction of this thesis, the author has supervised the following publication:\\
+
+	 \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
+	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+
+\newpage
+\section{List of Software Resources}
+
+As a result of the development of this thesis, the following open-source libraries and repositories have been produced.
+All of them are freely available through the author's GitHub page \cite{andresgithub} under open-source licenses.
+
+\begin{itemize}
+
+	\item Software tools:
+
+	\href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library} 
+		Tools for acoustic simulation and spherical array processing. \\
+
+	\href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions} Implementation of the SOFA convention in Python.\\
+
+	\href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}	 Ambisonic SOFA Convention proposal.\\
+
+	\href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper} Tool for automatic generation of annotated ambisonic datasets.\\
+
+	\item Method implementations:
+
+	\href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation} Contains the code implementing the methods described in Chapter~\ref{chap:rt60}.\\
+
+	\href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020} Full code implementing the SELD algorithm from Chapter~\ref{chap:seld2019}.\\
+
+	\href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3} 
+		Implementation of the method submitted to DCASE2019 (not included in this thesis). \\
+
+\end{itemize}

### 2020-07-05 manuscript.tex
beta_v3

--- 2020-06-29 manuscript.tex
@@
 %%%%%% Thanks
-\noindent {\Large \sffamily Thanks} thanks to....  \todo{thanks}
+\noindent {\Large \sffamily Thanks}\\
+
+First of all, I must thank my supervisors, Emilia and Adan, who have made possible the realization of this journey.
+
+Thanks to all my friends from the Audio Team at Eurecat: Umut, Julien, Tim, Niklas, Toni, Gerard, Andrés, John, Stefan, Marc.
+And also to  the people from Image and the Visualization teams.
+I spent last four years working with the finest of the companies.
+
+Thanks also to all the great people I met at the MTG, including Pritish, Xavier, and specially Edu for the logistic and intellectual help.
+I would like to give public acknowledgment to Xavier Serra for his successful task of leading and gathering such unique international group of talented people.
+
+Special thanks to Nikos and all the Signal Processing Lab crew for warmly welcoming me in Heraklion, and helping me in many different ways in your lovely place. I hope to come back there soon.
+
+Many thanks to Archontis and Tuomas for the visit opportunity, and for all the  facilities they gave me in my stay in this amazing place called Tampere.
+I take the opportunity to say hello and thanks to Kostas, Natalya, Annamaria, Toni, and all the nice people there.
+
+Thanks to Max, who has been always very receptive and supportive on numerous occasions, and has indirectly contributed to the success of this project. 
+Thanks also to Rafa, who decided to get on board for some unknown reason.
+
+Thanks to all the friends I made and preserved in Barcelona in the last four years.
+Special thanks to my musician family from Balkumbia, Fanfarrai and Telewawachi Kilili; I learned a lot from you.
+Special acknowledgements for the good times with the people in Can Collet (best farm-school in town) and the Valencian oldies.
+
+Thanks a lot to Chedar for his company in these final steps. And, of course, million thanks to my family and to Marta for their unconditional love and support.\\
+
+To all of you: thanks.
 
@@
 %PREFACE. 
-{\bf Preface \todo{TODO}}
+%{\bf Preface }
+\section*{\Large \sffamily  Preface}
+
+The present thesis has been carried out within the context of the Industrial Doctorate program from the Catalan government, as a collaboration between the Music Technology Group of the Pompeu Fabra University, and the Multimedia Unit of Eurecat, the Catalan Technology Center.
+It has been jointly supervised by Dr. Emilia Gómez and Dr. Adan Garriga.
+
+The industrial nature of the doctorate program implies a tendency towards the innovative application of the findings and conclusions generated during the development of the thesis. 
+Therefore, one of the targets of the research performed in the context of this thesis is the stress on real-world problems, scenarios and technologies, with the aim of developing methods that could be easily adapted into applications with a high level of technology readiness.
+Moreover, there has been also an stress on research reproducibility.
+This has not been motivated only by an academic perspective; but also by the open research and open innovation paradigms present in our today's society.\\
+
+The work presented here would not have been possible without the contributions from an excellent network of researchers.
+Apart from internal collaborations (UPF and Eurecat), two of the main contributions in the thesis followed two research visits to internationally recognized institutions: the Audio Research Group from Tampere University, Finland (Chapter~\ref{chap:rt60}), and the Signal Processing Group at the FORTH - Foundation for Research and Technology Hellas, Heraklion, Greece, (Chapter~\ref{chap:coherence}).
+
+The alphabetical list of academic collaborators is: Eduardo Fonseca, Emilia G\'omez, Rafael Ib\'añez-Usach,  Julien de Muynke, Archontis Politis, Xavier Serra, and Nikolaos Stefanakis.
 
@@
 \input{Chapters/Conclusions}
+
+%\begin{appendices}
+%\appendixpage
+%\noappendicestocpagenum
+%\addappheadtotoc
+
+%\chapter{List of Contributions}
+%
+%In the following list we show the main scientific contributions, as main author, related to this dissertation:
+%
+%\begin{itemize}
+%
+%	\item Peer-reviewed journal articles
+%	
+%	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
+%	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
+%	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
+%
+%	\item Peer-reviewed conference articles
+%
+%	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
+%	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
+%	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}\\
+%
+%	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
+%	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
+%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+%	
+%	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
+%	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
+%	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+%	
+%	\textbf{"Ambiscaper: A tool for automatic generation and annotation of reverberant ambisonics sound scenes".}\\
+%	\underline{A. P\'erez-L\'opez}.
+%	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
+%	
+%	
+%	\item Conference engineering briefs
+%
+%
+%	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
+%	\underline{A. P\'erez-L\'opez} and J. De Muynke.
+%	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}\\
+%	
+%	\textbf{"pysofaconventions, a Python API for SOFA".}
+%	\underline{A. P\'erez-L\'opez}.
+%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+%	
+%	\textbf{"A Python library for multichannel acoustic signal processing".}
+%	\underline{A. P\'erez-L\'opez} and A. Politis.
+%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+%	
+%\end{itemize}
+%
+%\newpage
+%
+%Moreover, although not strictly aligned with the research direction of this thesis, the author has supervised the following publication:\\
+%
+%	 \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
+%	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
+%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+%
+%
+%
+%\chapter{List of Software Resources}
+%
+%
+%As a result of the development of this thesis, the following open-source libraries and repositories have been produced.
+All of them are freely available through the author's GitHub page \cite{andresgithub} under open-source licenses.
+%
+%
+%\begin{itemize}
+%
+%	\item Software tools:
+%
+%	\begin{itemize}
+%	
+%		\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library} 
+%		Tools for acoustic simulation and spherical array processing. 
+%		 
+%		
+%		\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions} Implementation of the SOFA convention in Python. 
+%	
+%		\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}	 Ambisonic SOFA Convention proposal.
+%
+%		\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper} Tool for automatic generation of annotated ambisonic datasets.
+%
+%	\end{itemize}
+%	\newpage
+%	\item Method implementations: 
+%
+%	\begin{itemize}
+%		\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation} Contains the code implementing the methods described in Chapter~\ref{chap:rt60}.
+%	
+%		\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020} Full code implementing the SELD algorithm from Chapter~\ref{chap:seld2019}.
+%	
+%		\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3} 
+%		Implementation of the method submitted to DCASE2019 (not included in this thesis). 
+%
+%	\end{itemize}
+%
+%\end{itemize}
+%
+
+%\end{appendices}
 

### 2020-07-06 Chapters/ScientificBackground.tex
minor

@@
 Spherical harmonics are continuous functions defined on the sphere surface.
-Due to their mathematical properties, any spherical function can be decomposed as a combination of spherical harmonics, in what is known as the \textit{Spherical Harmonics Expansion} \cite{jarrett2017theory}.
+Due to their mathematical properties, any continuously differentiable spherical function can be decomposed as a combination of spherical harmonics, in what is known as the \textit{Spherical Harmonics Expansion} \cite{jarrett2017theory}.
 

### 2020-07-06 manuscript.tex
minor

@@
 
-Thanks a lot to Chedar for his company in these final steps. And, of course, million thanks to my family and to Marta for their unconditional love and support.\\
+Thanks a lot to Chedar for his company in the final steps. And, of course, million thanks to my family and to Marta for their unconditional love and support, powering all the way up here. \\
 
@@
 \section*{\Large \sffamily Abstract}
-
-%Recent years have witnessed a rise in immersive multimedia content, as a result of the developments in Virtual and Augmented Reality, where spatial audio plays a crucial role for the creation of a realistic experience.
-%Ambisonic microphones allow the recording of sound scenes while preserving their spatial characteristics.
-For that reason, ambisonic has turned into the \textit{de facto} standard for immersive audio transmission.
-Yet, few production tools exploit all the information available in such media.
-%In the present thesis we raise the research question of obtaining meaningful information from ambisonic recordings, based on the parametric analysis of the sound field.
-%In order to answer the question we have developed different state-of-the-art methods, which cover a broad range of applications within the parametric audio analysis scope. 
-%Specifically, we present novel contributions in the scope of blind reverberation time estimation, diffuseness estimation, and sound event localization and detection.
-Furthermore, several software tools developed for dataset generation and storage are also presented.
 
@@
 
-%Els micròfons ambisonics permeten gravar escenes sonores preservant les seues característiques espacials.
-A causa d'això, ambisonics s'ha convertit en l'estàndard de facto per a la transmissió d'àudio immersiu.
-No obstant això, encara son poques les eines de producció que permeten explotar tota la informació disponible en aquest mitjà.
-%En la present tesi plantegem la pregunta de com obtenir informació rellevant de gravacions ambisonics de manera automàtica, basant-nos en l'an\`alisi param\`etric del camp ac\'ustic. 
-%%d'una forma que siga significativa per als productors.
-%Per a això, hem desenvolupat una sèrie de mètodes i eines a nivell de l'estat de l'art, 
-%que cobreixen un ampli espectre d'aplicaciones a l'\`ambit de l'an\`alisi param\`etric del so. 
-%%que s'apliquen a diferents nivells de la cadena de producció de so immersiu; cadascun d'aquests elements representa una contribució nova a l'estat de l'art en aquest camp.
-%
-%%En los últimos años hemos asistido al auge del contenido multimedia immersivo, auspiciado por los avances en Realidad Virtual y Aumentada.
-En este medio, el sonido espacial juega un papel muy importante para la creación de una sensación de realismo.
-%%
-%%Los micrófonos ambisonics permiten grabar escenas sonoras preservando sus características espaciales.
-%%Debido a esto, ambisonics se ha convertido en el estándar de facto para la transmisión de audio immersivo.
-Sin embargo, todavía hay pocas herramientas de producción que permitan explotar toda la información disponible en este medio.
-%%
-%%En esta tesis planteamos la pregunta de cómo obtener información relevante de grabaciones ambisonics de manera automática, de una forma significativa para los productores.
-%%
-%%Para ello, hemos desarrollado una serie de métodos y herramientas que se aplican a diferentes niveles de la cadena de producción de sonido immersivo; cada uno de estos elementos representa una contribución novedosa al estado del arte en ese campo.
-
 \vspace*{\fill}
@@
 %PREFACE. 
-%{\bf Preface }
 \section*{\Large \sffamily  Preface}
@@
 
-%\begin{appendices}
-%\appendixpage
-%\noappendicestocpagenum
-%\addappheadtotoc
-
-%\chapter{List of Contributions}
-%
-%In the following list we show the main scientific contributions, as main author, related to this dissertation:
-%
-%\begin{itemize}
-%
-%	\item Peer-reviewed journal articles
-%	
-%	\textbf{"Analysis of spherical isotropic noise fields with an A-Format tetrahedral microphone"}.
-%	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
-%	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
-%
-%	\item Peer-reviewed conference articles
-%
-%	\textbf{"Blind reverberation time estimation from ambisonic recordings"}.
-%	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
-%	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}\\
-%
-%	\textbf{"PAPAFIL: a low complexity sound event localization and detection method with parametric particle filtering and gradient boosting".}
-%	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
-%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
-%	
-%	\textbf{"A hybrid parametric-deep learning approach for sound event localization and detection".}
-%	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
-%	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
-%	
-%	\textbf{"Ambiscaper: A tool for automatic generation and annotation of reverberant ambisonics sound scenes".}\\
-%	\underline{A. P\'erez-L\'opez}.
-%	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
-%	
-%	
-%	\item Conference engineering briefs
-%
-%
-%	\textbf{"Ambisonics directional room impulse response as a new convention of the spatially oriented format for acoustics".}
-%	\underline{A. P\'erez-L\'opez} and J. De Muynke.
-%	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}\\
-%	
-%	\textbf{"pysofaconventions, a Python API for SOFA".}
-%	\underline{A. P\'erez-L\'opez}.
-%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
-%	
-%	\textbf{"A Python library for multichannel acoustic signal processing".}
-%	\underline{A. P\'erez-L\'opez} and A. Politis.
-%	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
-%	
-%\end{itemize}
-%
-%\newpage
-%
-%Moreover, although not strictly aligned with the research direction of this thesis, the author has supervised the following publication:\\
-%
-%	 \textbf{"Sound event localization and detection based on CRNN using dense rectangular filters and channel rotation data augmentation".}
-%	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
-%	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
-%
-%
-%
-%\chapter{List of Software Resources}
-%
-%
-%As a result of the development of this thesis, the following open-source libraries and repositories have been produced.
-All of them are freely available through the author's GitHub page \cite{andresgithub} under open-source licenses.
-%
-%
-%\begin{itemize}
-%
-%	\item Software tools:
-%
-%	\begin{itemize}
-%	
-%		\item \href{https://github.com/andresperezlopez/masp}{masp: Multichannel Acoustic Signal Processing library} 
-%		Tools for acoustic simulation and spherical array processing. 
-%		 
-%		
-%		\item \href{https://github.com/andresperezlopez/pysofaconventions}{pysofaconventions} Implementation of the SOFA convention in Python. 
-%	
-%		\item \href{https://github.com/andresperezlopez/AmbisonicsDRIR}{AmbisonicsDRIR}	 Ambisonic SOFA Convention proposal.
-%
-%		\item \href{https://github.com/andresperezlopez/ambiscaper}{AmbiScaper} Tool for automatic generation of annotated ambisonic datasets.
-%
-%	\end{itemize}
-%	\newpage
-%	\item Method implementations: 
-%
-%	\begin{itemize}
-%		\item \href{https://github.com/andresperezlopez/ambisonic_rt_estimation}{ambisonic\_rt\_estimation} Contains the code implementing the methods described in Chapter~\ref{chap:rt60}.
-%	
-%		\item \href{https://github.com/andresperezlopez/DCASE2020}{DCASE2020} Full code implementing the SELD algorithm from Chapter~\ref{chap:seld2019}.
-%	
-%		\item \href{https://github.com/andresperezlopez/DCASE2019_task3}{DCASE2019\_task3} 
-%		Implementation of the method submitted to DCASE2019 (not included in this thesis). 
-%
-%	\end{itemize}
-%
-%\end{itemize}
-%
-
-%\end{appendices}
-
 %ADD BIBLIOGRAPHY

### 2020-07-28 Chapters/CoherenceEstimation.tex
minor

@@
 
-In this Chapter, we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
+In this Chapter, adapted from a previous academic publication \cite{stefanakis2019analysis},  we study diffuseness estimation by subjecting a tetrahedral microphone array to spherically isotropic noise fields.
 The motivation for this work is, first, that tetrahedral arrays are a well known type of microphone arrays, which have today become popular for applications related to Virtual and Augmented Reality. 

### 2020-07-28 Chapters/Conclusions.tex
minor

@@
 	\underline{A. P\'erez-L\'opez} and N. Stefanakis.
-	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.}
+	\textit{The Journal of the Acoustical Society of America 146.4 (2019): EL329-EL334.} \cite{stefanakis2019analysis}.
 
@@
 	\underline{A. P\'erez-L\'opez}, A. Politis and E. G\'omez.
-	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.}\\
+	Submitted to \textit{IEEE 22nd International Workshop on Multimedia Signal Processing, 2020.} \cite{rt}.
 
@@
 	\underline{A. P\'erez-L\'opez} and R. Ibañez-Usach.
-	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)} \cite{papafil}.
 
@@
 	\underline{A. P\'erez-L\'opez}, E. Fonseca and X. Serra.
-	In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)}.
+In \textit{Proceedings of the Detection and Classification of Acoustic Scenes and Events 2019 Workshop (DCASE2019)} \cite{perez2019hybrid}.
 
@@
 	\underline{A. P\'erez-L\'opez}.
-	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018.}
+	In \textit{Proceedings of the 16th International Workshop on Acoustic Signal Enhancement (IWAENC). IEEE, 2018} \cite{perez2018ambiscaper}.
 
@@
 	\underline{A. P\'erez-L\'opez} and J. De Muynke.
-	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018.}\\
+	In \textit{Proceedings of the 144th Audio Engineering Society Convention. Audio Engineering Society, 2018} \cite{perez2018ambisonics}.
 
@@
 	\underline{A. P\'erez-L\'opez}.
-	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020} \cite{pysofaconventions}.
 
@@
 	\underline{A. P\'erez-L\'opez} and A. Politis.
-	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020.}\\
+	In \textit{Proceedings of the 148th Audio Engineering Society Convention. Audio Engineering Society, 2020} \cite{perez2020python}.
 
@@
 	F. Ronchini, D. Arteaga and \underline{A. P\'erez-L\'opez}.
-	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)}.\\
+	Submitted to \textit{Detection and Classification of Acoustic Scenes and Events 2020 Workshop (DCASE2020)} \cite{ronchini}. \\
 

### 2020-07-28 Chapters/DataGeneration.tex
minor

@@
 
-The Multichannel Acoustic Signal Processing (\textit{MASP}) is a Python library consisting of a collection of methods related with acoustics and microphone array processing.
+The Multichannel Acoustic Signal Processing (\textit{MASP}) is a Python library consisting of a collection of methods related with acoustics and microphone array processing \cite{perez2020python}.
 The library is mostly a transcoding from several Matlab libraries by A. Politis\cite{politis2016microphone, github_politis}. 
@@
 
-Given that the described situation meets all requirements, the \textit{Ambisonics Directional Room Impulse Response} (AmbisonicsDRIR) convention has been therefore proposed as SOFA convention.\\
+Given that the described situation meets all requirements, the \textit{Ambisonics Directional Room Impulse Response} (AmbisonicsDRIR) convention has been therefore proposed as SOFA convention \cite{perez2018ambisonics}.\\
 
@@
 
-The library \textit{pysofaconventions} has been created with the aim to provide an alternative to the existing Matlab/Octave and C/C++ implementations. 
+The library \textit{pysofaconventions} has been created with the aim to provide an alternative to the existing Matlab/Octave and C/C++ implementations \cite{pysofaconventions}.
 For ease of installation, it is integrated in the standard python package manager, \textit{pip}.
@@
 
-\textit{AmbiScaper} is a tool designed to provide a flexible way of creating complex ambisonics sound scenes and their associated groundtruth annotations, to be used in the context of source localization and separation algorithms.
+\textit{AmbiScaper} \cite{perez2018ambiscaper} is a tool designed to provide a flexible way of creating complex ambisonics sound scenes and their associated groundtruth annotations, to be used in the context of source localization and separation algorithms.
 \textit{AmbiScaper} offers a high level control of the sound scene parameters, and provides a simple interface for the creation of large datasets with custom characteristics.

### 2020-07-28 Chapters/ReververationTimeEstimation.tex
minor

@@
 On the other hand, recent years have witnessed a growing interest in immersive audio for virtual and augmented reality.
-This situation has consolidated Ambisonics \cite{zotter2019ambisonics} as the \textit{de facto} standard for spatial audio.
+This situation has consolidated Ambisonics as the \textit{de facto} standard for spatial audio.
 Dedicated spherical microphone arrays have reached the market in last years; their multichannel nature makes possible spatial manipulations that
@@
 
-The rest of the chapter is organized as follows.
+The rest of the Chapter is organized as follows.
 Section~\ref{sec:signalmodel} introduces the nomenclature and the signal model.
@@
 Finally, a conclusion is presented in Section~\ref{sec:conclusion}.
+
+The research presented in this Chapter has been submitted for publication at the IEEE 22nd International Workshop on Multimedia Signal Processing \cite{rt}.
 
@@
 \begin{figure*}[htbp]
-	\centerline{\includegraphics[width=1.3\textwidth]{Figures/ReverberationTimeEstimation/96.png}}
+	\centerline{\includegraphics[width=1.1\textwidth]{Figures/ReverberationTimeEstimation/96.png}}
 	\caption{Experiment results for \textit{speech} (left column) and \textit{drums} (right column) datasets. Estimation error computed for each audio clip.}

### 2020-07-28 Chapters/SELD.tex
minor

@@
 
-Finally, it is important to remark that the method described here is the continuation of an algorithm presented at the DCASE 2019 Challenge \cite{perez2019hybrid}; both algorithms share a common structure and a similar approach to the SELD problem.
-However, the current proposal tries to solve some of the problems identified on our early approach, mainly related with a low frame recall derived from a naive approach to event segmentation.
-Moreover, the current method also presents a complexity reduction, regarding the single-source
-DOA estimation and the classifier.
+Finally, it is important to remark that the method described here has been reported in a conference paper submitted to the DCASE Workshop 2020 \cite{papafil},  which in turn describes the improvement of an algorithm presented at the DCASE 2019 Challenge \cite{perez2019hybrid}.
+
+While both algorithms share a common structure and a similar approach to the SELD problem, the current proposal tries to solve some of the problems identified on our early approach, mainly related with a low frame recall derived from a naive approach to event segmentation.
+Moreover, the current method also presents a complexity reduction, regarding the single-source DOA estimation and the classifier.
 
